### 2014-06-30 Paper/BOAST.tex
Added skeleton for Journal Paper. minimalistic.

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of Parallel and Distributed Computing}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

\title{BOAST: a Metaprogramming Framework to Write Portable and Efficient Computing Kernels for HPC Applications}

%% or include affiliations in footnotes:
\author[mymainaddress]{Brice Videau}
\ead{brice.videau@imag.fr}

\address[mymainaddress]{LIG/CNRS}

\begin{abstract}
The Abstract.
\end{abstract}

\begin{keyword}
BOAST \sep HPC \sep SPECFEM3D \sep BigDFT
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

Human ressources cost

\cite{videau2013boast}

\section{Problematic and State of the Art}

  \subsection{Performance Portability}

    An increasing concern for HPC Applications

  \subsection{One extreme case: the Mont-Blanc prototype}

  \subsection{State of the Art: Application Autotuning}

\section{Use Cases}

  \subsection{BigDFT}

  \subsection{SPECFEM3D}

\section{Using Code Generation in Application Autotuning}

  \subsection{General Principles}

  \subsection{Our Contribution: BOAST}
     + dumper pour tests de non régression

\section{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}

\section{Porting SPECFEM3D to OpenCL using BOAST}

\section{Feedback and Future Works}

\section{Conclusion}

\section*{References}

\bibliography{BOAST}

\end{document}

### 2014-07-07 Paper/BOAST.tex
update to paper

@@
 
-\title{BOAST: a Metaprogramming Framework to Write Portable and Efficient Computing Kernels for HPC Applications}
+\title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
 
@@
 \begin{keyword}
-BOAST \sep HPC \sep SPECFEM3D \sep BigDFT
+Code Generation \sep Portability \sep High Performance Computing \sep
+Autotuning \sep Non Regression Testing
 \end{keyword}
@@
 
-Human ressources cost
+Porting and tuning HPC applications to new platforms is tedious and costly
+in term of human resources. Portability efforts are often lost when
+migrating to a new architecture, or code lose maintainability because
+several versions of the code coexist, usually with a lot of duplication.
+Thus productivity of porting and tuning efforts is low as a huge fraction
+of those developments are never used after the platform they were intended
+for is decommissioned.
 
-\cite{videau2013boast}
+\section{Background and Motivation}
 
-\section{Problematic and State of the Art}
+  \subsection{Evolution of HPC Architectures}
 
-  \subsection{Performance Portability}
+Evolution of HPC Architectures is rapid and also diverse: in the last 5
+years no less than 6 architectures have been number one in the Top500:
+\begin{itemize}
+\item Intel Processor + Xeon Phi (Tianhe-2)
+\item AMD Processor + NVIDIA GPU (Titan)
+\item IBM BlueGene/Q (Sequoia)
+\item Fujitsu SPARC64 (K computer)
+\item Intel Processor + NVIDIA GPU (Tianhe-1)
+\item AMD Processor (Jaguar)
+\end{itemize}
+Being able to efficiently use those architectures on such a small
+time-frame is challenging.
 
-    An increasing concern for HPC Applications
+The race to exascale is not going to simplify the environment. All of the
+above architectures can be considered. For instance European FP7 project Deep
+considers using Accelerators (XEON Phi) while the European FP7 project
+Mont-Blanc considers using low-power embedded processor with integrated
+GPU. Network architectures are also part of the architecture and can be
+very diverse.
+
+  \subsection{HPC Applications}
+
+Developed by physicist, chemist, meteorologist. Usually in FORTRAN for
+historical and performance reason. Codes can be quite huge (several
+thousands lines of code) with lots of functionalities. Nonetheless they are
+usually based on computing kernels. Computing kernels are resource
+intensive well defined parts of a program.
 
@@
   \subsection{Our Contribution: BOAST}
+  \cite{videau2013boast}
      + dumper pour tests de non régression
@@
 
-\section{Feedback and Future Works}
+\section{Related Work}
 
-\section{Conclusion}
+\section{Conclusion and Future Works}
 

### 2014-07-07 Paper/BOAST.tex
update to paper

@@
 usually based on computing kernels. Computing kernels are resource
-intensive well defined parts of a program.
+intensive well defined parts of a program with a limited set of input and
+output. Those kernels are the time consuming part of an HPC application.
 
-  \subsection{One extreme case: the Mont-Blanc prototype}
-
-  \subsection{State of the Art: Application Autotuning}
+  \subsection{Performance Portability of Computing Kernels}
 
@@
 
+  \subsection{State of the Art: Application Autotuning}
+
 \section{Conclusion and Future Works}

### 2014-07-09 Paper/BOAST.tex
updated paper

@@
 \usepackage{lineno,hyperref}
+\usepackage{listings}
+\usepackage{color}
+\usepackage{url}
 \modulolinenumbers[5]
@@
 usually based on computing kernels. Computing kernels are resource
-intensive well defined parts of a program with a limited set of input and
-output. Those kernels are the time consuming part of an HPC application.
+intensive and well defined parts of a program with a limited set of input
+and output. Those kernels are the time consuming part of an HPC
+application. They are thus the prime target for optimization.
+
+In Section~\ref{use_cases} we will present two HPC application that we used
+as use cases: SPECFEM3D and BigDFT. They are both based on computing
+kernels.
 
@@
 
-\section{Use Cases}
+The problematic here is to obtain computing kernels that present good
+performances on the architectures encountered by the application while
+still being portable after the optimization process took place. Indeed the
+application might encounter one of the other architectures presented
+before. Investing manpower to optimize the application for a new
+architecture is reasonable, suffering hindrance from previous optimization
+work is not. Thus optimizations have to be as orthogonal as possible from
+one another so as to be easily activated and deactivated.
 
-  \subsection{BigDFT}
+If this paradigm is followed by developers then they will rapidly be
+confronted with a huge optimization space to search. They will need to
+be able to test easily the performance impact of the chosen optimizations
+without running the full application. The same reasoning implies that
+kernels should be tested for non regression without running the full
+application.
 
-  \subsection{SPECFEM3D}
+\section{BOAST: Using Code Generation in Application Autotuning}
 
-\section{Using Code Generation in Application Autotuning}
+  \cite{videau2013boast}
 
-  \subsection{General Principles}
+  \subsection{Kernel Description Language}
 
-  \subsection{Our Contribution: BOAST}
-  \cite{videau2013boast}
+  \subsection{BOAST Runtime}
      + dumper pour tests de non régression
 
-\section{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
+\section{Use Cases}
+\label{use_cases}
 
-\section{Porting SPECFEM3D to OpenCL using BOAST}
+  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
+
+    \subsubsection{BigDFT}
+
+BigDFT~\cite{bigdft} is a scientific application computing the electronic
+structure of systems using density functional
+theory (DFT). The application is mainly written in Fortran and currently
+includes some 193,000 lines of Fortran code, which accounts for more than 50 \% 
+% VANIA la plupart c'est Fortran et ensuite juste un peu plus de 50%, c'est quoi le reste?
+of the code base. It is a parallel application based on the standards
+MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports CUDA~\cite{cuda} and
+OpenCL~\cite{opencl}.  As it is used on systems that may
+have very different architectures, it is of great importance to be able to
+optimize and run the application according to the specific underlying platform.
+
+BigDFT is characterized by its extensive use of
+convolution operators~\cite{nussbaumer1982fast} applied on large arrays of
+data.  One example of a specific convolution, called
+MagicFilter~\cite{Genovese|CAS2010}, can be seen in Listing~\ref{lst:conv_example}. 
+It applies a filter $filt$ to the data set $in$ and then stores the result in the data set $out$ with a
+transposition~\cite{Goedecker1993}.
+
+\lstset{	language=C, 
+		basicstyle=\scriptsize, 
+		backgroundcolor=\color{white}, 
+		frame=single, 
+		captionpos=b,
+		caption={MagicFilter}, 
+		label={lst:conv_example}}
+
+\begin{lstlisting}
+double filt[16] = {F0, F1, ... , F15};
+void magicfilter(int n, int ndat, double* in, double* out){
+  double temp;
+  int m;
+  for( j = 0; j < ndat; j++) {
+    for( i = 0; i < n; i++) {
+      temp = 0;
+      for( k = 0; k < 16; k++) {
+        m = (i-7+k)%n
+        temp += in[m + j*n] * filt[k];
+      }
+      out [j + i*ndat] = temp ;
+    } 
+  }
+} 
+\end{lstlisting}
+
+As we can see, there are three nested loops working on arrays whose sizes
+vary.  Various optimizations can be applied to this treatment and
+may focus on the loop structure, as well as on the size of the data.  In this
+paper we focus on this particular convolution and explore different
+optimizations whose nature is presented in detail in the next section.
+
+  \subsection{Porting SPECFEM3D to OpenCL using BOAST}
+
+    \subsubsection{SPECFEM3D}
 

### 2014-07-11 Paper/BOAST.tex
bugfix
WIP paper with JFM.

@@
 \ead{brice.videau@imag.fr}
+\author{Kevin Pouget}
+\author{Luigi Genovese}
+\author{Thierry Deutsch}
+\author{Dimitri Komatitsch}
+\author{Jean-François Méhaut}
 
@@
 \begin{keyword}
-Code Generation \sep Portability \sep High Performance Computing \sep
-Autotuning \sep Non Regression Testing
+Code Generation \sep Portability \sep Genericity \sep Productivity and Software
+Design \sep High Performance Computing \sep Autotuning \sep Non Regression
+Testing
 \end{keyword}
@@
 Porting and tuning HPC applications to new platforms is tedious and costly
-in term of human resources. Portability efforts are often lost when
+in term of human resources.
+
+Portability efforts are often lost when
 migrating to a new architecture, or code lose maintainability because
 several versions of the code coexist, usually with a lot of duplication.
+
 Thus productivity of porting and tuning efforts is low as a huge fraction
@@
 for is decommissioned.
+
+Généricité
+ et
+fonctionalités
 
@@
 Being able to efficiently use those architectures on such a small
-time-frame is challenging.
+time-frame is challenging. Stress frequency of architecture changes.
 
@@
 
-  \subsection{HPC Applications}
+transition architecture application
+
+  \subsection{Scientific Computing Applications}
 
@@
 application. They are thus the prime target for optimization.
+
+Ajouter Multiplicité des devs et donc problèmatique de maintenabilité ++
 
@@
 
+  \begin{itemize}
+  \item Abstraction of programming concepts: variables, procedures
+  \item Operator overloading: simpler syntax
+  \item Two languages in one
+  \end{itemize}
+
   \subsection{BOAST Runtime}
-     + dumper pour tests de non régression
+
+   multitarget generation
+   compilation
+   execution
+  + dumper et replay pour tests de non régression
 
@@
 
+    \subsubsection{A Generic Convolution Library}
+
+  characteristique use case driven
+
+  bibliotheque premier pas vers un dsl
+
   \subsection{Porting SPECFEM3D to OpenCL using BOAST}
@@
     \subsubsection{SPECFEM3D}
+
+    \subsubsection{Porting to OpenCL}
 

### 2014-07-15 Paper/BOAST.tex
update paper with JF remarks

@@
 
-Porting and tuning HPC applications to new platforms is tedious and costly
-in term of human resources.
+Porting and tuning HPC applications to new platforms is tedious and costly in
+term of human resources.
 
-Portability efforts are often lost when
-migrating to a new architecture, or code lose maintainability because
-several versions of the code coexist, usually with a lot of duplication.
+Portability efforts are often lost when migrating to a new architecture, or code
+lose maintainability because several versions of the code coexist, usually with
+a lot of duplication.
 
-Thus productivity of porting and tuning efforts is low as a huge fraction
-of those developments are never used after the platform they were intended
-for is decommissioned.
+Thus productivity of porting and tuning efforts is low as a huge fraction of
+those developments are never used after the platform they were intended for is
+decommissioned.
 
-Généricité
- et
-fonctionalités
+Genericity of HPC codes is often limited. Producing generic code in FORTRAN
+90/95 is difficult as the language is not really  fit for it.
+
+Functionality of HPC codes is tied to the previous point. Without genericity,
+adding new functionalities can be quite costly.
 
@@
 
-The race to exascale is not going to simplify the environment. All of the
-above architectures can be considered. For instance European FP7 project Deep
+The race to exascale is not going to simplify the environment. All of the above
+architectures can be considered. Network architectures are also part of the
+architecture and can be very diverse.  For instance European FP7 project DEEP
 considers using Accelerators (XEON Phi) while the European FP7 project
-Mont-Blanc considers using low-power embedded processor with integrated
-GPU. Network architectures are also part of the architecture and can be
-very diverse.
+Mont-Blanc considers using low-power embedded processor with integrated GPU.
 
-transition architecture application
+Running axisting applications on those new architectures is an open project.
+Thus, those projects have work packages dedicated to applications. Those
+workpackages are dedicated to porting and optimizing Scientific applications on
+those new architectures. In the DEEP project six applications were selected for
+porting and optimizing, eleven were selected in the Mont-Blanc project.
 
@@
 
-Ajouter Multiplicité des devs et donc problèmatique de maintenabilité ++
+Those applications are often developed by several individuals. Sometimes some of
+those developers only work a few month on the application. Maintaining optimized
+code written by someone else is quite a challenge.
 
@@
 as use cases: SPECFEM3D and BigDFT. They are both based on computing
-kernels.
+kernels and were selected as candidate applications in the Mont-Blanc project.
 
-  \subsection{Performance Portability of Computing Kernels}
+  \subsection{How Should Computing Kernels be Written?}
 
@@
 application.
+
+What we want is computing kernels that are:
+\begin{itemize}
+\item Written in a portable manner,
+\item Written in a way that raise developer productivity,
+\item Written to present good performance.
+\end{itemize}
 

### 2014-07-15 BigDFT/GenericConvolution.rb

@@ -1263,7 +1263,7 @@ class GenericConvolutionOperator1d
       }
 
       print_call = lambda {
-        BOAST::print BOAST::Case( @bc[@idim], BC::PERIODIC, lambda {
+        BOAST::print BOAST::Case( @bc, BC::PERIODIC, lambda {
           print_call_param_a.call( BC::PERIODIC )
         }, BC::GROW, lambda {
           print_call_param_a.call( BC::GROW )

### 2014-07-15 BigDFT/Kinetic_test.rb

@@ -75,7 +75,7 @@ n[2] = n3
 
 optims = GenericOptimization::new(:unroll_range => [2,6,2], :mod_arr_test => true, :tt_arr_test => true, :unrolled_dim_index_test => true)
 k = Kinetic(conv_filter, optims)
-#k.print
+k.print
 k.build(:openmp => true)
 
 bc = NArray.int(3)

### 2014-07-15 BigDFT/WaveletFilters.rb

@@ -38,5 +38,5 @@ SYM8 = [-0.0033824159510050025955,
       0.0038087520138944894631,
      -0.014952258337062199118,
      -0.00030292051472413308126,
-      0.0018899503327676891843]
+      1.8899503327676891843e-3]
 

### 2014-07-15 Paper/BOAST.tex
improved BOAST description

@@
   \begin{itemize}
-  \item Abstraction of programming concepts: variables, procedures
-  \item Operator overloading: simpler syntax
-  \item Two languages in one
+  \item Abstraction of programming concepts: variables, procedures (
+portability, productivity )
+  \item Operator overloading: simpler syntax ( productivity, portability )
+  \item Two languages in one ( productivity, performance ) 
   \end{itemize}
@@
 
-   multitarget generation
-   compilation
-   execution
-  + dumper et replay pour tests de non régression
+   multi target language generation ( performance, portability )
+   compilation ( productivity, performance )
+   execution ( productivity, performance )
+  + dumper et replay pour tests de non régression ( productivity )
 

### 2014-07-15 Paper/BOAST.tex
update to paper

@@
 
-   multi target language generation ( performance, portability )
-   compilation ( productivity, performance )
-   execution ( productivity, performance )
-  + dumper et replay pour tests de non régression ( productivity )
+  \begin{itemize}
+  \item multi target language generation ( performance, portability )
+  \item compilation ( productivity, performance )
+  \item execution ( productivity, performance )
+  \item tracer, dumper and replay for non regression tests ( productivity )
+  \end{itemize}
 
@@
 includes some 193,000 lines of Fortran code, which accounts for more than 50 \% 
-% VANIA la plupart c'est Fortran et ensuite juste un peu plus de 50%, c'est quoi le reste?
 of the code base. It is a parallel application based on the standards

### 2014-07-29 Paper/BOAST.tex
updated paper

@@
 
-Running axisting applications on those new architectures is an open project.
+Running existing applications on those new architectures is an open project.
 Thus, those projects have work packages dedicated to applications. Those
@@
 \begin{itemize}
-\item Written in a portable manner,
-\item Written in a way that raise developer productivity,
-\item Written to present good performance.
+\item Written in a \emph{portable} manner,
+\item Written in a way that raise developer \emph{productivity},
+\item Written to present good \emph{performance}.
 \end{itemize}
@@
 \section{BOAST: Using Code Generation in Application Autotuning}
+
+  Provide scientific application developers with a framework to develop and test
+application computing kernels. Two aspects should be considered:
+\begin{itemize}
+\item code description,
+\item code generation and kernel execution runtime.
+\end{itemize}
 
@@
 
-  characteristique use case driven
+  BigDFT implementation of convolution/wavelets is a reference but is limited in scope (1 wavelet family ands its derived operators).
+  Generalizing to other wavelet families long and difficult if performance are to remain good.
 
-  bibliotheque premier pas vers un dsl
+Idea:
+\begin{itemize}
+\item write a generic convolution library in BOAST
+\item BLAS like interface
+\item Provide a wide variety of building blocks
+\item Optimize thos blocks using BOAST
+\item Build either from source or directly link with binary
+\item Release complete library
+\end{itemize}
+
+  Characteristic use case driven
+
+  Convolution Library: first step toward a dsl dedicated to wavelet processing for Physics computing.
 
@@
 
+\begin{itemize}
+
+\item Wave propagation research software
+
+\item HPC application (benchmark on supercomputers)
+
+\item Complex GPU (CUDA) code, 40+ kernels.
+
+\item 80+ parameters for kernels. Difficult to validate.
+\end{itemize}
+
     \subsubsection{Porting to OpenCL}
+
+  Idea:
+\begin{itemize}
+\item write kernels in BOAST,
+\item generate kernels in CUDA,
+\item validate kernels by comparing CUDA/generated CUDA executions using comparator
+\item generate kernels in OpenCL,
+\item develop OpenCL runtime backend using previously validated OpenCL kernels
+\item validate OpenCL version by comparing CUDA/OpenCL executions using comparator
+\end{itemize}
+
+  Evaluation:
+\begin{itemize}
+\item Similar CUDA/OpenCL
+\item Factoring : more compact code
+\item Integrated non regression support
+\item Integrated Kernel performance evaluation
+\end{itemize}
 
@@
 
-  \subsection{State of the Art: Application Autotuning}
+  \subsection{Kernel Description DSL}
+
+  Orio
+
+  Halide
+
+  Spiral
+
+  \subsection{Application Autotuning}
+
+  Atlas
+
+  Orio
+
+  Halide
+
+  \subsection{Optimization Space Pruner}
+
+  Collective Mind
+
+  Adaptive Sampling Kit (ASK)
 

### 2014-07-30 Paper/BOAST.tex
Improved Paper

@@
 The problematic here is to obtain computing kernels that present good
-performances on the architectures encountered by the application while
-still being portable after the optimization process took place. Indeed the
-application might encounter one of the other architectures presented
-before. Investing manpower to optimize the application for a new
-architecture is reasonable, suffering hindrance from previous optimization
-work is not. Thus optimizations have to be as orthogonal as possible from
-one another so as to be easily activated and deactivated.
+performances on the architectures encountered by the application while still
+being portable after the optimization process took place. Indeed the application
+might encounter one of the other architectures presented before. Investing
+manpower to optimize the application for a new architecture is reasonable,
+suffering hindrance from previous optimization work is not. Thus optimizations
+have to be as orthogonal as possible from one another so as to be easily
+activated and deactivated.
 
-If this paradigm is followed by developers then they will rapidly be
-confronted with a huge optimization space to search. They will need to
-be able to test easily the performance impact of the chosen optimizations
-without running the full application. The same reasoning implies that
-kernels should be tested for non regression without running the full
-application.
+If this paradigm is followed by developers then they will rapidly be confronted
+with a huge optimization space to search. They will need to be able to test
+easily the performance impact of the chosen optimizations without running the
+full application. The same reasoning implies that kernels should be tested for
+non regression without running the full application.
 
@@
   \end{itemize}
+
+Usually computing kernels are hotspots of an HPC application, and are most of
+the time based on a loop nest. A lot of effort is dedicated to their tuning and
+the obtained result is often quite different than the original procedure.
+Several transformations can be applied to such kernels. And those optimization
+are often applied manually as compiler may fail to recognise the opportunity.
+
+There are many different loop optimization techniques~\cite{wolf1991loop}. We
+can cite loop skewing~\cite{wolfe1986loops} (derives nested loops wavefronts)
+or loop interchange~\cite{allen1984automatic} (loop variables change places).
+
+The importance of  correct loop imbrication on BLAS~\cite{lawson1979basic}
+operations is studied in~\cite{soliman2009performance}, and shows performance
+increase of a factor up to 5 when using correct loop imbrication. The importance
+of code transformation is stressed in~\cite{ye2011porting}, where a selection of GPU
+kernels are ported to CPU and optimized.
+
+BOAST kernel description language should be able to express all these
+optimizations. This gives us a set of constraints to implement in the language:
+\begin{itemize}
+\item Arbitrary number of variables have to be created and manipulated (types,
+attributes...).
+\item Procedures have to be abstracted (reunite FORTRAN and C like languages,
+attributes...).
+\item Functions must be available.
+\item Variables, constants and functions must be composed in complex
+Expressions.
+\item Basic control structures (for, while, if/else...) have to be abstracted.
+\item Powerful array management (several dimensions, transformations,
+indexing...).
+\end{itemize}
+
+In order to manipulate those abstractions we want to have a syntax resembling
+what programmers use. For instance commonly used operator have to be available
+and behave as expected. It must also be possible to differentiate an action on
+an abstraction in the context of an expression and in the context of the
+management of this expression. This is why an embedded domain specific language
+approach was selected~\cite{hudak1996building}. This allows for the coexistence
+of two languages: the host language and the DSL. In our case the DSL allows the
+description of the kernel while the host language provides the meta programming
+of the kernel. Operator overloading of the host language will provide the
+familiar syntax programmers are accustomed to.
+
+It was also important to have our construct like for loops to have a syntax
+approaching those commonly found in programming languages. For this we needed a
+language which could seamlessly pass a block of code to a function.
+Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
+capabilities as well.
+
+%Considering source-to-source transformations that we investigate with the BOAST
+%tool in this paper, we can compare ourselves to POET~\cite{POET2012}  and
+%ABCLibScript~\cite{ABCLib2006}. However, both tools apply transformations
+%to the syntaxic tree of a program while BOAST intervenes well before in the
+%transformation process. Indeed, one may first use BOAST on order to generate a
+%program (a specific syntaxic tree) and then benefit from the transformation
+%facilities offered by one of the above tools.
 

### 2014-07-31 Paper/BOAST.tex
updated paper

@@
 
+    \subsubsection{BOAST keywords}
+
+  In order to clearly differentiate what is going to be generated from what is
+related to manipulations in the host language the \emph{print} method was
+redefined in the BOAST namespace. This print method only insures that the object
+it is called on implements a public print method. Each BOAST object is
+responsible for printing itself correctly depending on the BOAST configuration
+at the time the print public method is called.
+
+    \subsubsection{BOAST Abstractions}
+
+  Variable
+
+  Expression
+
+  If
+
+  For
+
+  While
+
+  Case
+
 %Considering source-to-source transformations that we investigate with the BOAST
@@
 
+  POET
+
   \subsection{Application Autotuning}

### 2014-08-26 Paper/BOAST.tex
updated article

@@
 
+\begin{center}
+\begin{figure}
+\includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
+\caption{Structure and Workflow of the BOAST framework.}
+\label{fig:boast_workflow}
+\end{figure}
+\end{center}
+
   \cite{videau2013boast}
@@
 
-It was also important to have our construct like for loops to have a syntax
-approaching those commonly found in programming languages. For this we needed a
-language which could seamlessly pass a block of code to a function.
+It was also important to have our constructs like \emph{for loops} to have a
+syntax approaching those commonly found in programming languages. For this we
+needed a language which could seamlessly pass a block of code to a function.
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection

### 2014-08-27 Paper/BOAST.tex
Improved paper

@@
 
-  Provide scientific application developers with a framework to develop and test
-application computing kernels. Two aspects should be considered:
-\begin{itemize}
-\item code description,
-\item code generation and kernel execution runtime.
-\end{itemize}
-
+\begin{figure}
 \begin{center}
-\begin{figure}
 \includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
@@
 \label{fig:boast_workflow}
+\end{center}
 \end{figure}
-\end{center}
+
+  The idea behind BOAST is to provide scientific application developers with a
+framework to develop and test application computing kernels. The envisioned
+workflow and program structure is presented in Figure~\ref{fig:boast_workflow}.
+The user should start from an application kernel (implemented or just designed)
+and implement it in a dedicated language (step 1). The language provides enough
+flexibility for the kernel to be meta-programmed with several orthogonal
+optimizations. Then the developer selects one configuration of the kernel and a
+target language. Those parameter define the output source code that will be
+generated by BOAST (step 2). The resulting code source is then built according to
+the user specified compiler and options (step 3). If input data are available
+then the kernel can be benchmarked and tested for non regression. Based on the
+results, other optimizations can be selected (step 4) or new optimizations can
+be added to the BOAST sources. The process can be repeated until a good
+candidate is found on the target platform. The resulting kernel is then added to
+the program (step 5).
+
+Several of the steps in the workflow can be automated: optimization and compiler
+flags exploration, non regression testing, benchmarking... This automation can
+be done by scripting by the user or by interfacing other dedicated tools. In
+order to achieve those goals three aspects should be considered:
+\begin{itemize}
+\item code description,
+\item code generation,
+\item and kernel execution runtime.
+\end{itemize}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.8]{BOAST_DSL.pdf}
+\caption{BOAST DSL intent.}
+\label{fig:dsl}
+\end{center}
+\end{figure}
+
 In order to manipulate those abstractions we want to have a syntax resembling
@@
 description of the kernel while the host language provides the meta programming
-of the kernel. Operator overloading of the host language will provide the
-familiar syntax programmers are accustomed to.
+of the kernel, cf Figure~\ref{fig:dsl}. Operator overloading of the host
+language will provide the familiar syntax programmers are accustomed to.
 
-It was also important to have our constructs like \emph{for loops} to have a
+It was also important to have our constructs like \emph{for} loops to have a
 syntax approaching those commonly found in programming languages. For this we
@@
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
-capabilities as well.
+capabilities as well. This is the main reason it was selected.
 
@@
 In order to clearly differentiate what is going to be generated from what is
-related to manipulations in the host language the \emph{print} method was
-redefined in the BOAST namespace. This print method only insures that the object
-it is called on implements a public print method. Each BOAST object is
-responsible for printing itself correctly depending on the BOAST configuration
-at the time the print public method is called.
+related to manipulations in the host language three keywords were defined. They
+are \emph{print}, \emph{decl} and \emph{close}. As the language is an EDSL,
+these three keywords are methods in the BOAST namespace.
+
+The \emph{print} method calls the public \emph{print} method of Objects it is
+called on. Each BOAST object is responsible for printing itself correctly
+depending on the BOAST configuration at the time the print public method is
+called. Calling directly the print method of a BOAST object yields the same
+result.
+
+The \emph{decl} method is used for two purposes. First, it can be used to
+declares a variable. Second, it can be used to print the beginning of a control
+structure without an associated code block.
+
+The \emph{close} method is the counterpart to the \emph{decl} method. It is used
+to close a control structure without an associated code block.
 
@@
   Case
+
+  Procedure
+
+  Function Call
 

### 2014-08-29 Paper/BOAST.tex
Updated article

@@
 \usepackage{color}
+
+\lstdefinestyle{customc}{
+  language=C,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={int64_t, int32_t}
+}
+
+\lstdefinestyle{BOAST}{
+  language=Ruby, 
+  basicstyle=\scriptsize, 
+  backgroundcolor=\color{white}, 
+  frame=single, 
+  captionpos=b,
+  morekeywords={BOAST, print, decl, close}
+}
+
+\lstdefinestyle{customf}{
+  language=FORTRAN,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b
+}
+
 \usepackage{url}
@@
 
-  Variable
+BOAST defines several class that are used to represent the structure of the
+code. These classes can be sorted in two groups:
 
-  Expression
+      \paragraph{Algebra}
+
+\begin{figure}
+\lstset{style=BOAST,
+        caption={BOAST code}, 
+        label={lst:BOAST_algebra}
+}
+
+\begin{lstlisting}
+i = BOAST::Int("i") # or BOAST::Variable::new(Int, "i")
+k = BOAST::Int("k", :size => 8)
+l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 20) ], :local => true )
+BOAST::decl i, k, l
+BOAST::print i === 5
+j = i + 5
+BOAST::print k === j * 2
+BOAST::print l[k] === 1.0
+\end{lstlisting}
+
+\begin{minipage}[b]{0.50\linewidth}
+\centering
+\lstset{style=customf,
+        caption={FORTRAN output.},
+        label={lst:BOAST_algebra_FORTRAN} }
+
+\begin{lstlisting}
+integer(kind=4) :: i
+integer(kind=8) :: k
+real(kind=8), dimension(-5:20) :: l
+i = 5
+k = (i + 5) * (2)
+l(k) = 1.0_wp
+\end{lstlisting}
+\end{minipage}
+\hspace{0.04\linewidth}
+\begin{minipage}[b]{0.44\linewidth}
+\centering
+\lstset{style=customc,
+        caption={C output.},
+        label={lst:BOAST_algebra_C} }
+
+\begin{lstlisting}
+int32_t i;
+int64_t k;
+double l[26];
+i = 5;
+k = (i + 5) * (2);
+l[k - (-5)] = 1.0;
+\end{lstlisting}
+\end{minipage}
+\caption{BOAST Code Snippet for Variables and Expressions}
+\label{fig:BOAST_algebra}
+\end{figure}
+The first and most fundamental abstraction is named Variable. Variables have a
+type, a name and a set of named attributes. The existing attributes are mainly
+inspired from FORTRAN. Those attributes are not limited and can be arbitrarily
+enriched, allowing a lot of flexibility in Variable management.
+
+The second abstraction is named Expression. It is used to combine variables into
+algebraic or logic expressions. Most of the classical operators are overloaded
+for those two abstractions and thus the syntax of the expressions are rather
+straightforward. The exception is the assignment operator as it is important to
+differentiate between assigning an Expression or a Variable in the ruby context
+and the assignment operator in the context of a BOAST Expression. Thus the
+assignment in a BOAST Expression is represented as the \emph{===} operator,
+while the classical assignment is kept as the \emph{=} operator.
+
+Figure~\ref{fig:BOAST_algebra} shows some basic usage of both Variables and
+Expressions as well as the \emph{print} and \emph{decl} keywords. For clarity we
+stayed out of BOAST namespace so BOAST related class and methods are prefixed by
+\emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that when
+run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
+(Listing~\ref{lst:BOAST_algebra_C}) output.
+
+      \paragraph{Control Structures}
 

### 2014-08-29 Tutorial-BOAST/Tutorial-BOAST.tex
Added BOAST tutorial sources

\documentclass{beamer}
\usepackage{etex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[francais]{babel}
\usepackage{amssymb}
\usepackage{epsfig,shadow}
\usepackage{verbatim}
\usepackage{pstricks}
\usepackage{beamerthemesplit}
\usepackage{graphicx}
%\usepackage{comment}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{figlatex}
\usepackage[french]{algorithmic}
\usepackage{multirow}

\graphicspath{{fig/}{campagne/fig}}
\newcommand{\talk}[2]{{#1} dit : {#2}}
\newcommand{\erik}[1]{\talk{Erik}{#1}}
\newcommand{\brice}[1]{\talk{Brice}{#1}}
\beamertemplatetransparentcovereddynamic

%\setbeamertemplate{background canvas}[vertical shading][bottom=red!10,top=blue!10]
%\usetheme{Warsaw}

\usecolortheme{progressbar}
\usefonttheme{progressbar}
\useoutertheme{progressbar}
\useinnertheme{progressbar}

\newcommand{\bordereau}{Bordereau\xspace}
\newcommand{\genepi}{Genepi\xspace}
\newcommand{\minelement}{{\em min\_element}\xspace}
\newcommand{\stablesort}{{\em stable\_sort}\xspace}
\newcommand{\merge}{{\em merge}\xspace}
\newcommand{\sort}{{\em sort}\xspace}
\newcommand{\find}{{\em find}\xspace}
\setbeamertemplate{navigation symbols}{}
\beamertemplatetransparentcovereddynamic
\newcommand{\pastel}{PaSTeL\xspace}

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % java

\title{BOAST}
\subtitle{Porting HPC applications to the Mont-Blanc prototype using BOAST }
\author{Kevin Pouget, \textbf{Brice Videau}, Jean-François Méhaut\\
(INRIA Grenoble - LIG - NANOSIM)}

\date{\textbf{Tutorial BOAST}\\June 25, 2014}

\begin{document}

\frame{\titlepage}

\section{Introduction}

\frame{
    \frametitle{Introduction}
}

\subsection{Context}

\begin{frame}
    \frametitle{Scientific Application Optimization}
\begin{itemize}
\item In the past, waiting for a new generation of hardware was enough to obtain performance gains.
\item Nowadays, architecture are so different that performance regress when a new architecture is released.
\item Sometimes the code is not fit anymore and cannot be compiled.
\item Few applications can harness the power of current computing platforms.
\item Thus, optimizing scientific application is of paramount importance.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Multiplicity of Architectures}
A High performance Computing application can encounter several types of architectures:
\begin{itemize}
\item Generalist Multicore CPU (AMD, Intel, PowerPC...)
\item Graphical Accelerators (ATI, NVIDIA...)
\item Computing Accelerators (CELL, MIC...)
\item \textbf{Low power CPUs (ARM...)}
\end{itemize}
Those architectures can present drastically different characteristics.
\end{frame}

\begin{frame}
\frametitle{Architectures Comparison}
\begin{table}
\centering
\footnotesize
\begin{tabular}{|l|r|r|r|r|}
\hline
Architecture & AMD/Intel CPUs & ARM & GPUs & Xeon Phi \\
\hline
Cores    & 4-12   & 2-4   & 512-2496    & 60  \\
\hline
Cache    & 3l & 2l & 2l incoherent & 2l \\
\hline
Memory (GiB)   & 2-4 (per core)   & 1 (per core)   & 2-6    & 6-16   \\
\hline
Vector Size   & 2-4    & 1-2    & 1-2    & 8 \\
\hline
Peak GFLOPS & 10-20 (per core)    & 2-4 (per core)   & 500-1500    & 1000\\
\hline
Peak GiB/s & 20-40 & 2.5-5 & 150-250 & 200\\
\hline
TDP W & 75  & 5 & 200 & 300\\
\hline
GFLOPS/W & 1-3 & 2-4 & 2-7 & 3\\
\hline
\end{tabular} \caption{Comparison between commonly found architectures in HPC.}
\end{table}
\end{frame}

\subsection{Problematic}

\begin{frame}
\frametitle{Exploiting Various Architectures}
Usually work is done on a class of architecture (CPUs or GPUs or accelerators).\\
\begin{block}{Well Known Examples}
\begin{itemize}
\item Atlas (Linear Algebra CPU)
\item PhiPAC (Linear Algebra CPU)
\item Spiral (FFT CPU)
\item FFTW (FFT CPU)
\item NukadaFFT(FFT GPU)
\end{itemize}
\end{block}
No work targeting several class of architectures.
What if the application is not based on a library?
\end{frame}

\subsection{Use Cases}

\begin{frame}
\frametitle{The Mont-Blanc European Project}
\begin{center}
\includegraphics[scale=0.3]{Mont-Blanc_logo-horizontal}
\end{center}
European project:
\begin{itemize}
\item Develop prototypes of HPC clusters using low power commercially available embedded technology (ARM CPUs, low power GPUs...).
\item Design the next generation in HPC systems based on embedded technologies and experiments on the prototypes.
\item Develop a portfolio of existing applications to test these systems and optimize their efficiency, using BSC's OmpSs programming model (11 existing applications were selected for this portfolio).
\end{itemize}
Prototype : based on Exynos 5250 : dual core Cortex A15 with T604 Mali GPU (OpenCL)
\end{frame}

\begin{frame}
    \frametitle{BigDFT a Tool for Nanotechnologies}
\begin{columns}
\column{6cm}
Ab initio simulation:
\begin{itemize}
  \item Simulates the properties of crystals and molecules,
  \item Computes the electronic density,
  \item Based on Daubechie wavelet.
\end{itemize}
The formalism was chosen because it is fit for HPC computations: 
\begin{itemize}
\item Each orbital can be treated independently most of the time,
\item Operator on orbitals are simple and straightforward.
\end{itemize}
\column{4.5cm}
\centering
\includegraphics[width=3.5cm]{CH4-grid.png}\\
Electronic density around a methane molecule.
\end{columns}
\end{frame}

\begin{frame}
\frametitle{BigDFT as an HPC application}
\begin{block}{Implementation details:}
\begin{itemize}
\item 200,000 lines of Fortran 90 and C
\item Supports MPI, OpenMP, CUDA and OpenCL
\item Uses BLAS
\item Scalability up to 16000 cores of Curie and 288GPUs
\end{itemize}
\end{block}

\begin{block}{Operators can be expressed as 3D convolutions:}
  \begin{itemize}
    \item Wavelet Transform
    \item Potential Energy
    \item Kinetic Energy
  \end{itemize}
  These convolutions are separable and filter are short (16 elements).\\
  Can take up to 90\% of the computation time on some systems.
\end{block}
\end{frame}

\begin{frame}
  \frametitle{SPECFEM3D a tool for wave propagation research}
\begin{columns}
\column{6cm}
Wave propagation simulation:
\begin{itemize}
  \item Used for geophysics and material research,
  \item Accurately simulate earthquakes,
  \item Based on spectral finite element.
\end{itemize}
Developped all around the world:
\begin{itemize}
\item Marseilles (CNRS),
\item Switzerland (ETH Zurich) CUDA,
\item United States (Princeton) Networking,
\item Grenoble (LIG/CNRS) OpenCL.
\end{itemize}
\column{5cm}
\centering
\includegraphics[width=5cm]{image_movie_Sichuan_009000_0.jpg}\\
Sichuan earthequake.
\end{columns}
\end{frame}

\begin{frame}
\frametitle{SPECFEM3D as an HPC application}
\begin{block}{Implementation details:}
\begin{itemize}
\item 80,000 lines of Fortran 90
\item Supports MPI, CUDA, OpenCL and an OMPSs + MPI miniapp
\item Scalability up to 693,600 cores on IBM BlueWaters
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
    \frametitle{Talk Outline}
    \tableofcontents[hideallsubsections,sections={2-}]
    %\includegraphics{plan_couleur.pdf}
\end{frame}

\section{Case Study}

\begin{frame}[fragile]
\frametitle{Case Study 1: BigDFT's MagicFilter}
The simplest convolution found in BigDFT, corresponds to the potential operator.
\begin{columns}
\column{4cm}
\begin{block}{Characteristics}
\begin{itemize}
\item Separable,
\item Filter length 16,
\item Transposition,
\item Periodic,
\item Only 32 operations per element.
\end{itemize}
\end{block}
\column{6.2cm}
\begin{block}{Pseudo code}
\tiny
\lstset{language=C,numbers=left}
\begin{lstlisting}
double filt[16] = {F0, F1, ... , F15};
void magicfilter(int n, int ndat, 
                 double *in, double *out){
  double temp;
  for(j=0; j<ndat; j++) {
    for(i=0; i<n; i++) {
       temp = 0;
       for(k=0; k<16; k++) {
         temp+= in[ ((i-7+k)%n) + j*n] 
                  * filt[k];
       }
       out[j + i*ndat] = temp;
} } }
\end{lstlisting}
\end{block}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Case study 2: SPECFEM3D port to OpenCL}
\begin{block}{Existing CUDA code:}
\begin{itemize}
\item 42 kernels and 15000 lines of code
\item kernels with 80+ parameters
\item $\sim{}$ 7500 lines of cuda code
\item $\sim{}$ 7500 lines of wrapper code
\end{itemize}
\end{block}
\begin{block}{Objectives:}
\begin{itemize}
\item Factorize the existing code,
\item Single OpenCL and CUDA description for the kernels,
\item Validate without unit tests, comparing native Cuda to generated Cuda executions
\item Keep similar performances.
\end{itemize}
\end{block}
\end{frame}

\section{A Parametrized Generator}
\begin{frame}
\frametitle{A Parametrized Generator}
\end{frame}

\begin{frame}
\frametitle{Classical Workflow}
\begin{figure}
\centering
\includegraphics<1>[scale=1]{Workflow1-1}
\includegraphics<2>[scale=1]{Workflow1-2}
\includegraphics<3>[scale=1]{Workflow1-3}
\includegraphics<4>[scale=1]{Workflow1-4}
\label{fig:2filters}
\end{figure}
\begin{itemize}
\item<1|only@1> Kernel optimization workflow
\item<1|only@1> Usually performed by a knowledgeable developer
\item<2|only@2> Compilers perform optimizations
\item<2|only@2> Architecture specific or generic optimizations
\item<3|only@3> Performance data hint at source transformations
\item<3|only@3> Architecture specific or generic hints
\item<4|only@4> Multiplication of kernel versions or loss of versions
\item<4|only@4> Difficulty to benchmark versions against each-other
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{BOAST Workflow}
\begin{figure}
\centering
\includegraphics<1>[scale=1]{Workflow2-1}
\includegraphics<2>[scale=1]{Workflow2-2}
\includegraphics<3>[scale=1]{Workflow2-3}
\label{fig:2filters}
\end{figure}
\begin{itemize}
\item<1|only@1> Meta-programming of optimizations in BOAST
\item<1|only@1> High level object oriented language
\item<2|only@2> Generate combination of optimizations
\item<2|only@2> C, OpenCL, FORTRAN and CUDA are supported
\item<3|only@3> Compilation and analysis are automated
\item<3|only@3> Selection of best version can also be automated
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{BOAST}
\begin{figure}
\centering
\includegraphics[scale=0.5]{BOAST}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Use Case Driven}
Parameters arising in a convolution:
\begin{itemize}
\item Filter: length, values, center.
\item Direction: forward or inverse convolution.
\item Boundary conditions: free or periodic.
\item Unroll factor: arbitrary.
\end{itemize}
How are those parameters constraining our tool?
\end{frame}

\begin{frame}
\frametitle{Features required}
Unroll factor:
\begin{itemize}
\item Create and manipulate an unknown number of variables,
\item Create loops with variable steps.
\end{itemize}
Boundary conditions:
\begin{itemize}
\item Manage arrays with parametrized size.
\end{itemize}
Filter and convolution direction:
\begin{itemize}
\item Transform arrays.
\end{itemize}
And of course be able to describe convolutions and output them in different languages.
\end{frame}

\begin{frame}
\frametitle{Proposed Generator}
Idea: use a high level language with support for operator overloading to describe the structure of the code, rather than trying to transform a decorated tree.\\
Define several abstractions:
\begin{itemize}
\item Variables: type (array, float, integer), size...
\item Operators: affect, multiply...
\item Procedure and functions: parameters, variables...
\item Constructs: for, while...
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Sample Code: Variables and Parameters}
\tiny
%\lstset{language=Ruby,numbers=left}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  #simple Variable
  i = Int "i"
  #simple constant
  lowfil = Int( "lowfil", :const => 1-center )
  #simple constant array
  fil = Real("fil", :const => arr, :dim => [ Dim(lowfil,upfil) ])
  #simple parameter
  ndat = Int("ndat", :dir => :in)
  #multidimensional array, an output parameter
  y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(dim_out_min, dim_out_max) ] )
\end{lstlisting}
\normalsize
Variables and Parameters are objects with a name, a type, and a set of named properties.
\end{frame}

\begin{frame}[fragile]
\frametitle{Sample Code: Procedure Declaration}
The following declaration:
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
p = Procedure("magic_filter", [n,ndat,x,y], [lowfil,upfil])
decl p
\end{lstlisting}
\normalsize 
Outputs Fortran:
\tiny
\lstset{language=Fortran,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
subroutine magicfilter(n, ndat, x, y)
  integer(kind=4), parameter :: lowfil = -8
  integer(kind=4), parameter :: upfil = 7
  integer(kind=4), intent(in) :: n
  integer(kind=4), intent(in) :: ndat
  real(kind=8), intent(in), dimension(0:n-1, ndat) :: x
  real(kind=8), intent(out), dimension(ndat, 0:n-1) :: y
\end{lstlisting}
\normalsize
Or C:
\tiny
\lstset{language=C,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
void magicfilter(const int32_t n, const int32_t ndat, const double * x, double * y){
  const int32_t lowfil = -8;
  const int32_t upfil = 7;
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Sample Code: Constructs and Arrays}
The following declaration:
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  unroll = 5
  print For(j,1,ndat-(unroll-1), unroll) {
  #.....
    (tt2 === tt2 + x[k,j+1]*fil[l]).print
  #.....
  }
\end{lstlisting}
\normalsize 
Outputs Fortran:
\tiny
\lstset{language=Fortran,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  do j=1, ndat-4, 5
    !......
    tt2=tt2+x(k,j+1)*fil(l)
    !......
  enddo
\end{lstlisting}
\normalsize
Or C:
\tiny
\lstset{language=C,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  for(j=1; j<=ndat-4; j+=5){
  /*...........*/
    tt2=tt2+x[k-0+(j+1-1)*(n-1-0+1)]*fil[l-lowfil];
  /*...........*/
  }
\end{lstlisting}
\end{frame}

\section{Evaluation}
\begin{frame}
\frametitle{Generator Evaluation}
Back to the test cases:
\begin{itemize}
\item The generator was used to unroll the Magicfilter an evaluate it's performance on an ARM processor and an Intel processor.
\item The generator was used to describe SPECFEM3D kernel.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Performance Results}
\centering
\begin{columns}
\column{5cm}
\begin{block}{Tegra2}
\includegraphics[scale=0.4]{cacc}\\
\includegraphics[scale=0.4]{totcyc}
\end{block}
\column{5cm}
\begin{block}{Intel T7500}
\includegraphics[scale=0.4]{nedni_cacc.png}\\
\includegraphics[scale=0.4]{nedni_totcyc.png}
\end{block}
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Improvement for BigDFT}
\begin{itemize}
\item Most of the convolutions have been ported to BOAST.
\item Results are encouraging: on the hardware BigDFT was hand optimized for, convolutions gained on average between 30 and 40\% of performance.
\item MagicFilter OpenCL versions tailored for problem size by BOAST gain 10 to 20\% of performance.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{SPECFEM3D OpenCL port}
Fully ported to OpenCL with comparable performances (using the \texttt{global\_s362ani\_small} test case):
\begin{itemize}
\item On a 2*6 cores (E5-2630) machine with 2 K40, using 12 MPI processes:
\begin{itemize}
\item OpenCL: 4m15s
\item CUDA: 3m10s
\end{itemize}
\item On an 2*4 cores (E5620) with a K20 using 6 MPI processes:
\begin{itemize}
\item OpenCL: 12m47s
\item CUDA: 11m23s
\end{itemize}
\end{itemize}
Difference comes from the capacity of cuda to specify the minimum number of blocks to launch on a multiprocessor.
Less than 4000 lines of BOAST code (7500 lines of cuda originally).
\end{frame}

\section{Conclusions and Future Work}

\begin{frame}
\frametitle{Conclusions and Future Work}
\end{frame}

\begin{frame}
\frametitle{Conclusions}
Generator has been used to test several loop unrolling strategies in BigDFT.\\
Highlights:
\begin{itemize}
\item Several output languages.
\item All constraints have been met.
\item Automatic benchmarking framework allows us to test several optimization levels and compilers.
\item Automatic non regression testing.
\item Several algorithmically different versions can be generated (changing the filter, boundary conditions...).
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Future Works and Considerations}
Future work:
\begin{itemize}
\item Produce an autotuning convolution library.
\item Implement a parametric space explorer or use an existing one (ASK: Adaptative Sampling Kit, Collective Mind...).
\item Vector code is supported, but needs improvements.
\item Test the OpenCL version of SPECFEM3D on the Mont-Blanc prototype.
\end{itemize}
Question raised:
\begin{itemize}
\item Is this approach extensible enough?
\item Can we improve the language used further?
\end{itemize}

\end{frame}

\end{document}

### 2014-08-29 Tutorial/Tutorial-BOAST-2.tex
renamed tutorial directory

\documentclass{beamer}
\usepackage{etex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amssymb}
\usepackage{epsfig,shadow}
\usepackage{verbatim}
\usepackage{pstricks}
\usepackage{beamerthemesplit}
\usepackage{graphicx}
\usepackage{comment}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{figlatex}
\usepackage{algorithmic}
\usepackage{multirow}

\graphicspath{{fig/}{campagne/fig}}
\newcommand{\talk}[2]{{#1} dit : {#2}}
\newcommand{\erik}[1]{\talk{Erik}{#1}}
\newcommand{\brice}[1]{\talk{Brice}{#1}}
\beamertemplatetransparentcovereddynamic

%\setbeamertemplate{background canvas}[vertical shading][bottom=red!10,top=blue!10]
%\usetheme{Warsaw}

\usecolortheme{progressbar}
\usefonttheme{progressbar}
\useoutertheme{progressbar}
\useinnertheme{progressbar}

\newcommand{\bordereau}{Bordereau\xspace}
\newcommand{\genepi}{Genepi\xspace}
\newcommand{\minelement}{{\em min\_element}\xspace}
\newcommand{\stablesort}{{\em stable\_sort}\xspace}
\newcommand{\merge}{{\em merge}\xspace}
\newcommand{\sort}{{\em sort}\xspace}
\newcommand{\find}{{\em find}\xspace}
\setbeamertemplate{navigation symbols}{}
\beamertemplatetransparentcovereddynamic
\newcommand{\pastel}{PaSTeL\xspace}

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.25,0.5,0.35} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % java

\title{BOAST}
\subtitle{Using BOAST: an Introduction }
\author{Brice Videau\\
(LIG - NANOSIM)}

\date{\textbf{BOAST Tutorial}\\June 25, 2014}

\begin{document}

\frame{\titlepage}

\section{Installing BOAST}

\begin{frame}[fragile]

    \frametitle{Installing BOAST}

Install ruby, version >= 1.9.3\\
On recent debian-based distributions:
\lstset{language=Bash,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
sudo apt-get install ruby ruby-dev
\end{lstlisting}
And then install the BOAST gem (ruby module):
\begin{lstlisting}
sudo gem install BOAST
\end{lstlisting}
If on a cluster frontend:
\begin{lstlisting}
gem install --user-install BOAST
\end{lstlisting}

\end{frame}

\section{First Steps}

\begin{frame}[fragile]
\frametitle{First interactive steps}
\lstset{language=Bash,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
Interactive Ruby:
\begin{lstlisting}
irb
\end{lstlisting}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
Simple BOAST commands:
\begin{lstlisting}
irb(main):001:0> require 'BOAST'
=> true
irb(main):002:0> a = BOAST::Int "a"
=> a
irb(main):003:0> b = BOAST::Real "b"
=> b
irb(main):004:0> BOAST::decl a, b
integer(kind=4) :: a
real(kind=8) :: b
=> [a, b]
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Defining a Procedure}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
Simple BOAST Procedure:
\footnotesize
\begin{lstlisting}
05:0> p = BOAST::Procedure( "test_proc", [a,b] )
06:0> BOAST::decl p
SUBROUTINE test_proc(a, b)
  integer, parameter :: wp=kind(1.0d0)
  integer(kind=4) :: a
  real(kind=8) :: b
007:0> BOAST::lang = BOAST::C
008:0> BOAST::decl p
void test_proc(int32_t a, double b){
009:0> BOAST::close p
}
\end{lstlisting}
Available languages are FORTRAN, C, CUDA, CL (OpenCL)
\end{frame}

\begin{frame}[fragile]
\frametitle{Defining a Full Procedure:}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\tiny
\begin{lstlisting}
010:0> a = BOAST::Real( "a", :dir => :in)
011:0> b = BOAST::Real( "b", :dir => :out)
012:0> p = BOAST::Procedure( "test_proc", [a,b] ) { BOAST::print b === a + 2 }
013:0> BOAST::lang = BOAST::FORTRAN
014:0> BOAST::print p
SUBROUTINE test_proc(a, b)
    integer, parameter :: wp=kind(1.0d0)
    real(kind=8), intent(in) :: a
    real(kind=8), intent(out) :: b
    b = a + 2
END SUBROUTINE test_proc
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Creating a Computing Kernel}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\tiny
\begin{lstlisting}
n = BOAST::Int(  "n", :dir => :in )
a = BOAST::Real( "a", :dir => :in,  :dim => [BOAST::Dim(n)] )
b = BOAST::Real( "b", :dir => :out, :dim => [BOAST::Dim(n)] )
p = BOAST::Procedure( "test_proc", [n, a, b] ) {
  BOAST::decl i = BOAST::Int( "i" )
  BOAST::For( i, 1, n ) {
    BOAST::print b[i] === a[i] + 2
  }.print
}
k = BOAST::CKernel::new
BOAST::print p
k.procedure = p
k.build
BOAST::verbose = true
k.build
> gcc -O2 -Wall -fPIC -I/usr/lib/ruby/1.9.1/x86_64-linux -I/usr/include/ruby-1.9.1 -I/usr/include/ruby-1.9.1/x86_64-linux -I/var/lib/gems/1.9.1/gems/narray-0.6.0.9 -DHAVE_NARRAY_H -c -o /tmp/Mod_test_proc20140624_19378_1qdep6u.o /tmp/Mod_test_proc20140624_19378_1qdep6u.c
> gfortran -O2 -Wall -fPIC -c -o /tmp/test_proc20140624-19378-1qdep6u.o /tmp/test_proc20140624-19378-1qdep6u.f90
> gcc -shared -o /tmp/Mod_test_proc20140624_19378_1qdep6u.so /tmp/Mod_test_proc20140624_19378_1qdep6u.o /tmp/test_proc20140624-19378-1qdep6u.o  -Wl,-Bsymbolic-functions -Wl,-z,relro -rdynamic -Wl,-export-dynamic  -L/usr/lib -lruby-1.9.1 -lrt
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Running a Computing Kernel}
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\tiny
\begin{lstlisting}
require 'narray'
input  = NArray.float(1024).random
output = NArray.float(1024)
n = BOAST::Int(  "n", :dir => :in )
a = BOAST::Real( "a", :dir => :in,  :dim => [BOAST::Dim(n)] )
b = BOAST::Real( "b", :dir => :out, :dim => [BOAST::Dim(n)] )
p = BOAST::Procedure( "test_proc", [n, a, b] ) {
  BOAST::decl i = BOAST::Int( "i" )
  BOAST::For( i, 1, n ) {
    BOAST::print b[i] === a[i] + 2
  }.print
}
k = BOAST::CKernel::new
BOAST::print p
k.procedure = p
k.run(input.length, input, output)
(output - input).each { |val| raise "Error!" if (val-2).abs > 1e-15 }
stats = k.run(input.length, input, output)
puts "#{stats[:duration]} s"
> 4.911e-06 s
\end{lstlisting}
\end{frame}

\section{BOAST Samples}
\begin{frame}
\frametitle{BOAST Samples}

\end{frame}

\subsection{The Canonic Case}
\begin{frame}[fragile]
\frametitle{The Canonic Case: Vector Addition Kernel}
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
def BOAST::vector_add
  kernel = CKernel::new
  function_name = "vector_add"
  n = Int("n",{:dir => :in, :signed => false})
  a = Real("a",{:dir => :in, :dim => [ Dim(0,n-1)] })
  b = Real("b",{:dir => :in, :dim => [ Dim(0,n-1)] })
  c = Real("c",{:dir => :out, :dim => [ Dim(0,n-1)] })
  i = Int("i",{:signed => false})
  print p = Procedure(function_name, [n,a,b,c]) {
    decl i
    if [CL, CUDA].include?(get_lang) then
      print i === get_global_id(0)
      print c[i] === a[i] + b[i]
    else
      print For(i,0,n-1) {
        print c[i] === a[i] + b[i]
      }
    end
  }
  kernel.procedure = p
  return kernel
end
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Running the Kernel}
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
n = 1024*1024
a = NArray.float(n).random
b = NArray.float(n).random
c = NArray.float(n)
c_ref = NArray.float(n)

epsilon = 10e-15

set_lang( FORTRAN )
k = vector_add
k.run(n,a,b,c_ref)

[C, CL, CUDA].each { |lang|
  set_lang( lang )
  c.random
  k = vector_add
  case lang
  when CL
    k.run(n,a,b,c, :global_work_size => [rndup(n,32), 1,1], :local_work_size => [32,1,1] )
  when CUDA
    k.run(n,a,b,c, :block_number => [rndup(n,32)/32, 1,1], :block_size => [32,1,1] )
  else
    k.run(n,a,b,c)
  end
  (c_ref - c).abs.each { |diff|
    raise "Warning: residue too big: #{elem}" if elem > epsilon
  }
}
\end{lstlisting}
\end{frame}
\subsection{Building Kernels}
\begin{frame}[fragile]
\frametitle{Building Kernels}
\begin{itemize}
\item Running a kernel builds it (if it is not already built).
\item Kernels can be built beforehand.
\item Usual build parameters can be specified.
\end{itemize}
Sample:
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  k.build({:FC => 'gfortran', :CC => 'gcc',\
           :FCFLAGS => "-O2", :LDFLAGS => ""})
  k.build({:FC => 'gfortran', :CC => 'gcc',\
           :FCFLAGS => "-O2 -fopenmp",:LDFLAGS => "-fopenmp"})
  k.build({:FC => 'ifort',    :CC => 'icc',
           :FCFLAGS => "-O2 -openmp" ,:LDFLAGS => "-openmp",:LD => "ifort"})
\end{lstlisting}
\normalsize
\begin{itemize}
\item Probes can be inserted at compile time.
\item Default: high resolution timer.
\end{itemize}
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  stats = k.run(...)
  puts stats[:duration]+" s"
\end{lstlisting}
\end{frame}

\subsection{Real Life Examples}
\begin{frame}
\frametitle {Real Life Examples}
\end{frame}

\begin{frame}[fragile]
\frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : Reference}
\tiny
\lstset{language=C,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
typedef float realw;
__global__ void assemble_boundary_potential_on_device(realw* d_potential_dot_dot_acoustic,
                                                      realw* d_send_potential_dot_dot_buffer,
                                                      int num_interfaces,
                                                      int max_nibool_interfaces,
                                                      int* d_nibool_interfaces,
                                                      int* d_ibool_interfaces) {

  int id = threadIdx.x + blockIdx.x*blockDim.x + blockIdx.y*gridDim.x*blockDim.x;
  int iglob,iloc;

  for( int iinterface=0; iinterface < num_interfaces; iinterface++) {
    if(id < d_nibool_interfaces[iinterface]) {

      iloc = id + max_nibool_interfaces*iinterface;

      iglob = d_ibool_interfaces[iloc] - 1;

      // assembles values
      atomicAdd(&d_potential_dot_dot_acoustic[iglob],d_send_potential_dot_dot_buffer[iloc]);
    }
  }
}
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]
\frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : BOAST (1)}
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
def BOAST::assemble_boundary_potential_on_device(ref = true)
  push_env( :array_start => 0 )
  kernel = CKernel::new
  function_name = "assemble_boundary_potential_on_device"
  num_interfaces                  = Int("num_interfaces",                   \
                                        :dir => :in)
  max_nibool_interfaces           = Int("max_nibool_interfaces",            \
                                        :dir => :in)
  d_potential_dot_dot_acoustic    = Real("d_potential_dot_dot_acoustic",    \
                                        :dir => :out,:dim => [ Dim() ])
  d_send_potential_dot_dot_buffer = Real("d_send_potential_dot_dot_buffer", \
                                        :dir => :in, :dim => [ Dim(num_interfaces*max_nibool_interfaces) ])
  d_nibool_interfaces             = Int("d_nibool_interfaces",              \
                                        :dir => :in, :dim => [ Dim(num_interfaces) ])
  d_ibool_interfaces              = Int("d_ibool_interfaces",               \
                                        :dir => :in, :dim => [ Dim(num_interfaces*max_nibool_interfaces) ])
  p = Procedure(function_name, [d_potential_dot_dot_acoustic,d_send_potential_dot_dot_buffer,num_interfaces,max_nibool_interfaces,d_nibool_interfaces,d_ibool_interfaces])
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]
\frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : BOAST (2)}
\tiny
\lstset{language=Ruby,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
  if(get_lang == CUDA and ref) then
    @@output.print File::read("specfem3D/#{function_name}.cu")
  elsif(get_lang == CUDA or get_lang == CL) then
    decl p
    id         = Int("id")
    iglob      = Int("iglob")
    iloc       = Int("iloc")
    iinterface = Int("iinterface")
    decl id, iglob, iloc, iinterface
    print id === get_global_id(0) + get_global_size(0)*get_global_id(1)
    print For(iinterface, 0, num_interfaces-1) {
      print If(id<d_nibool_interfaces[iinterface]) {
        print iloc === id + max_nibool_interfaces*iinterface
        print iglob === d_ibool_interfaces[iloc] - 1
        print atomicAdd(d_potential_dot_dot_acoustic + iglob, \
                        d_send_potential_dot_dot_buffer[iloc])
      }
    }
    close p
  else
    raise "Unsupported language!"
  end
  pop_env( :array_start )
  kernel.procedure = p
  return kernel
end
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]
\frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : Generated CUDA}
\tiny
\lstset{language=C,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}
__global__ void assemble_boundary_potential_on_device(float * d_potential_dot_dot_acoustic, const float * d_send_potential_dot_dot_buffer, const int num_interfaces, const int max_nibool_interfaces, const int * d_nibool_interfaces, const int * d_ibool_interfaces){
  int id;
  int iglob;
  int iloc;
  int iinterface;
  id = threadIdx.x + ((blockIdx.x) * (blockDim.x)) + (((gridDim.x) * (blockDim.x)) * (threadIdx.y + ((blockIdx.y) * (blockDim.y))));
  for(iinterface=0; iinterface<=num_interfaces - (1); iinterface+=1){
    if(id < d_nibool_interfaces[iinterface - 0]){
      iloc = id + ((max_nibool_interfaces) * (iinterface));
      iglob = d_ibool_interfaces[iloc - 0] - (1);
      atomicAdd(d_potential_dot_dot_acoustic + (iglob), d_send_potential_dot_dot_buffer[iloc - 0]);
    }
  }
}
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]
\frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : Generated OpenCL}
\tiny
\lstset{language=C,
basicstyle=\ttfamily,
keywordstyle=\color{javapurple}\bfseries,
stringstyle=\color{javared},
commentstyle=\color{javagreen},
numbers=left,
tabsize=4,
showspaces=false,
showstringspaces=false}
\begin{lstlisting}

__kernel void assemble_boundary_potential_on_device(__global float * d_potential_dot_dot_acoustic, const __global float * d_send_potential_dot_dot_buffer, const int num_interfaces, const int max_nibool_interfaces, const __global int * d_nibool_interfaces, const __global int * d_ibool_interfaces){
  int id;
  int iglob;
  int iloc;
  int iinterface;
  id = get_global_id(0) + ((get_global_size(0)) * (get_global_id(1)));
  for(iinterface=0; iinterface<=num_interfaces - (1); iinterface+=1){
    if(id < d_nibool_interfaces[iinterface - 0]){
      iloc = id + ((max_nibool_interfaces) * (iinterface));
      iglob = d_ibool_interfaces[iloc - 0] - (1);
      atomicAdd(d_potential_dot_dot_acoustic + (iglob), d_send_potential_dot_dot_buffer[iloc - 0]);
    }
  }
}
\end{lstlisting}
\end{frame}
\begin{frame}
\frametitle{BigDFT Synthesis Kernel}
\centering 
\includegraphics[scale=0.5]{Res_synthesis}\\

\end{frame}

\end{document}

### 2014-09-08 Paper/BOAST.tex
Improved Paper

@@
   captionpos=b,
-  morekeywords={BOAST, print, decl, close}
+  morekeywords={BOAST, print, decl, open, close}
 }
@@
 In order to clearly differentiate what is going to be generated from what is
-related to manipulations in the host language three keywords were defined. They
-are \emph{print}, \emph{decl} and \emph{close}. As the language is an EDSL,
-these three keywords are methods in the BOAST namespace.
+related to manipulations in the host language four keywords were defined. They
+are \emph{print}, \emph{decl}, \emph{open} and \emph{close}. As the language is
+an EDSL, these four keywords are methods in the BOAST namespace.
 
@@
 
-The \emph{decl} method is used for two purposes. First, it can be used to
-declares a variable. Second, it can be used to print the beginning of a control
+The \emph{decl} method is used to declare variables or procedures and functions.
+
+The \emph{open} method can be used to print the beginning of a control
 structure without an associated code block.
 
-The \emph{close} method is the counterpart to the \emph{decl} method. It is used
+The \emph{close} method is the counterpart to the \emph{open} method. It is used
 to close a control structure without an associated code block.
@@
 BOAST defines several class that are used to represent the structure of the
-code. These classes can be sorted in two groups:
+code. These classes can be sorted in two groups, algebraic related and control
+flow related:
 
@@
         caption={BOAST code}, 
-        label={lst:BOAST_algebra}
-}
-
+        label={lst:BOAST_algebra} }
 \begin{lstlisting}
@@
 BOAST::print l[k] === 1.0
+BOAST::print l[k-1] === BOAST::sin(l[k])
 \end{lstlisting}
@@
 l(k) = 1.0_wp
+l(k - (1)) = sin(l(k))
 \end{lstlisting}
@@
 l[k - (-5)] = 1.0;
+l[k - (1) - (-5)] = sin(l[k - (-5)]);
 \end{lstlisting}
@@
 assignment in a BOAST Expression is represented as the \emph{===} operator,
-while the classical assignment is kept as the \emph{=} operator.
+while the classical assignment is kept as the \emph{=} operator. Function calls
+(FuncCall) are also abstracted and can be used in Expressions.
 
@@
 Expressions as well as the \emph{print} and \emph{decl} keywords. For clarity we
-stayed out of BOAST namespace so BOAST related class and methods are prefixed by
-\emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that when
-run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
-(Listing~\ref{lst:BOAST_algebra_C}) output.
+stayed out of BOAST namespace so BOAST related class and methods are prefixed
+with \emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that
+when run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
+(Listing~\ref{lst:BOAST_algebra_C}) output. On the last line, a call to the
+\emph{sin} function is made through the automatic creation of a FuncCall object.
+This automatic behaviors is limited to BOAST namespace, and thus, here, the
+function call has to be prefixed with \emph{BOAST::}.
 
@@
 
-  If
+\begin{figure}
+\lstset{style=BOAST,
+        caption={BOAST code}, 
+        label={lst:BOAST_control} }
+\begin{lstlisting}
+\end{lstlisting}
 
-  For
+\begin{minipage}[b]{0.50\linewidth}
+\centering
+\lstset{style=customf,
+        caption={FORTRAN output.},
+        label={lst:BOAST_control_FORTRAN} }
 
-  While
+\begin{lstlisting}
+\end{lstlisting}
+\end{minipage}
+\hspace{0.04\linewidth}
+\begin{minipage}[b]{0.44\linewidth}
+\centering
+\lstset{style=customc,
+        caption={C output.},
+        label={lst:BOAST_control_C} }
 
-  Case
+\begin{lstlisting}
+\end{lstlisting}
+\end{minipage}
+\caption{BOAST Code Snippet for Variables and Expressions}
+\label{fig:BOAST_control}
+\end{figure}
+
+The classical control structures are implemented. If, For, While, Case are
+abstractions in BOAST matching the behavior of corresponding control structures
+in other languages. An exception is the For in BOAST more closely matches the
+for in FORTRAN than the one in C.
+
+Figure~\ref{fig:BOAST_control} show some basic usage of the control structures.
 
   Procedure
-
-  Function Call
 

### 2014-09-09 Paper/BOAST.tex
Refactoring

@@
   captionpos=b,
-  morekeywords={BOAST, print, decl, open, close}
+  morekeywords={BOAST, pr, decl, open, close}
 }
@@
 related to manipulations in the host language four keywords were defined. They
-are \emph{print}, \emph{decl}, \emph{open} and \emph{close}. As the language is
+are \emph{pr}, \emph{decl}, \emph{open} and \emph{close}. As the language is
 an EDSL, these four keywords are methods in the BOAST namespace.
 
-The \emph{print} method calls the public \emph{print} method of Objects it is
+The \emph{pr} method calls the public \emph{pr} method of Objects it is
 called on. Each BOAST object is responsible for printing itself correctly
@@
 BOAST::decl i, k, l
-BOAST::print i === 5
+BOAST::pr i === 5
 j = i + 5
-BOAST::print k === j * 2
-BOAST::print l[k] === 1.0
-BOAST::print l[k-1] === BOAST::sin(l[k])
+BOAST::pr k === j * 2
+BOAST::pr l[k] === 1.0
+BOAST::pr l[k-1] === BOAST::sin(l[k])
 \end{lstlisting}
@@
 Figure~\ref{fig:BOAST_algebra} shows some basic usage of both Variables and
-Expressions as well as the \emph{print} and \emph{decl} keywords. For clarity we
+Expressions as well as the \emph{pr} and \emph{decl} keywords. For clarity we
 stayed out of BOAST namespace so BOAST related class and methods are prefixed

### 2014-09-11 Paper/BOAST.tex
Improved article

@@
         caption={BOAST code}, 
-        label={lst:BOAST_algebra} }
+        label={lst:BOAST_algebra},
+        numbers=left }
 \begin{lstlisting}
-i = BOAST::Int("i") # or BOAST::Variable::new(Int, "i")
+i = BOAST::Int("i") # or BOAST::Variable::new("i", Int)
 k = BOAST::Int("k", :size => 8)
-l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 20) ], :local => true )
+l = BOAST::Real("l", :dim => [ BOAST::Dim(-5, 21) ], :local => true )
 BOAST::decl i, k, l
@@
 BOAST::pr l[k] === 1.0
-BOAST::pr l[k-1] === BOAST::sin(l[k])
+BOAST::register_funccall("sin")
+BOAST::pr l[k+1] === BOAST::sin(j)
 \end{lstlisting}
@@
 integer(kind=8) :: k
-real(kind=8), dimension(-5:20) :: l
+real(kind=8), dimension(-5:21) :: l
 i = 5
@@
 l(k) = 1.0_wp
-l(k - (1)) = sin(l(k))
+l(k + 1) = sin(i + 5)
 \end{lstlisting}
@@
 int64_t k;
-double l[26];
+double l[27];
 i = 5;
@@
 l[k - (-5)] = 1.0;
-l[k - (1) - (-5)] = sin(l[k - (-5)]);
+l[k + 1 - (-5)] = sin(i + 5);
 \end{lstlisting}
@@
 when run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
-(Listing~\ref{lst:BOAST_algebra_C}) output. On the last line, a call to the
-\emph{sin} function is made through the automatic creation of a FuncCall object.
-This automatic behaviors is limited to BOAST namespace, and thus, here, the
-function call has to be prefixed with \emph{BOAST::}.
+(Listing~\ref{lst:BOAST_algebra_C}) output. At first we define 2 Variables
+\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third variable
+named \emph{l} is a one dimensional local array of a length of $27$, it is indexed
+in the range $-5$ to $21$. All those Variables are affected to ruby variables
+of the corresponding name.
+
+On Line~4 we declare those three variables. Variable \emph{i} is then affected
+the value 5. On Line~6 the \emph{j} ruby variable is used to store the BOAST
+expression \emph{$i+5$}. This variable will the be used transparently through
+the rest of the program.
+
+On Line~8 we use the \emph{k} variable to index into the array \emph{l} using
+the bracket operator. Would the array be multidimensional the index would be
+coma separated, similar to FORTRAN notation.
+
+On the last line, a call to the \emph{sin} function is made through the creation
+of a FuncCall object.  The possibility to use this method is declared using the
+\emph{register\_funccall} method.
 
@@
         caption={BOAST code}, 
-        label={lst:BOAST_control} }
+        label={lst:BOAST_control},
+        numbers=left }
 \begin{lstlisting}
+BOAST::register_funccall("modulo")
+i = BOAST::Int("i")
+j = BOAST::Int("j")
+BOAST::pr j === 0
+BOAST::pr BOAST::For( i, 0, 100 ) {
+  BOAST::pr BOAST::If( BOAST::modulo(i,7) == 0 ) {
+    BOAST::pr j === j + 1
+  }
+}
 \end{lstlisting}
 
-\begin{minipage}[b]{0.50\linewidth}
+\begin{minipage}[b]{0.47\linewidth}
 \centering
@@
 \begin{lstlisting}
+j = 0
+do i = 0, 100, 1
+  if (modulo(i, 7) == 0) then
+    j = j + 1
+  end if
+end do
 \end{lstlisting}
@@
 \hspace{0.04\linewidth}
-\begin{minipage}[b]{0.44\linewidth}
+\begin{minipage}[b]{0.47\linewidth}
 \centering
@@
 \begin{lstlisting}
+j = 0;
+for (i = 0; i <= 100; i += 1) {
+  if (modulo(i, 7) == 0) {
+    j = j + 1;
+  }
+}
 \end{lstlisting}
 \end{minipage}
-\caption{BOAST Code Snippet for Variables and Expressions}
+\caption{BOAST Code Snippet for control structures}
 \label{fig:BOAST_control}
@@
 
-The classical control structures are implemented. If, For, While, Case are
-abstractions in BOAST matching the behavior of corresponding control structures
-in other languages. An exception is the For in BOAST more closely matches the
-for in FORTRAN than the one in C.
+The classical control structures are implemented. \emph{If}, \emph{For},
+\emph{While}, \emph{Case} are abstractions in BOAST matching the behavior of
+corresponding control structures in other languages. An exception is the For in
+BOAST more closely matches the for in FORTRAN than the one in C.
 
 Figure~\ref{fig:BOAST_control} show some basic usage of the control structures.
+The example a C macro or function that behaves similarly to the FORTRAN modulo
+intrinsic. The sample script (inefficiently) computes and stores in \emph{j} the
+number of multiples of $7$ in the 0 to 100 range. It uses the For and If control
+structures. A Ruby block is passed to each of those constructs. This block is
+evaluated at the time the construct is printed. If several such constructs are
+needed (in an \emph{if elsif else} case for instance) they can be explicitly
+passed as parameters using the \emph{lambda} Ruby keyword.
 
-  Procedure
+\begin{figure}
+\lstset{style=BOAST,
+        caption={BOAST code}, 
+        label={lst:BOAST_procedure},
+        numbers=left }
+\begin{lstlisting}
+n = Int("n")
+ndat = Int("ndat")
+x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
+y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(0, n-1) ])
+p = Procedure("magicfilter", [n, ndat, x, y])
+open p
+close p
+\end{lstlisting}
+
+\lstset{style=customf,
+        caption={FORTRAN output.},
+        label={lst:BOAST_procedure_FORTRAN} }
+
+\begin{lstlisting}
+SUBROUTINE magicfilter(n, ndat, x, y)
+  integer(kind=4) :: n
+  integer(kind=4) :: ndat
+  real(kind=8), intent(in), dimension(0:n - (1), 1:ndat + 1 - (1)) :: x
+  real(kind=8), intent(out), dimension(1:ndat + 1 - (1), 0:n - (1)) :: y
+END SUBROUTINE magicfilter
+\end{lstlisting}
+\lstset{style=customc,
+        caption={C output.},
+        label={lst:BOAST_procedure_C} }
+
+\begin{lstlisting}
+void magicfilter(int32_t n, int32_t ndat, const double * x, double * y){
+}
+\end{lstlisting}
+\caption{BOAST Code Snippet for Procedure}
+\label{fig:BOAST_procedure}
+\end{figure}
+
+The last control structure is Procedure. It is used to describe either procedure
+or functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
+abstraction. It is the signature of a real kernel from BigDFT. This kernel uses
+an input array \emph{x} and an output array \emph{y} of double precision
+numbers. Those array have two dimensions which depend on input variables
+\emph{n} and \emph{ndat}. We can see here the use of the \emph{open} and
+\emph{close} keywords that are used to print a control structure without an
+associated Ruby block. This time we placed ourselves inside of BOAST namespace.
 
@@
 %transformation process. Indeed, one may first use BOAST on order to generate a
-%program (a specific syntaxic tree) and then benefit from the transformation
+%program (a specific syntactic tree) and then benefit from the transformation
 %facilities offered by one of the above tools.
@@
   \subsection{BOAST Runtime}
+
+In the previous Section we presented BOAST's language. This allows us to
+describe procedures and functions and to meta-program them using Ruby. But in
+order to find the best version of a computing kernel each version has to be
+compiled, linked and then executed to assess it's performance. This can be very
+time consuming if this process cannot be automated. By enabling more versions to
+be evaluated the automation will bring improved portability, better performance
+and in the end improve the productivity of the developer.
+
+In this section we will present the different aspects of BOAST runtime that
+allow this automation.
 

### 2014-09-11 Tutorial/Tutorial-BOAST-2.tex
Updated tutorial

@@
 
+\lstdefinestyle{BOAST}{
+  language=Ruby,
+  basicstyle=\tiny\ttfamily,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={BOAST, pr, decl, open, close}
+}
+
 \title{BOAST}
@@
 
-\date{\textbf{BOAST Tutorial}\\June 25, 2014}
+\date{\textbf{BOAST Tutorial}\\September 19, 2014}
 
@@
 \end{lstlisting}
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 Simple BOAST commands:
+\lstset{style=BOAST}
 \begin{lstlisting}
@@
 \frametitle{Defining a Procedure}
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 Simple BOAST Procedure:
+\lstset{style=BOAST}
 \footnotesize
@@
 05:0> p = BOAST::Procedure( "test_proc", [a,b] )
-06:0> BOAST::decl p
+06:0> BOAST::open p
 SUBROUTINE test_proc(a, b)
@@
 \frametitle{Defining a Full Procedure:}
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
+\lstset{style=BOAST}
 \tiny
@@
 011:0> b = BOAST::Real( "b", :dir => :out)
-012:0> p = BOAST::Procedure( "test_proc", [a,b] ) { BOAST::print b === a + 2 }
+012:0> p = BOAST::Procedure( "test_proc", [a,b] ) {
+012:1>   BOAST::pr b === a + 2
+012:2> }
 013:0> BOAST::lang = BOAST::FORTRAN
-014:0> BOAST::print p
+014:0> BOAST::pr p
 SUBROUTINE test_proc(a, b)
@@
 \frametitle{Creating a Computing Kernel}
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
+\lstset{style=BOAST}
 \tiny
@@
   BOAST::decl i = BOAST::Int( "i" )
-  BOAST::For( i, 1, n ) {
-    BOAST::print b[i] === a[i] + 2
-  }.print
+  BOAST::pr BOAST::For( i, 1, n ) {
+    BOAST::pr b[i] === a[i] + 2
+  }
 }
 k = BOAST::CKernel::new
-BOAST::print p
+BOAST::pr p
 k.procedure = p
@@
 \frametitle{Running a Computing Kernel}
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
+\lstset{style=BOAST}
 \tiny
@@
   BOAST::decl i = BOAST::Int( "i" )
-  BOAST::For( i, 1, n ) {
-    BOAST::print b[i] === a[i] + 2
-  }.print
+  BOAST::pr BOAST::For( i, 1, n ) {
+    BOAST::pr b[i] === a[i] + 2
+  }
 }
 k = BOAST::CKernel::new
-BOAST::print p
+BOAST::pr p
 k.procedure = p
 k.run(input.length, input, output)
-(output - input).each { |val| raise "Error!" if (val-2).abs > 1e-15 }
+(output - input).each {
+  |val| raise "Error!" if (val-2).abs > 1e-15 }
 stats = k.run(input.length, input, output)
@@
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
+\lstset{style=BOAST}
 \begin{lstlisting}
@@
   i = Int("i",{:signed => false})
-  print p = Procedure(function_name, [n,a,b,c]) {
+  pr p = Procedure(function_name, [n,a,b,c]) {
     decl i
     if [CL, CUDA].include?(get_lang) then
-      print i === get_global_id(0)
-      print c[i] === a[i] + b[i]
+      pr i === get_global_id(0)
+      pr c[i] === a[i] + b[i]
     else
-      print For(i,0,n-1) {
-        print c[i] === a[i] + b[i]
+      pr For(i,0,n-1) {
+        pr c[i] === a[i] + b[i]
       }
@@
 \frametitle{Running the Kernel}
+\lstset{style=BOAST}
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 \begin{lstlisting}
@@
 c_ref = NArray.float(n)
-
 epsilon = 10e-15
-
 set_lang( FORTRAN )
@@
 k.run(n,a,b,c_ref)
-
 [C, CL, CUDA].each { |lang|
@@
 Sample:
+\lstset{style=BOAST}
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 \begin{lstlisting}
@@
 \end{itemize}
+\lstset{style=BOAST}
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 \begin{lstlisting}
@@
 \frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : BOAST (1)}
+\lstset{style=BOAST}
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 \begin{lstlisting}
@@
 \frametitle{SPECFEM3D assemble\_boundary\_potential\_on\_device : BOAST (2)}
+\lstset{style=BOAST}
 \tiny
-\lstset{language=Ruby,
-basicstyle=\ttfamily,
-keywordstyle=\color{javapurple}\bfseries,
-stringstyle=\color{javared},
-commentstyle=\color{javagreen},
-numbers=left,
-tabsize=4,
-showspaces=false,
-showstringspaces=false}
 \begin{lstlisting}
@@
   elsif(get_lang == CUDA or get_lang == CL) then
-    decl p
+    open p
     id         = Int("id")
@@
     decl id, iglob, iloc, iinterface
-    print id === get_global_id(0) + get_global_size(0)*get_global_id(1)
-    print For(iinterface, 0, num_interfaces-1) {
-      print If(id<d_nibool_interfaces[iinterface]) {
-        print iloc === id + max_nibool_interfaces*iinterface
-        print iglob === d_ibool_interfaces[iloc] - 1
-        print atomicAdd(d_potential_dot_dot_acoustic + iglob, \
+    pr id === get_global_id(0) + get_global_size(0)*get_global_id(1)
+    pr For(iinterface, 0, num_interfaces-1) {
+      pr If(id<d_nibool_interfaces[iinterface]) {
+        pr iloc === id + max_nibool_interfaces*iinterface
+        pr iglob === d_ibool_interfaces[iloc] - 1
+        pr atomicAdd(d_potential_dot_dot_acoustic + iglob, \
                      d_send_potential_dot_dot_buffer[iloc])
@@
 \end{frame}
-\begin{frame}
-\frametitle{BigDFT Synthesis Kernel}
-\centering 
-\includegraphics[scale=0.5]{Res_synthesis}\\
-
-\end{frame}
+%\begin{frame}
+%\frametitle{BigDFT Synthesis Kernel}
+%\centering 
+%\includegraphics[scale=0.5]{Res_synthesis}\\
+%
+%
+%\end{frame}
 

### 2014-09-11 Tutorial/Tutorial-BOAST.tex
Updated tutorial

--- 2014-08-29 Tutorial-BOAST/Tutorial-BOAST.tex
@@
 
-\date{\textbf{Tutorial BOAST}\\June 25, 2014}
+\date{\textbf{Tutorial BOAST}\\September 19, 2014}
 
@@
 p = Procedure("magic_filter", [n,ndat,x,y], [lowfil,upfil])
-decl p
+open p
 \end{lstlisting}
@@
   unroll = 5
-  print For(j,1,ndat-(unroll-1), unroll) {
+  pr For(j,1,ndat-(unroll-1), unroll) {
   #.....
-    (tt2 === tt2 + x[k,j+1]*fil[l]).print
+    pr tt2 === tt2 + x[k,j+1]*fil[l]
   #.....
@@
 \begin{frame}
+\frametitle{BigDFT Synthesis Kernel}
+\centering 
+\includegraphics[scale=0.5]{Res_synthesis}\\
+
+\end{frame}
+
+\begin{frame}
 \frametitle{Improvement for BigDFT}

### 2014-09-19 ARMclbench/Laplacian.rb
Improved ARM test

@@ -12,7 +12,6 @@ def rndup( val, div)
   return (val%div) == 0 ? val : val + div - (val%div)
 end
 
-set_lang(CL)
 set_array_start(0)
 register_funccall("clamp")
 
@@ -24,6 +23,34 @@ def laplacian_ref( input, output, width, height )
                           - input[x_i - 3, y_i + 1] - input[x_i, y_i + 1] - input[x_i + 3, y_i + 1] ).clamp(0,255)
     }
   }
+end
+
+def laplacian_c_ref
+  k = CKernel::new
+  height = Int("height", :dir => :in)
+  width = Int("width", :dir => :in)
+  w = Int("w")
+  pdst = Int("pdst", :dir => :out,  :signed => false, :size => 1, :dim => [ Dim(w), Dim(height) ] )
+  psrc = Int("psrc", :dir => :in, :signed => false, :size => 1, :dim => [ Dim(w), Dim(height) ] )
+  p = Procedure("math", [psrc, pdst, width, height]) {
+    decl i = Int("i")
+    decl j = Int("j")
+    decl tmp = Int("tmp")
+    decl w
+    pr w === width * 3
+    pr For(j, 1, height-2) {
+      pr For(i, 3, w-4) {
+         
+        pr tmp === ( -psrc[i - 3, j - 1] - psrc[i, j - 1] - psrc[i + 3, j - 1]\
+                    - psrc[i - 3, j]     + psrc[i, j] * 9 - psrc[i + 3, j]\
+                    - psrc[i - 3, j + 1] - psrc[i, j + 1] - psrc[i + 3, j + 1] )
+        pr pdst[i,j] === Ternary(tmp < 0, 0, Ternary(tmp>255, 255, tmp))
+      }
+    }
+  }
+  pr p
+  k.procedure = p
+  return k
 end
 
 def laplacian(x_component_number = 1, vector_length=1, y_component_number = 1, temporary_size = 4)
@@ -142,16 +169,28 @@ end
 #puts laplacian(4,2)
 #puts laplacian(3,2)
 
+sizes = [[768, 432], [2560, 1600], [2048, 2048], [5760, 3240], [7680, 4320]]
+inputs = []
+refs = []
+results = []
 width = 1024
 height = 512
 
-input = NArray.byte(width*3,height).random(256)
-output_ref = NArray.byte(width*3,height)
-output = NArray.byte(width*3,height)
+set_lang(C)
 
[…]

### 2014-09-19 Paper/BOAST.tex
Changed open in opn

@@
   captionpos=b,
-  morekeywords={BOAST, pr, decl, open, close}
+  morekeywords={BOAST, pr, decl, opn, close}
 }
@@
 related to manipulations in the host language four keywords were defined. They
-are \emph{pr}, \emph{decl}, \emph{open} and \emph{close}. As the language is
+are \emph{pr}, \emph{decl}, \emph{opn} and \emph{close}. As the language is
 an EDSL, these four keywords are methods in the BOAST namespace.
@@
 
-The \emph{open} method can be used to print the beginning of a control
+The \emph{opn} method can be used to print the beginning of a control
 structure without an associated code block.
 
-The \emph{close} method is the counterpart to the \emph{open} method. It is used
+The \emph{close} method is the counterpart to the \emph{opn} method. It is used
 to close a control structure without an associated code block.
@@
 p = Procedure("magicfilter", [n, ndat, x, y])
-open p
+opn p
 close p
@@
 numbers. Those array have two dimensions which depend on input variables
-\emph{n} and \emph{ndat}. We can see here the use of the \emph{open} and
+\emph{n} and \emph{ndat}. We can see here the use of the \emph{opn} and
 \emph{close} keywords that are used to print a control structure without an

### 2014-09-19 Tutorial/Tutorial-BOAST-2.tex
Changed open in opn

@@
 05:0> p = BOAST::Procedure( "test_proc", [a,b] )
-06:0> BOAST::open p
+06:0> BOAST::opn p
 SUBROUTINE test_proc(a, b)
@@
   elsif(get_lang == CUDA or get_lang == CL) then
-    open p
+    opn p
     id         = Int("id")

### 2014-09-25 Paper/BOAST.tex
Improved Paper

@@
 \begin{lstlisting}
-n = Int("n")
-ndat = Int("ndat")
+n = Int("n", :dir => :in)
+ndat = Int("ndat", :dir => :in)
 x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
@@
 SUBROUTINE magicfilter(n, ndat, x, y)
-  integer(kind=4) :: n
-  integer(kind=4) :: ndat
-  real(kind=8), intent(in), dimension(0:n - (1), 1:ndat + 1 - (1)) :: x
-  real(kind=8), intent(out), dimension(1:ndat + 1 - (1), 0:n - (1)) :: y
+  integer(kind=4), intent(in) :: n
+  integer(kind=4), intent(in) :: ndat
+  real(kind=8), intent(in), dimension(0:n - (1), ndat) :: x
+  real(kind=8), intent(out), dimension(ndat, 0:n - (1)) :: y
 END SUBROUTINE magicfilter
@@
 associated Ruby block. This time we placed ourselves inside of BOAST namespace.
+
+The generated outputs in FORTRAN~\ref{lst:BOAST_procedure_FORTRAN} and
+C~\ref{lst:BOAST_procedure_C} show the difference in meta-information
 

### 2014-09-25 Paper/BOAST.tex
Improved Paper

@@
 \begin{lstlisting}
-void magicfilter(int32_t n, int32_t ndat, const double * x, double * y){
+void magicfilter(const int32_t n, const int32_t ndat,
+                 const double * x, double * y)
 }
@@
 The generated outputs in FORTRAN~\ref{lst:BOAST_procedure_FORTRAN} and
-C~\ref{lst:BOAST_procedure_C} show the difference in meta-information
+C~\ref{lst:BOAST_procedure_C} show the difference in meta-information that is
+kept between both versions.
 
@@
 In this section we will present the different aspects of BOAST runtime that
-allow this automation.
-
+allow this automation. Those aspects are listed here:
   \begin{itemize}
@@
   \end{itemize}
+
+  \subsubsection{Multi-target Language Generation}
+
+Language availability and performance can vary between platforms. It is thus
+important to be able to express computing kernels in different languages, based
+on the availability and their respective merits on the target platform. Some
+languages can have additional features, such as languages that target GPUs
+(OpenCL, CUDA). The developer should be given tools to determine what kind of
+language is currently used in order to be able to use those additional features.
+
+Similarly the target language must be changed with ease in order to compare
+different alternatives. Two methods are dedicated to this task: \emph{set\_lang}
+and \emph{get\_lang}. The target language can also be set through environment
+variable before launching BOAST, allowing for easy command line scripting.
+
+  \subsubsection{Compilation}
+
+Compilation of the generated kernels must also be very flexible. Indeed,
+
+  \subsubsection{Execution}
+  \subsubsection{Kernel Replay}
 

### 2014-09-25 ARMclbench/Laplacian.rb
generated source with correct array declaration

@@ -272,7 +272,7 @@ opt_space = GenericOptimization::new( :x_component_number => [1,2,4,8,16],
                                       :temporary_size     => [2,4],
                                       :vector_recompute   => [true, false],
                                       :load_overlap       => [true, false] )
-check = true
+check = false
 opt_space.each_random { |opt|
   id = opt.to_s            
   puts id

### 2014-09-25 SPECFEM3D/output/compute_acoustic_kernel_cl.c

@@ -121,14 +121,14 @@ __kernel void compute_acoustic_kernel(const __global int * ibool, const __global
   int ijk_ispec;\n\
   int ijk_ispec_padded;\n\
   int iglob;\n\
-  float accel_elm[3];\n\
-  float b_displ_elm[3];\n\
+  float accel_elm[(3)];\n\
+  float b_displ_elm[(3)];\n\
   float rhol;\n\
   float kappal;\n\
   float div_displ;\n\
   float b_div_displ;\n\
-  __local float scalar_field_displ[NGLL3];\n\
-  __local float scalar_field_accel[NGLL3];\n\
+  __local float scalar_field_displ[(NGLL3)];\n\
+  __local float scalar_field_accel[(NGLL3)];\n\
   ispec = get_group_id(0) + (get_group_id(1)) * (get_num_groups(0));\n\
   if (ispec < NSPEC) {\n\
     ijk = get_local_id(0);\n\

### 2014-09-25 SPECFEM3D/output/compute_ani_kernel_cl.c

@@ -61,8 +61,8 @@ inline void atomicAdd(volatile __global float *source, const float val) {\n\
 #define BLOCKSIZE_TRANSFER 256\n\
 #endif\n\
 static void compute_strain_product(float * prod, const float eps_trace_over_3, const float * epsdev, const float b_eps_trace_over_3, const float * b_epsdev){\n\
-  float eps[6];\n\
-  float b_eps[6];\n\
+  float eps[(6)];\n\
+  float b_eps[(6)];\n\
   int p;\n\
   int i;\n\
   int j;\n\
@@ -101,9 +101,9 @@ __kernel void compute_ani_kernel(const __global float * epsilondev_xx, const __g
   int ijk_ispec;\n\
   float eps_trace_over_3;\n\
   float b_eps_trace_over_3;\n\
-  float prod[21];\n\
-  float epsdev[5];\n\
-  float b_epsdev[5];\n\
+  float prod[(21)];\n\
+  float epsdev[(5)];\n\
+  float b_epsdev[(5)];\n\
   ispec = get_group_id(0) + (get_group_id(1)) * (get_num_groups(0));\n\
   if (ispec < NSPEC) {\n\
     ijk_ispec = get_local_id(0) + (NGLL3) * (ispec);\n\

### 2014-09-25 SPECFEM3D/output/compute_ani_undo_att_kernel_cl.c

@@ -153,8 +153,8 @@ static void compute_element_strain_undo_att(const int ispec, const int ijk_ispec
   *(epsilon_trace_over_3) = templ;\n\
 }\n\
 static void compute_strain_product(float * prod, const float eps_trace_over_3, const float * epsdev, const float b_eps_trace_over_3, const float * b_epsdev){\n\
-  float eps[6];\n\
-  float b_eps[6];\n\
+  float eps[(6)];\n\
+  float b_eps[(6)];\n\
   int p;\n\
   int i;\n\
   int j;\n\
@@ -194,14 +194,14 @@ __kernel void compute_ani_undo_att_kernel(const __global float * epsilondev_xx, 
   int iglob;\n\
   float eps_trace_over_3;\n\
   float b_eps_trace_over_3;\n\
-  float prod[21];\n\
+  float prod[(21)];\n\
   int i;\n\
-  float epsdev[5];\n\
-  float b_epsdev[5];\n\
-  __local float s_dummyx_loc[NGLL3];\n\
-  __local float s_dummyy_loc[NGLL3];\n\
-  __local float s_dummyz_loc[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
+  float epsdev[(5)];\n\
+  float b_epsdev[(5)];\n\
+  __local float s_dummyx_loc[(NGLL3)];\n\
+  __local float s_dummyy_loc[(NGLL3)];\n\
+  __local float s_dummyz_loc[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
   ispec = get_group_id(0) + (get_group_id(1)) * (get_num_groups(0));\n\
   ijk_ispec = get_local_id(0) + (NGLL3) * (ispec);\n\
   tx = get_local_id(0);\n\

### 2014-09-25 SPECFEM3D/output/compute_iso_undo_att_kernel_cl.c

@@ -159,12 +159,12 @@ __kernel void compute_iso_undo_att_kernel(const __global float * epsilondev_xx, 
   int iglob;\n\
   float eps_trace_over_3;\n\
   float b_eps_trace_over_3;\n\
-  float epsdev[5];\n\
-  float b_epsdev[5];\n\
-  __local float s_dummyx_loc[NGLL3];\n\
-  __local float s_dummyy_loc[NGLL3];\n\
-  __local float s_dummyz_loc[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
+  float epsdev[(5)];\n\
+  float b_epsdev[(5)];\n\
+  __local float s_dummyx_loc[(NGLL3)];\n\
+  __local float s_dummyy_loc[(NGLL3)];\n\
+  __local float s_dummyz_loc[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
   ispec = get_group_id(0) + (get_group_id(1)) * (get_num_groups(0));\n\
   ijk_ispec = get_local_id(0) + (NGLL3) * (ispec);\n\
   tx = get_local_id(0);\n\

### 2014-09-25 SPECFEM3D/output/crust_mantle_impl_kernel_adjoint_cl.c

@@ -490,20 +490,20 @@ __kernel __attribute__((reqd_work_group_size(NGLL3_PADDED,1,1)))  void crust_man
   float rho_s_H1;\n\
   float rho_s_H2;\n\
   float rho_s_H3;\n\
-  __local float s_dummyx_loc[NGLL3];\n\
-  __local float s_dummyy_loc[NGLL3];\n\
-  __local float s_dummyz_loc[NGLL3];\n\
-  __local float s_tempx1[NGLL3];\n\
-  __local float s_tempx2[NGLL3];\n\
-  __local float s_tempx3[NGLL3];\n\
-  __local float s_tempy1[NGLL3];\n\
-  __local float s_tempy2[NGLL3];\n\
-  __local float s_tempy3[NGLL3];\n\
-  __local float s_tempz1[NGLL3];\n\
-  __local float s_tempz2[NGLL3];\n\
-  __local float s_tempz3[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
-  __local float sh_hprimewgll_xx[NGLL2];\n\
+  __local float s_dummyx_loc[(NGLL3)];\n\
+  __local float s_dummyy_loc[(NGLL3)];\n\
+  __local float s_dummyz_loc[(NGLL3)];\n\
+  __local float s_tempx1[(NGLL3)];\n\
+  __local float s_tempx2[(NGLL3)];\n\
+  __local float s_tempx3[(NGLL3)];\n\
+  __local float s_tempy1[(NGLL3)];\n\
+  __local float s_tempy2[(NGLL3)];\n\
+  __local float s_tempy3[(NGLL3)];\n\
+  __local float s_tempz1[(NGLL3)];\n\
+  __local float s_tempz2[(NGLL3)];\n\
+  __local float s_tempz3[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
+  __local float sh_hprimewgll_xx[(NGLL2)];\n\
   bx = (get_group_id(1)) * (get_num_groups(0)) + get_group_id(0);\n\
   tx = get_local_id(0);\n\
   K = (tx) / (NGLL2);\n\

### 2014-09-25 SPECFEM3D/output/get_maximum_scalar_kernel_cl.c

@@ -61,7 +61,7 @@ inline void atomicAdd(volatile __global float *source, const float val) {\n\
 #define BLOCKSIZE_TRANSFER 256\n\
 #endif\n\
 __kernel void get_maximum_scalar_kernel(const __global float * array, const int size, __global float * d_max){\n\
-  __local float sdata[BLOCKSIZE_TRANSFER];\n\
+  __local float sdata[(BLOCKSIZE_TRANSFER)];\n\
   int tid;\n\
   int bx;\n\
   int i;\n\

### 2014-09-25 SPECFEM3D/output/get_maximum_vector_kernel_cl.c

@@ -61,7 +61,7 @@ inline void atomicAdd(volatile __global float *source, const float val) {\n\
 #define BLOCKSIZE_TRANSFER 256\n\
 #endif\n\
 __kernel void get_maximum_vector_kernel(const __global float * array, const int size, __global float * d_max){\n\
-  __local float sdata[BLOCKSIZE_TRANSFER];\n\
+  __local float sdata[(BLOCKSIZE_TRANSFER)];\n\
   int tid;\n\
   int bx;\n\
   int i;\n\

### 2014-09-25 SPECFEM3D/output/inner_core_impl_kernel_forward_cl.c

@@ -283,20 +283,20 @@ __kernel __attribute__((reqd_work_group_size(NGLL3_PADDED,1,1)))  void inner_cor
   float rho_s_H1;\n\
   float rho_s_H2;\n\
   float rho_s_H3;\n\
-  __local float s_dummyx_loc[NGLL3];\n\
-  __local float s_dummyy_loc[NGLL3];\n\
-  __local float s_dummyz_loc[NGLL3];\n\
-  __local float s_tempx1[NGLL3];\n\
-  __local float s_tempx2[NGLL3];\n\
-  __local float s_tempx3[NGLL3];\n\
-  __local float s_tempy1[NGLL3];\n\
-  __local float s_tempy2[NGLL3];\n\
-  __local float s_tempy3[NGLL3];\n\
-  __local float s_tempz1[NGLL3];\n\
-  __local float s_tempz2[NGLL3];\n\
-  __local float s_tempz3[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
-  __local float sh_hprimewgll_xx[NGLL2];\n\
+  __local float s_dummyx_loc[(NGLL3)];\n\
+  __local float s_dummyy_loc[(NGLL3)];\n\
+  __local float s_dummyz_loc[(NGLL3)];\n\
+  __local float s_tempx1[(NGLL3)];\n\
+  __local float s_tempx2[(NGLL3)];\n\
+  __local float s_tempx3[(NGLL3)];\n\
+  __local float s_tempy1[(NGLL3)];\n\
+  __local float s_tempy2[(NGLL3)];\n\
+  __local float s_tempy3[(NGLL3)];\n\
+  __local float s_tempz1[(NGLL3)];\n\
+  __local float s_tempz2[(NGLL3)];\n\
+  __local float s_tempz3[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
+  __local float sh_hprimewgll_xx[(NGLL2)];\n\
   bx = (get_group_id(1)) * (get_num_groups(0)) + get_group_id(0);\n\
   tx = get_local_id(0);\n\
   K = (tx) / (NGLL2);\n\

### 2014-09-25 SPECFEM3D/output/outer_core_impl_kernel_adjoint_cl.c

@@ -135,12 +135,12 @@ __kernel void outer_core_impl_kernel_adjoint(const int nb_blocks_to_compute, con
   float grad_y_ln_rho;\n\
   float grad_z_ln_rho;\n\
   int int_radius;\n\
-  __local float s_dummy_loc[NGLL3];\n\
-  __local float s_temp1[NGLL3];\n\
-  __local float s_temp2[NGLL3];\n\
-  __local float s_temp3[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
-  __local float sh_hprimewgll_xx[NGLL2];\n\
+  __local float s_dummy_loc[(NGLL3)];\n\
+  __local float s_temp1[(NGLL3)];\n\
+  __local float s_temp2[(NGLL3)];\n\
+  __local float s_temp3[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
+  __local float sh_hprimewgll_xx[(NGLL2)];\n\
   bx = (get_group_id(1)) * (get_num_groups(0)) + get_group_id(0);\n\
   tx = get_local_id(0);\n\
   K = (tx) / (NGLL2);\n\

### 2014-09-25 SPECFEM3D/output/outer_core_impl_kernel_forward_cl.c

@@ -135,12 +135,12 @@ __kernel void outer_core_impl_kernel_forward(const int nb_blocks_to_compute, con
   float grad_y_ln_rho;\n\
   float grad_z_ln_rho;\n\
   int int_radius;\n\
-  __local float s_dummy_loc[NGLL3];\n\
-  __local float s_temp1[NGLL3];\n\
-  __local float s_temp2[NGLL3];\n\
-  __local float s_temp3[NGLL3];\n\
-  __local float sh_hprime_xx[NGLL2];\n\
-  __local float sh_hprimewgll_xx[NGLL2];\n\
+  __local float s_dummy_loc[(NGLL3)];\n\
+  __local float s_temp1[(NGLL3)];\n\
+  __local float s_temp2[(NGLL3)];\n\
+  __local float s_temp3[(NGLL3)];\n\
+  __local float sh_hprime_xx[(NGLL2)];\n\
+  __local float sh_hprimewgll_xx[(NGLL2)];\n\
   bx = (get_group_id(1)) * (get_num_groups(0)) + get_group_id(0);\n\
   tx = get_local_id(0);\n\
   K = (tx) / (NGLL2);\n\

### 2014-09-29 Paper/BOAST.tex
Improved paper

@@
 
-Compilation of the generated kernels must also be very flexible. Indeed,
+Compilation of the generated kernels must also be very flexible. Indeed, HPC
+applications may encounter platforms with very diverse compilation environment.
+Proprietary and dedicated compilers are common on HPC infrastructures. Thus
+BOAST build system shows similar behavior than common build system. Compilers
+and their compile/build options can be specified. This can be done at several
+places. Here is the list by increasing order of precedence: BOAST configuration
+file, environment variables and at kernel build time.
+
+This way the framework to test different compiler optimizations is completely
+available and performance study can include both kernel related parameters and
+compilation related parameters. This behavior is contained in the \emph{CKernel}
+class of BOAST. When instantiating this class a BOAST \emph{Procedure}
+representing the entry point of the kernel is specified.
 
   \subsubsection{Execution}
+
+The next logical step is to benchmark the built kernel. Boast offers a simple
+way to run a kernel that was successfully built while staying inside of BOAST.
+A built BOAST \emph{CKernel} exposes a run method that accepts arguments corresponding
+to the BOAST \emph{Procedure} used to instantiate the kernel. Arrays must be
+instances of \emph{NArray} which are numerical arrays that use C array
+underneath.
+
+Arrays which correspond to output parameters will be modified during the
+execution of the kernel so results can be checked. This allows for easy non
+regression testing. The run method also returns information (and result for
+kernels that are functions) about the run. For instance the run time of the
+kernel is one such information, it is obtained using the system wide realtime
+clock. Other probes can be inserted at compile time if needed.
+
   \subsubsection{Kernel Replay}
+
+The non regression methodology presented before is viable provided input data
+can be generated at runtime. Unfortunately it is not always possible, some
+applications have very complicated data patterns that are difficult to
+synthesize without running the full application. Thus BOAST offers a way to load
+binary data from the file system and use them as inputs of a kernel. Outputs can
+also be checked against those binary data thus enabling (almost) data oblivious
+non regression testing.
+
+In order to use this methodology one has to be able to trace an application to
+get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
+in the SPECFEM3D use case.
 

### 2014-09-29 Paper/BOAST.tex
Improved bibliography

@@
 
+Code generation and auto-tuning techniques are not new. Nonetheless, with recent
+developments in hardware, and the HPC landscape being as diverse as it is now,
+there is a renewed interest in the field. This related work Section is split in
+three parts, focusing first on tools that provide a DSL to describe computing
+kernels. Autotuning frameworks will then be presented. Last, optimization space
+pruners and their ties to auto-tuning will be introduced.
+
   \subsection{Kernel Description DSL}
 
-  Orio
+  The idea to describe computing kernels using a Domain Specific Language has
+already been used. SPIRAL~\cite{puschel2004spiral} is a decade old generation
+framework for signal processing. It uses a proprietary DSL, SPL (signal
+processing language), to describe a DSP algorithm. This DSL is then transformed
+into efficient programs in high level languages like C or FORTRAN.
+POET~\cite{yi2007poet} also uses a DSL to describe custom code transformations,
+like loop unrolling, loop blocking and loop interchange. Those transformation
+can be parametrized in order to tune the application. Orio~\cite{Hart2009:Orio}
+can be compared to POET as it aims at describing possible code transformations
+using a DSL.
 
-  Halide
-
-  Spiral
-
-  POET
+  Halide~\cite{ragan2013halide} is closer in some ways to our approach as it
+uses an embedded C++ DSL to describe image processing algorithm. This DSL allows
+decoupling the algorithm description from its scheduling. Each pixel in the
+resulting image has a completely defined dependency tree with regard to pixels
+in the input image (and intermediary results). During generation, depending on
+memory and computing cost some values can be recomputed rather than fetched from
+memory. We used Halide to implement the \emph{magicfilter} of BigDFT, but
+unfortunately results were 4 to 5 times slower than the one we obtain with
+BOAST. We speculate that the three-dimensional nature of our convolutions, as
+well as the filter length, can be considered edge cases in halide and quite far
+from the intended target.
 

### 2014-10-01 Paper/BOAST.tex
More biblio
Improved Article

@@
 there is a renewed interest in the field. This related work Section is split in
-three parts, focusing first on tools that provide a DSL to describe computing
-kernels. Autotuning frameworks will then be presented. Last, optimization space
+three parts, focusing first on Auto-tuning frameworks. Tools that provide a DSL
+to describe computing kernels will then be presented. Last, optimization space
 pruners and their ties to auto-tuning will be introduced.
+
+  \subsection{Application Auto-Tuning}
+
+  The most convenient way to obtain an application that can be auto-tuned on a
+given platform is to base this application on a widely used computing library.
+BLAS~\cite{dongarra1990set} and LAPACK~\cite{laug} are such libraries. These
+libraries are either hand tuned for selected platforms or have auto-tuned
+implementations. Atlas~\cite{whaley04} is an auto-tunes implementation of
+BLAS/LAPACK. ATLAS authors defined the \emph{Automated Empirical Optimization of
+Software} methodology that we implemented with BOAST. Their kernel generation is
+done using macro-functions in C.
+
+  Nonetheless many applications formalism cannot be reduced to standardized
+library or border on what could be considered edge cases for those libraries and
+not as optimized as more general cases. Orio~\cite{Hart2009:Orio} is an
+auto-tuning framework that has an approach close of that of BOAST but they are
+based on an annotated DSL describing loop transformation rather than a more
+generic EDSL. Halide~\cite{ragan2013halide} is an auto-tuning framework
+dedicated to image processing. It can also be used to describe other
+operations on memory buffers. Those two frameworks propose automated search
+space exploration to find the best version of a kernel.
 
@@
 can be compared to POET as it aims at describing possible code transformations
-using a DSL.
+using a DSL. All those approaches are very different from our as they put the
+emphasis on compilation techniques whereas boast relies on the user to express
+the different optimizations.
 
@@
 from the intended target.
-
-  \subsection{Application Autotuning}
-
-  Atlas
-
-  Orio
-
-  Halide
 

### 2014-10-14 Paper/BOAST.tex
improved bibliography

@@
 
-  Collective Mind
+  Once auto-tuning technique are used the parameter space explodes and the
+systematic sampling rapidly becomes impossible to achieve. The cost of finding
+the optimal kernel parameters and environment parameters (compiler flags, used
+language, ...) is rapidly prohibitive. Dedicated frameworks have been developed
+to address this problem. Adaptive Sampling Kit (ASK)~\cite{castro2013adaptive}
+is one such tool. It reduces the number of sample needed by creating a model of
+the performances and by minimizing the number of samples needed to find the
+parameters of this model. Several models and sampling techniques are
+implemented.
 
-  Adaptive Sampling Kit (ASK)
+  Collective Mind~\cite{fursin:hal-01054763} proposes similar techniques to
+solve the problem but also stores and the results and complete experimental
+setups in databases for future reference and reproducibility. This database
+approach also enables easy parallelization of the experimental process.
+Collective Mind also proposes collections of flags for many available
+compilers/versions easing the exploration process.
 

### 2014-10-16 BigDFT/Wavelet_test.rb
Added options to Wavelet generator.

@@ -41,10 +41,31 @@ L = [-0.0033824159510050025955,
      -0.014952258337062199118,
      -0.00030292051472413308126,
       0.0018899503327676891843]
-
+$options = { :unroll => 6, :step => 1 }
+$parser = OptionParser::new do |opts|
+  opts.on("-s","--step VAL","Unroll step") { |step|
+    $options[:step] = step.to_i
+  }
+  opts.on("-u","--unroll VAL","Unroll bound") { |unroll|
+    $options[:unroll] = unroll.to_i
+  }
+  opts.on("-f","--force_unroll","Force outer unroll") {
+    $options[:force_unroll] = true
+  }
+  opts.on("-o","--output","output code") {
+    $options[:output] = true
+  }
+  opts.parse!
+end
 epsilon = 10e-12
 
-optims = GenericOptimization::new(:unroll_range => 6, :mod_arr_test => true, :tt_arr_test => true)
+if $options[:force_unroll] then
+  unroll_range = [$options[:unroll],$options[:unroll]]
+else
+  unroll_range = [1,$options[:unroll],$options[:step]]
+end
+
+optims = GenericOptimization::new(:unroll_range => unroll_range, :mod_arr_test => true, :tt_arr_test => true)
 wave_filter = WaveletFilter::new("sym#{L.length/2}", L)
 n1 = 62
 n2 = 66
@@ -69,6 +90,15 @@ k1 = Wavelet(wave_filter, :decompose, optims)
 k1.build(:openmp => true)
 k2 = Wavelet(wave_filter, :recompose, optims)
 k2.build(:openmp => true)
+
+if $options[:output] then
+  suffix = ".c" if BOAST::get_lang == BOAST::C
+  suffix = ".f90" if BOAST::get_lang == BOAST::FORTRAN
+  File::open("wavelets#{suffix}","w") { |f|
+    f.puts k1
+    f.puts k2
+  }
+end
 
 repeat = 5
 begin

### 2014-10-21 Paper/specfem.tex
add first draft of SPECFEM use case

\newcommand{\productname}[1]{\textsc{#1}}
\newcommand{\latin}[1]{\textit{#1}}

\newcommand{\Specfem}{\productname{Specfem 3D}\xspace}
\newcommand{\Cuda}{\productname{Cuda}\xspace}
\newcommand{\eg}{\latin{e.g.,}\xspace}
\newcommand{\OCL}{\productname{OpenCL}\xspace}
\newcommand{\code}[1]{\texttt{#1}}

\newcommand{\etc}[1]{\latin{etc}}

\section{BOAST Use Cases}

\subsection{Porting SPECFEM3D to OpenCL using BOAST}

\subsubsection{SPECFEM3D}

In this last subsection, we study a free seismic wave propagation
simulator, \Specfem, in its \productname{Cartesian}
flavor\footnote{Specfem3d Cartesian --- CIG
  \url{http://www.geodynamics.org/cig/software/specfem3d}}. \Specfem
simulates seismic wave propagation at the local or regional scale
based upon spectral-element method (SEM), with very good accuracy and
convergence properties. ADD BENCHMARK APP FOR HPC SUPERCOMPUTERS.

When we started to work on the project (version v2.1 of July 2013) it
supported graphics card GPU acceleration through NVidia \Cuda. This
GPU support comes in addition to the MPI support implemented to enable
parallel computing. Most of \Specfem code-base is written in
\productname{Fortran2003}, only the GPU related parts are written in
C.  The split between CPU and GPU code was done at a rather fine
grain, as the application counted more than 40 GPU kernels. Some of
them were quite simple (\eg performing a few vector operations, but at
massively parallel scale), while at the other side of the spectrum,
some complex kernels took more than 80 parameters and perform very
specific physical transformations.

Because of the complexity of wave propagation (and of the application
architecture), it is hard to impossible to define unit tests. Hence,
the application is validated by the accuracy of the results it
produces, that is, the accuracy of seismograms of the simulated
earthquakes, in comparison with the actual ones. Non-regression
testing (after new developments) is also based on these seismograms,
with measurements of the relative error between two identical
simulations.

As part of the Mont-Blanc project, we had to port \Specfem to \OCL, so
that it could be use to benchmark Mont-Blanc HPC platforms.

\subsubsection{Porting to OpenCL}

\subparagraph{Porting kernels to BOAST} Nvidia \Cuda and \OCL are
based on the \emph{same} programming model: a massively parallel
accelerator running in a distinct memory environment. Thanks to that
proximity, we, BOAST developers, have been able to carry out most of
the porting task with only a limited knowledge about \Specfem internal
physics.

That lack of knowledge led us to be particularly attentive to the path
we undertook for the porting, as we would have been unable to
understand how and why the application was not operating properly, if
it was to fail.

Hence, our first milestone in the porting was translating \Specfem
\Cuda kernels to BOAST EDSL. This way, we could ask BOAST framework to
generate CUDA versions of the kernels, plug them back into \Specfem and
get (after fixing compile-time errors---prototypes and naming mistakes
mainly) a first set of \Specfem seismograms.

As we had expected, the seismograms were erroneous. But with the help
of shell scripts and BOAST framework ability to store and provide the
kernels' original source code, we built a set of \Specfem binaries
including only \emph{one} BOAST-generated kernel, with all others
reference-kernels. Running and validating all these binaries enabled
use to pinpoint the misbehaving kernels. We finished the debugging
with a side-by-side comparison that highlighted the coding mistakes.

\subparagraph{Porting runtime to \OCL} The second part of the porting
consisted in translating the CPU-side of the application from \Cuda
API to \OCL API. Most of the functions of the interfaces are very
similar, with only naming-convention and data-structure
distinctions. Hence, it was clear that automatic rewriting tools
(namely \code{sed} regexp and \code{emacs-list} functions) could be
useful. To give an idea of the cost of a \emph{manual} rewriting, we
can count (in \# of \OCL API function calls): 70 kernel ``function
calls'', 790 arguments to set, 230 memory transfers, 160 buffer
creations and 270 releases.

Once all the automatic (and many manual) transformations applied,
compilation errors fixed, and \OCL unsuccessful function calls solved,
the application managed to complete its execution and generated a
first set of seismograms. And again, as expected and feared, these
seismograms were not valid.

As we had already validated BOAST-generated kernels (and trusted \Cuda
and \OCL versions to be semantically identical), we knew that the bugs
were now in the CPU side, and we had to find a way to understand where
\Specfem's \Cuda version of the code diverged from its \OCL
counterpart. To help us in that purpose, we had a strong assumption:
both versions of the code were supposed to perform exactly the same
operations, with the same ``logical'' parameters (the APIs have
\emph{implementation} differences, for instance \OCL has two memory
transfer functions, \code{clEnqueueReadBuffer},
\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a
direction parameter \code{cudaMemcpy(..., dir)}, but above that it is
the same functionality nonetheless).

Hence, our idea to locate the execution problems was to make sure that
both execution actually did the same thing. As the \OCL results were
invalid, we knew the executions would diverge at one or several
points.

\subparagraph{Debugging \OCL Execution: \code{GPUTrace}} In order to
ensure that \Cuda and \OCL executions were doing the same operations,
we developed a tool that could intercept all the function calls
related to the GPU and store them for offline processing, in a format
common to \OCL and \Cuda. Our solution was inspired by \code{strace}
and \code{ltrace} tools: it dynamically preloads a library (with the
help \code{LD\_PRELOAD} in \productname{GNU/Linux}-based environments)
between the application and the GPU library, and prints the arguments
(input \emph{and} output) of the relevant function calls.

However, in contrast with \code{strace} and \code{ltrace},
\code{GPUTrace} has to be \emph{state-full}. Indeed, most of \OCL and
\Cuda parameters are handles to opaque types, so, in order to provide
more meaningful information, \code{GPUTrace} stores internally
information about these handles (buffer creation size and index,
kernel name and prototype, \etc). % more details here?

Once \code{GPUTrace} was operational, it was easy to confront \Cuda
and \OCL outputs with a graphical \code{diff} tool, and spot the
different porting mistakes: two parameters reversed, an offset
incorrectly applied, \etc. One last problem remained, impossible to
miss because of the seismograms not matching (but with just a slight
drift). We added more verbosity to \code{GPUTrace} output, first the
initial bits of the GPU memory buffers, then their full content. The
drift was visible in the trace, but it was nonetheless unclear where
it started. We finally got it after hours of code review, both on
BOAST kernels and \OCL code: one kernel was \emph{three}-dimensional,
whereas the others were two-dimensional. But for all of them, only two
dimensions were passed. As it was the same person who coded the \OCL
port and the GPU tracer, this third dimension was forgotten in both
side\ldots

\subparagraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
merged in \Specfem's development tree and under test and extension by
different research teams. On a platform with two K40x GPUs and x Intel
Xeon processors, we measured similar results between the original
\Cuda version and our BOAST-\Cuda version. With the same set of
optimization, BOAST \Cuda and \OCL version also reported similar
execution time spans. The best execution speed (25\% higher) was
achieved with \Cuda version though, as one optimization parameter
(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, as of version 1.1, to
\OCL runtime.

By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
shrank by a factor of 1.5 (from xxx LOC to xxx, in part because of
code duplication, but by removing manually unrolled loops). This is
beneficial for \Specfem as it improves the readability and
maintainability of its source code.

We have also been able to enhance \Specfem's non-regression testsuite
by adding per-kernel non-regression tests. This was done with the help
of \code{GPUTrace}, that we used to capture all the input parameters
of a particular \emph{valid} kernel execution, as well as the output
values. Then, during the non-regression testing, BOAST framework can
load these buffer files, allocate and write them on the GPU memory
through \Cuda or \OCL runtime, and trigger the kernel execution. A
comparison of the output values (for instance against a maximal error
level) validates the non-regression.

In the same mindset, we provided \Specfem testsuite with kernel
performance evaluation mechanisms. These tests will help developers to
try new optimization in kernel code and measure their impact, without
executing the whole application.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Local Variables:      %%%
%%% mode: latex           %%%
%%% TeX-master: "journal" %%%
%%% TeX-PDF-mode: t       %%%
%%% End:                  %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-11-21 Paper/BOAST.tex
add first draft of SPECFEM use case
Forst modifications in the paper for BigDFT

@@
-\documentclass[review]{elsarticle}
-
-\usepackage{lineno,hyperref}
+\documentclass%[review]
+              {elsarticle}
+\usepackage[utf8]{inputenc}
+\usepackage{hyperref}
 \usepackage{listings}
 \usepackage{color}
+\usepackage{xspace}
+
+%\usepackage{lineno}
+%\modulolinenumbers[5]
+%\linenumbers
 
@@
 \usepackage{url}
-\modulolinenumbers[5]
 
@@
 \end{frontmatter}
-
-\linenumbers
 
@@
   \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
+In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of four European
+laboratories (L\_Sim - CEA Grenoble, Basel University - Switzerland, Louvain-la-Neuve University - Belgium and
+Kiel University - Germany), with the aim of developing a novel approach for DFT calculations
+based on Daubechies wavelets.
+Rather than simply building a DFT code from scratch, the objective of this three-years
+project was to test the potential benefit of a new formalism in the context of electronic structure
+calculations.
+
+As a matter of fact, Daubechies wavelets exhibit a set of properties which make them ideal
+for a precise and optimized DFT approach. In particular, their systematicity allows to provide a
+reliable basis set for high-precision results, whereas their locality (both in real and reciprocal space)
+is highly desired to improve the efficiency and the flexibility of the treatment. Indeed, a localized
+basis set allows to optimize the number of degrees of freedom for a required accuracy (see \cite{bigdft}),
+which is highly desirable given the complexity and inhomogeneity of the systems under investigation
+nowadays.
+
+The application is mainly written in Fortran and currently
+includes about 360 thousands lines of Fortran code and 70 thousands lines of C language, which accounts for more than 50 \% 
+of the code base. It is a parallel application based on the standards
+MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports CUDA~\cite{cuda} and
+OpenCL~\cite{opencl}.  
+In the recent years this code has been used for many scientific applications,
+and its development and user consortium is continuously growing.
+Massively parallel computations are routinely executed with the BigDFT code, either in homogeneous or 
+hybrid architectures. 
+In 2009, the French Bull-Fourier was awarded for the implementation of the hybrid version of BigDFT.
+
+%As it is used on systems that may
+%have very different architectures, it is of great importance to be able to
+%optimize and run the application according to the specific underlying platform.
 
@@
 
-BigDFT~\cite{bigdft} is a scientific application computing the electronic
-structure of systems using density functional
-theory (DFT). The application is mainly written in Fortran and currently
-includes some 193,000 lines of Fortran code, which accounts for more than 50 \% 
-of the code base. It is a parallel application based on the standards
-MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports CUDA~\cite{cuda} and
-OpenCL~\cite{opencl}.  As it is used on systems that may
-have very different architectures, it is of great importance to be able to
-optimize and run the application according to the specific underlying platform.
+In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to wavefunctions (orbitals), which are
+represented arrays of floating point numbers. In wavelets formalism, the operators are written via convolutions with short,
+separable filters.
+The detailed description of how these operations are defined is beyond the scope of this paper and can be found
+in the BigDFT reference paper \cite{bigdft}.
 
-BigDFT is characterized by its extensive use of
-convolution operators~\cite{nussbaumer1982fast} applied on large arrays of
-data.  One example of a specific convolution, called
+% Convolutions are among the basic processing blocks of BigDFT. Special care has to be
+% taken regarding their performances. 
+The CPU convolutions of BigDFT have thus been thoroughly optimized.
+In a recent paper, the optimization of the CPU convolutions of BigDFT has been extensively considered.
+%BigDFT is characterized by its extensive use of
+%convolution operators~\cite{nussbaumer1982fast} applied on large arrays of
+%data.  
+One example of a specific convolution, called
 MagicFilter~\cite{Genovese|CAS2010}, can be seen in Listing~\ref{lst:conv_example}. 
@@
 vary.  Various optimizations can be applied to this treatment and
-may focus on the loop structure, as well as on the size of the data.  In this
-paper we focus on this particular convolution and explore different
-optimizations whose nature is presented in detail in the next section.
+may focus on the loop structure, as well as on the size of the data. 
+%In this
+%paper we focus on this particular convolution and explore different
+%optimizations whose nature is presented in detail in the next section.
 
     \subsubsection{A Generic Convolution Library}
+The number of convolution kernels needed in BigDFT has been continuously growing in the recent years.
+Various boundary conditions and functionalities have been added, making the BigDFT more and more powerful
+in terms of scientific applications.
+However, the cost of the maintainance and of development of the convolutions is always a 
+delicate point to be considered while including a new functionality.
+The convolution patterns are usually rather similar, leading to code duplication 
+and difficulties in code reading and modification.
 
-  BigDFT implementation of convolution/wavelets is a reference but is limited in scope (1 wavelet family ands its derived operators).
-  Generalizing to other wavelet families long and difficult if performance are to remain good.
+It appears therefore very interesting to benefit from a automatic tool to drive the implementation and the 
+generation of new convolutions.
+This would lead to optimized code, adapted to  different computing platforms, that
+is optimally factorized.
+In addition to this point, the help of such code generator is also important to
+build \emph{new science}: the cost of implementing new convolutions would become so little that other 
+functionalities (for example, the generalisation of the BigDFT convolutions to Neumann boundary 
+conditions  or the usage of wavelet-on-the-interval basis) can be added with limited manpower.
 
-Idea:
-\begin{itemize}
-\item write a generic convolution library in BOAST
-\item BLAS like interface
-\item Provide a wide variety of building blocks
-\item Optimize thos blocks using BOAST
-\item Build either from source or directly link with binary
-\item Release complete library
-\end{itemize}
+For these reasons, a convolution library has been engineered with the help of BOAST.
+The detailed API of the library will be described elsewhere.
+The spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled and called 
+from high-level operations.
+Each of the building blocks of these convolution library is optimally tuned by BOAST by choosing
+the sources providing the optimal kernel for the chosen computing platform.
+The sources of these kernels are then collected and compiled to meet the API of the library.
 
-  Characteristic use case driven
+A library written in this way might have an impact going largely beyond the community of BigDFT developers.
+Indeed, convolutions are basic operations of lots of scientific application codes, like for example
+finite differences approaches, which are universally used in computational physics.
 
-  Convolution Library: first step toward a dsl dedicated to wavelet processing for Physics computing.
+The interest in having a robuts and optimally tuned library in this scientific field is therefore evident.
+Techniques are under investigation to provide also fine-tuned binaries to the end-user rather than
+the code sources, such that more agressive interprocedural optimisation can be performed.
+This experience would be therefore a first step toward the release of a tunable optimised 
+convolution library oriented to computational physics communities.
+%   BigDFT implementation of convolution/wavelets is a reference but is limited in scope (1 wavelet family ands its derived operators).
+%   Generalizing to other wavelet families long and difficult if performance are to remain good.
+% 
+% Idea:
+% \begin{itemize}
+% \item write a generic convolution library in BOAST
+% \item BLAS like interface
+% \item Provide a wide variety of building blocks
+% \item Optimize thos blocks using BOAST
+% \item Build either from source or directly link with binary
+% \item Release complete library
+% \end{itemize}
+% 
+% 
+%   Characteristic use case driven
+% 
+%   Convolution Library: first step toward a dsl dedicated to wavelet processing for Physics computing.
 
-  \subsection{Porting SPECFEM3D to OpenCL using BOAST}
-
-    \subsubsection{SPECFEM3D}
-
-\begin{itemize}
-
-\item Wave propagation research software
-
-\item HPC application (benchmark on supercomputers)
-
-\item Complex GPU (CUDA) code, 40+ kernels.
-
-\item 80+ parameters for kernels. Difficult to validate.
-\end{itemize}
-
-    \subsubsection{Porting to OpenCL}
-
-  Idea:
-\begin{itemize}
-\item write kernels in BOAST,
-\item generate kernels in CUDA,
-\item validate kernels by comparing CUDA/generated CUDA executions using comparator
-\item generate kernels in OpenCL,
-\item develop OpenCL runtime backend using previously validated OpenCL kernels
-\item validate OpenCL version by comparing CUDA/OpenCL executions using comparator
-\end{itemize}
-
-  Evaluation:
-\begin{itemize}
-\item Similar CUDA/OpenCL
-\item Factoring : more compact code
-\item Integrated non regression support
-\item Integrated Kernel performance evaluation
-\end{itemize}
+\input{specfem}
 

### 2014-12-01 Paper/BOAST.tex
Improved Paper.

@@
 
-\section{Use Cases}
+\section{BOAST Use Cases}
 \label{use_cases}
 
-  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
-In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of four European
-laboratories (L\_Sim - CEA Grenoble, Basel University - Switzerland, Louvain-la-Neuve University - Belgium and
-Kiel University - Germany), with the aim of developing a novel approach for DFT calculations
-based on Daubechies wavelets.
-Rather than simply building a DFT code from scratch, the objective of this three-years
-project was to test the potential benefit of a new formalism in the context of electronic structure
-calculations.
+In this section we will present two scientific applications that use BOAST. The
+first one, BigDFT, uses boast in order to develop new functionalities with
+performance portability in mind. The second one, SPECFEM3D, uses BOAST to
+factorize OpenCL and CUDA development, while having robust non regression
+tests.
 
-As a matter of fact, Daubechies wavelets exhibit a set of properties which make them ideal
-for a precise and optimized DFT approach. In particular, their systematicity allows to provide a
-reliable basis set for high-precision results, whereas their locality (both in real and reciprocal space)
-is highly desired to improve the efficiency and the flexibility of the treatment. Indeed, a localized
-basis set allows to optimize the number of degrees of freedom for a required accuracy (see \cite{bigdft}),
-which is highly desirable given the complexity and inhomogeneity of the systems under investigation
-nowadays.
+  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using
+BOAST}
 
-The application is mainly written in Fortran and currently
-includes about 360 thousands lines of Fortran code and 70 thousands lines of C language, which accounts for more than 50 \% 
-of the code base. It is a parallel application based on the standards
-MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports CUDA~\cite{cuda} and
-OpenCL~\cite{opencl}.  
-In the recent years this code has been used for many scientific applications,
-and its development and user consortium is continuously growing.
-Massively parallel computations are routinely executed with the BigDFT code, either in homogeneous or 
-hybrid architectures. 
-In 2009, the French Bull-Fourier was awarded for the implementation of the hybrid version of BigDFT.
+In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of
+four European laboratories (L\_Sim - CEA Grenoble, Basel University -
+Switzerland, Louvain-la-Neuve University - Belgium and Kiel University -
+Germany), with the aim of developing a novel approach for DFT calculations
+based on Daubechies wavelets.  Rather than simply building a DFT code from
+scratch, the objective of this three-years project was to test the potential
+benefit of a new formalism in the context of electronic structure calculations.
+
+As a matter of fact, Daubechies wavelets exhibit a set of properties which make
+them ideal for a precise and optimized DFT approach. In particular, their
+systematicity allows to provide a reliable basis set for high-precision
+results, whereas their locality (both in real and reciprocal space) is highly
+desired to improve the efficiency and the flexibility of the treatment. Indeed,
+a localized basis set allows to optimize the number of degrees of freedom for a
+required accuracy (see \cite{bigdft}), which is highly desirable given the
+complexity and inhomogeneity of the systems under investigation nowadays.
+
+The application is mainly written in Fortran and currently includes about 360
+thousands lines of Fortran code and 70 thousands lines of C language, which
+accounts for more than 50 \% of the code base. It is a parallel application
+based on the standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also
+supports CUDA~\cite{cuda} and OpenCL~\cite{opencl}.  In the recent years this
+code has been used for many scientific applications, and its development and
+user consortium is continuously growing.  Massively parallel computations are
+routinely executed with the BigDFT code, either in homogeneous or hybrid
+architectures.  In 2009, the French Bull-Fourier was awarded for the
+implementation of the hybrid version of BigDFT.
 
@@
 
-In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to wavefunctions (orbitals), which are
-represented arrays of floating point numbers. In wavelets formalism, the operators are written via convolutions with short,
-separable filters.
-The detailed description of how these operations are defined is beyond the scope of this paper and can be found
-in the BigDFT reference paper \cite{bigdft}.
+In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to
+wavefunctions (orbitals), which are represented arrays of floating point
+numbers. In wavelets formalism, the operators are written via convolutions with
+short, separable filters.  The detailed description of how these operations are
+defined is beyond the scope of this paper and can be found in the BigDFT
+reference paper \cite{bigdft}.
 
-% Convolutions are among the basic processing blocks of BigDFT. Special care has to be
-% taken regarding their performances. 
-The CPU convolutions of BigDFT have thus been thoroughly optimized.
-In a recent paper, the optimization of the CPU convolutions of BigDFT has been extensively considered.
-%BigDFT is characterized by its extensive use of
-%convolution operators~\cite{nussbaumer1982fast} applied on large arrays of
-%data.  
+% Convolutions are among the basic processing blocks of BigDFT. Special care
+% has to be taken regarding their performances. 
+The CPU convolutions of BigDFT have thus been thoroughly optimized.  In a
+recent paper, the optimization of the CPU convolutions of BigDFT has been
+extensively considered.
+%BigDFT is characterized by its extensive use of convolution
+%operators~\cite{nussbaumer1982fast} applied on large arrays of data.  
 One example of a specific convolution, called
-MagicFilter~\cite{Genovese|CAS2010}, can be seen in Listing~\ref{lst:conv_example}. 
-It applies a filter $filt$ to the data set $in$ and then stores the result in the data set $out$ with a
+MagicFilter~\cite{Genovese|CAS2010}, can be seen in
+Listing~\ref{lst:conv_example}.  It applies a filter $filt$ to the data set
+$in$ and then stores the result in the data set $out$ with a
 transposition~\cite{Goedecker1993}.
@@
 
-As we can see, there are three nested loops working on arrays whose sizes
-vary.  Various optimizations can be applied to this treatment and
-may focus on the loop structure, as well as on the size of the data. 
-%In this
-%paper we focus on this particular convolution and explore different
+As we can see, there are three nested loops working on arrays whose sizes vary.
+Various optimizations can be applied to this treatment and may focus on the
+loop structure, as well as on the size of the data. 
+%In this paper we focus on this particular convolution and explore different
 %optimizations whose nature is presented in detail in the next section.
@@
     \subsubsection{A Generic Convolution Library}
-The number of convolution kernels needed in BigDFT has been continuously growing in the recent years.
-Various boundary conditions and functionalities have been added, making the BigDFT more and more powerful
-in terms of scientific applications.
-However, the cost of the maintainance and of development of the convolutions is always a 
-delicate point to be considered while including a new functionality.
-The convolution patterns are usually rather similar, leading to code duplication 
-and difficulties in code reading and modification.
 
-It appears therefore very interesting to benefit from a automatic tool to drive the implementation and the 
-generation of new convolutions.
-This would lead to optimized code, adapted to  different computing platforms, that
-is optimally factorized.
-In addition to this point, the help of such code generator is also important to
-build \emph{new science}: the cost of implementing new convolutions would become so little that other 
-functionalities (for example, the generalisation of the BigDFT convolutions to Neumann boundary 
-conditions  or the usage of wavelet-on-the-interval basis) can be added with limited manpower.
+The number of convolution kernels needed in BigDFT has been continuously
+growing in the recent years.  Various boundary conditions and functionalities
+have been added, making the BigDFT more and more powerful in terms of
+scientific applications.  However, the cost of the maintainance and of
+development of the convolutions is always a delicate point to be considered
+while including a new functionality.  The convolution patterns are usually
+rather similar, leading to code duplication and difficulties in code reading
+and modification.
 
-For these reasons, a convolution library has been engineered with the help of BOAST.
-The detailed API of the library will be described elsewhere.
-The spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled and called 
-from high-level operations.
-Each of the building blocks of these convolution library is optimally tuned by BOAST by choosing
-the sources providing the optimal kernel for the chosen computing platform.
-The sources of these kernels are then collected and compiled to meet the API of the library.
+It appears therefore very interesting to benefit from a automatic tool to drive
+the implementation and the generation of new convolutions.  This would lead to
+optimized code, adapted to  different computing platforms, that is optimally
+factorized.  In addition to this point, the help of such code generator is also
+important to build \emph{new science}: the cost of implementing new
+convolutions would become so little that other functionalities (for example,
+the generalisation of the BigDFT convolutions to Neumann boundary conditions
+or the usage of wavelet-on-the-interval basis) can be added with limited
+manpower.
 
-A library written in this way might have an impact going largely beyond the community of BigDFT developers.
-Indeed, convolutions are basic operations of lots of scientific application codes, like for example
-finite differences approaches, which are universally used in computational physics.
+For these reasons, a convolution library has been engineered with the help of
+BOAST.  The detailed API of the library will be described elsewhere.  The
+spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
+and called from high-level operations.  Each of the building blocks of these
+convolution library is optimally tuned by BOAST by choosing the sources
+providing the optimal kernel for the chosen computing platform.  The sources of
+these kernels are then collected and compiled to meet the API of the library.
 
-The interest in having a robuts and optimally tuned library in this scientific field is therefore evident.
-Techniques are under investigation to provide also fine-tuned binaries to the end-user rather than
-the code sources, such that more agressive interprocedural optimisation can be performed.
-This experience would be therefore a first step toward the release of a tunable optimised 
+A library written in this way might have an impact going largely beyond the
+community of BigDFT developers.  Indeed, convolutions are basic operations of
+lots of scientific application codes, like for example finite differences
+approaches, which are universally used in computational physics.
+
+The interest in having a robuts and optimally tuned library in this scientific
+field is therefore evident.  Techniques are under investigation to provide also
+fine-tuned binaries to the end-user rather than the code sources, such that
+more agressive interprocedural optimisation can be performed.  This experience
+would be therefore a first step toward the release of a tunable optimised
 convolution library oriented to computational physics communities.
-%   BigDFT implementation of convolution/wavelets is a reference but is limited in scope (1 wavelet family ands its derived operators).
-%   Generalizing to other wavelet families long and difficult if performance are to remain good.
+%   BigDFT implementation of convolution/wavelets is a reference but is limited
+%   in scope (1 wavelet family ands its derived operators).  Generalizing to
+%   other wavelet families long and difficult if performance are to remain
+%   good.
 % 
-% Idea:
-% \begin{itemize}
-% \item write a generic convolution library in BOAST
-% \item BLAS like interface
-% \item Provide a wide variety of building blocks
-% \item Optimize thos blocks using BOAST
-% \item Build either from source or directly link with binary
-% \item Release complete library
-% \end{itemize}
+% Idea: \begin{itemize} \item write a generic convolution library in BOAST
+% \item BLAS like interface \item Provide a wide variety of building blocks
+% \item Optimize thos blocks using BOAST \item Build either from source or
+% directly link with binary \item Release complete library \end{itemize}
 % 
@@
 % 
-%   Convolution Library: first step toward a dsl dedicated to wavelet processing for Physics computing.
+%   Convolution Library: first step toward a dsl dedicated to wavelet
+%   processing for Physics computing.
 

### 2014-12-02 Tutorial/Tutorial-BOAST.tex
Changed open in opn
Added option to For class creation. :step is now an option.

@@
 p = Procedure("magic_filter", [n,ndat,x,y], [lowfil,upfil])
-open p
+opn p
 \end{lstlisting}
@@
   unroll = 5
-  pr For(j,1,ndat-(unroll-1), unroll) {
+  pr For(j,1,ndat-(unroll-1), step: unroll) {
   #.....

### 2014-12-02 BigDFT/GenericConvolution.rb

@@ -795,10 +795,10 @@ class ConvolutionOperator1d
       ises0 = startendpoints(@dims[@dim_indexes[0]], unro == @dim_indexes[0], unrolling_length, reliq)
       BOAST::get_output.print("!$omp do\n") if BOAST::get_lang == BOAST::FORTRAN
       BOAST::get_output.print("#pragma omp for\n") if BOAST::get_lang == BOAST::C
-      BOAST::For(iters[@dim_indexes[0]], *ises0 ) {
+      BOAST::For(iters[@dim_indexes[0]], ises0[0], ises0[1], step: ises0[2] ) {
         if @dim_indexes.length == 3 then
           ises1 = startendpoints(@dims[@dim_indexes[1]], unro == @dim_indexes[1], unrolling_length, reliq)
-          BOAST::For(iters[@dim_indexes[1]], *ises1) {
+          BOAST::For(iters[@dim_indexes[1]], ises1[0], ises1[1], step: ises1[2]) {
             conv_lines(iters, l, t, tlen, unro, mods, unroll_inner)
           }.pr
         else

### 2014-12-02 BigDFT/old/AnaRotPer-2.rb

@@ -385,7 +385,7 @@ EOF
       inner_op = lambda { |ci,di,fBC|
         ci.each{ |s| (s === 0.0).print }
         di.each{ |s| (s === 0.0).print }
-        For::new(l,lowfil,upfil,2) {
+        For::new(l,lowfil,upfil,step: 2) {
           (k === l + i*2).print if fBC
           ci.each_index{ |ind|
             #(k === FuncCall::new( "modulo", l+ i*2, n*2)).print if not fBC
@@ -418,8 +418,8 @@ EOF
         #external loop, with unrolling. the rest is excluded
         @@output.print("!$omp do\n") if BOAST::get_lang == BOAST::FORTRAN
         if unro > 1 then
-          #forJ1 = For::new(j,js,ntot-(unro-1), unro) #do not forget to add the rest
-          forJ1 = For::new(j,js,(ntot/unro)*unro, unro) #do not forget to add the rest
+          #forJ1 = For::new(j,js,ntot-(unro-1), step: unro) #do not forget to add the rest
+          forJ1 = For::new(j,js,(ntot/unro)*unro, step: unro) #do not forget to add the rest
         else
           forJ1 = For::new(j,js,ntot)
         end

### 2014-12-02 BigDFT/old/AnaRotPer.rb

@@ -109,7 +109,7 @@ module BOAST
       se.each{ |s| s.decl }
 
       if unroll > 0 then
-        for1 = For::new(j,1,ndat-(unroll-1), unroll)
+        for1 = For::new(j,1,ndat-(unroll-1), step: unroll)
       else
         for1 = For::new(j,1,ndat)
       end

### 2014-12-02 BigDFT/old/Kinetic-2.rb

@@ -100,7 +100,7 @@ module ConvolutionGenerator
     kinetic_1d = lambda { |t,js|
       unro=t.length
       @@output.print("!$omp do\n") if ConvolutionGenerator::get_lang == ConvolutionGenerator::FORTRAN
-      For::new(j,js,(ndat/unro)*unro, unro) {
+      For::new(j,js,(ndat/unro)*unro, step: unro) {
         For::new(i, 0, -lowfil) {
           for_uniq.call(t,true)
         }.print

### 2014-12-02 BigDFT/old/MagicFilter_GPU.rb

@@ -144,7 +144,7 @@ EOF
       (igtf === Ternary::new(igt < 0, igt + size_n, Ternary::new(igt >= n, igt - size_n, igt))).print
       (tmp_buff[j2+0,i2] === x[jgt, igtf]).print
       (igt === igt + wgs).print
-      f1 = For::new(k, filt.length/2, size_n+filt.length/2-1-1-(wgs-1), wgs){
+      f1 = For::new(k, filt.length/2, size_n+filt.length/2-1-1-(wgs-1), step: wgs){
         (igtf === igt).print
         (tmp_buff[j2+k,i2] === x[jgt, igtf]).print
         (igt === igt + wgs).print
@@ -164,7 +164,7 @@ EOF
       tt.decl
       l = Variable::new( "l",Int)
       l.decl
-      f1 = For::new(k, 0, size_n-1-(wgs-1), wgs) {
+      f1 = For::new(k, 0, size_n-1-(wgs-1), step: wgs) {
         (tt === 0.0).print
         f2 = For::new(l, 0, filt.length-1) {
           (tt === tt + tmp_buff[k+l+i2, j2]*fil[l]).print
@@ -245,7 +245,7 @@ EOF
       (igt === j2).print
       (tmp_buff[igt,i2] === x[jgt, igt]).print
       (igt === igt + wgs).print
-      For::new(k, wgs, size_n-wgs-1, wgs){
+      For::new(k, wgs, size_n-wgs-1, step: wgs){
         (tmp_buff[igt,i2] === x[jgt, igt]).print
         (igt === igt + wgs).print
       }.unroll
@@ -379,7 +379,7 @@ EOF
       (igt === j2).print
       (tmp_buff[igt,i2] === x[jgt, igt]).print
       (igt === igt + wgs).print
-#      For::new(k, wgs, size_n-wgs-1, wgs){
+#      For::new(k, wgs, size_n-wgs-1, step: wgs){
 #        (tmp_buff[igt,i2] === x[jgt, igt]).print
 #        (igt === igt + wgs).print
 #      }.unroll

### 2014-12-02 BigDFT/old/Synthesis.rb

@@ -311,7 +311,7 @@ EOF
 
 
       if unroll > 0 then
-        for1 = For::new(j,1,ndat-(unroll-1), unroll)
+        for1 = For::new(j,1,ndat-(unroll-1), step: unroll)
       else
         for1 = For::new(j,1,ndat)
       end

### 2014-12-02 CacheTest/KernelRead.rb

@@ -53,7 +53,7 @@ module BOAST
         @@output.print "  _mm256_set1_pd(0);\n"
       end
       For::new(i, 1, m_cycles*m_stride) {
-        For::new(j, m_start, buffer_size + m_start - m_stride*unrolled, m_stride*unrolled) {
+        For::new(j, m_start, buffer_size + m_start - m_stride*unrolled, step: m_stride*unrolled) {
           unrolled.times { |k|
              if machine == "sse3" then
                (sumV === FuncCall::new("_mm_add_epi64", sumV, buffer[j+m_stride*k])).print
@@ -64,7 +64,7 @@ module BOAST
              end
                 }
         }.print
-        For::new(j, buffer_size + m_start - FuncCall::new( "modulo", buffer_size+m_start, m_stride*unrolled),  buffer_size + m_start - 1, m_stride) {
+        For::new(j, buffer_size + m_start - FuncCall::new( "modulo", buffer_size+m_start, m_stride*unrolled),  buffer_size + m_start - 1, step: m_stride) {
            if machine == "sse3" then
              (sumV === FuncCall::new("_mm_add_epi64", sumV, buffer[j])).print
            elsif machine == "neon" then
@@ -112,12 +112,12 @@ module BOAST
       sum.decl
       (sum === 0).print
       For::new(i, 1, m_cycles*m_stride) {
-        For::new(j, m_start, buffer_size + m_start - m_stride*unrolled, m_stride*unrolled) {
+        For::new(j, m_start, buffer_size + m_start - m_stride*unrolled, step: m_stride*unrolled) {
           unrolled.times { |k|
             (sum === sum + buffer[j+m_stride*k]).print
           }
         }.print
-        For::new(j, buffer_size + m_start - FuncCall::new( "modulo", buffer_size+m_start, m_stride*unrolled),  buffer_size + m_start - 1, m_stride) {
+        For::new(j, buffer_size + m_start - FuncCall::new( "modulo", buffer_size+m_start, m_stride*unrolled),  buffer_size + m_start - 1, step: m_stride) {
             (sum === sum + buffer[j]).print
         }.print
       }.print

### 2014-12-02 CacheTest/KernelRead2.rb

@@ -41,12 +41,12 @@ module BOAST
       print sumV.set(0)
 
       print For(i, 1, m_cycles*m_stride) {
-        print For(j, m_start, elem_n + m_start - m_stride*unrolled, m_stride*unrolled) {
+        print For(j, m_start, elem_n + m_start - m_stride*unrolled, step: m_stride*unrolled) {
           unrolled.times { |k|
             print sumV === sumV + buffer[j+m_stride*k]
           }
         }
-        print For(j, elem_n + m_start - modulo(elem_n+m_start, m_stride*unrolled),  elem_n + m_start - 1, m_stride) {
+        print For(j, elem_n + m_start - modulo(elem_n+m_start, m_stride*unrolled),  elem_n + m_start - 1, step: m_stride) {
           print sumV === sumV + buffer[j]
         }
       }

### 2014-12-15 Paper/BOAST.tex
spelling
strange formulation, where will it be described?

@@
 
-It appears therefore very interesting to benefit from a automatic tool to drive
+It appears therefore very interesting to benefit from an automatic tool to drive
 the implementation and the generation of new convolutions.  This would lead to
@@
 For these reasons, a convolution library has been engineered with the help of
-BOAST.  The detailed API of the library will be described elsewhere.  The
+BOAST.  *The detailed API of the library will be described elsewhere*.  The
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled

### 2014-12-15 Paper/BOAST.tex
plural (those)

@@
 For these reasons, a convolution library has been engineered with the help of
-BOAST.  *The detailed API of the library will be described elsewhere*.  The
+BOAST.  \textbf{The detailed API of the library will be described elsewhere}.  The
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
 and called from high-level operations.  Each of the building blocks of these
-convolution library is optimally tuned by BOAST by choosing the sources
+convolution libraries is optimally tuned by BOAST by choosing the sources
 providing the optimal kernel for the chosen computing platform.  The sources of

### 2014-12-16 Paper/BOAST.tex

@@
 fine-tuned binaries to the end-user rather than the code sources, such that
-more agressive interprocedural optimisation can be performed.  This experience
+more agressive interprocedural optimization can be performed.  This experience
 would be therefore a first step toward the release of a tunable optimised

### 2014-12-16 Paper/specfem.tex
Improved Paper.
Removed useless section.

@@
 \newcommand{\etc}[1]{\latin{etc}}
-
-\section{BOAST Use Cases}
 
@@
 supported graphics card GPU acceleration through NVidia \Cuda. This
-GPU support comes in addition to the MPI support implemented to enable
-parallel computing. Most of \Specfem code-base is written in
+GPU support came in addition to the MPI support implemented to enable
+multi-CPU parallel computing. Most of \Specfem code-base is written in
 \productname{Fortran2003}, only the GPU related parts are written in
@@
 \subparagraph{Porting kernels to BOAST} Nvidia \Cuda and \OCL are
-based on the \emph{same} programming model: a massively parallel
-accelerator running in a distinct memory environment. Thanks to that
-proximity, we, BOAST developers, have been able to carry out most of
-the porting task with only a limited knowledge about \Specfem internal
-physics.
+based on the same programming model: a massively parallel accelerator
+running in another memory environment. Thanks to that proximity, we,
+BOAST developers, have been able to carry out most of the porting task
+with only a limited knowledge about \Specfem internal physics.
 
@@
 
-Hence, our first milestone in the porting was translating \Specfem
-\Cuda kernels to BOAST EDSL. This way, we could ask BOAST framework to
-generate CUDA versions of the kernels, plug them back into \Specfem and
-get (after fixing compile-time errors---prototypes and naming mistakes
-mainly) a first set of \Specfem seismograms.
+Hence, our first milestone in the porting was the translation of
+\Specfem's \Cuda kernels into BOAST EDSL. This way, we could ask BOAST
+framework to generate a CUDA version of the kernels, plug them back
+into \Specfem and get (after fixing compile-time errors---prototypes
+and naming mistakes mainly) a first set of \Specfem seismograms.
 
@@
 \subparagraph{Porting runtime to \OCL} The second part of the porting
-consisted in translating the CPU-side of the application from \Cuda
-API to \OCL API. Most of the functions of the interfaces are very
-similar, with only naming-convention and data-structure
+consisted in the translation of the CPU-side of the application, from
+\Cuda API to \OCL API. Most of the functions of the interfaces are
+very similar, with only naming-convention and data-structure
 distinctions. Hence, it was clear that automatic rewriting tools
-(namely \code{sed} regexp and \code{emacs-list} functions) could be
+(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
 useful. To give an idea of the cost of a \emph{manual} rewriting, we
@@
 
-Once all the automatic (and many manual) transformations applied,
+Once all the automatic (and many manual) transformations were applied,
 compilation errors fixed, and \OCL unsuccessful function calls solved,
-the application managed to complete its execution and generated a
-first set of seismograms. And again, as expected and feared, these
+the application managed to complete its execution and generate a first
+set of seismograms. And again, as expected and feared, these
 seismograms were not valid.
@@
 \code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a
-direction parameter \code{cudaMemcpy(..., dir)}, but above that it is
-the same functionality nonetheless).
+direction parameter \code{cudaMemcpy(..., dir)}, but above that, it is
+the same functionalities).
 
@@
 information about these handles (buffer creation size and index,
-kernel name and prototype, \etc). % more details here?
+kernel name and prototype, \etc{}). % more details here?
 
@@
 different porting mistakes: two parameters reversed, an offset
-incorrectly applied, \etc. One last problem remained, impossible to
+incorrectly applied, \etc{}. One last problem remained, impossible to
 miss because of the seismograms not matching (but with just a slight
-drift). We added more verbosity to \code{GPUTrace} output, first the
+drift). We added more verbosity to \code{GPUTrace} output: first the
 initial bits of the GPU memory buffers, then their full content. The
 drift was visible in the trace, but it was nonetheless unclear where
-it started. We finally got it after hours of code review, both on
-BOAST kernels and \OCL code: one kernel was \emph{three}-dimensional,
+it started. We finally got it after hours of code review of BOAST
+kernels and \OCL code: one kernel was \emph{three}-dimensional,
 whereas the others were two-dimensional. But for all of them, only two
-dimensions were passed. As it was the same person who coded the \OCL
-port and the GPU tracer, this third dimension was forgotten in both
-side\ldots
+dimensions were passed, and one was missing. As it was the same person
+who coded the \OCL port and the GPU tracer, this third dimension was
+forgotten in both codes\ldots
 
@@
 \Cuda version and our BOAST-\Cuda version. With the same set of
-optimization, BOAST \Cuda and \OCL version also reported similar
-execution time spans. The best execution speed (25\% higher) was
-achieved with \Cuda version though, as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, as of version 1.1, to
-\OCL runtime.
+optimization flags, BOAST \Cuda and \OCL version reported similar
+execution time spans. The best execution speed was achieved with \Cuda
+version though (25\% higher than \OCL), as one optimization parameter
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of
+version 1.1.
 
 By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
-shrank by a factor of 1.5 (from xxx LOC to xxx, in part because of
-code duplication, but by removing manually unrolled loops). This is
+shrank by a factor of 1.5 (from xxx LOC to xxx, partly because of code
+duplication, but by removing manually unrolled loops). This is
 beneficial for \Specfem as it improves the readability and
@@
 values. Then, during the non-regression testing, BOAST framework can
-load these buffer files, allocate and write them on the GPU memory
-through \Cuda or \OCL runtime, and trigger the kernel execution. A
-comparison of the output values (for instance against a maximal error
-level) validates the non-regression.
+load these buffer files, allocate GPU memory and initialize it through
+\Cuda or \OCL runtime, and trigger the kernel execution. A comparison
+of the output values (for instance against a maximal error level)
+validates the non-regression.
 

### 2014-12-16 Paper/BOAST.tex
fix commit mistake
work on specfem/gputrace section

--- 2014-12-15 Paper/BOAST.tex
@@
 get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
-in the SPECFEM3D use case.
+in the next subsection.
+
+\subsection{Trace Debugging of BOAST-based GPU Applications}
+\newcommand{\code}[1]{\texttt{#1}}
+
+BOAST support of multi-target code generation can be leveraged to port an
+application to another environment. As an example, we present in Section
+\ref{subsec:specfem} the portage of a Cuda application to OpenCL environment.
+
+Debugging applications running on such GPU environments is well-recognized as a
+hard and time consuming activity. In complement with BOAST, we designed a
+trace-based debugging tool that simplifies this porting operation.
+
+The tools relies on the assumption that Cuda and OpenCL versions of the code
+should the very same operations on the GPU, at the logical level at least (the
+APIs have \emph{implementation} differences, but the offer same
+functionalities). The usage of BOAST framework sustains this assumption, as both
+sets of kernels will be generated from the same source code.
+
+Hence, the debugging of the newly developed (ie, OpenCL) version can be narrowed
+down to making sure that both versions apply the same operations to the GPU, and
+understanding what diverges when they do not. To that purpose, we developed
+\code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
+\code{GPUTrace} dynamically preloads a library (with the help \code{LD\_PRELOAD}
+in \productname{GNU/Linux}-based environments) between the application and the
+GPU library, and prints function name, execution range and arguments (input
+\emph{and} output) of the relevant function calls. These information are printed
+in a unified format for both APIs.
+
+However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
+to be \emph{state-full}. Indeed, most of OpenCL and Cuda parameters are handles
+to opaque types, so, in order to provide meaningful information, \code{GPUTrace}
+stores internally information about these handles (buffer creation size and
+index, kernel name and prototype, \etc{}).
+
+Once two call traces have been generated by \code{GPUTrace}, it is easy to
+compare them with a graphical \code{diff} tool, and spot the different porting
+mistakes: two parameters reversed, an offset incorrectly applied, \etc{}.
+
+By default, \code{GPUTrace} only prints a unique identifier for the memory
+buffers. A first flag can be activated to append the first bits of the buffer to
+the trace, printed in the right format, for visual inspection. Or another flag
+can be set to dump the whole content of the buffer into a file, for a full
+inspection.
+
+This last option is also useful for generating replay buffers for BOAST kernel
+execution. With a set of filters based on kernel names and execution counters,
+developers can precisely select which kernel execution parameters should be
+dumped, for further reuse as real-case benchmarks and non-regression testing.
 
@@
 
-It appears therefore very interesting to benefit from an automatic tool to drive
+It appears therefore very interesting to benefit from a automatic tool to drive
 the implementation and the generation of new convolutions.  This would lead to
@@
 For these reasons, a convolution library has been engineered with the help of
-BOAST.  *The detailed API of the library will be described elsewhere*.  The
+BOAST.  The detailed API of the library will be described elsewhere.  The
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled

### 2014-12-16 Paper/BOAST.tex

@@
 and called from high-level operations.  Each of the building blocks of these
-convolution library is optimally tuned by BOAST by choosing the sources
+convolution libraries is optimally tuned by BOAST by choosing the sources
 providing the optimal kernel for the chosen computing platform.  The sources of

### 2015-01-07 Paper/BOAST.tex
rework debugging section

--- 2014-12-16 Paper/BOAST.tex
@@
 \usepackage{xspace}
+
+\newcommand{\productname}[1]{\textsc{#1}}
+\newcommand{\latin}[1]{\textit{#1}}
+
+\newcommand{\Specfem}{\productname{Specfem 3D}\xspace}
+\newcommand{\Cuda}{\productname{Cuda}\xspace}
+\newcommand{\eg}{\latin{e.g.,}\xspace}
+\newcommand{\OCL}{\productname{OpenCL}\xspace}
+\newcommand{\code}[1]{\texttt{#1}}
+
+\newcommand{\etc}[1]{\latin{etc}}
 
@@
   \subsubsection{Multi-target Language Generation}
-
+\label{sec:multitarget}
 Language availability and performance can vary between platforms. It is thus
@@
 get input data. Such a tracer, dedicated to CUDA and OpenCL, will be presented
-in the SPECFEM3D use case.
+in the next subsection.
+
+\subsection{Trace Debugging of BOAST-based GPU Applications}
+\label{sec:gputrace}
+
+Debugging applications running on GPU environments is well-recognized
+as a hard and time-consuming activity. In complement with BOAST, we
+designed a trace-based debugging tool that simplifies this porting
+operation. The tool relies on BOAST support of multi-target code
+generation (Section~\ref{sec:multitarget}), used to port an
+application from one GPU-programming framework to another. A
+case-study of the portage of a Cuda application to OpenCL is presented
+in Section~\ref{subsec:specfem}.
+
+The idea behind the tool is based on the assumption that the different
+GPU ports of the code should do the very same operations, at least at
+the logical the logical level (the APIs will have
+\emph{implementation} differences, but they should offer nonetheless
+same functionalities). The usage of BOAST framework sustains this
+assumption, as both sets of kernels should generated from the same
+source code.
+
+Hence, the verificatation and validation of the new port can be
+narrowed down to asserting that both codes apply the same operations
+on the GPU. And the debugging part will consist in understanding what
+diverges. To that purpose, we developed \code{GPUTrace}, inspired by
+\code{strace} and \code{ltrace} tools: \code{GPUTrace} dynamically
+preloads a library between the application and the GPU library, and
+collects the function name, execution range and argument values (input
+\emph{and} output) of the relevant function calls. These information
+are traced in a unified format for all the APIs.
+
+However, in contrast with \code{strace} and \code{ltrace},
+\code{GPUTrace} has to be \emph{state-full}. Indeed, most of API
+parameters are handles to opaque types, so, in order to provide
+meaningful information, \code{GPUTrace} stores information about these
+objects (handle value, buffer creation size and attributes, kernel
+name and prototype, \etc{}).
+
+Once two call traces have been generated by \code{GPUTrace}, it is
+easy to compare them with a graphical \code{diff} tool, and spot the
+different porting mistakes: two parameters reversed, an offset
+incorrectly applied, \etc{}.
+
+By default, \code{GPUTrace} only prints a unique identifier (the
+creation index) for the memory buffers. A first flag can be activated
+to append the first bits of the buffer to the trace, printed in the
+right format, for visual inspection. Or another flag can be set to
+dump the whole content of the buffer into a file, for a full
+inspection.
+
+This last option is also useful for generating \emph{replay buffers}
+for BOAST kernel execution. With a set of filters based on kernel
+names and execution counters, developers can precisely select which
+kernel execution parameters should be dumped, for further reuse as
+real-case benchmarks and non-regression testing.
 
@@
 For these reasons, a convolution library has been engineered with the help of
-BOAST.  \textbf{The detailed API of the library will be described elsewhere}.  The
+BOAST.  The detailed API of the library will be described elsewhere.  The
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled

### 2015-01-07 Paper/specfem.tex
work on specfem/gputrace section
rework debugging section

@@
-\newcommand{\productname}[1]{\textsc{#1}}
-\newcommand{\latin}[1]{\textit{#1}}
-
-\newcommand{\Specfem}{\productname{Specfem 3D}\xspace}
-\newcommand{\Cuda}{\productname{Cuda}\xspace}
-\newcommand{\eg}{\latin{e.g.,}\xspace}
-\newcommand{\OCL}{\productname{OpenCL}\xspace}
-\newcommand{\code}[1]{\texttt{#1}}
-
-\newcommand{\etc}[1]{\latin{etc}}
-
 \subsection{Porting SPECFEM3D to OpenCL using BOAST}
+\label{subsec:specfem}
 
@@
 As part of the Mont-Blanc project, we had to port \Specfem to \OCL, so
-that it could be use to benchmark Mont-Blanc HPC platforms.
+that it could be use to benchmark Mont-Blanc HPC platform.
 
@@
 
-\subparagraph{Porting kernels to BOAST} Nvidia \Cuda and \OCL are
-based on the same programming model: a massively parallel accelerator
-running in another memory environment. Thanks to that proximity, we,
-BOAST developers, have been able to carry out most of the porting task
-with only a limited knowledge about \Specfem internal physics.
+\subparagraph{Porting kernels to BOAST}
 
-That lack of knowledge led us to be particularly attentive to the path
-we undertook for the porting, as we would have been unable to
-understand how and why the application was not operating properly, if
-it was to fail.
+Nvidia \Cuda and \OCL are based on the same programming model: a
+massively parallel accelerator running in disjoint, non addressable
+memory environment. Thanks to that proximity, we, BOAST developers,
+have been able to carry out most of the porting task with only a
+limited knowledge about \Specfem internal physics.
+
+This lack of \Specfem internal knowledge led us to be particularly
+attentive to the path we undertook for the porting, as we would have
+been unable to understand how and why the application was not
+operating properly, if it was to fail.
 
@@
 
-Hence, our idea to locate the execution problems was to make sure that
-both execution actually did the same thing. As the \OCL results were
-invalid, we knew the executions would diverge at one or several
+Hence, our idea for locating the execution problems was to make sure
+that both execution actually did the same thing. As the \OCL results
+were invalid, we knew the executions would diverge at one or several
 points.
 
-\subparagraph{Debugging \OCL Execution: \code{GPUTrace}} In order to
-ensure that \Cuda and \OCL executions were doing the same operations,
-we developed a tool that could intercept all the function calls
-related to the GPU and store them for offline processing, in a format
-common to \OCL and \Cuda. Our solution was inspired by \code{strace}
-and \code{ltrace} tools: it dynamically preloads a library (with the
-help \code{LD\_PRELOAD} in \productname{GNU/Linux}-based environments)
-between the application and the GPU library, and prints the arguments
-(input \emph{and} output) of the relevant function calls.
+\subparagraph{Debugging \OCL Execution: \code{GPUTrace}}
 
-However, in contrast with \code{strace} and \code{ltrace},
-\code{GPUTrace} has to be \emph{state-full}. Indeed, most of \OCL and
-\Cuda parameters are handles to opaque types, so, in order to provide
-more meaningful information, \code{GPUTrace} stores internally
-information about these handles (buffer creation size and index,
-kernel name and prototype, \etc{}). % more details here?
+With the help of \code{GPUTrace} (Section~\ref{sec:gputrace}), we
+could to confront \Cuda and \OCL execution traces with a graphical
+\code{diff} tool, and spot the different porting mistakes: some
+parameters reversed, offsets incorrectly applied, \etc{}.
 
-Once \code{GPUTrace} was operational, it was easy to confront \Cuda
-and \OCL outputs with a graphical \code{diff} tool, and spot the
-different porting mistakes: two parameters reversed, an offset
-incorrectly applied, \etc{}. One last problem remained, impossible to
-miss because of the seismograms not matching (but with just a slight
-drift). We added more verbosity to \code{GPUTrace} output: first the
-initial bits of the GPU memory buffers, then their full content. The
-drift was visible in the trace, but it was nonetheless unclear where
-it started. We finally got it after hours of code review of BOAST
-kernels and \OCL code: one kernel was \emph{three}-dimensional,
-whereas the others were two-dimensional. But for all of them, only two
-dimensions were passed, and one was missing. As it was the same person
-who coded the \OCL port and the GPU tracer, this third dimension was
-forgotten in both codes\ldots
+One last problem remained, clearly highlighted by the seismograms not
+matching perfectly (they had a similar shape, but with a reduced
+intensity). We added more verbosity to \code{GPUTrace} output: first
+the initial bits of the GPU memory buffers, then their full
+content. The drift was visible in the trace, but it was nonetheless
+unclear where it started. We finally got it after hours of code review
+of BOAST kernels and \OCL code: one kernel was
+\emph{three}-dimensional, whereas the others were two-dimensional. But
+for all of them, only two dimensions were passed, and one was
+missing. As it was the same person who coded the \OCL port and the GPU
+tracer, this third dimension was forgotten in both codes\ldots
 
@@
 merged in \Specfem's development tree and under test and extension by
-different research teams. On a platform with two K40x GPUs and x Intel
+different research teams. On a platform with two K40x GPUs and N Intel
 Xeon processors, we measured similar results between the original

### 2015-02-23 Paper/BOAST.tex
Finished integrating Kevin's work.

@@
 
-\input{specfem}
+\subsection{Porting SPECFEM3D to OpenCL using BOAST}
+\label{subsec:specfem}
+
+\subsubsection{SPECFEM3D}
+
+In this last subsection, we study a free seismic wave propagation
+simulator, \Specfem, in its \productname{Cartesian}
+flavor\footnote{Specfem3d Cartesian --- CIG
+  \url{http://www.geodynamics.org/cig/software/specfem3d}}. \Specfem
+simulates seismic wave propagation at the local or regional scale
+based upon spectral-element method (SEM), with very good accuracy and
+convergence properties. ADD BENCHMARK APP FOR HPC SUPERCOMPUTERS.
+
+When we started to work on the project (version v2.1 of July 2013) it
+supported graphics card GPU acceleration through NVidia \Cuda. This
+GPU support came in addition to the MPI support implemented to enable
+multi-CPU parallel computing. Most of \Specfem code-base is written in
+\productname{Fortran2003}, only the GPU related parts are written in
+C.  The split between CPU and GPU code was done at a rather fine
+grain, as the application counted more than 40 GPU kernels. Some of
+them were quite simple (\eg performing a few vector operations, but at
+massively parallel scale), while at the other side of the spectrum,
+some complex kernels took more than 80 parameters and perform very
+specific physical transformations.
+
+Because of the complexity of wave propagation (and of the application
+architecture), it is hard to impossible to define unit tests. Hence,
+the application is validated by the accuracy of the results it
+produces, that is, the accuracy of seismograms of the simulated
+earthquakes, in comparison with the actual ones. Non-regression
+testing (after new developments) is also based on these seismograms,
+with measurements of the relative error between two identical
+simulations.
+
+As part of the Mont-Blanc project, we had to port \Specfem to \OCL, so
+that it could be use to benchmark Mont-Blanc HPC platforms.
+
+\subsubsection{Porting to OpenCL}
+
+\subparagraph{Porting kernels to BOAST}
+
+Nvidia \Cuda and \OCL are based on the same programming model: a
+massively parallel accelerator running in disjoint, non addressable
+memory environment. Thanks to that proximity, we, BOAST developers,
+have been able to carry out most of the porting task with only a
+limited knowledge about \Specfem internal physics.
+
+This lack of \Specfem internal knowledge led us to be particularly
+attentive to the path we undertook for the porting, as we would have
+been unable to understand how and why the application was not
+operating properly, if it was to fail.
+
+Hence, our first milestone in the porting was the translation of
+\Specfem's \Cuda kernels into BOAST EDSL. This way, we could ask BOAST
+framework to generate a CUDA version of the kernels, plug them back
+into \Specfem and get (after fixing compile-time errors---prototypes
+and naming mistakes mainly) a first set of \Specfem seismograms.
+
+As we had expected, the seismograms were erroneous. But with the help
+of shell scripts and BOAST framework ability to store and provide the
+kernels' original source code, we built a set of \Specfem binaries
+including only \emph{one} BOAST-generated kernel, with all others
+reference-kernels. Running and validating all these binaries enabled
+use to pinpoint the misbehaving kernels. We finished the debugging
+with a side-by-side comparison that highlighted the coding mistakes.
+
+\subparagraph{Porting runtime to \OCL} The second part of the porting
+consisted in the translation of the CPU-side of the application, from
+\Cuda API to \OCL API. Most of the functions of the interfaces are
+very similar, with only naming-convention and data-structure
+distinctions. Hence, it was clear that automatic rewriting tools
+(namely \code{sed} regexp and \code{emacs-lisp} functions) could be
+useful. To give an idea of the cost of a \emph{manual} rewriting, we
+can count (in \# of \OCL API function calls): 70 kernel ``function
+calls'', 790 arguments to set, 230 memory transfers, 160 buffer
+creations and 270 releases.
+
+Once all the automatic (and many manual) transformations were applied,
+compilation errors fixed, and \OCL unsuccessful function calls solved,
+the application managed to complete its execution and generate a first
+set of seismograms. And again, as expected and feared, these
+seismograms were not valid.
+
+As we had already validated BOAST-generated kernels (and trusted \Cuda
+and \OCL versions to be semantically identical), we knew that the bugs
+were now in the CPU side, and we had to find a way to understand where
+\Specfem's \Cuda version of the code diverged from its \OCL
+counterpart. To help us in that purpose, we had a strong assumption:
+both versions of the code were supposed to perform exactly the same
+operations, with the same ``logical'' parameters (the APIs have
+\emph{implementation} differences, for instance \OCL has two memory
+transfer functions, \code{clEnqueueReadBuffer},
+\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a
+direction parameter \code{cudaMemcpy(..., dir)}, but above that, it is
+the same functionalities).
+
+Hence, our idea for locating the execution problems was to make sure
+that both execution actually did the same thing. As the \OCL results
+were invalid, we knew the executions would diverge at one or several
+points.
+
+\subparagraph{Debugging \OCL Execution: \code{GPUTrace}}
+
+With the help of \code{GPUTrace} (Section~\ref{sec:gputrace}), we
+could to confront \Cuda and \OCL execution traces with a graphical
+\code{diff} tool, and spot the different porting mistakes: some
+parameters reversed, offsets incorrectly applied, \etc{}.
+
+One last problem remained, clearly highlighted by the seismograms not
+matching perfectly (they had a similar shape, but with a reduced
+intensity). We added more verbosity to \code{GPUTrace} output: first
+the initial bits of the GPU memory buffers, then their full
+content. The drift was visible in the trace, but it was nonetheless
+unclear where it started. We finally got it after hours of code review
+of BOAST kernels and \OCL code: one kernel was
+\emph{three}-dimensional, whereas the others were two-dimensional. But
+for all of them, only two dimensions were passed, and one was
+missing. As it was the same person who coded the \OCL port and the GPU
+tracer, this third dimension was forgotten in both codes\ldots
+
+\subparagraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
+merged in \Specfem's development tree and under test and extension by
+different research teams. On a platform with two K40x GPUs and N Intel
+Xeon processors, we measured similar results between the original
+\Cuda version and our BOAST-\Cuda version. With the same set of
+optimization flags, BOAST \Cuda and \OCL version reported similar
+execution time spans. The best execution speed was achieved with \Cuda
+version though (25\% higher than \OCL), as one optimization parameter
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of
+version 1.1.
+
+By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
+shrank by a factor of 1.5 (from xxx LOC to xxx, partly because of code
+duplication, but by removing manually unrolled loops). This is
+beneficial for \Specfem as it improves the readability and
+maintainability of its source code.
+
+We have also been able to enhance \Specfem's non-regression testsuite
+by adding per-kernel non-regression tests. This was done with the help
+of \code{GPUTrace}, that we used to capture all the input parameters
+of a particular \emph{valid} kernel execution, as well as the output
+values. Then, during the non-regression testing, BOAST framework can
+load these buffer files, allocate GPU memory and initialize it through
+\Cuda or \OCL runtime, and trigger the kernel execution. A comparison
+of the output values (for instance against a maximal error level)
+validates the non-regression.
+
+In the same mindset, we provided \Specfem testsuite with kernel
+performance evaluation mechanisms. These tests will help developers to
+try new optimization in kernel code and measure their impact, without
+executing the whole application.
 

### 2015-02-23 Paper/BOAST.tex
fill missing blanks

@@
 In this last subsection, we study a free seismic wave propagation
-simulator, \Specfem, in its \productname{Cartesian}
-flavor\footnote{Specfem3d Cartesian --- CIG
-  \url{http://www.geodynamics.org/cig/software/specfem3d}}. \Specfem
+simulator, \Specfem, in its \productname{Globe}
+flavor\footnote{Specfem3d Globe --- CIG
+  \url{http://www.geodynamics.org/cig/software/specfem3d-globe}}. \Specfem
 simulates seismic wave propagation at the local or regional scale
 based upon spectral-element method (SEM), with very good accuracy and
-convergence properties. ADD BENCHMARK APP FOR HPC SUPERCOMPUTERS.
+convergence properties. It is a reference application for super
+computer benchmarking, thanks to its good scaling capabilities.
 
@@
 By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
-shrank by a factor of 1.5 (from xxx LOC to xxx, partly because of code
-duplication, but by removing manually unrolled loops). This is
-beneficial for \Specfem as it improves the readability and
-maintainability of its source code.
+shrank by a factor of 1.8 (from 7500 LOC to less than 4000, partly
+because of code duplication, but by removing manually unrolled
+loops). This is beneficial for \Specfem as it improves the readability
+and maintainability of its source code.
 

### 2015-02-23 Paper/D5.5/D5.5.tex
Initial report version.

--- 2015-02-23 Paper/BOAST.tex
@@
-\documentclass%[review]
-              {elsarticle}
+\documentclass[11pt, a4paper, twoside]{montblanc}
+
 \usepackage[utf8]{inputenc}
@@
 \newcommand{\productname}[1]{\textsc{#1}}
-\newcommand{\latin}[1]{\textit{#1}}
+%\newcommand{\latin}[1]{\textit{#1}}
 
@@
 \newcommand{\Cuda}{\productname{Cuda}\xspace}
-\newcommand{\eg}{\latin{e.g.,}\xspace}
+%\newcommand{\eg}{\latin{e.g.,}\xspace}
 \newcommand{\OCL}{\productname{OpenCL}\xspace}
@@
 
-\newcommand{\etc}[1]{\latin{etc}}
+%\newcommand{\etc}[1]{\latin{etc}}
 
@@
 
-\journal{Journal of Parallel and Distributed Computing}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
-%%%%%%%%%%%%%%%%%%%%%%%
-
+\bibliographystyle{plain}
 \begin{document}
-
-\begin{frontmatter}
-
+\devnum{5.5}
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
-
-%% or include affiliations in footnotes:
-\author[mymainaddress]{Brice Videau}
-\ead{brice.videau@imag.fr}
-\author{Kevin Pouget}
-\author{Luigi Genovese}
-\author{Thierry Deutsch}
-\author{Dimitri Komatitsch}
-\author{Jean-François Méhaut}
-
-\address[mymainaddress]{LIG/CNRS}
-
-\begin{abstract}
-The Abstract.
-\end{abstract}
-
-\begin{keyword}
-Code Generation \sep Portability \sep Genericity \sep Productivity and Software
-Design \sep High Performance Computing \sep Autotuning \sep Non Regression
-Testing
-\end{keyword}
-
-\end{frontmatter}
+\version{0.1}
+\deadline{6}
+\level{RE}
+\nature{Report}
+\coordinator{ (CNRS)}
+\contributors{Thierry Deutsch, Luigi Genovese, Jean-François Méhaut, Kevin Pouget, Brice Videau}
+\reviewers{None}
+\keywords{BOAST, Deliverable}
+
+\maketitle
+
+\begin{changelog}
+\change{0.1}{Initial version of the deliverable}
+\end{changelog}
+
+\frontmatter
+
+\begin{executive}
+This document describes BOAST a metaprogramming framework to produce portable and efficient computing kernels for HPC application.
+\end{executive}
 
@@
 
-\section*{References}
-
-\bibliography{BOAST}
+%\section*{References}
+
+\bibliography{D5.5}
 
 \end{document}
+
+% vim: set spell ft=tex fo=aw2t expandtab sw=2 tw=100:

### 2015-02-23 Paper/D5.5/example.tex
Initial report version.

\documentclass[11pt, a4paper, twoside]{montblanc}

\begin{document}
\devnum{0.0}
\title{Mont-Blanc \LaTeX Template User Guide}
\version{0.3}
\deadline{6}
\level{RE}
\nature{Report}
\coordinator{Isaac Gelado and Alex Ramirez (BSC)}
\contributors{All Mont-Blanc Partners}
\reviewers{None}
\keywords{\LaTeX, Deliverable}

\maketitle

\begin{changelog}
\change{0.1}{Initial version of the Mont-blanc \LaTeX template}
\change{0.2}{Fixed font colors and headers}
\change{0.3}{Added new \LaTeX commands for commonly used keywords}
\end{changelog}

\frontmatter

\begin{executive}
This document describes how to use the \LaTeX template for Mont-Blanc deliverables.
\end{executive}

\section{Introduction}
Writing deliverables is a tedious task that can easily become a nightmare if you word processor 
starts messing stuff around. Fortunately, almighty Donald Knuth invented \LaTeX to simplify our life 
when reporting to the European Commission. What Dr. Knuth could not figure out was the format our 
belove commissioners like deliverables to be submitted, so he provided the means for us to define 
\LaTeX templates to specify the format of the documents we need to produce. This document presents 
the main features of the Mont-Blanc \LaTeX template for deliverables.

The template here presented is far from being complete, so do not hesitate to contact the author to 
request whatever you might need.

\section{Using the Mont-Blanc \LaTeX Template}
To use the Mont-Blanc \LaTeX template you need to use the \texttt{montblanc} document class. This 
can be done by setting the document header as follows:
\begin{verbatim}
\documentclass[11pt, a4paper, twoside]{montblanc}
\end{verbatim}

\subsection{Front Page Commands}
The Mont-Blanc \LaTeX template produces a fancy front-page with the project logo, deliverable 
number, version, and name, as well as a document information table. Although many breakthroughs have 
been done in the field of Artificial Intelligence, we have incorporated none of theses in this 
\LaTeX template, so you will need to specify all this information in the top of the document. This 
is done as follows:
\begin{itemize}
\item \verb|\devnum{}|: deliverable number.
\item \verb|\title{}|: deliverable title.
\item \verb|\version{}|: deliverable version number.
\item \verb|\deadline{}|: month number (\eg M6) when the deliverable is due.
\item \verb|\level{}|: dissemination level for the deliverable.
\item \verb|\nature{}|: type of deliverable (\eg Report).
\item \verb|\coordiantor{}|: name of the buddy coordinating the deliverable.
\item \verb|\contributors{}|: name of the folks contributing to the deliverable.
\item \verb|\reviewers{}|: names of our beloved reviewers.
\item \verb|\keywords{}|: keywords for the deliverable.
\end{itemize}

After setting all these parameters, the front page is produced with the command
\begin{verbatim}
\maketitle
\end{verbatim}

\subsection{Front Matter Commands}
Deliverables typically have some stuff before the actual meat, namely: change log, table of 
contents, and executive summary. Each of this section can be appended to the document using the 
corresponding environments.

The Change Log environment also defines a specific command (\verb|\change{}{}|) to add changes to 
the document. The Change Log is added to the deliverable as follows:
\begin{verbatim}
\begin{chagelog}
\change{version number}{description of change}
\change{version number}{description of change}
. . .
\end{changelog}
\end{verbatim}

The table of contents is created by including the \verb|\frontmatter| command. Finally, the 
executive summary is included using the \texttt{executive} environment as follows:
\begin{verbatim}
\begin{executive}
Text for the executive summary
\end{executive}
\end{verbatim}

\subsection{Miscellaneous Commands for the Body}
Besides the standard \LaTeX commands, some commodity commands are provided by this template:

\subsubsection{Safe Abbreviations}
Some commonly used abbreviations can easily cause trouble on \LaTeX, so we have included safe 
versions of this abbreviations:
\begin{itemize}
\item \verb|\eg|: \eg
\item \verb|\ie|: \ie
\item \verb|\vs|: \vs
\item \verb|\etc|: \etc
\item \verb|\etal|: \etal
\item \verb|\MB|: \MB
\item \verb|\IO|: \IO
\item \verb|GFW|: \GFW
\end{itemize}

\subsubsection{Annotations}
Quite often we need to include some annotations for others to read during the preparation of a 
deliverable. The following commands have been added to simplify adding such annotations:
\begin{itemize}
\item \verb|\todo{}| is used to indicate that some text is missing and somebody has to add it. This 
command produces text as follows: \todo{This is a TODO example}.
\item \verb|\fixme{}| is used to indicate that some part is wrong and needs to be corrected. This 
command produces text as follows: \fixme{This is a FIXME example}.
\end{itemize}

\section{Reporting Bugs and Wishes}
If you find any bug and\slash or feature that is missing, please report it to 
\texttt{isaac.gelado@bsc.es}. We cannot guarantee that the bug\slash feature will get fixed 
immediately, but at least we will know about it.

\end{document}

% vim: set spell ft=tex fo=aw2t expandtab sw=2 tw=100:

### 2015-03-09 Paper/D5.5/D5.5.tex
Updated deliverable.

@@
   language=FORTRAN,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b
+}
+
+\lstdefinestyle{custombash}{
+  language=Bash,
   basicstyle=\scriptsize,
@@
 \begin{executive}
-This document describes BOAST a metaprogramming framework to produce portable and efficient computing kernels for HPC application.
+This document describes BOAST, a metaprogramming framework to produce portable and efficient computing kernels for HPC application.
 \end{executive}
@@
 Porting and tuning HPC applications to new platforms is tedious and costly in
-term of human resources.
-
-Portability efforts are often lost when migrating to a new architecture, or code
-lose maintainability because several versions of the code coexist, usually with
-a lot of duplication.
+term of human resources. Nonetheless it is a very important aspect of the
+Mont-Blanc project. Indeed, for the project, more than 10 applications were
+selected to be ported and optimized for the prototype platform.
+
+Unfortunately, portability efforts are often lost when migrating to a new
+architecture. Worse, code may lose maintainability because several versions of
+some functionalities coexist, usually with a lot of duplication.
 
@@
 
-Genericity of HPC codes is often limited. Producing generic code in FORTRAN
-90/95 is difficult as the language is not really  fit for it.
+Genericity of HPC codes is often limited. One of the reason is that producing
+generic code in FORTRAN 90/95 is difficult as the language is not really fit
+for it. Sometimes, adding genericity will degrade performance as optimization
+opportunities that come from over-specification are lost.
 
@@
 regression testing. The run method also returns information (and result for
-kernels that are functions) about the run. For instance the run time of the
-kernel is one such information, it is obtained using the system wide realtime
-clock. Other probes can be inserted at compile time if needed.
+kernels that are functions as well as output scalars) about the run. For
+instance the run time of the kernel is one such information, it is obtained
+using the system wide realtime clock. Other probes can be inserted at compile
+time if needed.
+
+PAPI~\cite{mucci1999papi} is supported and different counters can be specified
+during each execution of a kernel.
 
@@
 short, separable filters.  The detailed description of how these operations are
-defined is beyond the scope of this paper and can be found in the BigDFT
+defined is beyond the scope of this report and can be found in the BigDFT
 reference paper \cite{bigdft}.
@@
 
+In this report we presented the BOAST infrastructure as well as its
+application to two use cases from the Mont-Blanc project. Results are
+encouraging as BOAST proved to be a powerful and flexible tool.
+
 %\section*{References}
@@
 
+\section*{Annexes}
+
+This section will present some simple examples to familiarize the user with
+BOAST. More samples can be found in the git repository.
+
+\subsection*{Installation}
+
+BOAST is ruby based, so ruby needs to be installed on the machine.
+Installation of boast can be done using the ruby built-in package manager:
+\textit{gem}. See Listing~\ref{lst:BOAST_install} for reference.
+
+\lstset{style=custombash,
+        caption={BOAST installation}, 
+        label={lst:BOAST_install},
+        numbers=left }
+\begin{lstlisting}
+sudo apt-get install ruby ruby-dev
+gem install --user-install BOAST
+\end{lstlisting}
+
+\subsection*{Variable and Procedure Declaration}
+
+The following samples are presented using \textit{irb} ruby interactive
+interpreter. It can be launched using the \textit{irb} command in a terminal.
+Listing~\ref{lst:variable_declaration} shows the declaration of two variables
+of different kind.
+
+\lstset{style=BOAST,
+        caption={Variable Declaration},
+        label={lst:variable_declaration}}
+\begin{lstlisting}
+irb(main):001:0> require 'BOAST'
+=> true
+irb(main):002:0> a = BOAST::Int "a"
+=> a
+irb(main):003:0> b = BOAST::Real "b"
+=> b
+irb(main):004:0> BOAST::decl a, b
+integer(kind=4) :: a
+real(kind=8) :: b
+=> [a, b]
+\end{lstlisting}
+
+Listing~\ref{lst:procedure_declaration} shows the declaration of a procedure
+using the two previous variables as parameters. For clarity irb echoes have
+been suppressed.
+
+\lstset{style=BOAST,
+        caption={Procedure Declaration},
+        label={lst:procedure_declaration}}
+\begin{lstlisting}
+005:0> p = BOAST::Procedure( "test_proc", [a,b] )
+006:0> BOAST::opn p
+SUBROUTINE test_proc(a, b)
+  integer, parameter :: wp=kind(1.0d0)
+  integer(kind=4) :: a
+  real(kind=8) :: b
+007:0> BOAST::close p
+END SUBROUTINE test_proc
+\end{lstlisting}
+
+\subsection*{Switching Language}
+
+Listing~\ref{lst:switching_language} shows how to switch BOAST to C. Available
+languages are \textit{FORTRAN}, \textit{C}, \textit{CUDA} and \textit{CL}.
+
+\lstset{style=BOAST,
+        caption={Switching Language},
+        label={lst:switching_language}}
+\begin{lstlisting}
+008:0> BOAST::lang = BOAST::C
+009:0> BOAST::opn p
+void test_proc(int32_t a, double b){
+010:0> BOAST::close p
+}
+\end{lstlisting}
+
+\subsection*{Defining a Complete Procedure}
+
+Listing~\ref{lst:complete_procedure} shows how to define a procedure and the
+associated code. Note that here the parameters of the procedure have been
+associated a direction: one, \textit{a}, is an input parameter while the other
+is an output parameter.
+
+\lstset{style=BOAST,
+        caption={Complete Procedure},
+        label={lst:complete_procedure}}
+\begin{lstlisting}
+011:0> BOAST::lang = BOAST::FORTRAN
+012:0> a = BOAST::Real( "a", :dir => :in)
+013:0> b = BOAST::Real( "b", :dir => :out)
+014:0> p = BOAST::Procedure( "plus_two", [a,b] ) {
+015:1*   BOAST::pr b === a + 2
+016:1> }
+017:0> BOAST::pr p
+SUBROUTINE plus_two(a, b)
+  integer, parameter :: wp=kind(1.0d0)
+  real(kind=8), intent(in) :: a
+  real(kind=8), intent(out) :: b
+  b = a + 2
+END SUBROUTINE plus_two
+018:0> BOAST::lang = BOAST::C
+019:0> BOAST::pr p
+void plus_two(const double a, double * b){
+  (*b) = a + 2;
+}
+\end{lstlisting}
+
+\subsection*{Creating, Building and Running a Computing Kernel}
+
+Listing~\ref{lst:computing_kernel} shows how to create a Computing kernel
+(\textit{CKernel}) and build it. Once a computing kernel is instancied the
+output of BOAST will be redirected to the computing kernel source code. Line 4
+sets the entry point of the computing kernel to the procedure we just defined.
+By default compilation commands are not shown unless an error occurs. This
+behavior can be change by switching to verbose mode.
+
+When running the kernel all the arguments have to be specified. Running a
+kernel returns a hash table containing information about the procedure
+execution. In this simple case two information are returned, first the value
+of the output parameter \textit{b} and second the time the kernel execution
+took.
+
+\lstset{style=BOAST,
+        caption={Computing Kernel},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:computing_kernel}}
+\begin{lstlisting}
+020:0> BOAST::lang = BOAST::FORTRAN
+021:0> k = BOAST::CKernel::new
+022:0> BOAST::pr p
+023:0> k.procedure = p
+024:0> puts k
+SUBROUTINE plus_two(a, b)
+  integer, parameter :: wp=kind(1.0d0)
+  real(kind=8), intent(in) :: a
+  real(kind=8), intent(out) :: b
+  b = a + 2
+END SUBROUTINE plus_two
+025:0> k.build
+026:0> BOAST::verbose = true
+027:0> k.build
+gcc -O2 -Wall -fPIC -I/usr/lib/x86_64-linux-gnu/ruby/2.1.0 -I/usr/include/ruby-2.1.0 -I/usr/include/ruby-2.1.0/x86_64-linux-gnu -I/usr/include/x86_64-linux-gnu/ruby-2.1.0 -I/var/lib/gems/2.1.0/gems/narray-0.6.1.1 -DHAVE_NARRAY_H -c -o /tmp/Mod_plus_two20150309_4611_5a129k.o /tmp/Mod_plus_two20150309_4611_5a129k.c
+gfortran -O2 -Wall -fPIC -c -o /tmp/plus_two20150309-4611-5a129k.o /tmp/plus_two20150309-4611-5a129k.f90
+gcc -shared -o /tmp/Mod_plus_two20150309_4611_5a129k.so /tmp/Mod_plus_two20150309_4611_5a129k.o /tmp/plus_two20150309-4611-5a129k.o -Wl,-Bsymbolic-functions -Wl,-z,relro -rdynamic -Wl,-export-dynamic -L/usr/lib -lruby-2.1 -lrt
+028:0> r = k.run(5,0)
+029:0> puts r
+{:reference_return=>{:b=>7.0}, :duration=>5.84e-07}
+\end{lstlisting}
+
+\subsection*{Using Arrays in Procedures}
+
+Most computing kernels don't work on scalar values but rather on array of
+data. Listing~\ref{lst:arrays} shows how to use arrays in computing kernels.
+In this case we place ourselves in BOAST namespace to reduce the syntax
+overhead. Variables \textit{a} and \textit{b} are one-dimensional arrays of
+size \textit{n}. Arrays in BOAST start at index 1 unless specified otherwise.
+For instance \lstinline!Dim(0,n-1)! would have created a dimension starting at
+0. Array bounds can also be negative and several dimensions can be specified
+to obtain muti-dimensional arrays.
+For self contained procedures/kernels one can sue the shortcut written on
+line~13 to create a CKernel object. As we are not specifying build options the
+build command can also be ommited and will be automatically called when
+running the kernel the first time. Lines~17 to 19 are used to check the result
+of the kernel.
+
+\lstset{style=BOAST,
+        caption={Array Usage},
+        label={lst:arrays}}
+\begin{lstlisting}
+001:0> require 'BOAST'
+002:0> require 'narray'
+003:0> include BOAST
+004:0> n = Int(  "n", :dir => :in )
+005:0> a = Real( "a", :dir => :in,  :dim => [Dim(n)] )
+006:0> b = Real( "b", :dir => :out, :dim => [Dim(n)] )
+007:0> p = Procedure( "plus_two", [n, a, b] ) {
+008:1*   decl i = Int( "i" )
+009:1>   pr For( i, 1, n ) {
+010:2*     pr b[i] === a[i] + 2.0
+011:2>   }
+012:1> }
+013:0> k = p.ckernel
+014:0> input  = NArray.float(1024).random
+015:0> output = NArray.float(1024)
+016:0> k.run(input.length, input, output)
+017:0> (output - input).each { |val|
+018:1*   raise "Error!" if (val-2).abs > 1e-15
+019:1> }
+020:0> stats = k.run(input.length, input, output)
+021:0> puts "Success, duration: #{stats[:duration]} s"
+Success, duration: 3.79e-06 s
+\end{lstlisting}
+
+\subsection*{The Canonical Case: Vector Addition}
+
+Listing~\ref{lst:vector_add_decl} shows the addition of two vector in a third
+one. Here BOAST is configured to have arrays starting at 0 and to use single
+precision reals by default (Lines~5 and 6).  The kernel declaration is
+encapsulated inside a method to avoid cluttering the global namespace. Line 15
+the expression \lstinline!c[i] === a[i] + b[i]! is stored inside a variable
+\textit{expr} for later use. Lines~16 to 23 show that the kernel differs
+depending on the target language, in CUDA and OpenCL each thread will process
+one element.
+
+\lstset{style=BOAST,
+        caption={Vector Addition Declaration},
+        label={lst:vector_add_decl}}
+\begin{lstlisting}
+require 'narray'
+require 'BOAST'
+include BOAST
+
+set_array_start(0)
+set_default_real_size(4)
+
+def vector_add
+  n = Int("n",:dir => :in)
+  a = Real("a",:dir => :in, :dim => [ Dim(n)] )
+  b = Real("b",:dir => :in, :dim => [ Dim(n)] )
+  c = Real("c",:dir => :out, :dim => [ Dim(n)] )
+  p = Procedure("vector_add", [n,a,b,c]) {
+    decl i = Int("i")
+    expr = c[i] === a[i] + b[i]
+    if (get_lang == CL or get_lang == CUDA) then
+      pr i === get_global_id(0)
+      pr expr
+    else
+      pr For(i,0,n-1) {
+        pr expr
+      }
+    end
+  }
+  return p.ckernel
+end
+\end{lstlisting}
+
+Listing~\ref{lst:vector_add_run} shows the a way to check the validity of the
+previous kernel over the available range of languages. The options that are
+passed to run are only relevant for GPU languages and are thus ignored in
+FORTRAN and C (Line~16). Success is only printed if results are validated,
+else an exception is raised (Lines~17 to 20).
+
+\lstset{style=BOAST,
+        caption={Vector Addition Test},
+        label={lst:vector_add_run}}
+\begin{lstlisting}
+n = 1024*1024
+a = NArray.sfloat(n).random
+b = NArray.sfloat(n).random
+c = NArray.sfloat(n)
+
+epsilon = 10e-15
+
+c_ref = a + b
+
+[:FORTRAN, :C, :CL, :CUDA].each { |l|
+  set_lang( BOAST.const_get(l)  )
+  puts "#{l}:"
+  k = vector_add
+  puts k.print
+  c.random!
+  k.run(n, a, b, c, :global_work_size => [n,1,1], :local_work_size => [32,1,1])
+  diff = (c_ref - c).abs
+  diff.each { |elem|
+    raise "Warning: residue too big: #{elem}" if elem > epsilon
+  }
+}
+puts "Success!"
+\end{lstlisting}
+
 \end{document}

### 2015-03-09 Paper/D5.5/D5.5.tex
More deliverable.

@@
 
+  \subsection{Scientific Computing Applications}
+
+Scientific Computing Applications are usually developed by physicist, chemist or
+meteorologist. Usually in FORTRAN for historical and performance reason. Codes
+can be quite huge (several thousands lines of code) with lots of
+functionalities. Nonetheless they are usually based on computing kernels.
+Computing kernels are resource intensive and well defined parts of a program
+with a limited set of input and output. Those kernels are the time consuming
+part of an HPC application. They are thus the prime target for optimization.
+
+Those applications are often developed by several individuals. Sometimes some of
+those developers only work a few month on the application. Maintaining optimized
+code written by someone else is quite a challenge.
+
+In Section~\ref{use_cases} we will present two HPC application that we used
+as use cases: SPECFEM3D and BigDFT. They are both based on computing
+kernels and were selected as candidate applications in the Mont-Blanc project.
+
+  \subsection{How Should Computing Kernels be Written?}
+
+The problematic here is to obtain computing kernels that present good
+performances on the architectures encountered by the application while still
+being portable after the optimization process took place. Indeed the application
+might encounter one of the other architectures presented before. Investing
+manpower to optimize the application for a new architecture is reasonable,
+suffering hindrance from previous optimization work is not. Thus optimizations
+have to be as orthogonal as possible from one another so as to be easily
+activated and deactivated.
+
+If this paradigm is followed by developers then they will rapidly be confronted
+with a huge optimization space to search. They will need to be able to test
+easily the performance impact of the chosen optimizations without running the
+full application. The same reasoning implies that kernels should be tested for
+non regression without running the full application.
+
+What we want is computing kernels that are:
+\begin{itemize}
+\item Written in a \emph{portable} manner,
+\item Written in a way that raise developer \emph{productivity},
+\item Written to present good \emph{performance}.
+\end{itemize}
+
   \subsection{Evolution of HPC Architectures}
@@
 Thus, those projects have work packages dedicated to applications. Those
-workpackages are dedicated to porting and optimizing Scientific applications on
+work packages are dedicated to porting and optimizing Scientific applications on
 those new architectures. In the DEEP project six applications were selected for
@@
 
-  \subsection{Scientific Computing Applications}
-
-Developed by physicist, chemist, meteorologist. Usually in FORTRAN for
-historical and performance reason. Codes can be quite huge (several
-thousands lines of code) with lots of functionalities. Nonetheless they are
-usually based on computing kernels. Computing kernels are resource
-intensive and well defined parts of a program with a limited set of input
-and output. Those kernels are the time consuming part of an HPC
-application. They are thus the prime target for optimization.
-
-Those applications are often developed by several individuals. Sometimes some of
-those developers only work a few month on the application. Maintaining optimized
-code written by someone else is quite a challenge.
-
-In Section~\ref{use_cases} we will present two HPC application that we used
-as use cases: SPECFEM3D and BigDFT. They are both based on computing
-kernels and were selected as candidate applications in the Mont-Blanc project.
-
-  \subsection{How Should Computing Kernels be Written?}
-
-The problematic here is to obtain computing kernels that present good
-performances on the architectures encountered by the application while still
-being portable after the optimization process took place. Indeed the application
-might encounter one of the other architectures presented before. Investing
-manpower to optimize the application for a new architecture is reasonable,
-suffering hindrance from previous optimization work is not. Thus optimizations
-have to be as orthogonal as possible from one another so as to be easily
-activated and deactivated.
-
-If this paradigm is followed by developers then they will rapidly be confronted
-with a huge optimization space to search. They will need to be able to test
-easily the performance impact of the chosen optimizations without running the
-full application. The same reasoning implies that kernels should be tested for
-non regression without running the full application.
-
-What we want is computing kernels that are:
-\begin{itemize}
-\item Written in a \emph{portable} manner,
-\item Written in a way that raise developer \emph{productivity},
-\item Written to present good \emph{performance}.
-\end{itemize}
+\section{Motivating Example}
 

### 2015-03-10 Paper/D5.5/D5.5.tex
Added a motivating example.

@@
   morekeywords={int64_t, int32_t}
+}
+
+\lstdefinestyle{customcl}{
+  language=C,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={__global, __local, global, local, __kernel, kernel, short,
+uchar, uchar16, int16, short16}
 }
@@
 
-\section{Motivating Example}
+\section{Motivating Example: OpenCL Laplace}
+
+During one of the Mont-Blac face to face meeting a talk on OpenCL optimization
+on the Mali GPU was given. One of the case study was a Laplace filter on the
+GPU~\cite{opencl_arm_training}.
+
+This is a good example of an algorithm that is simple in its formulation but can
+be complex to optimize because many different optimizations are available and
+can be combined together. The correct combination of optimization will be
+dependant on the targeted architecture.
+
+\subsection{The Laplace Filter}
+
+Listing~\ref{lst:laplace_c_filter} show a simple implementation of the Laplace
+filter in C99. For clarity, boundary conditions have been omitted.
+
+\lstset{style=customc,
+        caption={Laplace C99 Filter},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_c_filter}}
+\begin{lstlisting}
+void laplace(const int width,
+             const int height,
+             const unsigned char src[height][width][3],
+                   unsigned char dst[height][width][3]){
+  for (int j = 1; j < height-1; j++) {
+    for (int i = 1; i < width-1; i++) {
+      for (int c = 0; c < 3; c++) {
+        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
+                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
+                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
+        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
+      }
+    }
+  }
+}
+\end{lstlisting}
+
+A naive implementation in OpenCL is proposed in
+Listing~\ref{lst:laplace_opencl_filter}. Each work item processes one pixel of
+the resulting image.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter}}
+\begin{lstlisting}
+kernel laplace(const int width,
+               const int height,
+               global const uchar *src,
+               global       uchar *dst){
+  int i = get_global_id(0);
+  int j = get_global_id(1);
+  for (int c = 0; c < 3; c++) {
+    int tmp = -src[3*width*(j-1) + 3*(i-1) + c]\
+             - src[3*width*(j-1) + 3*(i  ) + c]\
+             - src[3*width*(j-1) + 3*(i+1) + c]\
+             - src[3*width*(j  ) + 3*(i-1) + c]\
+           + 9*src[3*width*(j  ) + 3*(i  ) + c]\
+             - src[3*width*(j  ) + 3*(i+1) + c]\
+             - src[3*width*(j+1) + 3*(i-1) + c]\
+             - src[3*width*(j+1) + 3*(i  ) + c]\
+             - src[3*width*(j+1) + 3*(i+1) + c];
+    dst[3*width*j + 3*i + c] = clamp(tmp, 0, 255);
+  }
+}
+\end{lstlisting}
+
+\subsection{Possible Optimizations of the OpenCL Version}
+
+Several optimizations were proposed and successively applied to the previous
+implementation.
+
+\paragraph{Vectorization}
+The first proposed optimization is to compute 5 pixels instead of one using the
+vectors available to the Mali architecture. To be efficient the vectors have to
+be directly loaded from memory. Using vectors of 16 elements 15 useful component
+are simultaneously loaded and can be used in computation. This optimization
+yields speedups between 1.5 and 6 depending on the image size.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_vector}}
+\begin{lstlisting}
+kernel laplace(const int width,
+               const int height,
+               global const uchar *src,
+               global       uchar *dst){
+  int i = get_global_id(0);
+  int j = get_global_id(1);
+
+  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
+  uchar16 v12_ = vload16( 0, src + 3*width*(j-1) + 3*5*i     );
+  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
+  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
+  uchar16 v22_ = vload16( 0, src + 3*width*(j  ) + 3*5*i     );
+  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
+  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
+  uchar16 v32_ = vload16( 0, src + 3*width*(j+1) + 3*5*i     );
+  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
+
+  int16 v11 = convert_int16(v11_);
+  int16 v12 = convert_int16(v12_);
+  int16 v13 = convert_int16(v13_);
+  int16 v21 = convert_int16(v21_);
+  int16 v22 = convert_int16(v22_);
+  int16 v23 = convert_int16(v23_);
+  int16 v31 = convert_int16(v31_);
+  int16 v32 = convert_int16(v32_);
+  int16 v33 = convert_int16(v33_);
+
+  int16 res = v22 * (int)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
+        res = clamp(res, (int16)0, (int16)255);
+  uchar res_ = convert_uchar16(res);
+
+  vstore8(res_.s01234567, 0, dst + 3*width*j + 3*5*i);
+  vstore8(res_.s89ab,     0, dst + 3*width*j + 3*5*i + 8);
+  vstore8(res_.scd,       0, dst + 3*width*j + 3*5*i + 12);
+  dst[3*width*j + 3*5*i + 14] = res_.se;
+}
+\end{lstlisting}
+
+\paragraph{Synthesizing Loads} It is possible to reduce the number of loads
+since the vectors are overlapping.
+Listing~\ref{lst:laplace_opencl_filter_vector_synth} shows how the vectors are
+loaded in this case. This optimization yields marginal improvements.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_vector_synth}}
+\begin{lstlisting}
+  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
+  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
+  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
+  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
+  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
+  uchar16 v22_ = vload16( v21_.s3456789a, v23_.s56789abc );
+  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
+  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
+  uchar16 v32_ = vload16( v31_.s3456789a, v33_.s56789abc );
+\end{lstlisting}
+
+\paragraph{Temporary Variables Size} Using \textit{int} to store intermediary
+results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
+the code is modified to use smaller types. This yields a speedup of 1.3 for most
+image sizes.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized Using \textit{short}},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_short}}
+\begin{lstlisting}
+  short16 v11 = convert_short16(v11_);
+  short16 v12 = convert_short16(v12_);
+  short16 v13 = convert_short16(v13_);
+  short16 v21 = convert_short16(v21_);
+  short16 v22 = convert_short16(v22_);
+  short16 v23 = convert_short16(v23_);
+  short16 v31 = convert_short16(v31_);
+  short16 v32 = convert_short16(v32_);
+  short16 v33 = convert_short16(v33_);
+
+  short16 res = v22 * (short)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
+        res = clamp(res, (short16)0, (short16)255);
+\end{lstlisting}
+
+\paragraph{More Optimizations} Several other optimizations can be attempted at
+this point. For instance reducing or increasing the number of pixel each work
+item is processing. They can yield improved performance in some cases, especially
+when avoiding memory alignment problems. These optimizations will not be shown
+here but can be found in~\cite{opencl_arm_training}.
+
+\subsection{Improving the Methodology}
+
+The process described in the previous Subsection is quite tedious and requires
+some intimate knowledge of the target architecture. The results are impressive,
+as speedups of almost 10 are observed compared to the naive OpenCL
+implementation.
+
+Nonetheless, the process is frustrating. The optimizations are never evaluated
+independently from one another. Some were arbitrarily configured (the number of
+pixel chosen for instance). Testing the different combination of those
+optimizations and the different parameters would prove very costly in developer
+time and like all repetitive and tedious tasks, error prone. Thus every created
+kernel would have to be thoroughly tested to insure no error was done.
+
+In the next Section we will present our answer to this problematic: BOAST.
+BOAST is a framework dedicated to kernel description, optimization, regression testing
+and autotuning.
 

### 2015-03-10 ARMclbench/Laplacian.rb
Slightly modified the Laplace reference c implementation for improved readability (especially using VLAs)

@@ -30,21 +30,23 @@ def laplacian_c_ref
   height = Int("height", :dir => :in)
   width = Int("width", :dir => :in)
   w = Int("w")
-  pdst = Int("pdst", :dir => :out,  :signed => false, :size => 1, :dim => [ Dim(w), Dim(height) ] )
-  psrc = Int("psrc", :dir => :in, :signed => false, :size => 1, :dim => [ Dim(w), Dim(height) ] )
-  p = Procedure("math", [psrc, pdst, width, height]) {
+  pdst = Int("pdst", :dir => :out,  :signed => false, :size => 1, :dim => [ Dim(3), Dim(width), Dim(height) ] )
+  psrc = Int("psrc", :dir => :in, :signed => false, :size => 1, :dim => [ Dim(3), Dim(width), Dim(height) ] )
+  p = Procedure("math", [width, height, psrc, pdst]) {
     decl i = Int("i")
     decl j = Int("j")
+    decl c = Int("c")
     decl tmp = Int("tmp")
     decl w
     pr w === width * 3
     pr For(j, 1, height-2) {
-      pr For(i, 3, w-4) {
-         
-        pr tmp === ( -psrc[i - 3, j - 1] - psrc[i, j - 1] - psrc[i + 3, j - 1]\
-                    - psrc[i - 3, j]     + psrc[i, j] * 9 - psrc[i + 3, j]\
-                    - psrc[i - 3, j + 1] - psrc[i, j + 1] - psrc[i + 3, j + 1] )
-        pr pdst[i,j] === Ternary(tmp < 0, 0, Ternary(tmp>255, 255, tmp))
+      pr For(i, 1, width-2) {
+        pr For(c, 0, 2) {
+          pr tmp === (  -psrc[c, i-1, j-1] - psrc[c, i, j-1] -   psrc[c, i+1, j-1]\
+                       - psrc[c, i-1, j]   + psrc[c, i, j] * 9 - psrc[c, i+1, j]\
+                       - psrc[c, i-1, j+1] - psrc[c, i, j+1] -   psrc[c, i+1, j+1] )
+          pr pdst[c,i,j] === Ternary(tmp < 0, 0, Ternary(tmp>255, 255, tmp))
+        }
       }
     }
   }
@@ -262,7 +264,7 @@ sizes.each { |width, height|
   input = NArray.byte(width*3,height+1).random(256)
   output_ref = NArray.byte(width*3,height)
 
[…]

### 2015-03-12 Paper/D5.5/D5.5.tex
Added results for Laplace and BigDFT.

@@
 \section{Motivating Example: OpenCL Laplace}
-
-During one of the Mont-Blac face to face meeting a talk on OpenCL optimization
+\label{sec:laplace}
+
+During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
 on the Mali GPU was given. One of the case study was a Laplace filter on the
@@
   uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
-  uchar16 v22_ = vload16( v21_.s3456789a, v23_.s56789abc );
+  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
   uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
   uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
-  uchar16 v32_ = vload16( v31_.s3456789a, v33_.s56789abc );
+  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
 \end{lstlisting}
@@
 void magicfilter(const int32_t n, const int32_t ndat,
-                 const double * x, double * y)
+                 const double * x, double * y) {
 }
@@
 
-In this section we will present two scientific applications that use BOAST. The
+In this section we will present the benefits of using BOAST on the Laplace
+motivating example as well as two scientific applications that use BOAST. The
 first one, BigDFT, uses boast in order to develop new functionalities with
@@
 tests.
+
+  \subsection{Laplace Filter Kernel}
+
+Section~\ref{sec:laplace} presented the Laplace motivating example. From this
+section we know that a number of optimization can have an impact on the
+performance of the kernel on the Mali architecture. But what of other
+architectures? And are there other optimizations that can impact the
+performance?
+
+    \subsubsection{Optimization Space}
+
+The list of already identified optimizations are:
+\begin{itemize}
+\item vectorization,
+\item intermediary data type,
+\item number of pixels processed,
+\item and synthesizing loads.
+\end{itemize}
+To create our generic implementation we decided to work at the component level
+rather than the pixel level. This approach leads to more flexibility and
+genericity when applying optimizations. We also decided to study the impact of
+another parameter which is the number of component to process on the column
+direction. This leads to being able to process tiles instead of only rows.
+
+Here are the parameters we finally selected for our kernel optimization and
+their possible values:
+\begin{itemize}
+\item \textit{x\_component\_number}: a positive integer
+\item \textit{y\_component\_number}: a positive integer
+\item \textit{vector\_length}: 1, 2, 4, 8 or 16
+\item \textit{temporary\_size}: 2 or 4
+\item \textit{synthesize\_loads}: \textit{true} or \textit{false}
+\item \textit{vector\_recompute}: \textit{true} or \textit{false}
+\end{itemize}
+The last parameter is used when \textit{x\_component\_number} is not divisible by
+\textit{vector\_length}. Two solutions are possible then, divide the remainder
+of the division in vectors of smaller sizes (\textit{vector\_recompute} =
+\textit{false}) or load more data and compute useless values in vectors of the
+specified size. This last option is mimicking the behavior of ARM
+implementation, although when working at the component level it may seem an
+absurd thing to try.
+
+    \subsubsection{Performance Results}
+
+Table~\ref{tbl:ARM} are the best results obtained by ARM on different image
+compared to the naive implementation and to the best version version BOAST
+found. It shows that the generated version systematically beats the hand
+optimized version. As far as the optimization options are concerned the result
+are a disappointment: the same kernel configuration is the best for all image
+sizes. This kernel is using \textit{x\_component\_number} = 16,
+\textit{y\_component\_number} = 1, \textit{vector\_length} = 16,
+\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false.
+\textit{vector\_recompute} has no meaning here as \textit{x\_component\_number}
+== \textit{vector\_length}. Those results also shows that, when working on full
+vectors, synthesizing the loads is harmful to performance and the programmer is
+better of trusting the cache to load each vector in 1 cycle without compromising
+the bandwidth.
+
+\begin{table}
+\centering
+\begin{tabular}{|l|r|r|r|r|r|}
+\hline
+Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\
+\hline
+768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
+\hline
+2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
+\hline
+2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
+\hline
+5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
+\hline
+7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
+\hline
+\end{tabular}
+\caption{Best performance of ARM Laplace kernel.}
+\label{tbl:ARM}
+\end{table}
+
+But what if we run out benchmark on other architectures? Table~\ref{tbl:other}
+the results obtained when running our BOAST implementation on other
+architectures. The chosen architectures were an Intel i7-2760QM CPU (Sandy
+Bridge architecture) that supports OpenCL 1.2 and an NVIDIA gtx680 GPU that
+support OpenCL 1.1. We can see that the performance ratio between the different
+architectures is rather stable across image sizes.
+
+The optimization parameters selected are not the same for those architectures.
+Indeed, the Intel CPU favors kernels that have the parameters:
+\textit{x\_component\_number} = 16,
+\textit{vector\_length} = 8,
+\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false. Once again
+\textit{vector\_recompute} has no meaning. \textit{y\_component\_number} varies
+from 4 to 2 when image size increase, thus decreasing task granularity as the
+global work size increase. This result is unexpected and understanding why this
+happen could be interesting. The NVIDIA GPU favors processing square tiles:
+\textit{x\_component\_number} = 4, \textit{y\_component\_number} = 4,
+\textit{vector\_length} = 4, \textit{temporary\_size} = 2 and
+\textit{synthesize\_loads} = false. Once more \textit{vector\_recompute} has no
+meaning.
+
+\begin{table}
+\centering
+\begin{tabular}{|l|r|r|r|r|r|}
+\hline
+Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\
+\hline
+768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
+\hline
+2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
+\hline
+2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
+\hline
+5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
+\hline
+7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
+\hline
+\end{tabular}
+\caption{Best performance of Laplace Kernel on several architectures.}
+\label{tbl:other}
+\end{table}
+
+    \subsubsection{Laplace Conclusion}
+
+In this Subsection we have shown the interest of BOAST in optimizing a well
+known algorithm across different architectures. Chosen optimization combinations
+are highly dependant on the targeted architecture. In the next Subsections we
+will show that our methodology also applies to real applications.
 
@@
 convolution library oriented to computational physics communities.
+
+    \subsubsection{Performance report}
+
+Several kernels have already been implemented in BOAST for the convolution
+library. Figure~\ref{fig:synthesis} show the performance of the wavelet
+transform operation as a factor of the unrolling length of the outer loop, the
+language used to implement it as well as the activation or not of the OpenMP
+parallelization. Results are given as a speedup compared to the sequential hand
+tuned implementation that can be found in BigDFT. These tests were run on the
+Inte Xeon X5550 that was used to hand optimize the code.
+
+We can see that this function is better optimized using FORTRAN and small
+unrolling factors. In the hand optimized version the unrolling factor was chosen
+much too high (a factor of 12 was used). This factor might have been optimal at
+the point the procedure was optimized (compiler version changed in the meantime) but
+since then the environment changed. Other optimization have also been
+incorporated in the BAOST sources, like the systematic inner loop unrolling, and
+those could also help increase performance while limiting the interest of the
+outer loop unrolling.
+
+Nonetheless what is interesting from the physicist point of view is that the
+generated source will give her better performance on a whole range of
+architectures/compiler combination than that of the hand tuned code.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Res_synthesis}
+\caption{Impact of Unrolling, Language and OpenMP on a Wavelet Transform}
+\end{center}
+\label{fig:synthesis}
+\end{figure}
 %   BigDFT implementation of convolution/wavelets is a reference but is limited

### 2015-03-12 Paper/D5.5/D5.5.tex
Implemented most of Frédéric's suggestions. Missing Capitalization of bibliography + API description of convolution library.

@@
 Scientific Computing Applications are usually developed by physicist, chemist or
-meteorologist. Usually in FORTRAN for historical and performance reason. Codes
-can be quite huge (several thousands lines of code) with lots of
-functionalities. Nonetheless they are usually based on computing kernels.
-Computing kernels are resource intensive and well defined parts of a program
-with a limited set of input and output. Those kernels are the time consuming
-part of an HPC application. They are thus the prime target for optimization.
+meteorologist. Those codes are usually written in FORTRAN for historical and
+performance reason. Codes can be quite huge (several thousands lines of code)
+with lots of functionalities. Nonetheless they are usually based on computing
+kernels.  Computing kernels are resource intensive and well defined parts of a
+program that usually work on precisely defined data. Those kernels are the time
+consuming part of an HPC application. They are thus the prime target for
+optimization.
 
@@
 those developers only work a few month on the application. Maintaining optimized
-code written by someone else is quite a challenge.
+code written by someone else is quite a challenge. Several languages and
+programming paradigms can also be used in a project and thus the maintainer must
+be knowledgeable in several area of expertise. Portability problems can also be
+cause by the availability of an optimizing compiler for a specific language and
+architecture. C compilers are usually the first available while FORTRAN may
+sometimes arrive later.
 
@@
 Being able to efficiently use those architectures on such a small
-time-frame is challenging. Stress frequency of architecture changes.
+time-frame is challenging.
 
@@
 
-Running existing applications on those new architectures is an open project.
-Thus, those projects have work packages dedicated to applications. Those
-work packages are dedicated to porting and optimizing Scientific applications on
-those new architectures. In the DEEP project six applications were selected for
-porting and optimizing, eleven were selected in the Mont-Blanc project.
+Running existing applications on those new architectures is an open research
+subject as well as an ongoing porting effort.  Thus, those projects have work
+packages dedicated to applications. Those work packages are dedicated to porting
+and optimizing Scientific applications on those new architectures. In the DEEP
+project six applications were selected for porting and optimizing, eleven were
+selected in the Mont-Blanc project.
 
@@
 
-  The idea behind BOAST is to provide scientific application developers with a
-framework to develop and test application computing kernels. The envisioned
-workflow and program structure is presented in Figure~\ref{fig:boast_workflow}.
-The user should start from an application kernel (implemented or just designed)
-and implement it in a dedicated language (step 1). The language provides enough
-flexibility for the kernel to be meta-programmed with several orthogonal
-optimizations. Then the developer selects one configuration of the kernel and a
-target language. Those parameter define the output source code that will be
-generated by BOAST (step 2). The resulting code source is then built according to
-the user specified compiler and options (step 3). If input data are available
-then the kernel can be benchmarked and tested for non regression. Based on the
-results, other optimizations can be selected (step 4) or new optimizations can
-be added to the BOAST sources. The process can be repeated until a good
-candidate is found on the target platform. The resulting kernel is then added to
-the program (step 5).
+The ideas behind BOAST were first presented in~\cite{videau2013boast}, and are
+to provide scientific application developers with a framework to develop and
+test application computing kernels. The envisioned workflow and program
+structure is presented in Figure~\ref{fig:boast_workflow}.  The user should
+start from an application kernel (implemented or just designed) and implement it
+in a dedicated language (step 1). The language provides enough flexibility for
+the kernel to be meta-programmed with several orthogonal optimizations. Then the
+developer selects one configuration of the kernel and a target language. Those
+parameter define the output source code that will be generated by BOAST (step
+2). The resulting code source is then built according to the user specified
+compiler and options (step 3). If input data are available then the kernel can
+be benchmarked and tested for non regression. Based on the results, other
+optimizations can be selected (step 4) or new optimizations can be added to the
+BOAST sources. The process can be repeated until a good candidate is found on
+the target platform. The resulting kernel is then added to the program (step 5).
 
@@
 
-  \cite{videau2013boast}
-
   \subsection{Kernel Description Language}
 
-  \begin{itemize}
-  \item Abstraction of programming concepts: variables, procedures (
-portability, productivity )
-  \item Operator overloading: simpler syntax ( productivity, portability )
-  \item Two languages in one ( productivity, performance ) 
-  \end{itemize}
-
-Usually computing kernels are hotspots of an HPC application, and are most of
-the time based on a loop nest. A lot of effort is dedicated to their tuning and
-the obtained result is often quite different than the original procedure.
+%  \begin{itemize}
+%  \item Abstraction of programming concepts: variables, procedures (
+%portability, productivity )
+%  \item Operator overloading: simpler syntax ( productivity, portability )
+%  \item Two languages in one ( productivity, performance ) 
+%  \end{itemize}
+
+Usually computing kernels are hotspots of an HPC application, and they are most
+of the time based on a loop nest. A lot of efforts are dedicated to their tuning
+and the obtained result is often quite different from the original procedure.
 Several transformations can be applied to such kernels. And those optimization
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{BOAST_DSL.pdf}
-\caption{BOAST DSL intent.}
-\label{fig:dsl}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.8]{BOAST_DSL.pdf}
+%\caption{BOAST DSL intent.}
+%\label{fig:dsl}
+%\end{center}
+%\end{figure}
 
 In order to manipulate those abstractions we want to have a syntax resembling
-what programmers use. For instance commonly used operator have to be available
+what programmers use. For instance commonly used operators have to be available
 and behave as expected. It must also be possible to differentiate an action on
 an abstraction in the context of an expression and in the context of the
-management of this expression. This is why an embedded domain specific language
-approach was selected~\cite{hudak1996building}. This allows for the coexistence
-of two languages: the host language and the DSL. In our case the DSL allows the
-description of the kernel while the host language provides the meta programming
-of the kernel, cf Figure~\ref{fig:dsl}. Operator overloading of the host
-language will provide the familiar syntax programmers are accustomed to.
+management of this expression. $c = a + b$, an expression that affects the
+results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves the
+expression $a + b$ to a variable $c$. This is why an embedded domain specific
+language approach was selected~\cite{hudak1996building}. This allows for the
+coexistence of two languages: the host language and the DSL. In our case the DSL
+allows the description of the kernel ($c = a + b$) while the host language
+provides the meta programming of the kernel ($c \leftarrow a + b$)%, cf Figure~\ref{fig:dsl}
+. Operator overloading of the host language will provide the familiar syntax
+programmers are accustomed to. The host language will also provide easy
+interfacing with libraries we might need during the development of our framework.
 
@@
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
-capabilities as well. This is the main reason it was selected.
+capabilities as well. This is the main reason why it was selected.
 
@@
 related to manipulations in the host language four keywords were defined. They
-are \emph{pr}, \emph{decl}, \emph{opn} and \emph{close}. As the language is
+are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is
 an EDSL, these four keywords are methods in the BOAST namespace.
+%Usage of these keywords can be found on Figure~\ref{fig:keywords}.
+
+The \emph{decl} method is used to declare variables or procedures and functions.
 
@@
 
-The \emph{decl} method is used to declare variables or procedures and functions.
-
 The \emph{opn} method can be used to print the beginning of a control
@@
 
-BOAST defines several class that are used to represent the structure of the
+BOAST defines several classes that are used to represent the structure of the
 code. These classes can be sorted in two groups, algebraic related and control
@@
 with \emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that
-when run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
+produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
 (Listing~\ref{lst:BOAST_algebra_C}) output. At first we define 2 Variables
-\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third variable
-named \emph{l} is a one dimensional local array of a length of $27$, it is indexed
-in the range $-5$ to $21$. All those Variables are affected to ruby variables
-of the corresponding name.
+\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third
+variable named \emph{l} is a one dimensional local array of a length of $27$, it
+is indexed in the range $-5$ to $21$. All those Variables are affected to ruby
+variables of the corresponding name.
 
@@
 
-Figure~\ref{fig:BOAST_control} show some basic usage of the control structures.
+Figure~\ref{fig:BOAST_control} shows some basic usages of the control structures.
 The example a C macro or function that behaves similarly to the FORTRAN modulo
@@
 describe procedures and functions and to meta-program them using Ruby. But in
-order to find the best version of a computing kernel each version has to be
-compiled, linked and then executed to assess it's performance. This can be very
+order to find the best version of a computing kernel, each version has to be
+compiled, linked and then executed to assess its performance. This can be very
 time consuming if this process cannot be automated. By enabling more versions to
-be evaluated the automation will bring improved portability, better performance
-and in the end improve the productivity of the developer.
+be evaluated the automation will bring improved portability, better performance,
+and in the end will improve the productivity of the developer.
 
@@
 Proprietary and dedicated compilers are common on HPC infrastructures. Thus
-BOAST build system shows similar behavior than common build system. Compilers
+BOAST build system exhibits similar behavior than common build system. Compilers
 and their compile/build options can be specified. This can be done at several
 places. Here is the list by increasing order of precedence: BOAST configuration
-file, environment variables and at kernel build time.
+file, environment variables, and at kernel build time.
 
@@
 
-The next logical step is to benchmark the built kernel. Boast offers a simple
+The next logical step is to benchmark the built kernel. BOAST offers a simple
 way to run a kernel that was successfully built while staying inside of BOAST.
@@
 The non regression methodology presented before is viable provided input data
-can be generated at runtime. Unfortunately it is not always possible, some
+can be generated at runtime. For instance in the case of the Laplace kernel,
+generating an input image and the corresponding reference output image, using a
+reference implementation, is easy. Unfortunately it is not always possible, some
 applications have very complicated data patterns that are difficult to
@@
 
-\subsection{Trace Debugging of BOAST-based GPU Applications}
+\subsection{Non Regression Testing Using Trace Debugging}
 \label{sec:gputrace}
 
-Debugging applications running on GPU environments is well-recognized
-as a hard and time-consuming activity. In complement with BOAST, we
-designed a trace-based debugging tool that simplifies this porting
-operation. The tool relies on BOAST support of multi-target code
-generation (Section~\ref{sec:multitarget}), used to port an
-application from one GPU-programming framework to another. A
-case-study of the portage of a Cuda application to OpenCL is presented
-in Section~\ref{subsec:specfem}.
+Debugging applications running on GPU environments is well-recognized as a hard
+and time-consuming activity. In complement with BOAST, we designed a trace-based
+debugging tool that simplifies this porting operation. The tool relies on BOAST
+support of multi-target code generation (Section~\ref{sec:multitarget}), used to
+validate an application from one GPU-programming framework to another or between
+differetn implementation using the same programming model. A case-study of the
+port of a Cuda application to OpenCL is presented in
+Section~\ref{subsec:specfem}.
 
@@
 
-Hence, the verificatation and validation of the new port can be
+Hence, the verification and validation of the new port can be
 narrowed down to asserting that both codes apply the same operations
@@
 motivating example as well as two scientific applications that use BOAST. The
-first one, BigDFT, uses boast in order to develop new functionalities with
-performance portability in mind. The second one, SPECFEM3D, uses BOAST to
-factorize OpenCL and CUDA development, while having robust non regression
-tests.
+first one, BigDFT~\cite{genovese2008daubechies}, uses boast in order to develop new
+functionalities with performance portability in mind. The second one,
+SPECFEM3D~\cite{komatitsch2011fluid}, uses BOAST to factorize OpenCL and CUDA
+development, while having robust non regression tests.
 
@@
 
-In 2005, the EU FP6-STREP-NEST BigDFT project funded a consortium of
-four European laboratories (L\_Sim - CEA Grenoble, Basel University -
-Switzerland, Louvain-la-Neuve University - Belgium and Kiel University -
-Germany), with the aim of developing a novel approach for DFT calculations
-based on Daubechies wavelets.  Rather than simply building a DFT code from
-scratch, the objective of this three-years project was to test the potential
-benefit of a new formalism in the context of electronic structure calculations.
+In 2005, the EU FP6-STREP-NEST BigDFT~\cite{genovese2008daubechies} project
+funded a consortium of four European laboratories (L\_Sim - CEA Grenoble, Basel
+University - Switzerland, Louvain-la-Neuve University - Belgium and Kiel
+University - Germany), with the aim of developing a novel approach for DFT
+calculations based on Daubechies wavelets.  Rather than simply building a DFT
+code from scratch, the objective of this three-years project was to test the
+potential benefit of a new formalism in the context of electronic structure
+calculations.
 
@@
 a localized basis set allows to optimize the number of degrees of freedom for a
-required accuracy (see \cite{bigdft}), which is highly desirable given the
-complexity and inhomogeneity of the systems under investigation nowadays.
+required accuracy~\cite{genovese2008daubechies}, which is highly desirable given
+the complexity and inhomogeneity of the systems under investigation nowadays.
 
@@
 architectures.  In 2009, the French Bull-Fourier was awarded for the
-implementation of the hybrid version of BigDFT.
+implementation of the hybrid version of BigDFT~\cite{genovese2009density}.
 
@@
 defined is beyond the scope of this report and can be found in the BigDFT
-reference paper \cite{bigdft}.
+reference paper~\cite{genovese2008daubechies}.
 
@@
 % has to be taken regarding their performances. 
-The CPU convolutions of BigDFT have thus been thoroughly optimized.  In a
-recent paper, the optimization of the CPU convolutions of BigDFT has been
-extensively considered.
+The CPU convolutions of BigDFT have thus been thoroughly optimized.  In a recent
+paper~\cite{videau2013optimizing}, the optimization of the CPU convolutions of
+BigDFT has been extensively considered.
 %BigDFT is characterized by its extensive use of convolution
@@
 		captionpos=b,
-		caption={MagicFilter}, 
+		caption={MagicFilter.}, 
 		label={lst:conv_example}}
@@
 have been added, making the BigDFT more and more powerful in terms of
-scientific applications.  However, the cost of the maintainance and of
+scientific applications.  However, the cost of the maintenance and of
 development of the convolutions is always a delicate point to be considered
 while including a new functionality.  The convolution patterns are usually
-rather similar, leading to code duplication and difficulties in code reading
-and modification.
+rather similar, leading to code duplication and difficulties in code
+maintainability.
 
@@
 the implementation and the generation of new convolutions.  This would lead to
-optimized code, adapted to  different computing platforms, that is optimally
+an optimized code, adapted to different computing platforms, that is optimally
 factorized.  In addition to this point, the help of such code generator is also
@@
 
-The interest in having a robuts and optimally tuned library in this scientific
+The interest in having a robust and optimally tuned library in this scientific
 field is therefore evident.  Techniques are under investigation to provide also
 fine-tuned binaries to the end-user rather than the code sources, such that
-more agressive interprocedural optimization can be performed.  This experience
-would be therefore a first step toward the release of a tunable optimised
+more aggressive inter procedural optimization can be performed.  This experience
+therefore would be a first step toward the release of a tunable optimised
 convolution library oriented to computational physics communities.
@@
 Inte Xeon X5550 that was used to hand optimize the code.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Res_synthesis}
+\end{center}
+\caption{Impact of Unrolling, Language and OpenMP on a Wavelet Transform}
+\label{fig:synthesis}
+\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.7]{Res_synthesis}
-\caption{Impact of Unrolling, Language and OpenMP on a Wavelet Transform}
-\end{center}
-\label{fig:synthesis}
-\end{figure}
 %   BigDFT implementation of convolution/wavelets is a reference but is limited
@@
 
-    \subsubsection{SPECFEM3D}
-
 In this last subsection, we study a free seismic wave propagation
@@
 computer benchmarking, thanks to its good scaling capabilities.
+
+    \subsubsection{SPECFEM3D}
 
@@
 As part of the Mont-Blanc project, we had to port \Specfem to \OCL, so
-that it could be use to benchmark Mont-Blanc HPC platforms.
+that it could be used to benchmark Mont-Blanc HPC platforms.
 
@@
 This lack of \Specfem internal knowledge led us to be particularly
-attentive to the path we undertook for the porting, as we would have
+attentive to the path we undertook for the port, as we would have
 been unable to understand how and why the application was not
@@
 
-Hence, our first milestone in the porting was the translation of
+Hence, our first milestone in the port was the translation of
 \Specfem's \Cuda kernels into BOAST EDSL. This way, we could ask BOAST
@@
 
-\subparagraph{Porting runtime to \OCL} The second part of the porting
+\subparagraph{Porting runtime to \OCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
@@
 calls'', 790 arguments to set, 230 memory transfers, 160 buffer
-creations and 270 releases.
+creations, and 270 releases.
 
@@
 set of seismograms. And again, as expected and feared, these
-seismograms were not valid.
+seismograms were not valid as their shape was completely different from the
+reference ones.
 
@@
 With the help of \code{GPUTrace} (Section~\ref{sec:gputrace}), we
-could to confront \Cuda and \OCL execution traces with a graphical
+could confront \Cuda and \OCL execution traces with a graphical
 \code{diff} tool, and spot the different porting mistakes: some
@@
 unclear where it started. We finally got it after hours of code review
-of BOAST kernels and \OCL code: one kernel was
+of BOAST kernels and \OCL code. One kernel was
 \emph{three}-dimensional, whereas the others were two-dimensional. But
@@
 missing. As it was the same person who coded the \OCL port and the GPU
-tracer, this third dimension was forgotten in both codes\ldots
+tracer, this third dimension was forgotten in both codes.
 
@@
 
-  Nonetheless many applications formalism cannot be reduced to standardized
+  Nonetheless many application formalisms cannot be reduced to standardized
 library or border on what could be considered edge cases for those libraries and
@@
 using a DSL. All those approaches are very different from our as they put the
-emphasis on compilation techniques whereas boast relies on the user to express
+emphasis on compilation techniques whereas BOAST relies on the user to express
 the different optimizations.
@@
 in the input image (and intermediary results). During generation, depending on
-memory and computing cost some values can be recomputed rather than fetched from
+memory and computing cost, some values can be recomputed rather than fetched from
 memory. We used Halide to implement the \emph{magicfilter} of BigDFT, but
-unfortunately results were 4 to 5 times slower than the one we obtain with
+unfortunately results were 4 to 5 times slower than the one we obtained with
 BOAST. We speculate that the three-dimensional nature of our convolutions, as
@@
 
-  Once auto-tuning technique are used the parameter space explodes and the
+  Once auto-tuning technique are used, the parameter space explodes and the
 systematic sampling rapidly becomes impossible to achieve. The cost of finding
@@
 is one such tool. It reduces the number of sample needed by creating a model of
-the performances and by minimizing the number of samples needed to find the
+the performance and by minimizing the number of samples needed to find the
 parameters of this model. Several models and sampling techniques are

### 2015-03-13 Paper/D5.5/D5.5.tex
Added summary.

@@
 \begin{executive}
-This document describes BOAST, a metaprogramming framework to produce portable and efficient computing kernels for HPC application.
+This document describes BOAST, a metaprogramming framework to produce portable
+and efficient computing kernels for HPC application. BOAST offers an embedded
+domain specific language to describe the kernels and their possible
+optimization. BOAST also supplies a complete runtime to compile, run, benchmark,
+and check the validity of the generated kernels. BOAST is being used in two
+flagship HPC applications BigDFT and SPECFEM3D, to improve performance
+portability of those codes.
 \end{executive}

### 2015-03-13 Paper/D5.5/D5.5.tex
Added a bit more description of the convolution library.

@@
 performances on the architectures encountered by the application while still
-being portable after the optimization process took place. Indeed the application
-might encounter one of the other architectures presented before. Investing
-manpower to optimize the application for a new architecture is reasonable,
-suffering hindrance from previous optimization work is not. Thus optimizations
-have to be as orthogonal as possible from one another so as to be easily
-activated and deactivated.
+being portable after the optimization process took place. Indeed the
+application might encounter one of the many architectures that can be found in
+HPC. Investing manpower to optimize the application for a new architecture is
+reasonable, suffering hindrance from previous optimization work is not. Thus
+optimizations have to be as orthogonal as possible from one another so as to be
+easily activated and deactivated.
 
@@
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
-and called from high-level operations.  Each of the building blocks of these
-convolution libraries is optimally tuned by BOAST by choosing the sources
-providing the optimal kernel for the chosen computing platform.  The sources of
-these kernels are then collected and compiled to meet the API of the library.
+and called from high-level operations. The basic blocks will be comprised of
+unidimensional wavelet transforms and convolutions applied to multidimensional
+arrays. Combining those blocks will yield multidimensional transforms.
+
+Each of the building blocks of these convolution libraries is optimally tuned
+by BOAST by choosing the sources providing the optimal kernel for the chosen
+computing platform.  The sources of these kernels are then collected and
+compiled to meet the API of the library.
 

### 2015-03-13 Paper/D5.5/D5.5.tex
Added more conclusion.

@@
 In this report we presented the BOAST infrastructure as well as its
-application to two use cases from the Mont-Blanc project. Results are
-encouraging as BOAST proved to be a powerful and flexible tool.
+application to three use cases from the Mont-Blanc project. Results are
+encouraging as BOAST proved to be a powerful and flexible tool that allowed
+gains in performance and performance portability.
+
+Future development will focus on three goals. Interfacing with binary analysis
+tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop to
+guide optimization. Interfacing with search space modellers/pruners in order to
+optimize the search of the optimal version of a kernel. Continue working on
+vector code. For instance producing a collection of small to medium useful
+vector patterns (transposition for instance) in BOAST could really help develop
+vectorized version of our algorithm.
 

### 2015-03-23 Paper/D5.5/D5.5.tex
Small fix.
V 0.2 sent to internal reviewers.
Added Olivier's corrections.

@@
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
-\version{0.1}
-\deadline{6}
+\version{0.2}
+\deadline{18}
 \level{RE}
 \nature{Report}
-\coordinator{ (CNRS)}
+\coordinator{CNRS}
 \contributors{Thierry Deutsch, Luigi Genovese, Jean-François Méhaut, Kevin Pouget, Brice Videau}
@@
 \change{0.1}{Initial version of the deliverable}
+\change{0.2}{First version sent to reviewers}
 \end{changelog}
@@
 meteorologist. Those codes are usually written in FORTRAN for historical and
-performance reason. Codes can be quite huge (several thousands lines of code)
+performance reasons. Codes can be quite huge (several thousands lines of code)
 with lots of functionalities. Nonetheless they are usually based on computing
@@
 
-Those applications are often developed by several individuals. Sometimes some of
-those developers only work a few month on the application. Maintaining optimized
-code written by someone else is quite a challenge. Several languages and
-programming paradigms can also be used in a project and thus the maintainer must
-be knowledgeable in several area of expertise. Portability problems can also be
-cause by the availability of an optimizing compiler for a specific language and
-architecture. C compilers are usually the first available while FORTRAN may
-sometimes arrive later.
-
-In Section~\ref{use_cases} we will present two HPC application that we used
+Those applications are often developed by several individuals. Sometimes some
+of those developers only work a few monthes on the application. Maintaining
+optimized code written by someone else is quite a challenge. Several languages
+and programming paradigms can also be used in a project and thus the maintainer
+must be knowledgeable in several areas of expertise. Portability problems can
+also be caused by the availability of an optimizing compiler for a specific
+language and architecture. C compilers are usually the first available while
+FORTRAN may sometimes arrive later.
+
+In Section~\ref{use_cases} we will present two HPC applications that we used
 as use cases: SPECFEM3D and BigDFT. They are both based on computing
@@
 The race to exascale is not going to simplify the environment. All of the above
-architectures can be considered. Network architectures are also part of the
-architecture and can be very diverse.  For instance European FP7 project DEEP
-considers using Accelerators (XEON Phi) while the European FP7 project
-Mont-Blanc considers using low-power embedded processor with integrated GPU.
+architectures can be considered. Network architectures also can be very
+diverse.  For instance European FP7 project DEEP considers using Accelerators
+(XEON Phi) while the European FP7 project Mont-Blanc considers using low-power
+embedded processor with integrated GPU.
 
@@
 
-This is a good example of an algorithm that is simple in its formulation but can
-be complex to optimize because many different optimizations are available and
-can be combined together. The correct combination of optimization will be
-dependant on the targeted architecture.
+This is a good example of an algorithm that is simple in its formulation but
+can be complex to optimize because many different optimizations are available
+and can be combined together. The correct combination of optimizations will
+depend on the targeted architecture.
 
@@
 
-Listing~\ref{lst:laplace_c_filter} show a simple implementation of the Laplace
+Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace
 filter in C99. For clarity, boundary conditions have been omitted.
@@
 vectors available to the Mali architecture. To be efficient the vectors have to
-be directly loaded from memory. Using vectors of 16 elements 15 useful component
+be directly loaded from memory. Using vectors of 16 elements 15 useful components
 are simultaneously loaded and can be used in computation. This optimization
@@
 \paragraph{More Optimizations} Several other optimizations can be attempted at
-this point. For instance reducing or increasing the number of pixel each work
+this point. For instance reducing or increasing the number of pixels each work
 item is processing. They can yield improved performance in some cases, especially
@@
 independently from one another. Some were arbitrarily configured (the number of
-pixel chosen for instance). Testing the different combination of those
+pixels chosen for instance). Testing the different combinations of those
 optimizations and the different parameters would prove very costly in developer
 time and like all repetitive and tedious tasks, error prone. Thus every created
-kernel would have to be thoroughly tested to insure no error was done.
+kernel would have to be thoroughly tested to ensure no error was done.
 
@@
 
-Several of the steps in the workflow can be automated: optimization and compiler
-flags exploration, non regression testing, benchmarking... This automation can
-be done by scripting by the user or by interfacing other dedicated tools. In
-order to achieve those goals three aspects should be considered:
+Several steps in the workflow can be automated: optimization and compiler flags
+exploration, non regression testing, benchmarking... This automation can be
+done by scripting by the user or by interfacing other dedicated tools. In order
+to achieve those goals three aspects should be considered:
 \begin{itemize}
@@
 and the obtained result is often quite different from the original procedure.
-Several transformations can be applied to such kernels. And those optimization
+Several transformations can be applied to such kernels. And those optimizations
 are often applied manually as compiler may fail to recognise the opportunity.
@@
 related to manipulations in the host language four keywords were defined. They
-are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is
-an EDSL, these four keywords are methods in the BOAST namespace.
+are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
+EDSL, these four keywords are methods in the BOAST namespace. Sample usage of
+these keywords will be found in the next Figures.
 %Usage of these keywords can be found on Figure~\ref{fig:keywords}.
@@
 \emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third
-variable named \emph{l} is a one dimensional local array of a length of $27$, it
+variable named \emph{l} is a one dimensional local array of length $27$, it
 is indexed in the range $-5$ to $21$. All those Variables are affected to ruby
@@
 the bracket operator. Would the array be multidimensional the index would be
-coma separated, similar to FORTRAN notation.
+comma separated, similar to FORTRAN notation.
 
@@
 Figure~\ref{fig:BOAST_control} shows some basic usages of the control structures.
-The example a C macro or function that behaves similarly to the FORTRAN modulo
+The example shows a C macro or function that behaves similarly to the FORTRAN modulo
 intrinsic. The sample script (inefficiently) computes and stores in \emph{j} the
@@
 an input array \emph{x} and an output array \emph{y} of double precision
-numbers. Those array have two dimensions which depend on input variables
+numbers. Those arrays have two dimensions which depend on input variables
 \emph{n} and \emph{ndat}. We can see here the use of the \emph{opn} and
@@
 different alternatives. Two methods are dedicated to this task: \emph{set\_lang}
-and \emph{get\_lang}. The target language can also be set through environment
+and \emph{get\_lang}. The target language can also be set through an environment
 variable before launching BOAST, allowing for easy command line scripting.
@@
 Compilation of the generated kernels must also be very flexible. Indeed, HPC
-applications may encounter platforms with very diverse compilation environment.
+applications may encounter platforms with very diverse compilation environments.
 Proprietary and dedicated compilers are common on HPC infrastructures. Thus
-BOAST build system exhibits similar behavior than common build system. Compilers
+BOAST build system exhibits similar behavior than common build systems. Compilers
 and their compile/build options can be specified. This can be done at several
@@
 to the BOAST \emph{Procedure} used to instantiate the kernel. Arrays must be
-instances of \emph{NArray} which are numerical arrays that use C array
+instances of \emph{NArray} which are numerical arrays that use C arrays
 underneath.
@@
 validate an application from one GPU-programming framework to another or between
-differetn implementation using the same programming model. A case-study of the
+different implementations using the same programming model. A case-study of the
 port of a Cuda application to OpenCL is presented in
@@
 
-The idea behind the tool is based on the assumption that the different
-GPU ports of the code should do the very same operations, at least at
-the logical the logical level (the APIs will have
-\emph{implementation} differences, but they should offer nonetheless
-same functionalities). The usage of BOAST framework sustains this
-assumption, as both sets of kernels should generated from the same
+The idea behind the tool is based on the assumption that the different GPU
+ports of the code should do the very same operations, at least at the logical
+level (the APIs will have \emph{implementation} differences, but they should
+offer nonetheless same functionalities). The usage of BOAST framework sustains
+this assumption, as both sets of kernels should be generated from the same
 source code.
@@
 objects (handle value, buffer creation size and attributes, kernel
-name and prototype, \etc{}).
+name and prototype, \etc{}). This is a custom trace format.
 
@@
 different porting mistakes: two parameters reversed, an offset
-incorrectly applied, \etc{}.
+incorrectly applied, \etc{}
 
@@
 Section~\ref{sec:laplace} presented the Laplace motivating example. From this
-section we know that a number of optimization can have an impact on the
+section we know that a number of optimizations can have an impact on the
 performance of the kernel on the Mali architecture. But what of other
@@
 genericity when applying optimizations. We also decided to study the impact of
-another parameter which is the number of component to process on the column
+another parameter which is the number of components to process on the column
 direction. This leads to being able to process tiles instead of only rows.
@@
 
-Table~\ref{tbl:ARM} are the best results obtained by ARM on different image
+Table~\ref{tbl:ARM} are the best results obtained by ARM on different images
 compared to the naive implementation and to the best version version BOAST
@@
 vectors, synthesizing the loads is harmful to performance and the programmer is
-better of trusting the cache to load each vector in 1 cycle without compromising
-the bandwidth.
+better of trusting the cache to load each vector in 1 cycle without
+compromising the bandwidth. The results shown here are the best of 4 runs for
+each configuration. As the test environment was well controlled kernel
+durations were very stable.
 
@@
 
-But what if we run out benchmark on other architectures? Table~\ref{tbl:other}
-the results obtained when running our BOAST implementation on other
+But what if we run our benchmark on other architectures? Table~\ref{tbl:other}
+shows the results obtained when running our BOAST implementation on other
 architectures. The chosen architectures were an Intel i7-2760QM CPU (Sandy
@@
 known algorithm across different architectures. Chosen optimization combinations
-are highly dependant on the targeted architecture. In the next Subsections we
+are highly dependent on the targeted architecture. In the next Subsections we
 will show that our methodology also applies to real applications.
@@
 results, whereas their locality (both in real and reciprocal space) is highly
-desired to improve the efficiency and the flexibility of the treatment. Indeed,
+desired to improve the efficiency and the flexibility of the processing. Indeed,
 a localized basis set allows to optimize the number of degrees of freedom for a
@@
 routinely executed with the BigDFT code, either in homogeneous or hybrid
-architectures.  In 2009, the French Bull-Fourier was awarded for the
+architectures.  In 2009, the French Bull-Fourier award was attributed for the
 implementation of the hybrid version of BigDFT~\cite{genovese2009density}.
@@
 fine-tuned binaries to the end-user rather than the code sources, such that
-more aggressive inter procedural optimization can be performed.  This experience
-therefore would be a first step toward the release of a tunable optimised
-convolution library oriented to computational physics communities.
+more aggressive inter procedural optimization can be performed. Indeed, BOAST
+finds the optimal source code for a given kernel and compiler configuration but
+one could imagine using binary optimizer to the compiled binaries. Those
+optimizers could be coupled to BOAST and the output of the process would be the
+final binary rather than the source code. This experience therefore would be a
+first step toward the release of a tunable optimised convolution library
+oriented to computational physics communities.
 
@@
 Several kernels have already been implemented in BOAST for the convolution
-library. Figure~\ref{fig:synthesis} show the performance of the wavelet
+library. Figure~\ref{fig:synthesis} shows the performance of the wavelet
 transform operation as a factor of the unrolling length of the outer loop, the
@@
 tuned implementation that can be found in BigDFT. These tests were run on the
-Inte Xeon X5550 that was used to hand optimize the code.
+Intel Xeon X5550 that was used to hand optimize the code.
 
@@
 much too high (a factor of 12 was used). This factor might have been optimal at
-the point the procedure was optimized (compiler version changed in the meantime) but
-since then the environment changed. Other optimization have also been
+the time the procedure was optimized (compiler version changed in the meantime) but
+since then the environment changed. Other optimizations have also been
 incorporated in the BAOST sources, like the systematic inner loop unrolling, and
@@
 Nonetheless what is interesting from the physicist point of view is that the
-generated source will give her better performance on a whole range of
-architectures/compiler combination than that of the hand tuned code.
+generated source will give its better performance on a whole range of
+architectures/compilers combinations than that of the hand tuned code.
 
@@
 This lack of \Specfem internal knowledge led us to be particularly
-attentive to the path we undertook for the port, as we would have
+careful to the path we undertook for the port, as we would have
 been unable to understand how and why the application was not
@@
 
-As we had already validated BOAST-generated kernels (and trusted \Cuda
-and \OCL versions to be semantically identical), we knew that the bugs
-were now in the CPU side, and we had to find a way to understand where
-\Specfem's \Cuda version of the code diverged from its \OCL
-counterpart. To help us in that purpose, we had a strong assumption:
-both versions of the code were supposed to perform exactly the same
-operations, with the same ``logical'' parameters (the APIs have
-\emph{implementation} differences, for instance \OCL has two memory
-transfer functions, \code{clEnqueueReadBuffer},
-\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a
-direction parameter \code{cudaMemcpy(..., dir)}, but above that, it is
-the same functionalities).
+As we had already validated BOAST-generated kernels (and trusted \Cuda and \OCL
+versions to be semantically identical), we knew that the bugs were now in the
+usage of the runtime, and we had to find a way to understand where \Specfem's
+\Cuda version of the code diverged from its \OCL counterpart. To help us in
+that purpose, we had a strong assumption: both versions of the code were
+supposed to perform exactly the same operations, with the same ``logical''
+parameters (the APIs have \emph{implementation} differences, for instance \OCL
+has two memory transfer functions, \code{clEnqueueReadBuffer},
+\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a direction
+parameter \code{cudaMemcpy(..., dir)}, but above that, it is the same
+functionalities).
 
@@
 version though (25\% higher than \OCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of
-version 1.1.
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of version
+1.1. This parameter, in addition to specifying the work group size (which can
+be done in \OCL), also constrains the number of work group that must run in
+parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
+the compiler must be very conservative on register usage in order to allow this
+parallelism which allows better overlapping of communications end computations.
+This functionality is not supported in \OCL.
 
@@
 shrank by a factor of 1.8 (from 7500 LOC to less than 4000, partly
-because of code duplication, but by removing manually unrolled
+because of code duplication, also removing manually unrolled
 loops). This is beneficial for \Specfem as it improves the readability
@@
 performance evaluation mechanisms. These tests will help developers to
-try new optimization in kernel code and measure their impact, without
+try new optimizations in kernel code and measure their impact, without
 executing the whole application.
@@
 libraries are either hand tuned for selected platforms or have auto-tuned
-implementations. Atlas~\cite{whaley04} is an auto-tunes implementation of
+implementations. Atlas~\cite{whaley04} is an auto-tuned implementation of
 BLAS/LAPACK. ATLAS authors defined the \emph{Automated Empirical Optimization of
@@
 space exploration to find the best version of a kernel.
+LGen~\cite{spampinato2014basic} is a compiler that generates linear algebra
+programs for small fixed size problems. Knowing the problem size it can fully
+unroll and vectorize some loops, yielding better performance than state of the
+art generic implementations.
 
@@
 
-  Once auto-tuning technique are used, the parameter space explodes and the
+  Once auto-tuning techniques are used, the parameter space explodes and the
 systematic sampling rapidly becomes impossible to achieve. The cost of finding
@@
 to address this problem. Adaptive Sampling Kit (ASK)~\cite{castro2013adaptive}
-is one such tool. It reduces the number of sample needed by creating a model of
+is one such tool. It reduces the number of samples needed by creating a model of
 the performance and by minimizing the number of samples needed to find the
@@
 Listing~\ref{lst:computing_kernel} shows how to create a Computing kernel
-(\textit{CKernel}) and build it. Once a computing kernel is instancied the
+(\textit{CKernel}) and build it. Once a computing kernel is instantiated the
 output of BOAST will be redirected to the computing kernel source code. Line 4
@@
 By default compilation commands are not shown unless an error occurs. This
-behavior can be change by switching to verbose mode.
+behavior can be changed by switching to verbose mode.
 
@@
 kernel returns a hash table containing information about the procedure
-execution. In this simple case two information are returned, first the value
+execution. In this simple case two informations are returned, first the value
 of the output parameter \textit{b} and second the time the kernel execution
@@
 
-Most computing kernels don't work on scalar values but rather on array of
+Most computing kernels don't work on scalar values but rather on arrays of
 data. Listing~\ref{lst:arrays} shows how to use arrays in computing kernels.
@@
 to obtain muti-dimensional arrays.
-For self contained procedures/kernels one can sue the shortcut written on
+For self contained procedures/kernels one can use the shortcut written on
 line~13 to create a CKernel object. As we are not specifying build options the
-build command can also be ommited and will be automatically called when
+build command can also be omitted and will be automatically called when
 running the kernel the first time. Lines~17 to 19 are used to check the result
@@
 
-Listing~\ref{lst:vector_add_decl} shows the addition of two vector in a third
+Listing~\ref{lst:vector_add_decl} shows the addition of two vectors in a third
 one. Here BOAST is configured to have arrays starting at 0 and to use single

### 2015-03-25 Paper/D5.5/D5.5.tex
Small fixes to deliverable.
Version 0.3.

@@
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
-\version{0.2}
+\version{0.3}
 \deadline{18}
@@
 \coordinator{CNRS}
-\contributors{Thierry Deutsch, Luigi Genovese, Jean-François Méhaut, Kevin Pouget, Brice Videau}
-\reviewers{None}
+\contributors{Thierry Deutsch (CEA), Luigi Genovese (CEA), Jean-François Méhaut (UJF), Kevin Pouget (CNRS), Brice Videau (CNRS)}
+\reviewers{Olivier Aumage (INRIA), Frédéric Desprez (INRIA)}
 \keywords{BOAST, Deliverable}
@@
 \change{0.2}{First version sent to reviewers}
+\change{0.3}{With corrections from reviewers}
 \end{changelog}
@@
 \item Written in a \emph{portable} manner,
-\item Written in a way that raise developer \emph{productivity},
+\item Written in a way that raises developer \emph{productivity},
 \item Written to present good \emph{performance}.

### 2015-03-27 Paper/D5.5/D5.5.tex
Minor indent fix.
Small fix before sending to BSC.
Added Harald Servat corrections.

@@
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
-\version{0.3}
+\version{0.5}
 \deadline{18}
@@
 \contributors{Thierry Deutsch (CEA), Luigi Genovese (CEA), Jean-François Méhaut (UJF), Kevin Pouget (CNRS), Brice Videau (CNRS)}
-\reviewers{Olivier Aumage (INRIA), Frédéric Desprez (INRIA)}
+\reviewers{Olivier Aumage (INRIA), Frédéric Desprez (INRIA), Harald Servat (BSC)}
 \keywords{BOAST, Deliverable}
@@
 \change{0.3}{With corrections from reviewers}
+\change{0.4}{Additional corrections from reviewers}
 \end{changelog}
@@
 domain specific language to describe the kernels and their possible
-optimization. BOAST also supplies a complete runtime to compile, run, benchmark,
+optimization. BOAST also supplies a complete run-time to compile, run, benchmark,
 and check the validity of the generated kernels. BOAST is being used in two
@@
 Porting and tuning HPC applications to new platforms is tedious and costly in
-term of human resources. Nonetheless it is a very important aspect of the
-Mont-Blanc project. Indeed, for the project, more than 10 applications were
+terms of human resources. Nonetheless, it is a very important aspect of the
+Mont-Blanc project. Indeed, for the project, more than ten applications were
 selected to be ported and optimized for the prototype platform.
@@
 Genericity of HPC codes is often limited. One of the reason is that producing
-generic code in FORTRAN 90/95 is difficult as the language is not really fit
-for it. Sometimes, adding genericity will degrade performance as optimization
+generic code in Fortran 90/95 is difficult as the language does not really fit
+for it. Sometimes, adding genericity degrades performance as optimization
 opportunities that come from over-specification are lost.
@@
 Scientific Computing Applications are usually developed by physicist, chemist or
-meteorologist. Those codes are usually written in FORTRAN for historical and
-performance reasons. Codes can be quite huge (several thousands lines of code)
-with lots of functionalities. Nonetheless they are usually based on computing
-kernels.  Computing kernels are resource intensive and well defined parts of a
-program that usually work on precisely defined data. Those kernels are the time
-consuming part of an HPC application. They are thus the prime target for
+meteorologist. Those codes are usually written in Fortran for historical and
+performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
+with lots of functionalities. Nonetheless, they are usually based on computing
+kernels.  Computing kernels are resource intensive and well-defined parts of a
+program that usually work on precisely defined data. Those kernels represent the most
+time-consuming part of an HPC application, and consequently, they are the prime target for
 optimization.
 
-Those applications are often developed by several individuals. Sometimes some
-of those developers only work a few monthes on the application. Maintaining
+Those applications are often developed by several individuals. Sometimes, some
+of those developers only work a few months on the application. Maintaining
 optimized code written by someone else is quite a challenge. Several languages
@@
 language and architecture. C compilers are usually the first available while
-FORTRAN may sometimes arrive later.
+Fortran may sometimes arrive later.
 
@@
 
-What we want is computing kernels that are:
+What we want is computing kernels that are written:
 \begin{itemize}
-\item Written in a \emph{portable} manner,
-\item Written in a way that raises developer \emph{productivity},
-\item Written to present good \emph{performance}.
+\item in a \emph{portable} manner,
+\item in a way that raises developer \emph{productivity},
+\item and, present good \emph{performance}.
 \end{itemize}
@@
 \paragraph{Vectorization}
-The first proposed optimization is to compute 5 pixels instead of one using the
-vectors available to the Mali architecture. To be efficient the vectors have to
-be directly loaded from memory. Using vectors of 16 elements 15 useful components
+The first proposed optimization involves computing five pixels instead of one using the
+vectors available to the Mali architecture. The vectors have to reside in memory to
+ensure an efficient execution. Using vectors of 16 elements, 15 useful components
 are simultaneously loaded and can be used in computation. This optimization
@@
 \includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
-\caption{Structure and Workflow of the BOAST framework.}
+\caption{Structure and Work-flow of the BOAST framework.}
 \label{fig:boast_workflow}
@@
 
-The ideas behind BOAST were first presented in~\cite{videau2013boast}, and are
-to provide scientific application developers with a framework to develop and
-test application computing kernels. The envisioned workflow and program
-structure is presented in Figure~\ref{fig:boast_workflow}.  The user should
-start from an application kernel (implemented or just designed) and implement it
-in a dedicated language (step 1). The language provides enough flexibility for
+BOAST provides scientific application developers with a framework to develop and
+test application computing kernels~\cite{videau2013boast}.
+Figure~\ref{fig:boast_workflow} illustrates the envisioned work-flow and program
+structure.
+The user starts from an application kernel (either designed or implemented), and
+write it in a dedicated language (step 1).
+The language provides enough flexibility for
 the kernel to be meta-programmed with several orthogonal optimizations. Then the
-developer selects one configuration of the kernel and a target language. Those
-parameter define the output source code that will be generated by BOAST (step
+developer selects a kernel configuration and a target language. Those
+parameters define the output source code that will be generated by BOAST (step
 2). The resulting code source is then built according to the user specified
@@
 
-Several steps in the workflow can be automated: optimization and compiler flags
+Several steps in the work-flow can be automated: optimization and compiler flags
 exploration, non regression testing, benchmarking... This automation can be
-done by scripting by the user or by interfacing other dedicated tools. In order
+scripted by the user or by interfacing other dedicated tools. In order
 to achieve those goals three aspects should be considered:
@@
 \item code generation,
-\item and kernel execution runtime.
+\item and kernel execution run-time.
 \end{itemize}
@@
 attributes...).
-\item Procedures have to be abstracted (reunite FORTRAN and C like languages,
+\item Procedures have to be abstracted (reunite Fortran and C like languages,
 attributes...).
@@
 In order to manipulate those abstractions we want to have a syntax resembling
-what programmers use. For instance commonly used operators have to be available
+what programmers use. For instance, commonly used operators have to be available
 and behave as expected. It must also be possible to differentiate an action on
@@
 language approach was selected~\cite{hudak1996building}. This allows for the
-coexistence of two languages: the host language and the DSL. In our case the DSL
-allows the description of the kernel ($c = a + b$) while the host language
+coexistence of two languages: the host language and the Domain Specific Language (DSL).
+In our case, DSL allows the description of the kernel ($c = a + b$) while the host language
 provides the meta programming of the kernel ($c \leftarrow a + b$)%, cf Figure~\ref{fig:dsl}
@@
 It was also important to have our constructs like \emph{for} loops to have a
-syntax approaching those commonly found in programming languages. For this we
+syntax approaching those commonly found in programming languages. To this end, we
 needed a language which could seamlessly pass a block of code to a function.
@@
 
-    \subsubsection{BOAST keywords}
+    \subsubsection{BOAST Keywords}
 
@@
 \lstset{style=customf,
-        caption={FORTRAN output.},
+        caption={Fortran output.},
         label={lst:BOAST_algebra_FORTRAN} }
@@
 type, a name and a set of named attributes. The existing attributes are mainly
-inspired from FORTRAN. Those attributes are not limited and can be arbitrarily
+inspired from Fortran. Those attributes are not limited and can be arbitrarily
 enriched, allowing a lot of flexibility in Variable management.
 
-The second abstraction is named Expression. It is used to combine variables into
+The second abstraction is named Expression. It combines variables into
 algebraic or logic expressions. Most of the classical operators are overloaded
@@
 straightforward. The exception is the assignment operator as it is important to
-differentiate between assigning an Expression or a Variable in the ruby context
+differentiate between assigning an Expression or a Variable in the Ruby context
 and the assignment operator in the context of a BOAST Expression. Thus the
@@
 with \emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that
-produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
+produces the Fortran (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
 (Listing~\ref{lst:BOAST_algebra_C}) output. At first we define 2 Variables
@@
 variable named \emph{l} is a one dimensional local array of length $27$, it
-is indexed in the range $-5$ to $21$. All those Variables are affected to ruby
+is indexed in the range $-5$ to $21$. All those Variables are affected to Ruby
 variables of the corresponding name.
@@
 On Line~4 we declare those three variables. Variable \emph{i} is then affected
-the value 5. On Line~6 the \emph{j} ruby variable is used to store the BOAST
+the value 5. On Line~6 the \emph{j} Ruby variable is used to store the BOAST
 expression \emph{$i+5$}. This variable will the be used transparently through
@@
 the bracket operator. Would the array be multidimensional the index would be
-comma separated, similar to FORTRAN notation.
+comma separated, similar to Fortran notation.
 
@@
 \lstset{style=customf,
-        caption={FORTRAN output.},
+        caption={Fortran output.},
         label={lst:BOAST_control_FORTRAN} }
@@
 corresponding control structures in other languages. An exception is the For in
-BOAST more closely matches the for in FORTRAN than the one in C.
+BOAST more closely matches the for in Fortran than the one in C.
 
 Figure~\ref{fig:BOAST_control} shows some basic usages of the control structures.
-The example shows a C macro or function that behaves similarly to the FORTRAN modulo
+The example shows a C macro or function that behaves similarly to the Fortran modulo
 intrinsic. The sample script (inefficiently) computes and stores in \emph{j} the
@@
 evaluated at the time the construct is printed. If several such constructs are
-needed (in an \emph{if elsif else} case for instance) they can be explicitly
+needed (in an \emph{if elsif else} case, for instance) they can be explicitly
 passed as parameters using the \emph{lambda} Ruby keyword.
@@
 \lstset{style=customf,
-        caption={FORTRAN output.},
+        caption={Fortran output.},
         label={lst:BOAST_procedure_FORTRAN} }
@@
 
-The last control structure is Procedure. It is used to describe either procedure
+The last control structure is Procedure. It describes either procedures
 or functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
-abstraction. It is the signature of a real kernel from BigDFT. This kernel uses
+abstraction. It illustrates the signature of a real kernel from BigDFT. This kernel uses
 an input array \emph{x} and an output array \emph{y} of double precision
@@
 
-The generated outputs in FORTRAN~\ref{lst:BOAST_procedure_FORTRAN} and
-C~\ref{lst:BOAST_procedure_C} show the difference in meta-information that is
+The generated outputs in Fortran (Listing~\ref{lst:BOAST_procedure_FORTRAN}) and
+C (Listing~\ref{lst:BOAST_procedure_C}) show the difference in meta-information that is
 kept between both versions.
@@
 
-\subsection{BOAST Runtime}
+\subsection{BOAST Run-time}
 
 In the previous Section we presented BOAST's language. This allows us to
-describe procedures and functions and to meta-program them using Ruby. But in
-order to find the best version of a computing kernel, each version has to be
-compiled, linked and then executed to assess its performance. This can be very
-time consuming if this process cannot be automated. By enabling more versions to
-be evaluated the automation will bring improved portability, better performance,
+describe procedures and functions and to meta-program them using Ruby.
+Each version has to be compiled, linked and executed to assess its performance
+in order to find the best version of a computing kernel.  This can be very
+time consuming if this process cannot be automated. By enabling more versions
+for evaluation, the automation will bring improved portability, better performance,
 and in the end will improve the productivity of the developer.
 
-In this section we will present the different aspects of BOAST runtime that
-allow this automation. Those aspects are listed here:
+In this section we will present the different aspects of BOAST run-time that
+allow this automation. Those aspects include:
   \begin{itemize}
@@
 \label{sec:multitarget}
-Language availability and performance can vary between platforms. It is thus
-important to be able to express computing kernels in different languages, based
+Language availability and performance varies between platforms. It is thus
+important to express computing kernels in different languages, based
 on the availability and their respective merits on the target platform. Some
-languages can have additional features, such as languages that target GPUs
+languages have additional features, such as languages that target GPUs
 (OpenCL, CUDA). The developer should be given tools to determine what kind of
@@
 
-Compilation of the generated kernels must also be very flexible. Indeed, HPC
-applications may encounter platforms with very diverse compilation environments.
+Compilation of the generated kernels must also be very flexible because HPC
+application developers may encounter platforms with very diverse compilation environments.
 Proprietary and dedicated compilers are common on HPC infrastructures. Thus
-BOAST build system exhibits similar behavior than common build systems. Compilers
-and their compile/build options can be specified. This can be done at several
-places. Here is the list by increasing order of precedence: BOAST configuration
+BOAST build system exhibits similar behavior to common build systems. Compilers
+and their compile/build options can be specified at several
+places. The list by increasing order of precedence includes: BOAST configuration
 file, environment variables, and at kernel build time.
@@
 The next logical step is to benchmark the built kernel. BOAST offers a simple
-way to run a kernel that was successfully built while staying inside of BOAST.
+way to run a kernel that was successfully built while staying within BOAST.
 A built BOAST \emph{CKernel} exposes a run method that accepts arguments corresponding
@@
 kernels that are functions as well as output scalars) about the run. For
-instance the run time of the kernel is one such information, it is obtained
-using the system wide realtime clock. Other probes can be inserted at compile
-time if needed.
-
-PAPI~\cite{mucci1999papi} is supported and different counters can be specified
-during each execution of a kernel.
+instance, one such information includes the run time of the kernel and it is
+obtained using the system-wide real-time clock. Other probes can be inserted
+at compile-time, if needed.
+BOAST also supports PAPI~\cite{mucci1999papi} to capture hardware performance 
+counters during each kernel execution.
 
@@
 The non regression methodology presented before is viable provided input data
-can be generated at runtime. For instance in the case of the Laplace kernel,
+can be generated at run-time. For instance, in the case of the Laplace kernel,
 generating an input image and the corresponding reference output image, using a
 reference implementation, is easy. Unfortunately it is not always possible, some
-applications have very complicated data patterns that are difficult to
+applications have complex data patterns that are difficult to
 synthesize without running the full application. Thus BOAST offers a way to load
@@
 different implementations using the same programming model. A case-study of the
-port of a Cuda application to OpenCL is presented in
+port of a CUDA application to OpenCL is presented in
 Section~\ref{subsec:specfem}.
@@
 
+%
+% HSG Question: why GPUTrace has to be state-full?
+%
 However, in contrast with \code{strace} and \code{ltrace},
@@
 
-Once two call traces have been generated by \code{GPUTrace}, it is
-easy to compare them with a graphical \code{diff} tool, and spot the
+Once two call traces have been generated by \code{GPUTrace}, the user
+can compare them with a graphical \code{diff} tool, and spot the
 different porting mistakes: two parameters reversed, an offset
@@
 By default, \code{GPUTrace} only prints a unique identifier (the
-creation index) for the memory buffers. A first flag can be activated
+creation index) for the memory buffers.  Additionally, \code{GPUTrace}
+supports several modifier flags. One flag can be activated
 to append the first bits of the buffer to the trace, printed in the
-right format, for visual inspection. Or another flag can be set to
+right format, for visual inspection. Another flag can be set to
 dump the whole content of the buffer into a file, for a full
@@
 motivating example as well as two scientific applications that use BOAST. The
-first one, BigDFT~\cite{genovese2008daubechies}, uses boast in order to develop new
+first one, BigDFT~\cite{genovese2008daubechies}, uses BOAST in order to develop new
 functionalities with performance portability in mind. The second one,
@@
 section we know that a number of optimizations can have an impact on the
-performance of the kernel on the Mali architecture. But what of other
-architectures? And are there other optimizations that can impact the
+performance of the kernel on the Mali architecture. But, what is the impact in other
+architectures? And, are there additional optimizations that can impact the
 performance?
@@
 \textit{false}) or load more data and compute useless values in vectors of the
-specified size. This last option is mimicking the behavior of ARM
-implementation, although when working at the component level it may seem an
-absurd thing to try.
+specified size (\textit{vector\_recompute} = \textit{true}).
+This last option mimics the behavior of the ARM implementation, although when
+working at the component level it may not be a valuable thing to try.
 
@@
 
-Table~\ref{tbl:ARM} are the best results obtained by ARM on different images
+Table~\ref{tbl:ARM} show the best results obtained by ARM on different images
 compared to the naive implementation and to the best version version BOAST
-found. It shows that the generated version systematically beats the hand
-optimized version. As far as the optimization options are concerned the result
-are a disappointment: the same kernel configuration is the best for all image
-sizes. This kernel is using \textit{x\_component\_number} = 16,
+found. It shows that the generated version systematically outperforms the hand
+optimized version. As far as the optimization options are concerned, the results
+are disappointing: the same kernel configuration is the best for all image
+sizes. This kernel uses \textit{x\_component\_number} = 16,
 \textit{y\_component\_number} = 1, \textit{vector\_length} = 16,
 \textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false.
-\textit{vector\_recompute} has no meaning here as \textit{x\_component\_number}
-== \textit{vector\_length}. Those results also shows that, when working on full
+\textit{vector\_recompute} because \textit{x\_component\_number}
+== \textit{vector\_length}. Those results show that, when working on full
 vectors, synthesizing the loads is harmful to performance and the programmer is
-better of trusting the cache to load each vector in 1 cycle without
-compromising the bandwidth. The results shown here are the best of 4 runs for
-each configuration. As the test environment was well controlled kernel
-durations were very stable.
+better of trusting the cache to load each vector in one cycle without
+compromising the bandwidth. The results shown here are the best of four runs for
+each configuration.
 
@@
 shows the results obtained when running our BOAST implementation on other
-architectures. The chosen architectures were an Intel i7-2760QM CPU (Sandy
-Bridge architecture) that supports OpenCL 1.2 and an NVIDIA gtx680 GPU that
-support OpenCL 1.1. We can see that the performance ratio between the different
-architectures is rather stable across image sizes.
+architectures. The chosen architectures include an Intel i7-2760QM CPU (Sandy
+Bridge architecture) that supports OpenCL 1.2 and a system with an NVIDIA gtx680 GPU that
+supports OpenCL 1.1. We see that the performance ratio between the different
+architectures is stable across image sizes.
 
@@
 \textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false. Once again
-\textit{vector\_recompute} has no meaning. \textit{y\_component\_number} varies
+\textit{vector\_recompute} does not apply. \textit{y\_component\_number} varies
 from 4 to 2 when image size increase, thus decreasing task granularity as the
@@
 \textit{vector\_length} = 4, \textit{temporary\_size} = 2 and
-\textit{synthesize\_loads} = false. Once more \textit{vector\_recompute} has no
+\textit{synthesize\_loads} = false. Once more, \textit{vector\_recompute} has no 
 meaning.
@@
 
-In this Subsection we have shown the interest of BOAST in optimizing a well
-known algorithm across different architectures. Chosen optimization combinations
+In this Subsection we have shown the interest of BOAST in optimizing a well-known
+algorithm across different architectures. Chosen optimization combinations
 are highly dependent on the targeted architecture. In the next Subsections we
@@
 
-The application is mainly written in Fortran and currently includes about 360
-thousands lines of Fortran code and 70 thousands lines of C language, which
-accounts for more than 50 \% of the code base. It is a parallel application
+Despite the application is mainly written in Fortran, it currently includes
+360 kLOC and 70kLOC of Fortran and C languages, respectively, accounting for more
+than 50\% of the code base.  It is a parallel application
 based on the standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also
@@
 convolutions would become so little that other functionalities (for example,
-the generalisation of the BigDFT convolutions to Neumann boundary conditions
+the generalization of the BigDFT convolutions to Neumann boundary conditions
 or the usage of wavelet-on-the-interval basis) can be added with limited
@@
 
-We can see that this function is better optimized using FORTRAN and small
+We can see that this function is better optimized using Fortran and small
 unrolling factors. In the hand optimized version the unrolling factor was chosen
@@
 since then the environment changed. Other optimizations have also been
-incorporated in the BAOST sources, like the systematic inner loop unrolling, and
+incorporated in the BOAST sources, like the systematic inner loop unrolling, and
 those could also help increase performance while limiting the interest of the
@@
 massively parallel accelerator running in disjoint, non addressable
-memory environment. Thanks to that proximity, we, BOAST developers,
-have been able to carry out most of the porting task with only a
-limited knowledge about \Specfem internal physics.
+memory environment. Thanks to that proximity, we have been able to carry
+out most of the porting task with only a limited knowledge about
+\Specfem internal physics.
 
@@
 
-\subparagraph{Porting runtime to \OCL} The second part of the port
+\subparagraph{Porting run-time to \OCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
@@
 
-Once all the automatic (and many manual) transformations were applied,
+Once the transformations were applied,
 compilation errors fixed, and \OCL unsuccessful function calls solved,
@@
 versions to be semantically identical), we knew that the bugs were now in the
-usage of the runtime, and we had to find a way to understand where \Specfem's
+usage of the run-time, and we had to find a way to understand where \Specfem's
 \Cuda version of the code diverged from its \OCL counterpart. To help us in
@@
 parameter \code{cudaMemcpy(..., dir)}, but above that, it is the same
-functionalities).
+functionality).
 
@@
 for all of them, only two dimensions were passed, and one was
-missing. As it was the same person who coded the \OCL port and the GPU
-tracer, this third dimension was forgotten in both codes.
+missing.
+%As it was the same person who coded the \OCL port and the GPU
+%tracer, this third dimension was forgotten in both codes.
 
@@
 version though (25\% higher than \OCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of version
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
 1.1. This parameter, in addition to specifying the work group size (which can
@@
 By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
-shrank by a factor of 1.8 (from 7500 LOC to less than 4000, partly
+shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, partly
 because of code duplication, also removing manually unrolled
@@
 
-We have also been able to enhance \Specfem's non-regression testsuite
+We have also been able to enhance \Specfem's non-regression test-suite
 by adding per-kernel non-regression tests. This was done with the help
@@
 of a particular \emph{valid} kernel execution, as well as the output
-values. Then, during the non-regression testing, BOAST framework can
-load these buffer files, allocate GPU memory and initialize it through
-\Cuda or \OCL runtime, and trigger the kernel execution. A comparison
+values. Then, during the non-regression testing, BOAST framework
+loads these buffer files, allocates GPU memory and initializes it through
+\Cuda or \OCL run-time, and triggers the kernel execution. A comparison
 of the output values (for instance against a maximal error level)
@@
 
-In the same mindset, we provided \Specfem testsuite with kernel
+In the same mindset, we provided \Specfem test-suite with kernel
 performance evaluation mechanisms. These tests will help developers to
@@
 LGen~\cite{spampinato2014basic} is a compiler that generates linear algebra
-programs for small fixed size problems. Knowing the problem size it can fully
-unroll and vectorize some loops, yielding better performance than state of the
+programs for small fixed size problems. Knowing the problem size it fully
+unrolls and vectorizes loops, yielding better performance than state of the
 art generic implementations.
@@
 
-  The idea to describe computing kernels using a Domain Specific Language has
-already been used. SPIRAL~\cite{puschel2004spiral} is a decade old generation
-framework for signal processing. It uses a proprietary DSL, SPL (signal
-processing language), to describe a DSP algorithm. This DSL is then transformed
-into efficient programs in high level languages like C or FORTRAN.
+  The idea to describe computing kernels using a DSL has been already explored.
+SPIRAL~\cite{puschel2004spiral} is a decade old generation
+framework for signal processing. It uses a proprietary DSL, SPL (Signal
+Processing Language), to describe a DSP algorithm. This DSL is then transformed
+into efficient programs in high level languages like C or Fortran.
 POET~\cite{yi2007poet} also uses a DSL to describe custom code transformations,
@@
 in the input image (and intermediary results). During generation, depending on
-memory and computing cost, some values can be recomputed rather than fetched from
+memory and computing cost, some values are recomputed rather than fetched from
 memory. We used Halide to implement the \emph{magicfilter} of BigDFT, but
@@
 BOAST. We speculate that the three-dimensional nature of our convolutions, as
-well as the filter length, can be considered edge cases in halide and quite far
+well as the filter length, can be considered edge cases in Halide and quite far
 from the intended target.
@@
 vector code. For instance producing a collection of small to medium useful
-vector patterns (transposition for instance) in BOAST could really help develop
+vector patterns (transposition, for instance) in BOAST could really help develop
 vectorized version of our algorithm.
@@
 This section will present some simple examples to familiarize the user with
-BOAST. More samples can be found in the git repository.
+BOAST. More samples can be found with source code from the program in the git
+repository: \url{https://github.com/Nanosim-LIG/boast}.
 
@@
 
-BOAST is ruby based, so ruby needs to be installed on the machine.
-Installation of boast can be done using the ruby built-in package manager:
+BOAST is Ruby based, so Ruby needs to be installed on the machine.
+Installation of boast can be done using the Ruby built-in package manager:
 \textit{gem}. See Listing~\ref{lst:BOAST_install} for reference.
@@
 
-The following samples are presented using \textit{irb} ruby interactive
+The following samples are presented using \textit{irb} Ruby interactive
 interpreter. It can be launched using the \textit{irb} command in a terminal.
@@
 Listing~\ref{lst:procedure_declaration} shows the declaration of a procedure
-using the two previous variables as parameters. For clarity irb echoes have
+using the two previous variables as parameters. For clarity, \textit{irb} echoes have
 been suppressed.
@@
 passed to run are only relevant for GPU languages and are thus ignored in
-FORTRAN and C (Line~16). Success is only printed if results are validated,
+Fortran and C (Line~16). Success is only printed if results are validated,
 else an exception is raised (Lines~17 to 20).

### 2015-03-30 Paper/D5.5/D5.5.tex
add details about GPUTrace

@@
   morekeywords={int64_t, int32_t}
+}
+
+\lstdefinestyle{customgputrace}{
+  language=C,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={New, kernel, buffer, Mb, READ_WRITE, Buffer, written,
+    b, at, out}
 }
@@
 \emph{and} output) of the relevant function calls. These information
-are traced in a unified format for all the APIs.
-
-%
-% HSG Question: why GPUTrace has to be state-full?
-%
+are traced in a unified format for all the APIs. This is a custom
+trace format. Listing~\ref{lst:GPUTrace} presents a sample output
+generated during \Specfem tracing.
+
 However, in contrast with \code{strace} and \code{ltrace},
 \code{GPUTrace} has to be \emph{state-full}. Indeed, most of API
-parameters are handles to opaque types, so, in order to provide
-meaningful information, \code{GPUTrace} stores information about these
-objects (handle value, buffer creation size and attributes, kernel
-name and prototype, \etc{}). This is a custom trace format.
+parameters are handles to opaque types. So, in order to generate
+meaningful traces, \code{GPUTrace} gathers information about these
+objects at creation time (handle value, buffer creation size and
+attributes, kernel name and prototype, \etc{}) and re-injects this
+information when the objects used. A state-less implementation of
+\code{GPUTrace} would highly rely on the GPU libraries introspection
+capabilities, which seem not possible at the moment.
+
+\lstset{style=customgputrace,
+        caption={GPUTrace sample trace}, 
+        label={lst:GPUTrace},
+        numbers=left}
+\begin{lstlisting}
+New kernel: update_potential_kernel
+...
+New buffer #94, 2Mb, READ_WRITE (0x74d0420)
+New buffer #95, 2Mb, READ_WRITE (0x74d08d0)
+New buffer #96, 2Mb, READ_WRITE (0x74d0d80)
+...
+Buffer #94 written, 2314764b at +0b 
+Buffer #95 written, 2314764b at +0b
+Buffer #96 written, 2314764b at +0b
+...
+update_disp_veloc_kernel<128,1><4522,1>(
+     float *displ=<buffer #94 1.00e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
+     float *veloc=<buffer #95 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     float *accel=<buffer #96 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     const int size=<578691>
+     const float deltat=<1.365914e-04>
+     const float deltatsqover2=<9.328606e-09>
+     const float deltatover2=<6.829570e-05>
+     ----
+     <out> float *displ=<buffer #94 1.000e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
+     <out> float *veloc=<buffer #95 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     <out> float *accel=<buffer #96 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+);
+\end{lstlisting}
 

### 2015-03-31 Paper/D5.5/D5.5.tex
Added Florent to authors.
Incremented version number for release.

@@
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
-\version{0.5}
+\version{1.0}
 \deadline{18}
@@
 \coordinator{CNRS}
-\contributors{Thierry Deutsch (CEA), Luigi Genovese (CEA), Jean-François Méhaut (UJF), Kevin Pouget (CNRS), Brice Videau (CNRS)}
+\contributors{Florent Bouchez Tichadou (UJF), Thierry Deutsch (CEA), Luigi Genovese (CEA), Jean-François Méhaut (UJF), Kevin Pouget (CNRS), Brice Videau (CNRS)}
 \reviewers{Olivier Aumage (INRIA), Frédéric Desprez (INRIA), Harald Servat (BSC)}
@@
 \change{0.4}{Additional corrections from reviewers}
+\change{1.0}{Version to be sent to the EU}
 \end{changelog}

### 2015-03-31 Paper/D5.5/D5.5.tex
style des tables

@@
 \centering
-\begin{tabular}{|l|r|r|r|r|r|}
-\hline
-Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\
-\hline
+\begin{tabular}{c|r|r|c|r|c}
+% \hline
+  Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\[4pt]
+  \hline&&&&\\[-8pt]
 768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
-\hline
+% \hline
 2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
-\hline
+% \hline
 2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
-\hline
+% \hline
 5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
-\hline
+% \hline
 7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
-\hline
+% \hline
 \end{tabular}
@@
 \centering
-\begin{tabular}{|l|r|r|r|r|r|}
-\hline
-Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\
-\hline
+\begin{tabular}{c|r|r|c|r|c}
+% \hline
+  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\[4pt]
+  \hline&&&&\\[-8pt]
 768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
-\hline
+% \hline
 2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
-\hline
+% \hline
 2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
-\hline
+% \hline
 5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
-\hline
+% \hline
 7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
-\hline
+% \hline
 \end{tabular}

### 2015-03-31 Paper/BOAST.tex
Paper updated with D5.5 modifications.

@@
 \newcommand{\latin}[1]{\textit{#1}}
+\newcommand{\eg}{\latin{e.g.,}\xspace}
+\newcommand{\etc}[1]{\latin{etc}}
 
@@
 \newcommand{\Cuda}{\productname{Cuda}\xspace}
-\newcommand{\eg}{\latin{e.g.,}\xspace}
 \newcommand{\OCL}{\productname{OpenCL}\xspace}
 \newcommand{\code}[1]{\texttt{#1}}
-
-\newcommand{\etc}[1]{\latin{etc}}
-
-%\usepackage{lineno}
-%\modulolinenumbers[5]
-%\linenumbers
 
@@
 
+\lstdefinestyle{customgputrace}{
+  language=C,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={New, kernel, buffer, Mb, READ_WRITE, Buffer, written,
+    b, at, out}
+}
+
+\lstdefinestyle{customcl}{
+  language=C,
+  basicstyle=\scriptsize,
+  backgroundcolor=\color{white},
+  frame=single,
+  captionpos=b,
+  morekeywords={__global, __local, global, local, __kernel, kernel, short,
+uchar, uchar16, int16, short16}
+}
+
 \lstdefinestyle{BOAST}{
@@
 Porting and tuning HPC applications to new platforms is tedious and costly in
-term of human resources.
-
-Portability efforts are often lost when migrating to a new architecture, or code
-lose maintainability because several versions of the code coexist, usually with
-a lot of duplication.
+terms of human resources. Nonetheless, it is a very important aspect of the
+Mont-Blanc project. Indeed, for the project, more than ten applications were
+selected to be ported and optimized for the prototype platform.
+
+Unfortunately, portability efforts are often lost when migrating to a new
+architecture. Worse, code may lose maintainability because several versions of
+some functionalities coexist, usually with a lot of duplication.
 
@@
 
-Genericity of HPC codes is often limited. Producing generic code in FORTRAN
-90/95 is difficult as the language is not really  fit for it.
+Genericity of HPC codes is often limited. One of the reason is that producing
+generic code in Fortran 90/95 is difficult as the language does not really fit
+for it. Sometimes, adding genericity degrades performance as optimization
+opportunities that come from over-specification are lost.
 
@@
 \section{Background and Motivation}
+
+  \subsection{Scientific Computing Applications}
+
+Scientific Computing Applications are usually developed by physicist, chemist or
+meteorologist. Those codes are usually written in Fortran for historical and
+performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
+with lots of functionalities. Nonetheless, they are usually based on computing
+kernels.  Computing kernels are resource intensive and well-defined parts of a
+program that usually work on precisely defined data. Those kernels represent the most
+time-consuming part of an HPC application, and consequently, they are the prime target for
+optimization.
+
+Those applications are often developed by several individuals. Sometimes, some
+of those developers only work a few months on the application. Maintaining
+optimized code written by someone else is quite a challenge. Several languages
+and programming paradigms can also be used in a project and thus the maintainer
+must be knowledgeable in several areas of expertise. Portability problems can
+also be caused by the availability of an optimizing compiler for a specific
+language and architecture. C compilers are usually the first available while
+Fortran may sometimes arrive later.
+
+In Section~\ref{use_cases} we will present two HPC applications that we used
+as use cases: SPECFEM3D and BigDFT. They are both based on computing
+kernels and were selected as candidate applications in the Mont-Blanc project.
+
+  \subsection{How Should Computing Kernels be Written?}
+
+The problematic here is to obtain computing kernels that present good
+performances on the architectures encountered by the application while still
+being portable after the optimization process took place. Indeed the
+application might encounter one of the many architectures that can be found in
+HPC. Investing manpower to optimize the application for a new architecture is
+reasonable, suffering hindrance from previous optimization work is not. Thus
+optimizations have to be as orthogonal as possible from one another so as to be
+easily activated and deactivated.
+
+If this paradigm is followed by developers then they will rapidly be confronted
+with a huge optimization space to search. They will need to be able to test
+easily the performance impact of the chosen optimizations without running the
+full application. The same reasoning implies that kernels should be tested for
+non regression without running the full application.
+
+What we want is computing kernels that are written:
+\begin{itemize}
+\item in a \emph{portable} manner,
+\item in a way that raises developer \emph{productivity},
+\item and, present good \emph{performance}.
+\end{itemize}
 
@@
 Being able to efficiently use those architectures on such a small
-time-frame is challenging. Stress frequency of architecture changes.
+time-frame is challenging.
 
 The race to exascale is not going to simplify the environment. All of the above
-architectures can be considered. Network architectures are also part of the
-architecture and can be very diverse.  For instance European FP7 project DEEP
-considers using Accelerators (XEON Phi) while the European FP7 project
-Mont-Blanc considers using low-power embedded processor with integrated GPU.
-
-Running existing applications on those new architectures is an open project.
-Thus, those projects have work packages dedicated to applications. Those
-workpackages are dedicated to porting and optimizing Scientific applications on
-those new architectures. In the DEEP project six applications were selected for
-porting and optimizing, eleven were selected in the Mont-Blanc project.
-
-  \subsection{Scientific Computing Applications}
-
-Developed by physicist, chemist, meteorologist. Usually in FORTRAN for
-historical and performance reason. Codes can be quite huge (several
-thousands lines of code) with lots of functionalities. Nonetheless they are
-usually based on computing kernels. Computing kernels are resource
-intensive and well defined parts of a program with a limited set of input
-and output. Those kernels are the time consuming part of an HPC
-application. They are thus the prime target for optimization.
-
-Those applications are often developed by several individuals. Sometimes some of
-those developers only work a few month on the application. Maintaining optimized
-code written by someone else is quite a challenge.
-
-In Section~\ref{use_cases} we will present two HPC application that we used
-as use cases: SPECFEM3D and BigDFT. They are both based on computing
-kernels and were selected as candidate applications in the Mont-Blanc project.
-
-  \subsection{How Should Computing Kernels be Written?}
-
-The problematic here is to obtain computing kernels that present good
-performances on the architectures encountered by the application while still
-being portable after the optimization process took place. Indeed the application
-might encounter one of the other architectures presented before. Investing
-manpower to optimize the application for a new architecture is reasonable,
-suffering hindrance from previous optimization work is not. Thus optimizations
-have to be as orthogonal as possible from one another so as to be easily
-activated and deactivated.
-
-If this paradigm is followed by developers then they will rapidly be confronted
-with a huge optimization space to search. They will need to be able to test
-easily the performance impact of the chosen optimizations without running the
-full application. The same reasoning implies that kernels should be tested for
-non regression without running the full application.
-
-What we want is computing kernels that are:
-\begin{itemize}
-\item Written in a \emph{portable} manner,
-\item Written in a way that raise developer \emph{productivity},
-\item Written to present good \emph{performance}.
-\end{itemize}
+architectures can be considered. Network architectures also can be very
+diverse.  For instance European FP7 project DEEP considers using Accelerators
+(XEON Phi) while the European FP7 project Mont-Blanc considers using low-power
+embedded processor with integrated GPU.
+
+Running existing applications on those new architectures is an open research
+subject as well as an ongoing porting effort.  Thus, those projects have work
+packages dedicated to applications. Those work packages are dedicated to porting
+and optimizing Scientific applications on those new architectures. In the DEEP
+project six applications were selected for porting and optimizing, eleven were
+selected in the Mont-Blanc project.
+
+\section{Motivating Example: OpenCL Laplace}
+\label{sec:laplace}
+
+During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
+on the Mali GPU was given. One of the case study was a Laplace filter on the
+GPU~\cite{opencl_arm_training}.
+
+This is a good example of an algorithm that is simple in its formulation but
+can be complex to optimize because many different optimizations are available
+and can be combined together. The correct combination of optimizations will
+depend on the targeted architecture.
+
+\subsection{The Laplace Filter}
+
+Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace
+filter in C99. For clarity, boundary conditions have been omitted.
+
+\lstset{style=customc,
+        caption={Laplace C99 Filter},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_c_filter}}
+\begin{lstlisting}
+void laplace(const int width,
+             const int height,
+             const unsigned char src[height][width][3],
+                   unsigned char dst[height][width][3]){
+  for (int j = 1; j < height-1; j++) {
+    for (int i = 1; i < width-1; i++) {
+      for (int c = 0; c < 3; c++) {
+        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
+                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
+                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
+        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
+      }
+    }
+  }
+}
+\end{lstlisting}
+
+A naive implementation in OpenCL is proposed in
+Listing~\ref{lst:laplace_opencl_filter}. Each work item processes one pixel of
+the resulting image.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter}}
+\begin{lstlisting}
+kernel laplace(const int width,
+               const int height,
+               global const uchar *src,
+               global       uchar *dst){
+  int i = get_global_id(0);
+  int j = get_global_id(1);
+  for (int c = 0; c < 3; c++) {
+    int tmp = -src[3*width*(j-1) + 3*(i-1) + c]\
+             - src[3*width*(j-1) + 3*(i  ) + c]\
+             - src[3*width*(j-1) + 3*(i+1) + c]\
+             - src[3*width*(j  ) + 3*(i-1) + c]\
+           + 9*src[3*width*(j  ) + 3*(i  ) + c]\
+             - src[3*width*(j  ) + 3*(i+1) + c]\
+             - src[3*width*(j+1) + 3*(i-1) + c]\
+             - src[3*width*(j+1) + 3*(i  ) + c]\
+             - src[3*width*(j+1) + 3*(i+1) + c];
+    dst[3*width*j + 3*i + c] = clamp(tmp, 0, 255);
+  }
+}
+\end{lstlisting}
+
+\subsection{Possible Optimizations of the OpenCL Version}
+
+Several optimizations were proposed and successively applied to the previous
+implementation.
+
+\paragraph{Vectorization}
+The first proposed optimization involves computing five pixels instead of one using the
+vectors available to the Mali architecture. The vectors have to reside in memory to
+ensure an efficient execution. Using vectors of 16 elements, 15 useful components
+are simultaneously loaded and can be used in computation. This optimization
+yields speedups between 1.5 and 6 depending on the image size.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_vector}}
+\begin{lstlisting}
+kernel laplace(const int width,
+               const int height,
+               global const uchar *src,
+               global       uchar *dst){
+  int i = get_global_id(0);
+  int j = get_global_id(1);
+
+  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
+  uchar16 v12_ = vload16( 0, src + 3*width*(j-1) + 3*5*i     );
+  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
+  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
+  uchar16 v22_ = vload16( 0, src + 3*width*(j  ) + 3*5*i     );
+  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
+  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
+  uchar16 v32_ = vload16( 0, src + 3*width*(j+1) + 3*5*i     );
+  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
+
+  int16 v11 = convert_int16(v11_);
+  int16 v12 = convert_int16(v12_);
+  int16 v13 = convert_int16(v13_);
+  int16 v21 = convert_int16(v21_);
+  int16 v22 = convert_int16(v22_);
+  int16 v23 = convert_int16(v23_);
+  int16 v31 = convert_int16(v31_);
+  int16 v32 = convert_int16(v32_);
+  int16 v33 = convert_int16(v33_);
+
+  int16 res = v22 * (int)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
+        res = clamp(res, (int16)0, (int16)255);
+  uchar res_ = convert_uchar16(res);
+
+  vstore8(res_.s01234567, 0, dst + 3*width*j + 3*5*i);
+  vstore8(res_.s89ab,     0, dst + 3*width*j + 3*5*i + 8);
+  vstore8(res_.scd,       0, dst + 3*width*j + 3*5*i + 12);
+  dst[3*width*j + 3*5*i + 14] = res_.se;
+}
+\end{lstlisting}
+
+\paragraph{Synthesizing Loads} It is possible to reduce the number of loads
+since the vectors are overlapping.
+Listing~\ref{lst:laplace_opencl_filter_vector_synth} shows how the vectors are
+loaded in this case. This optimization yields marginal improvements.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_vector_synth}}
+\begin{lstlisting}
+  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
+  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
+  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
+  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
+  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
+  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
+  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
+  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
+  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
+\end{lstlisting}
+
+\paragraph{Temporary Variables Size} Using \textit{int} to store intermediary
+results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
+the code is modified to use smaller types. This yields a speedup of 1.3 for most
+image sizes.
+
+\lstset{style=customcl,
+        caption={Laplace OpenCL Filter Vectorized Using \textit{short}},
+        breaklines=true,
+        breakautoindent=false,
+        label={lst:laplace_opencl_filter_short}}
+\begin{lstlisting}
+  short16 v11 = convert_short16(v11_);
+  short16 v12 = convert_short16(v12_);
+  short16 v13 = convert_short16(v13_);
+  short16 v21 = convert_short16(v21_);
+  short16 v22 = convert_short16(v22_);
+  short16 v23 = convert_short16(v23_);
+  short16 v31 = convert_short16(v31_);
+  short16 v32 = convert_short16(v32_);
+  short16 v33 = convert_short16(v33_);
+
+  short16 res = v22 * (short)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
+          res = clamp(res, (short16)0, (short16)255);
+\end{lstlisting}
+
+\paragraph{More Optimizations} Several other optimizations can be attempted at
+this point. For instance reducing or increasing the number of pixels each work
+item is processing. They can yield improved performance in some cases, especially
+when avoiding memory alignment problems. These optimizations will not be shown
+here but can be found in~\cite{opencl_arm_training}.
+
+\subsection{Improving the Methodology}
+
+The process described in the previous Subsection is quite tedious and requires
+some intimate knowledge of the target architecture. The results are impressive,
+as speedups of almost 10 are observed compared to the naive OpenCL
+implementation.
+
+Nonetheless, the process is frustrating. The optimizations are never evaluated
+independently from one another. Some were arbitrarily configured (the number of
+pixels chosen for instance). Testing the different combinations of those
+optimizations and the different parameters would prove very costly in developer
+time and like all repetitive and tedious tasks, error prone. Thus every created
+kernel would have to be thoroughly tested to ensure no error was done.
+
+In the next Section we will present our answer to this problematic: BOAST.
+BOAST is a framework dedicated to kernel description, optimization, regression testing
+and autotuning.
 
@@
 
-  The idea behind BOAST is to provide scientific application developers with a
-framework to develop and test application computing kernels. The envisioned
-workflow and program structure is presented in Figure~\ref{fig:boast_workflow}.
-The user should start from an application kernel (implemented or just designed)
-and implement it in a dedicated language (step 1). The language provides enough
-flexibility for the kernel to be meta-programmed with several orthogonal
-optimizations. Then the developer selects one configuration of the kernel and a
-target language. Those parameter define the output source code that will be
-generated by BOAST (step 2). The resulting code source is then built according to
-the user specified compiler and options (step 3). If input data are available
-then the kernel can be benchmarked and tested for non regression. Based on the
-results, other optimizations can be selected (step 4) or new optimizations can
-be added to the BOAST sources. The process can be repeated until a good
-candidate is found on the target platform. The resulting kernel is then added to
-the program (step 5).
-
-Several of the steps in the workflow can be automated: optimization and compiler
-flags exploration, non regression testing, benchmarking... This automation can
-be done by scripting by the user or by interfacing other dedicated tools. In
-order to achieve those goals three aspects should be considered:
+BOAST provides scientific application developers with a framework to develop and
+test application computing kernels~\cite{videau2013boast}.
+Figure~\ref{fig:boast_workflow} illustrates the envisioned work-flow and program
+structure.
+The user starts from an application kernel (either designed or implemented), and
+write it in a dedicated language (step 1).
+The language provides enough flexibility for
+the kernel to be meta-programmed with several orthogonal optimizations. Then the
+developer selects a kernel configuration and a target language. Those
+parameters define the output source code that will be generated by BOAST (step
+2). The resulting code source is then built according to the user specified
+compiler and options (step 3). If input data are available then the kernel can
+be benchmarked and tested for non regression. Based on the results, other
+optimizations can be selected (step 4) or new optimizations can be added to the
+BOAST sources. The process can be repeated until a good candidate is found on
+the target platform. The resulting kernel is then added to the program (step 5).
+
+Several steps in the work-flow can be automated: optimization and compiler flags
+exploration, non regression testing, benchmarking... This automation can be
+scripted by the user or by interfacing other dedicated tools. In order
+to achieve those goals three aspects should be considered:
 \begin{itemize}
@@
 \item code generation,
-\item and kernel execution runtime.
+\item and kernel execution run-time.
 \end{itemize}
 
-  \cite{videau2013boast}
-
   \subsection{Kernel Description Language}
 
-  \begin{itemize}
-  \item Abstraction of programming concepts: variables, procedures (
-portability, productivity )
-  \item Operator overloading: simpler syntax ( productivity, portability )
-  \item Two languages in one ( productivity, performance ) 
-  \end{itemize}
-
-Usually computing kernels are hotspots of an HPC application, and are most of
-the time based on a loop nest. A lot of effort is dedicated to their tuning and
-the obtained result is often quite different than the original procedure.
-Several transformations can be applied to such kernels. And those optimization
+%  \begin{itemize}
+%  \item Abstraction of programming concepts: variables, procedures (
+%portability, productivity )
+%  \item Operator overloading: simpler syntax ( productivity, portability )
+%  \item Two languages in one ( productivity, performance ) 
+%  \end{itemize}
+
+Usually computing kernels are hotspots of an HPC application, and they are most
+of the time based on a loop nest. A lot of efforts are dedicated to their tuning
+and the obtained result is often quite different from the original procedure.
+Several transformations can be applied to such kernels. And those optimizations
 are often applied manually as compiler may fail to recognise the opportunity.
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.8]{BOAST_DSL.pdf}
-\caption{BOAST DSL intent.}
-\label{fig:dsl}
-\end{center}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[scale=0.8]{BOAST_DSL.pdf}
+%\caption{BOAST DSL intent.}
+%\label{fig:dsl}
+%\end{center}
+%\end{figure}
 
 In order to manipulate those abstractions we want to have a syntax resembling
-what programmers use. For instance commonly used operator have to be available
+what programmers use. For instance, commonly used operators have to be available
 and behave as expected. It must also be possible to differentiate an action on
 an abstraction in the context of an expression and in the context of the
-management of this expression. This is why an embedded domain specific language
-approach was selected~\cite{hudak1996building}. This allows for the coexistence
-of two languages: the host language and the DSL. In our case the DSL allows the
-description of the kernel while the host language provides the meta programming
-of the kernel, cf Figure~\ref{fig:dsl}. Operator overloading of the host
-language will provide the familiar syntax programmers are accustomed to.
+management of this expression. $c = a + b$, an expression that affects the
+results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves the
+expression $a + b$ to a variable $c$. This is why an embedded domain specific
+language approach was selected~\cite{hudak1996building}. This allows for the
+coexistence of two languages: the host language and the Domain Specific Language (DSL).
+In our case, DSL allows the description of the kernel ($c = a + b$) while the host language
+provides the meta programming of the kernel ($c \leftarrow a + b$)%, cf Figure~\ref{fig:dsl}
+. Operator overloading of the host language will provide the familiar syntax
+programmers are accustomed to. The host language will also provide easy
+interfacing with libraries we might need during the development of our framework.
 
 It was also important to have our constructs like \emph{for} loops to have a
-syntax approaching those commonly found in programming languages. For this we
+syntax approaching those commonly found in programming languages. To this end, we
 needed a language which could seamlessly pass a block of code to a function.
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
-capabilities as well. This is the main reason it was selected.
-
-    \subsubsection{BOAST keywords}
+capabilities as well. This is the main reason why it was selected.
+
+    \subsubsection{BOAST Keywords}
 
@@
 related to manipulations in the host language four keywords were defined. They
-are \emph{pr}, \emph{decl}, \emph{opn} and \emph{close}. As the language is
-an EDSL, these four keywords are methods in the BOAST namespace.
+are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
+EDSL, these four keywords are methods in the BOAST namespace. Sample usage of
+these keywords will be found in the next Figures.
+%Usage of these keywords can be found on Figure~\ref{fig:keywords}.
+
+The \emph{decl} method is used to declare variables or procedures and functions.
 
@@
 with \emph{BOAST::}. Listing~\ref{lst:BOAST_algebra} shows the BOAST code that
-when run produces the FORTRAN (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
+produces the Fortran (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
 (Listing~\ref{lst:BOAST_algebra_C}) output. At first we define 2 Variables
-\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third variable
-named \emph{l} is a one dimensional local array of a length of $27$, it is indexed
-in the range $-5$ to $21$. All those Variables are affected to ruby variables
-of the corresponding name.
+\emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third
+variable named \emph{l} is a one dimensional local array of length $27$, it
+is indexed in the range $-5$ to $21$. All those Variables are affected to Ruby
+variables of the corresponding name.
 
 On Line~4 we declare those three variables. Variable \emph{i} is then affected
-the value 5. On Line~6 the \emph{j} ruby variable is used to store the BOAST
+the value 5. On Line~6 the \emph{j} Ruby variable is used to store the BOAST
 expression \emph{$i+5$}. This variable will the be used transparently through
@@
 the bracket operator. Would the array be multidimensional the index would be
-coma separated, similar to FORTRAN notation.
+comma separated, similar to Fortran notation.
 
@@
 corresponding control structures in other languages. An exception is the For in
-BOAST more closely matches the for in FORTRAN than the one in C.
-
-Figure~\ref{fig:BOAST_control} show some basic usage of the control structures.
-The example a C macro or function that behaves similarly to the FORTRAN modulo
+BOAST more closely matches the for in Fortran than the one in C.
+
+Figure~\ref{fig:BOAST_control} shows some basic usages of the control structures.
+The example shows a C macro or function that behaves similarly to the Fortran modulo
 intrinsic. The sample script (inefficiently) computes and stores in \emph{j} the
@@
 
-The last control structure is Procedure. It is used to describe either procedure
+The last control structure is Procedure. It describes either procedures
 or functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
-abstraction. It is the signature of a real kernel from BigDFT. This kernel uses
+abstraction. It illustrates the signature of a real kernel from BigDFT. This kernel uses
 an input array \emph{x} and an output array \emph{y} of double precision
-numbers. Those array have two dimensions which depend on input variables
+numbers. Those arrays have two dimensions which depend on input variables
 \emph{n} and \emph{ndat}. We can see here the use of the \emph{opn} and
@@
 
-\subsection{BOAST Runtime}
+\subsection{BOAST Run-time}
 
 In the previous Section we presented BOAST's language. This allows us to
-describe procedures and functions and to meta-program them using Ruby. But in
-order to find the best version of a computing kernel each version has to be
-compiled, linked and then executed to assess it's performance. This can be very
-time consuming if this process cannot be automated. By enabling more versions to
-be evaluated the automation will bring improved portability, better performance
-and in the end improve the productivity of the developer.
-
-In this section we will present the different aspects of BOAST runtime that
-allow this automation. Those aspects are listed here:
+describe procedures and functions and to meta-program them using Ruby.
+Each version has to be compiled, linked and executed to assess its performance
+in order to find the best version of a computing kernel.  This can be very
+time consuming if this process cannot be automated. By enabling more versions
+for evaluation, the automation will bring improved portability, better performance,
+and in the end will improve the productivity of the developer.
+
+In this section we will present the different aspects of BOAST run-time that
+allow this automation. Those aspects include:
   \begin{itemize}
@@
 
-Compilation of the generated kernels must also be very flexible. Indeed, HPC
-applications may encounter platforms with very diverse compilation environment.
+Compilation of the generated kernels must also be very flexible because HPC
+application developers may encounter platforms with very diverse compilation environments.
 Proprietary and dedicated compilers are common on HPC infrastructures. Thus
-BOAST build system shows similar behavior than common build system. Compilers
-and their compile/build options can be specified. This can be done at several
-places. Here is the list by increasing order of precedence: BOAST configuration
-file, environment variables and at kernel build time.
+BOAST build system exhibits similar behavior to common build systems. Compilers
+and their compile/build options can be specified at several
+places. The list by increasing order of precedence includes: BOAST configuration
+file, environment variables, and at kernel build time.
 
@@
 
-The next logical step is to benchmark the built kernel. Boast offers a simple
-way to run a kernel that was successfully built while staying inside of BOAST.
+The next logical step is to benchmark the built kernel. BOAST offers a simple
+way to run a kernel that was successfully built while staying within BOAST.
 A built BOAST \emph{CKernel} exposes a run method that accepts arguments corresponding
 to the BOAST \emph{Procedure} used to instantiate the kernel. Arrays must be
-instances of \emph{NArray} which are numerical arrays that use C array
+instances of \emph{NArray} which are numerical arrays that use C arrays
 underneath.
@@
 regression testing. The run method also returns information (and result for
-kernels that are functions) about the run. For instance the run time of the
-kernel is one such information, it is obtained using the system wide realtime
-clock. Other probes can be inserted at compile time if needed.
+kernels that are functions as well as output scalars) about the run. For
+instance, one such information includes the run time of the kernel and it is
+obtained using the system-wide real-time clock. Other probes can be inserted
+at compile-time, if needed.
+BOAST also supports PAPI~\cite{mucci1999papi} to capture hardware performance 
+counters during each kernel execution.
 
@@
 The non regression methodology presented before is viable provided input data
-can be generated at runtime. Unfortunately it is not always possible, some
-applications have very complicated data patterns that are difficult to
+can be generated at run-time. For instance, in the case of the Laplace kernel,
+generating an input image and the corresponding reference output image, using a
+reference implementation, is easy. Unfortunately it is not always possible, some
+applications have complex data patterns that are difficult to
 synthesize without running the full application. Thus BOAST offers a way to load
@@
 
-\subsection{Trace Debugging of BOAST-based GPU Applications}
+\subsection{Non Regression Testing Using Trace Debugging}
 \label{sec:gputrace}
 
-Debugging applications running on GPU environments is well-recognized
-as a hard and time-consuming activity. In complement with BOAST, we
-designed a trace-based debugging tool that simplifies this porting
-operation. The tool relies on BOAST support of multi-target code
-generation (Section~\ref{sec:multitarget}), used to port an
-application from one GPU-programming framework to another. A
-case-study of the portage of a Cuda application to OpenCL is presented
-in Section~\ref{subsec:specfem}.
-
-The idea behind the tool is based on the assumption that the different
-GPU ports of the code should do the very same operations, at least at
-the logical the logical level (the APIs will have
-\emph{implementation} differences, but they should offer nonetheless
-same functionalities). The usage of BOAST framework sustains this
-assumption, as both sets of kernels should generated from the same
+Debugging applications running on GPU environments is well-recognized as a hard
+and time-consuming activity. In complement with BOAST, we designed a trace-based
+debugging tool that simplifies this porting operation. The tool relies on BOAST
+support of multi-target code generation (Section~\ref{sec:multitarget}), used to
+validate an application from one GPU-programming framework to another or between
+different implementations using the same programming model. A case-study of the
+port of a CUDA application to OpenCL is presented in
+Section~\ref{subsec:specfem}.
+
+The idea behind the tool is based on the assumption that the different GPU
+ports of the code should do the very same operations, at least at the logical
+level (the APIs will have \emph{implementation} differences, but they should
+offer nonetheless same functionalities). The usage of BOAST framework sustains
+this assumption, as both sets of kernels should be generated from the same
 source code.
 
-Hence, the verificatation and validation of the new port can be
+Hence, the verification and validation of the new port can be
 narrowed down to asserting that both codes apply the same operations
@@
 \emph{and} output) of the relevant function calls. These information
-are traced in a unified format for all the APIs.
+are traced in a unified format for all the APIs. This is a custom
+trace format. Listing~\ref{lst:GPUTrace} presents a sample output
+generated during \Specfem tracing.
 
@@
 \code{GPUTrace} has to be \emph{state-full}. Indeed, most of API
-parameters are handles to opaque types, so, in order to provide
-meaningful information, \code{GPUTrace} stores information about these
-objects (handle value, buffer creation size and attributes, kernel
-name and prototype, \etc{}).
-
-Once two call traces have been generated by \code{GPUTrace}, it is
-easy to compare them with a graphical \code{diff} tool, and spot the
+parameters are handles to opaque types. So, in order to generate
+meaningful traces, \code{GPUTrace} gathers information about these
+objects at creation time (handle value, buffer creation size and
+attributes, kernel name and prototype, \etc{}) and re-injects this
+information when the objects used. A state-less implementation of
+\code{GPUTrace} would highly rely on the GPU libraries introspection
+capabilities, which seem not possible at the moment.
+
+\lstset{style=customgputrace,
+        caption={GPUTrace sample trace}, 
+        label={lst:GPUTrace},
+        numbers=left}
+\begin{lstlisting}
+New kernel: update_potential_kernel
+...
+New buffer #94, 2Mb, READ_WRITE (0x74d0420)
+New buffer #95, 2Mb, READ_WRITE (0x74d08d0)
+New buffer #96, 2Mb, READ_WRITE (0x74d0d80)
+...
+Buffer #94 written, 2314764b at +0b 
+Buffer #95 written, 2314764b at +0b
+Buffer #96 written, 2314764b at +0b
+...
+update_disp_veloc_kernel<128,1><4522,1>(
+     float *displ=<buffer #94 1.00e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
+     float *veloc=<buffer #95 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     float *accel=<buffer #96 0.00e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     const int size=<578691>
+     const float deltat=<1.365914e-04>
+     const float deltatsqover2=<9.328606e-09>
+     const float deltatover2=<6.829570e-05>
+     ----
+     <out> float *displ=<buffer #94 1.000e-24, 1.00e-24, 1.00e-24, 1.00e-24, ...>
+     <out> float *veloc=<buffer #95 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+     <out> float *accel=<buffer #96 0.000e+00, 0.00e+00, 0.00e+00, 0.00e+00, ...>
+);
+\end{lstlisting}
+
+Once two call traces have been generated by \code{GPUTrace}, the user
+can compare them with a graphical \code{diff} tool, and spot the
 different porting mistakes: two parameters reversed, an offset
-incorrectly applied, \etc{}.
+incorrectly applied, \etc{}
 
 By default, \code{GPUTrace} only prints a unique identifier (the
-creation index) for the memory buffers. A first flag can be activated
+creation index) for the memory buffers.  Additionally, \code{GPUTrace}
+supports several modifier flags. One flag can be activated
 to append the first bits of the buffer to the trace, printed in the
-right format, for visual inspection. Or another flag can be set to
+right format, for visual inspection. Another flag can be set to
 dump the whole content of the buffer into a file, for a full
@@
 
-In this section we will present two scientific applications that use BOAST. The
-first one, BigDFT, uses boast in order to develop new functionalities with
-performance portability in mind. The second one, SPECFEM3D, uses BOAST to
-factorize OpenCL and CUDA development, while having robust non regression
-tests.
+In this section we will present the benefits of using BOAST on the Laplace
+motivating example as well as two scientific applications that use BOAST. The
+first one, BigDFT~\cite{genovese2008daubechies}, uses BOAST in order to develop new
+functionalities with performance portability in mind. The second one,
+SPECFEM3D~\cite{komatitsch2011fluid}, uses BOAST to factorize OpenCL and CUDA
+development, while having robust non regression tests.
+
+  \subsection{Laplace Filter Kernel}
+
+Section~\ref{sec:laplace} presented the Laplace motivating example. From this
+section we know that a number of optimizations can have an impact on the
+performance of the kernel on the Mali architecture. But, what is the impact in other
+architectures? And, are there additional optimizations that can impact the
+performance?
+
+    \subsubsection{Optimization Space}
+
+The list of already identified optimizations are:
+\begin{itemize}
+\item vectorization,
+\item intermediary data type,
+\item number of pixels processed,
+\item and synthesizing loads.
+\end{itemize}
+To create our generic implementation we decided to work at the component level
+rather than the pixel level. This approach leads to more flexibility and
+genericity when applying optimizations. We also decided to study the impact of
+another parameter which is the number of components to process on the column
+direction. This leads to being able to process tiles instead of only rows.
+
+Here are the parameters we finally selected for our kernel optimization and
+their possible values:
+\begin{itemize}
+\item \textit{x\_component\_number}: a positive integer
+\item \textit{y\_component\_number}: a positive integer
+\item \textit{vector\_length}: 1, 2, 4, 8 or 16
+\item \textit{temporary\_size}: 2 or 4
+\item \textit{synthesize\_loads}: \textit{true} or \textit{false}
+\item \textit{vector\_recompute}: \textit{true} or \textit{false}
+\end{itemize}
+The last parameter is used when \textit{x\_component\_number} is not divisible by
+\textit{vector\_length}. Two solutions are possible then, divide the remainder
+of the division in vectors of smaller sizes (\textit{vector\_recompute} =
+\textit{false}) or load more data and compute useless values in vectors of the
+specified size (\textit{vector\_recompute} = \textit{true}).
+This last option mimics the behavior of the ARM implementation, although when
+working at the component level it may not be a valuable thing to try.
+
+    \subsubsection{Performance Results}
+
+Table~\ref{tbl:ARM} show the best results obtained by ARM on different images
+compared to the naive implementation and to the best version version BOAST
+found. It shows that the generated version systematically outperforms the hand
+optimized version. As far as the optimization options are concerned, the results
+are disappointing: the same kernel configuration is the best for all image
+sizes. This kernel uses \textit{x\_component\_number} = 16,
+\textit{y\_component\_number} = 1, \textit{vector\_length} = 16,
+\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false.
+\textit{vector\_recompute} because \textit{x\_component\_number}
+== \textit{vector\_length}. Those results show that, when working on full
+vectors, synthesizing the loads is harmful to performance and the programmer is
+better of trusting the cache to load each vector in one cycle without
+compromising the bandwidth. The results shown here are the best of four runs for
+each configuration.
+
+\begin{table}
+\centering
+\begin{tabular}{c|r|r|c|r|c}
+% \hline
+  Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\[4pt]
+  \hline&&&&\\[-8pt]
+768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
+% \hline
+2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
+% \hline
+2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
+% \hline
+5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
+% \hline
+7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7 \\
+% \hline
+\end{tabular}
+\caption{Best performance of ARM Laplace kernel.}
+\label{tbl:ARM}
+\end{table}
+
+But what if we run our benchmark on other architectures? Table~\ref{tbl:other}
+shows the results obtained when running our BOAST implementation on other
+architectures. The chosen architectures include an Intel i7-2760QM CPU (Sandy
+Bridge architecture) that supports OpenCL 1.2 and a system with an NVIDIA gtx680 GPU that
+supports OpenCL 1.1. We see that the performance ratio between the different
+architectures is stable across image sizes.
+
+The optimization parameters selected are not the same for those architectures.
+Indeed, the Intel CPU favors kernels that have the parameters:
+\textit{x\_component\_number} = 16,
+\textit{vector\_length} = 8,
+\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false. Once again
+\textit{vector\_recompute} does not apply. \textit{y\_component\_number} varies
+from 4 to 2 when image size increase, thus decreasing task granularity as the
+global work size increase. This result is unexpected and understanding why this
+happen could be interesting. The NVIDIA GPU favors processing square tiles:
+\textit{x\_component\_number} = 4, \textit{y\_component\_number} = 4,
+\textit{vector\_length} = 4, \textit{temporary\_size} = 2 and
+\textit{synthesize\_loads} = false. Once more, \textit{vector\_recompute} has no 
+meaning.
+
+\begin{table}
+\centering
+\begin{tabular}{c|r|r|c|r|c}
+% \hline
+  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\[4pt]
+  \hline&&&&\\[-8pt]
+768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
+% \hline
+2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
+% \hline
+2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
+% \hline
+5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
+% \hline
+7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3  \\
+% \hline
+\end{tabular}
+\caption{Best performance of Laplace Kernel on several architectures.}
+\label{tbl:other}
+\end{table}
+
+    \subsubsection{Laplace Conclusion}
+
+In this Subsection we have shown the interest of BOAST in optimizing a well-known
+algorithm across different architectures. Chosen optimization combinations
+are highly dependent on the targeted architecture. In the next Subsections we
+will show that our methodology also applies to real applications.
 
@@
 results, whereas their locality (both in real and reciprocal space) is highly
-desired to improve the efficiency and the flexibility of the treatment. Indeed,
+desired to improve the efficiency and the flexibility of the processing. Indeed,
 a localized basis set allows to optimize the number of degrees of freedom for a
-required accuracy (see \cite{bigdft}), which is highly desirable given the
-complexity and inhomogeneity of the systems under investigation nowadays.
-
-The application is mainly written in Fortran and currently includes about 360
-thousands lines of Fortran code and 70 thousands lines of C language, which
-accounts for more than 50 \% of the code base. It is a parallel application
+required accuracy~\cite{genovese2008daubechies}, which is highly desirable given
+the complexity and inhomogeneity of the systems under investigation nowadays.
+
+Despite the application is mainly written in Fortran, it currently includes
+360 kLOC and 70kLOC of Fortran and C languages, respectively, accounting for more
+than 50\% of the code base.  It is a parallel application
 based on the standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also
@@
 routinely executed with the BigDFT code, either in homogeneous or hybrid
-architectures.  In 2009, the French Bull-Fourier was awarded for the
-implementation of the hybrid version of BigDFT.
+architectures.  In 2009, the French Bull-Fourier award was attributed for the
+implementation of the hybrid version of BigDFT~\cite{genovese2009density}.
 
@@
 short, separable filters.  The detailed description of how these operations are
-defined is beyond the scope of this paper and can be found in the BigDFT
-reference paper \cite{bigdft}.
+defined is beyond the scope of this report and can be found in the BigDFT
+reference paper~\cite{genovese2008daubechies}.
 
@@
 
-The interest in having a robuts and optimally tuned library in this scientific
+The interest in having a robust and optimally tuned library in this scientific
 field is therefore evident. Techniques are under investigation to provide also
 fine-tuned binaries to the end-user rather than the code sources, such that
-more agressive interprocedural optimization can be performed.  This experience
-would be therefore a first step toward the release of a tunable optimised
-convolution library oriented to computational physics communities.
+more aggressive inter procedural optimization can be performed. Indeed, BOAST
+finds the optimal source code for a given kernel and compiler configuration but
+one could imagine using binary optimizer to the compiled binaries. Those
+optimizers could be coupled to BOAST and the output of the process would be the
+final binary rather than the source code. This experience therefore would be a
+first step toward the release of a tunable optimised convolution library
+oriented to computational physics communities.
+
+    \subsubsection{Performance report}
+
+Several kernels have already been implemented in BOAST for the convolution
+library. Figure~\ref{fig:synthesis} shows the performance of the wavelet
+transform operation as a factor of the unrolling length of the outer loop, the
+language used to implement it as well as the activation or not of the OpenMP
+parallelization. Results are given as a speedup compared to the sequential hand
+tuned implementation that can be found in BigDFT. These tests were run on the
+Intel Xeon X5550 that was used to hand optimize the code.
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.7]{Res_synthesis}
+\end{center}
+\caption{Impact of Unrolling, Language and OpenMP on a Wavelet Transform}
+\label{fig:synthesis}
+\end{figure}
+
+We can see that this function is better optimized using Fortran and small
+unrolling factors. In the hand optimized version the unrolling factor was chosen
+much too high (a factor of 12 was used). This factor might have been optimal at
+the time the procedure was optimized (compiler version changed in the meantime) but
+since then the environment changed. Other optimizations have also been
+incorporated in the BOAST sources, like the systematic inner loop unrolling, and
+those could also help increase performance while limiting the interest of the
+outer loop unrolling.
+
+Nonetheless what is interesting from the physicist point of view is that the
+generated source will give its better performance on a whole range of
+architectures/compilers combinations than that of the hand tuned code.
+
 %   BigDFT implementation of convolution/wavelets is a reference but is limited
@@
 
-\subsubsection{SPECFEM3D}
-
 In this last subsection, we study a free seismic wave propagation
@@
 massively parallel accelerator running in disjoint, non addressable
-memory environment. Thanks to that proximity, we, BOAST developers,
-have been able to carry out most of the porting task with only a
-limited knowledge about \Specfem internal physics.
+memory environment. Thanks to that proximity, we have been able to carry
+out most of the porting task with only a limited knowledge about
+\Specfem internal physics.
 
 This lack of \Specfem internal knowledge led us to be particularly
-attentive to the path we undertook for the porting, as we would have
+careful to the path we undertook for the port, as we would have
 been unable to understand how and why the application was not
@@
 
-\subparagraph{Porting runtime to \OCL} The second part of the porting
+\subparagraph{Porting run-time to \OCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
@@
 calls'', 790 arguments to set, 230 memory transfers, 160 buffer
-creations and 270 releases.
-
-Once all the automatic (and many manual) transformations were applied,
+creations, and 270 releases.
+
+Once the transformations were applied,
 compilation errors fixed, and \OCL unsuccessful function calls solved,
@@
 set of seismograms. And again, as expected and feared, these
-seismograms were not valid.
-
-As we had already validated BOAST-generated kernels (and trusted \Cuda
-and \OCL versions to be semantically identical), we knew that the bugs
-were now in the CPU side, and we had to find a way to understand where
-\Specfem's \Cuda version of the code diverged from its \OCL
-counterpart. To help us in that purpose, we had a strong assumption:
-both versions of the code were supposed to perform exactly the same
-operations, with the same ``logical'' parameters (the APIs have
-\emph{implementation} differences, for instance \OCL has two memory
-transfer functions, \code{clEnqueueReadBuffer},
-\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a
-direction parameter \code{cudaMemcpy(..., dir)}, but above that, it is
-the same functionalities).
+seismograms were not valid as their shape was completely different from the
+reference ones.
+
+As we had already validated BOAST-generated kernels (and trusted \Cuda and \OCL
+versions to be semantically identical), we knew that the bugs were now in the
+usage of the run-time, and we had to find a way to understand where \Specfem's
+\Cuda version of the code diverged from its \OCL counterpart. To help us in
+that purpose, we had a strong assumption: both versions of the code were
+supposed to perform exactly the same operations, with the same ``logical''
+parameters (the APIs have \emph{implementation} differences, for instance \OCL
+has two memory transfer functions, \code{clEnqueueReadBuffer},
+\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a direction
+parameter \code{cudaMemcpy(..., dir)}, but above that, it is the same
+functionality).
 
@@
 unclear where it started. We finally got it after hours of code review
-of BOAST kernels and \OCL code: one kernel was
+of BOAST kernels and \OCL code. One kernel was
 \emph{three}-dimensional, whereas the others were two-dimensional. But
 for all of them, only two dimensions were passed, and one was
-missing. As it was the same person who coded the \OCL port and the GPU
-tracer, this third dimension was forgotten in both codes\ldots
+missing.
+%As it was the same person who coded the \OCL port and the GPU
+%tracer, this third dimension was forgotten in both codes.
 
@@
 version though (25\% higher than \OCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL runtime, as of
-version 1.1.
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
+1.1. This parameter, in addition to specifying the work group size (which can
+be done in \OCL), also constrains the number of work group that must run in
+parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
+the compiler must be very conservative on register usage in order to allow this
+parallelism which allows better overlapping of communications end computations.
+This functionality is not supported in \OCL.
 
 By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
-shrank by a factor of 1.8 (from 7500 LOC to less than 4000, partly
-because of code duplication, but by removing manually unrolled
+shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, partly
+because of code duplication, also removing manually unrolled
 loops). This is beneficial for \Specfem as it improves the readability
@@
 
-In the same mindset, we provided \Specfem testsuite with kernel
+In the same mindset, we provided \Specfem test-suite with kernel
 performance evaluation mechanisms. These tests will help developers to
-try new optimization in kernel code and measure their impact, without
+try new optimizations in kernel code and measure their impact, without
 executing the whole application.
@@
 space exploration to find the best version of a kernel.
+LGen~\cite{spampinato2014basic} is a compiler that generates linear algebra
+programs for small fixed size problems. Knowing the problem size it fully
+unrolls and vectorizes loops, yielding better performance than state of the
+art generic implementations.
 
@@
 in the input image (and intermediary results). During generation, depending on
-memory and computing cost some values can be recomputed rather than fetched from
+memory and computing cost, some values are recomputed rather than fetched from
 memory. We used Halide to implement the \emph{magicfilter} of BigDFT, but
-unfortunately results were 4 to 5 times slower than the one we obtain with
+unfortunately results were 4 to 5 times slower than the one we obtained with
 BOAST. We speculate that the three-dimensional nature of our convolutions, as
-well as the filter length, can be considered edge cases in halide and quite far
+well as the filter length, can be considered edge cases in Halide and quite far
 from the intended target.
@@
 
-  Once auto-tuning technique are used the parameter space explodes and the
+  Once auto-tuning techniques are used, the parameter space explodes and the
 systematic sampling rapidly becomes impossible to achieve. The cost of finding
@@
 to address this problem. Adaptive Sampling Kit (ASK)~\cite{castro2013adaptive}
-is one such tool. It reduces the number of sample needed by creating a model of
-the performances and by minimizing the number of samples needed to find the
+is one such tool. It reduces the number of samples needed by creating a model of
+the performance and by minimizing the number of samples needed to find the
 parameters of this model. Several models and sampling techniques are
@@
 
-\section*{References}
+In this report we presented the BOAST infrastructure as well as its
+application to three use cases from the Mont-Blanc project. Results are
+encouraging as BOAST proved to be a powerful and flexible tool that allowed
+gains in performance and performance portability.
+
+Future development will focus on three goals. Interfacing with binary analysis
+tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop to
+guide optimization. Interfacing with search space modellers/pruners in order to
+optimize the search of the optimal version of a kernel. Continue working on
+vector code. For instance producing a collection of small to medium useful
+vector patterns (transposition, for instance) in BOAST could really help develop
+vectorized version of our algorithm.
+
+%\section*{References}
 

### 2015-07-06 Paper/BOAST.tex
With corrections from Frédéric.

@@
 \newcommand{\eg}{\latin{e.g.,}\xspace}
-\newcommand{\etc}[1]{\latin{etc}}
+\newcommand{\etc}[1]{\latin{etc...}}
 
@@
 \section{Background and Motivation}
-
-  \subsection{Scientific Computing Applications}
-
-Scientific Computing Applications are usually developed by physicist, chemist or
-meteorologist. Those codes are usually written in Fortran for historical and
-performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
-with lots of functionalities. Nonetheless, they are usually based on computing
-kernels.  Computing kernels are resource intensive and well-defined parts of a
-program that usually work on precisely defined data. Those kernels represent the most
-time-consuming part of an HPC application, and consequently, they are the prime target for
-optimization.
-
-Those applications are often developed by several individuals. Sometimes, some
-of those developers only work a few months on the application. Maintaining
-optimized code written by someone else is quite a challenge. Several languages
-and programming paradigms can also be used in a project and thus the maintainer
-must be knowledgeable in several areas of expertise. Portability problems can
-also be caused by the availability of an optimizing compiler for a specific
-language and architecture. C compilers are usually the first available while
-Fortran may sometimes arrive later.
-
-In Section~\ref{use_cases} we will present two HPC applications that we used
-as use cases: SPECFEM3D and BigDFT. They are both based on computing
-kernels and were selected as candidate applications in the Mont-Blanc project.
-
-  \subsection{How Should Computing Kernels be Written?}
-
-The problematic here is to obtain computing kernels that present good
-performances on the architectures encountered by the application while still
-being portable after the optimization process took place. Indeed the
-application might encounter one of the many architectures that can be found in
-HPC. Investing manpower to optimize the application for a new architecture is
-reasonable, suffering hindrance from previous optimization work is not. Thus
-optimizations have to be as orthogonal as possible from one another so as to be
-easily activated and deactivated.
-
-If this paradigm is followed by developers then they will rapidly be confronted
-with a huge optimization space to search. They will need to be able to test
-easily the performance impact of the chosen optimizations without running the
-full application. The same reasoning implies that kernels should be tested for
-non regression without running the full application.
-
-What we want is computing kernels that are written:
-\begin{itemize}
-\item in a \emph{portable} manner,
-\item in a way that raises developer \emph{productivity},
-\item and, present good \emph{performance}.
-\end{itemize}
 
@@
 
+  \subsection{Scientific Computing Applications}
+
+Scientific Computing Applications are usually developed by physicist, chemist or
+meteorologist. Those codes are usually written in Fortran for historical and
+performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
+with lots of functionalities. Nonetheless, they are usually based on computing
+kernels.  Computing kernels are resource intensive and well-defined parts of a
+program that usually work on precisely defined data. Those kernels represent the most
+time-consuming part of an HPC application, and consequently, they are the prime target for
+optimization.
+
+Those applications are often developed by several individuals. Sometimes, some
+of those developers only work a few months on the application. Maintaining
+optimized code written by someone else is quite a challenge. Several languages
+and programming paradigms can also be used in a project and thus the maintainer
+must be knowledgeable in several areas of expertise. Portability problems can
+also be caused by the availability of an optimizing compiler for a specific
+language and architecture. C compilers are usually the first available while
+Fortran may sometimes arrive later.
+
+In Section~\ref{use_cases} we will present two HPC applications that we used
+as use cases: SPECFEM3D and BigDFT. They are both based on computing
+kernels and were selected as candidate applications in the Mont-Blanc project.
+
+  \subsection{How Should Computing Kernels be Written?}
+
+The problematic here is to obtain computing kernels that present good
+performances on the architectures encountered by the application while still
+being portable after the optimization process took place. Indeed the
+application might encounter one of the many architectures that can be found in
+HPC. Investing manpower to optimize the application for a new architecture is
+reasonable, suffering hindrance from previous optimization work is not. Thus
+optimizations have to be as orthogonal as possible from one another so as to be
+easily activated and deactivated.
+
+If this paradigm is followed by developers then they will rapidly be confronted
+with a huge optimization space to search. They will need to be able to test
+easily the performance impact of the chosen optimizations without running the
+full application. The same reasoning implies that kernels should be tested for
+non regression without running the full application.
+
+What we want is computing kernels that are written:
+\begin{itemize}
+\item in a \emph{portable} manner,
+\item in a way that raises developer \emph{productivity},
+\item and, present good \emph{performance}.
+\end{itemize}
+
 \section{Motivating Example: OpenCL Laplace}
@@
 and can be combined together. The correct combination of optimizations will
-depend on the targeted architecture.
+depend on the target architecture.
 
@@
 when avoiding memory alignment problems. These optimizations will not be shown
-here but can be found in~\cite{opencl_arm_training}.
+here but can be found in~\cite{opencl_arm_training}. Others consist in tiling
+the problem space or improving the loads synthesis.
 
@@
 pixels chosen for instance). Testing the different combinations of those
-optimizations and the different parameters would prove very costly in developer
-time and like all repetitive and tedious tasks, error prone. Thus every created
-kernel would have to be thoroughly tested to ensure no error was done.
-
-In the next Section we will present our answer to this problematic: BOAST.
-BOAST is a framework dedicated to kernel description, optimization, regression testing
+optimizations and the different parameters would be very costly in developer
+time and, moreover, like all repetitive and tedious tasks, error prone. Thus
+every created kernel would have to be thoroughly tested to ensure no error was
+done.
+
+In the next Section we will present our answer to these issues: BOAST.  BOAST is
+a framework dedicated to kernel description, optimization, regression testing
 and autotuning.
@@
 \includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
-\caption{Structure and Work-flow of the BOAST framework.}
+\caption{Structure and Workflow of the BOAST framework.}
 \label{fig:boast_workflow}
@@
 test application computing kernels~\cite{videau2013boast}.
-Figure~\ref{fig:boast_workflow} illustrates the envisioned work-flow and program
-structure.
-The user starts from an application kernel (either designed or implemented), and
-write it in a dedicated language (step 1).
-The language provides enough flexibility for
-the kernel to be meta-programmed with several orthogonal optimizations. Then the
-developer selects a kernel configuration and a target language. Those
-parameters define the output source code that will be generated by BOAST (step
-2). The resulting code source is then built according to the user specified
-compiler and options (step 3). If input data are available then the kernel can
-be benchmarked and tested for non regression. Based on the results, other
-optimizations can be selected (step 4) or new optimizations can be added to the
-BOAST sources. The process can be repeated until a good candidate is found on
-the target platform. The resulting kernel is then added to the program (step 5).
-
-Several steps in the work-flow can be automated: optimization and compiler flags
+Figure~\ref{fig:boast_workflow} illustrates the envisioned workflow and program
+structure.  The user starts from an application kernel (either designed or
+implemented), and writes it in a dedicated language (step 1).  The language
+provides enough flexibility for the kernel to be meta-programmed with several
+orthogonal optimizations. Then the developer selects a kernel configuration and
+a target language. Those parameters define the output source code that will be
+generated by BOAST (step 2). The resulting code source is then built according
+to the user specified compiler and options (step 3). If input data are available
+then the kernel can be benchmarked and tested for non regression. Based on the
+results, other optimizations can be selected (step 4) or new optimizations can
+be added to the BOAST sources. The process can be repeated until a good
+candidate is found on the target platform. The resulting kernel is then added to
+the program (step 5).
+
+Several steps in the workflow can be automated: optimization and compiler flags
 exploration, non regression testing, benchmarking... This automation can be
 scripted by the user or by interfacing other dedicated tools. In order
-to achieve those goals three aspects should be considered:
-\begin{itemize}
-\item code description,
-\item code generation,
-\item and kernel execution run-time.
-\end{itemize}
+to achieve those goals three aspects should be considered: code description,
+code generation, and kernel execution run-time.
 
@@
 Usually computing kernels are hotspots of an HPC application, and they are most
-of the time based on a loop nest. A lot of efforts are dedicated to their tuning
+of the time based on loop nests. A lot of efforts are dedicated to their tuning
 and the obtained result is often quite different from the original procedure.
@@
 \item Functions must be available.
-\item Variables, constants and functions must be composed in complex
-Expressions.
+\item Variables, constants and functions should be able to be composed in
+complex Expressions.
 \item Basic control structures (for, while, if/else...) have to be abstracted.
@@
 expression $a + b$ to a variable $c$. This is why an embedded domain specific
-language approach was selected~\cite{hudak1996building}. This allows for the
-coexistence of two languages: the host language and the Domain Specific Language (DSL).
-In our case, DSL allows the description of the kernel ($c = a + b$) while the host language
-provides the meta programming of the kernel ($c \leftarrow a + b$)%, cf Figure~\ref{fig:dsl}
-. Operator overloading of the host language will provide the familiar syntax
-programmers are accustomed to. The host language will also provide easy
-interfacing with libraries we might need during the development of our framework.
+language approach was selected~\cite{hudak1996building}. This feature allows for
+the coexistence of two languages: the host language and the Domain Specific
+Language (DSL).  In our case, DSL allows the description of the kernel ($c = a +
+b$) while the host language provides the meta programming of the kernel ($c
+\leftarrow a + b$). Operator overloading of the host language will provide the
+familiar syntax programmers are accustomed to.
 
@@
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
-capabilities as well. This is the main reason why it was selected.
+capabilities as well. This is the main reason why it was selected for BOAST.
+
+One of the added advantage of using a high-level scripting language like Ruby
+as the host language if its interfacing capabilities, providing an easy way
+to use the libraries needed during the development of the framework.
 
@@
 
-The \emph{pr} method calls the public \emph{pr} method of Objects it is
+The \emph{pr} method calls the public \emph{pr} method of objects it is
 called on. Each BOAST object is responsible for printing itself correctly
 depending on the BOAST configuration at the time the print public method is
-called. Calling directly the print method of a BOAST object yields the same
+called. Calling directly the \emph{pr} method of a BOAST object yields the same
 result.
@@
 code. These classes can be sorted in two groups, algebraic related and control
-flow related:
+flow related.
 
@@
 \end{minipage}
-\caption{BOAST Code Snippet for Variables and Expressions}
+\caption{BOAST Code Snippet for Variables and Expressions.}
 \label{fig:BOAST_algebra}
@@
 The first and most fundamental abstraction is named Variable. Variables have a
-type, a name and a set of named attributes. The existing attributes are mainly
+type, a name, and a set of named attributes. The existing attributes are mainly
 inspired from Fortran. Those attributes are not limited and can be arbitrarily
@@
 
-On Line~8 we use the \emph{k} variable to index into the array \emph{l} using
-the bracket operator. Would the array be multidimensional the index would be
-comma separated, similar to Fortran notation.
+On Line~8 we use variable \emph{k} to index into array \emph{l} using the
+bracket operator. Would the array be multidimensional, the index would be comma
+separated, similar to Fortran notation.
 
@@
 \end{minipage}
-\caption{BOAST Code Snippet for control structures}
+\caption{BOAST Code Snippet for control structures.}
 \label{fig:BOAST_control}
@@
 \emph{While}, \emph{Case} are abstractions in BOAST matching the behavior of
-corresponding control structures in other languages. An exception is the For in
-BOAST more closely matches the for in Fortran than the one in C.
+corresponding control structures in other languages. An exception is the
+\emph{For} in BOAST matches more closely the \emph{For} in Fortran than the one
+in C.
 
@@
 \end{lstlisting}
-\caption{BOAST Code Snippet for Procedure}
+\caption{BOAST Code Snippet for Procedure.}
 \label{fig:BOAST_procedure}
@@
 
-The last control structure is Procedure. It describes either procedures
-or functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
-abstraction. It illustrates the signature of a real kernel from BigDFT. This kernel uses
-an input array \emph{x} and an output array \emph{y} of double precision
-numbers. Those arrays have two dimensions which depend on input variables
-\emph{n} and \emph{ndat}. We can see here the use of the \emph{opn} and
-\emph{close} keywords that are used to print a control structure without an
-associated Ruby block. This time we placed ourselves inside of BOAST namespace.
+The last control structure is Procedure. It describes either procedures or
+functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
+abstraction. It illustrates the signature of a real kernel from BigDFT. This
+kernel uses an input array \emph{x} and an output array \emph{y}, both composed
+of double precision numbers. Those arrays have two dimensions which depend on
+input variables \emph{n} and \emph{ndat}. We can see here the use of the
+\emph{opn} and \emph{close} keywords that are used to print a control structure
+without an associated Ruby block. This time we placed ourselves inside the BOAST
+namespace.
 
@@
 describe procedures and functions and to meta-program them using Ruby.
-Each version has to be compiled, linked and executed to assess its performance
+Each version has to be compiled, linked, and executed to assess its performance
 in order to find the best version of a computing kernel.  This can be very
@@
 Language availability and performance varies between platforms. It is thus
-important to express computing kernels in different languages, based
-on the availability and their respective merits on the target platform. Some
-languages have additional features, such as languages that target GPUs
-(OpenCL, CUDA). The developer should be given tools to determine what kind of
-language is currently used in order to be able to use those additional features.
+important to express computing kernels in different languages, based on the
+availability and their respective merits on the target platform. Some languages
+have additional features, such as languages that target GPUs (OpenCL, CUDA), or
+languages that support threading paradigms (OpenMP). The developer should be
+given tools to determine what kind of language is currently used in order to be
+able to use those additional features.
 
@@
 compilation related parameters. This behavior is contained in the \emph{CKernel}
-class of BOAST. When instantiating this class a BOAST \emph{Procedure}
+class of BOAST. When instantiating this class, a BOAST \emph{Procedure}
 representing the entry point of the kernel is specified.
@@
 The next logical step is to benchmark the built kernel. BOAST offers a simple
-way to run a kernel that was successfully built while staying within BOAST.
-A built BOAST \emph{CKernel} exposes a run method that accepts arguments corresponding
-to the BOAST \emph{Procedure} used to instantiate the kernel. Arrays must be
-instances of \emph{NArray} which are numerical arrays that use C arrays
-underneath.
+way to run a kernel that was successfully built without the need to fork another
+process.  A built BOAST \emph{CKernel} exposes a run method that accepts
+arguments corresponding to the BOAST \emph{Procedure} used to instantiate the
+kernel. Arrays must be instances of \emph{NArray} which are numerical arrays
+that use C arrays underneath.
 
@@
 generating an input image and the corresponding reference output image, using a
-reference implementation, is easy. Unfortunately it is not always possible, some
-applications have complex data patterns that are difficult to
+reference implementation, is easy. Unfortunately it is not always possible,
+because some applications have complex data patterns that are difficult to
 synthesize without running the full application. Thus BOAST offers a way to load
@@
 
-The idea behind the tool is based on the assumption that the different GPU
-ports of the code should do the very same operations, at least at the logical
-level (the APIs will have \emph{implementation} differences, but they should
-offer nonetheless same functionalities). The usage of BOAST framework sustains
-this assumption, as both sets of kernels should be generated from the same
-source code.
+The idea behind this tool is based on the assumption that the different GPU
+ports of the code should perform the very same operations, at least at the
+logical level (the APIs will have \emph{implementation} differences, but they
+should offer nonetheless same functionalities). The usage of BOAST framework
+sustains this assumption, as both sets of kernels should be generated from the
+same source code.
 
@@
 
-Once two call traces have been generated by \code{GPUTrace}, the user
-can compare them with a graphical \code{diff} tool, and spot the
-different porting mistakes: two parameters reversed, an offset
-incorrectly applied, \etc{}
-
-By default, \code{GPUTrace} only prints a unique identifier (the
-creation index) for the memory buffers.  Additionally, \code{GPUTrace}
-supports several modifier flags. One flag can be activated
-to append the first bits of the buffer to the trace, printed in the
-right format, for visual inspection. Another flag can be set to
-dump the whole content of the buffer into a file, for a full
-inspection.
-
-This last option is also useful for generating \emph{replay buffers}
-for BOAST kernel execution. With a set of filters based on kernel
-names and execution counters, developers can precisely select which
-kernel execution parameters should be dumped, for further reuse as
-real-case benchmarks and non-regression testing.
+Once two call traces have been generated by \code{GPUTrace}, the user can
+compare them with a graphical \code{diff} tool, and spot the different porting
+mistakes: two parameters reversed, an offset incorrectly applied, \etc{}
+
+By default, \code{GPUTrace} only prints a unique identifier (the creation index)
+for the memory buffers.  Additionally, \code{GPUTrace} supports several modifier
+flags. One flag can be activated to append the first bits of the buffer to the
+trace, printed in the right format, for visual inspection. Another flag can be
+set to dump the whole content of the buffer into a file, for a full inspection.
+
+This last option is also useful to generate \emph{replay buffers} for BOAST
+kernel execution. With a set of filters based on kernel names and execution
+counters, developers can precisely select which kernel execution parameters
+should be dumped, for further reuse as real-case benchmarks and non-regression
+testing.
 
@@
 
-The list of already identified optimizations are:
-\begin{itemize}
-\item vectorization,
-\item intermediary data type,
-\item number of pixels processed,
-\item and synthesizing loads.
-\end{itemize}
-To create our generic implementation we decided to work at the component level
-rather than the pixel level. This approach leads to more flexibility and
-genericity when applying optimizations. We also decided to study the impact of
-another parameter which is the number of components to process on the column
-direction. This leads to being able to process tiles instead of only rows.
+The list of already identified optimizations are vectorization, intermediary
+data type, number of pixels processed, and synthesizing loads.  To create our
+generic implementation we decided to work at the component level rather than the
+pixel level. This approach leads to more flexibility and genericity when
+applying optimizations. We also decided to study the impact of another parameter
+which is the number of components to process on the column direction. This leads
+to being able to process tiles instead of only rows.
 
@@
 % \hline
-  Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\[4pt]
+  Image Size  & Naive (s) & Best (s) & Accel. & BOAST (s) & Accel. \\[4pt]
   \hline&&&&\\[-8pt]
@@
 The optimization parameters selected are not the same for those architectures.
-Indeed, the Intel CPU favors kernels that have the parameters:
-\textit{x\_component\_number} = 16,
-\textit{vector\_length} = 8,
+Indeed, the Intel CPU favors kernels that have the parameters,
+\textit{x\_compo-nent\_number} = 16, \textit{vector\_length} = 8,
 \textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false. Once again
@@
 % \hline
-  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\[4pt]
+  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NV & Ratio \\[4pt]
   \hline&&&&\\[-8pt]
@@
 
-In this Subsection we have shown the interest of BOAST in optimizing a well-known
-algorithm across different architectures. Chosen optimization combinations
-are highly dependent on the targeted architecture. In the next Subsections we
-will show that our methodology also applies to real applications.
+In this Subsection we have shown the interest of BOAST in optimizing a
+well-known algorithm across different architectures. Chosen optimization
+combinations are highly dependent on the target architecture. But BOAST
+simplifies and speeds up the identification of the correct combination at the
+cost of metaprogramming. In the next Subsections we will show that our
+methodology also applies to real applications.
 
@@
 them ideal for a precise and optimized DFT approach. In particular, their
-systematicity allows to provide a reliable basis set for high-precision
-results, whereas their locality (both in real and reciprocal space) is highly
-desired to improve the efficiency and the flexibility of the processing. Indeed,
-a localized basis set allows to optimize the number of degrees of freedom for a
+systematicity allows to provide a reliable basis set for high-precision results,
+whereas their locality (both in real and reciprocal space) is highly desired to
+improve the efficiency and the flexibility of the processing. Indeed, a
+localized basis set allows to optimize the number of degrees of freedom for a
 required accuracy~\cite{genovese2008daubechies}, which is highly desirable given
@@
 
-Despite the application is mainly written in Fortran, it currently includes
-360 kLOC and 70kLOC of Fortran and C languages, respectively, accounting for more
-than 50\% of the code base.  It is a parallel application
-based on the standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also
-supports CUDA~\cite{cuda} and OpenCL~\cite{opencl}.  In the recent years this
-code has been used for many scientific applications, and its development and
-user consortium is continuously growing.  Massively parallel computations are
+Despite that the application is mainly written in Fortran, it currently includes
+360 kLOC and 70 kLOC of Fortran and C languages, respectively, accounting for
+more than 50\% of the code base.  It is a parallel application based on the
+standards MPI~\cite{mpi} and OpenMP~\cite{openmp}.  It also supports
+CUDA~\cite{cuda} and OpenCL~\cite{opencl}.  In the recent years this code has
+been used for many scientific applications, and its development and user
+consortium is continuously growing.  Massively parallel computations are
 routinely executed with the BigDFT code, either in homogeneous or hybrid
@@
 In the Kohn-Sham (KS) formulation of DFT, the electrons are associated to
-wavefunctions (orbitals), which are represented arrays of floating point
+wavefunctions (orbitals), which are represented by arrays of floating point
 numbers. In wavelets formalism, the operators are written via convolutions with
 short, separable filters.  The detailed description of how these operations are
-defined is beyond the scope of this report and can be found in the BigDFT
-reference paper~\cite{genovese2008daubechies}.
+defined is beyond the scope of this paper and can be found in the BigDFT
+reference paper~\cite{genovese2008daubechies}. Convolutions are basic operations
+of lots of scientific application codes, like for example finite differences
+approaches, which are universally used in computational physics.
 
@@
 For these reasons, a convolution library has been engineered with the help of
-BOAST.  The detailed API of the library will be described elsewhere.  The
+BOAST. The detailed API of the library is beyond the scope of this article. The
 spirit is similar to BLAS-LAPACK API, where low-level operations are scheduled
@@
 
-Each of the building blocks of these convolution libraries is optimally tuned
-by BOAST by choosing the sources providing the optimal kernel for the chosen
+Each of the building blocks of these convolution libraries is optimally tuned by
+BOAST by choosing the sources providing the optimal kernel for the chosen
 computing platform.  The sources of these kernels are then collected and
-compiled to meet the API of the library.
-
-A library written in this way might have an impact going largely beyond the
-community of BigDFT developers.  Indeed, convolutions are basic operations of
-lots of scientific application codes, like for example finite differences
-approaches, which are universally used in computational physics.
+compiled to meet the API of the library. A library written in this way might
+have an impact going largely beyond the community of BigDFT developers.
 
@@
 \begin{center}
-\includegraphics[scale=0.7]{Res_synthesis}
+\includegraphics[width=\linewidth]{Res_synthesis}
 \end{center}
@@
 
-In this report we presented the BOAST infrastructure as well as its
+In this paper we presented the BOAST infrastructure as well as its
 application to three use cases from the Mont-Blanc project. Results are

### 2015-07-22 Tutorial/Tutorial-BOAST-2.tex
Small improvements.

@@
 007:0> BOAST::lang = BOAST::C
-008:0> BOAST::decl p
+008:0> BOAST::opn p
 void test_proc(int32_t a, double b){
@@
 require 'narray'
-input  = NArray.float(1024).random
-output = NArray.float(1024)
 n = BOAST::Int(  "n", :dir => :in )
@@
 k.procedure = p
+
+input  = NArray.float(1024).random
+output = NArray.float(1024)
 k.run(input.length, input, output)
-(output - input).each {
-  |val| raise "Error!" if (val-2).abs > 1e-15 }
+(output - input).each { |val|
+  raise "Error!" if (val-2).abs > 1e-15
+}
 stats = k.run(input.length, input, output)
@@
 n = 1024*1024
-a = NArray.float(n).random
-b = NArray.float(n).random
-c = NArray.float(n)
-c_ref = NArray.float(n)
+a = NArray.sfloat(n).random
+b = NArray.sfloat(n).random
+c = NArray.sfloat(n)
+
 epsilon = 10e-15
-set_lang( FORTRAN )
+
+c_ref = a + b
+
+[:FORTRAN, :C, :CL, :CUDA].each { |l|
+  set_lang( BOAST.const_get(l)  )
+  puts "#{l}:"
   k = vector_add
-k.run(n,a,b,c_ref)
-[C, CL, CUDA].each { |lang|
-  set_lang( lang )
-  c.random
-  k = vector_add
-  case lang
-  when CL
-    k.run(n,a,b,c, :global_work_size => [rndup(n,32), 1,1], :local_work_size => [32,1,1] )
-  when CUDA
-    k.run(n,a,b,c, :block_number => [rndup(n,32)/32, 1,1], :block_size => [32,1,1] )
-  else
-    k.run(n,a,b,c)
-  end
-  (c_ref - c).abs.each { |diff|
+  puts k.print
+  c.random!
+  k.run(n, a, b, c, :global_work_size => [n,1,1], :local_work_size => [32,1,1])
+  diff = (c_ref - c).abs
+  diff.each { |elem|
     raise "Warning: residue too big: #{elem}" if elem > epsilon
@@
 }
+puts "Success!"
 \end{lstlisting}

### 2015-09-24 BigDFT/Wavelet_test.rb
Added option to dump the binaries.

@@ -46,6 +46,9 @@ $parser = OptionParser::new do |opts|
   opts.on("-s", "--step VAL", "Unroll step") { |step|
     $options[:step] = step.to_i
   }
+  opts.on("-d", "--dump", "Dump binary") { |step|
+    $options[:dump] = true
+  }
   opts.on("-u", "--unroll VAL", "Unroll bound") { |unroll|
     $options[:unroll] = unroll.to_i
   }
@@ -113,6 +116,10 @@ if $options[:output] then
     f.puts k1
     f.puts k2
   }
+end
+if $options[:dump] then
+  k1.dump_binary
+  k2.dump_binary
 end
 
 repeat = 5

### 2015-09-29 BigDFT/MFVec.rb
vector microkernel for n1

@@ -19,47 +19,75 @@ FILTER = [ "8.4334247333529341094733325815816e-7",
            "2.72734492911979659657715313017228e-6" ]
 
 set_lang( C )
-set_replace_constants(false)
+#set_replace_constants(false)
+set_boast_inspect( true )
 
-def mfvec(use_sub = false)
-  tile_x = 1
-  tile_y = 1
+def comp_tile_n1( n1 = 1, n2 = 1, vector_length = 1, real_size = 8 )
+  raise "Invalid vector size" if n1%vector_length != 0
+  arr = BOAST::ConstArray::new(FILTER, BOAST::Real)
+  filter = Real( "filt", :const => arr, :dim => Dim(0,FILTER.length-1) )
+  n1_stride_input = Int( "n1_stride_in", :dir => :in)
+  n1_stride_output = Int( "n1_stride_out", :dir => :in)
+  input =  Real( "input", :dir => :in, :align => vector_length * real_size, :dim => Dim() )
+  output = Real( "output", :dir => :out, :align => vector_length * real_size, :dim => Dim() )
+  p = Procedure( "comp_n1_#{n1}_#{n2}_#{vector_length}_#{real_size}", [n1_stride_input, n1_stride_output, input, output], [filter], :inline => true, :local => true ) {
+    f = Real( "f", :vector_length => vector_length )
+    decl f
+    tt = (n1/vector_length).times.collect { |i|
+      n2.times.collect { |j|
+        Real( "tt#{i}#{j}", :vector_length => vector_length )
+      }
+    }
+    xx = (n1/vector_length).times.collect { |i|
+      n2.times.collect { |j|
+        Real( "xx#{i}#{j}", :vector_length => vector_length )
+      }
+    }
+    decl *tt.flatten
+    decl *xx.flatten
+    (n1/vector_length).times.collect { |i|
+      n2.times.collect { |j|
+        pr tt[i][j].set( 0.0 )
+      }
+    }
+    FILTER.length.times { |k|
+      pr f.set(filter[k])
+      (n1/vector_length).times.collect { |i|
+        n2.times.collect { |j|
+          pr xx[i][j] === input[n1_stride_input*j+vector_length*i+k]
+          pr tt[i][j] === tt[i][j] + xx[i][j]*f
+        }
+      }
+    }
+    (n1/vector_length).times.collect { |i|
+      n2.times.collect { |j|
[…]

### 2015-09-29 CacheTest/KernelRead2.rb
Force alignment of result variable.

@@ -51,7 +51,9 @@ module BOAST
         }
       }
 
-      print resultV.dereference === sumV
+      rV = resultV.dereference
+      rV.align = length * elem_size
+      print rV === sumV
 
       print sum === 0
       (0...length).each { |indx|

### 2015-10-13 BigDFT/Poisson_test.rb
Put back some precision changes, because we still have issues in non periodic case

@@ -14,7 +14,7 @@ hgrids = NArray.float(3)
 hgrids[0] = 0.1
 hgrids[1] = 0.1
 hgrids[2] = 0.1
-geocode = 2
+geocode = 3
 epsilon = 10e-11
 $options = {}
 $parser = OptionParser::new do |opts|
@@ -179,6 +179,9 @@ k = nabla_u_and_square_ref
 stats = k.run(geocode, n01, n02, n03, u, du_ref,du2_ref , nord, hgrids, filter)
 puts "#{k.procedure.name}: #{stats[:duration]*1.0e3} #{3*59*n01*n02*n03 / (stats[:duration]*1.0e9)} GFlops"
 
+#annoyingly, too much computation to get a good precision here ...
+epsilon = 10e-6
+
 k3 = nabla_u_and_square(n01,n02,n03,kconv)
 
 k3.build(:openmp => true)
@@ -271,7 +274,7 @@ rhopol_boast = nil
 
 
 #now the fssnord3dmatnabla3var_lg variant
-
+epsilon = 10e-11
 
 u = NArray.float(n01,n02,n03).random
 du_boast = NArray.float(n01,n02,n03,3)
@@ -375,7 +378,7 @@ stats = k.run(geocode, n01, n02, n03, u, du2_ref , nord, hgrids, filter)
 puts "#{k.procedure.name}: #{stats[:duration]*1.0e3} #{3*59*n01*n02*n03 / (stats[:duration]*1.0e9)} GFlops"
 
 #annoyingly, too much computation to get a good precision here ...
-epsilon = 10e-9
+epsilon = 10e-6
 
 
 #cheat to use and square ?

### 2015-10-13 ARMclbench/Laplacian.rb
Correct a bug that is being revealed by BOAST v1.0.7.

@@ -202,7 +202,7 @@ def laplacian(options)
             start_indx = v_i * vector_length + x_i * 3
             end_indx = start_indx + vector_length - 1
             merge_expr = merge_vectors(tempload, ranges, start_indx, end_indx)
-            pr tempcnn[x_i][v_i][y_i] === Int( "(#{out_vec_type})(#{merge_expr.join(",")})", :size => temporary_size, :vector_length => vector_length)
+            pr tempcnn[x_i][v_i][y_i] === Int( "(#{out_vec_type})(#{merge_expr.join(",")})", :size => 1, :vector_length => vector_length)
           }
         }
       else

### 2015-10-22 Paper/BOAST.tex
Rewrote introduction with Thierry.

@@
 
-Porting and tuning HPC applications to new platforms is tedious and costly in
-terms of human resources. Nonetheless, it is a very important aspect of the
-Mont-Blanc project. Indeed, for the project, more than ten applications were
-selected to be ported and optimized for the prototype platform.
-
-Unfortunately, portability efforts are often lost when migrating to a new
+Porting and tuning HPC applications to new platforms is tedious and costly in term of human resources.
+Many developer teams are more concentrated on solving the physical problems faster and better with new algorithms
+and simulating new physics. 
+This is why optimization even if it is important is a task which is sometimes neglected or hidden for many developers.
+In the same way, the domain specific language (DSL) has exactly this goal,  the complete separation of the high-level code dedicated to simulate new physics
+and the low-level one which is more dependent to the HPC architecture. 
+Nevertheless, decoupling the high and low levels don't avoid the need of optimization which becomes more and more unavoidable 
+with the new architectures based on many cores and accelerators.
+
+Unfortunately, portability efforts of the low level i.e. computing kernels are often lost when migrating to a new
 architecture. Worse, code may lose maintainability because several versions of
 some functionalities coexist, usually with a lot of duplication.
-
 Thus productivity of porting and tuning efforts is low as a huge fraction of
@@
 
-Genericity of HPC codes is often limited. One of the reason is that producing
+The primary goal of the metaprogramming framework BOAST is to simplify the optimization of the computing kernels for HPC
+applications. The main idea is to use a high level language to define an API of computing kernels i.e. subroutines and
+their related arguments, but also describe the way to optimize it (unrolling, vectorization, parametrization, templating, tiling).
+BOAST will test automatically the many possibilities of optimizations and choose the best one after validation from a
+reference version.
+Then BOAST can generate pieces of Fortran, C, OpenCL or CUDA codes, test if for a specific target architecture.
+
+With the definition of an API, BOAST can hide the details of the implementation at high level but also is the way to
+factorize different kernels depending only on parameters as datatypes, the length of filters or the boundary conditions.
+The maintainability and genericity of HPC codes is often limited. One of the reason is that producing
 generic code in Fortran 90/95 is difficult as the language does not really fit
-for it. Sometimes, adding genericity degrades performance as optimization
+for it. Sometimes, adding genericity or functionalities can be quite costly and degrades performance as optimization
 opportunities that come from over-specification are lost.
-
-Functionality of HPC codes is tied to the previous point. Without genericity,
-adding new functionalities can be quite costly.
+BOAST can provide a solution to handle genericity and add new functionalities at low levels.
+For instance, the optimization of short convolutions depends strongly
+on the length of the filters. This is why HPC codes using finite differences handle only one order. BOAST allows
+handling any order and even more generalizing to another numerical scheme.
+
+In this paper, we illustrate the optimization problems related to scientific computing with the Laplace equation.
+Then we present our solution, BOAST, its architecture and its functionalities. We apply BOAST to three use cases, the
+Laplace equation, short multi-dimensional convolutions to perform differential operators and finally
+the wavelet transform which are used in a quantum chemistry code, BigDFT.
+Then BOAST is applied to a full application, SPECFEM3D dedicated to the seismic wave propagation simulator.
+80 GPU computing kernels have been ported improving considerably the maintanability and providing a single code base for
+OpenCL and CUDA kernel versions.
+Finally we compare BOAST with other related works.
+
+% Porting and tuning HPC applications to new platforms is tedious and costly in
+% terms of human resources. Nonetheless, it is a very important aspect of the
+% Mont-Blanc project. Indeed, for the project, more than ten applications were
+% selected to be ported and optimized for the prototype platform.
+% 
+% Unfortunately, portability efforts are often lost when migrating to a new
+% architecture. Worse, code may lose maintainability because several versions of
+% some functionalities coexist, usually with a lot of duplication.
+% 
+% Thus productivity of porting and tuning efforts is low as a huge fraction of
+% those developments are never used after the platform they were intended for is
+% decommissioned.
+% 
+% Genericity of HPC codes is often limited. One of the reason is that producing
+% generic code in Fortran 90/95 is difficult as the language does not really fit
+% for it. Sometimes, adding genericity degrades performance as optimization
+% opportunities that come from over-specification are lost.
+% 
+% Functionality of HPC codes is tied to the previous point. Without genericity,
+% adding new functionalities can be quite costly.
 
 \section{Background and Motivation}
+
+The motivation of the BOAST project comes from the need of the HPC communities which have full and large applications
+difficult to port and maintain for emerging architectures based on accelerators, longer vector lengths, or low power
+processors.
 
@@
 
-\section{Motivating Example: OpenCL Laplace}
+\section{Basic Example: OpenCL Laplace}
 \label{sec:laplace}

### 2015-10-22 BigDFT/MFVec.rb
WIP.

@@ -31,69 +31,135 @@ def comp_tile_n1( n1 = 1, n2 = 1, vector_length = 1, real_size = 8 )
   n1_stride_output = Int( "n1_stride_out", :dir => :in)
   input =  Real( "input", :dir => :in, :align => vector_length * real_size, :dim => Dim() )
   output = Real( "output", :dir => :out, :align => vector_length * real_size, :dim => Dim() )
-  p = Procedure( "comp_n1_#{n1}_#{n2}_#{vector_length}_#{real_size}", [n1_stride_input, n1_stride_output, input, output], [filter], :inline => true, :local => true ) {
-    f = Real( "f", :vector_length => vector_length )
-    decl f
-    tt = (n1/vector_length).times.collect { |i|
-      n2.times.collect { |j|
-        Real( "tt#{i}#{j}", :vector_length => vector_length )
-      }
-    }
-    xx = (n1/vector_length).times.collect { |i|
-      n2.times.collect { |j|
-        Real( "xx#{i}#{j}", :vector_length => vector_length )
-      }
-    }
-    decl *tt.flatten
-    decl *xx.flatten
-    (n1/vector_length).times.collect { |i|
-      n2.times.collect { |j|
-        pr tt[i][j].set( 0.0 )
-      }
-    }
-    FILTER.length.times { |k|
-      pr f.set(filter[k])
-      (n1/vector_length).times.collect { |i|
+  return (1..n1).collect { |n|
+    Procedure( "comp_n1_#{n}_#{n2}_#{vector_length}_#{real_size}", [n1_stride_input, n1_stride_output, input, output], [filter], :inline => true, :local => true ) {
+      f = Real( "f", :vector_length => vector_length )
+      decl f
+      tt = ((n+vector_length-1)/vector_length).times.collect { |i|
         n2.times.collect { |j|
-          pr xx[i][j] === input[n1_stride_input*j+vector_length*i+k]
-          pr tt[i][j] === FMA(xx[i][j], f, tt[i][j])
+          Real( "tt#{i}#{j}", :vector_length => vector_length )
+        }
+      }
+      xx = ((n+vector_length-1)/vector_length).times.collect { |i|
+        n2.times.collect { |j|
+          Real( "xx#{i}#{j}", :vector_length => vector_length )
+        }
+      }
[…]

### 2015-12-22 Paper2/BOAST.tex
Added preliminary IEEE version.

--- 2015-10-22 Paper/BOAST.tex
@@
-\documentclass%[review]
-              {elsarticle}
+\documentclass{IEEEtran}
 \usepackage[utf8]{inputenc}
+\usepackage{graphicx}
 \usepackage{hyperref}
@@
 
-\journal{Journal of Parallel and Distributed Computing}
+\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
+{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
+%\journal{Journal of Parallel and Distributed Computing}
 
@@
 %% `Elsevier LaTeX' style
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{IEEEtran}
 %%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{frontmatter}
+%\begin{frontmatter}
 
@@
 %% or include affiliations in footnotes:
-\author[mymainaddress]{Brice Videau}
-\ead{brice.videau@imag.fr}
-\author{Kevin Pouget}
-\author{Luigi Genovese}
-\author{Thierry Deutsch}
-\author{Dimitri Komatitsch}
-\author{Jean-François Méhaut}
-
-\address[mymainaddress]{LIG/CNRS}
+\author{Brice Videau, Kevin Pouget, Luigi Genovese, Thierry Deutsch, Dimitri Komatitsch, Frédéric Desprez, Jean-François Méhaut}
+
+%\address[mymainaddress]{LIG/CNRS}
+
+%\begin{keyword}
+%Code Generation \sep Portability \sep Genericity \sep Productivity and Software
+%Design \sep High Performance Computing \sep Autotuning \sep Non Regression
+%Testing
+%\end{keyword}
+
+%\end{frontmatter}
+\maketitle
 
@@
 \end{abstract}
-
-\begin{keyword}
-Code Generation \sep Portability \sep Genericity \sep Productivity and Software
-Design \sep High Performance Computing \sep Autotuning \sep Non Regression
-Testing
-\end{keyword}
-
-\end{frontmatter}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{BOAST_Workflow.pdf}
+\includegraphics[width=\columnwidth]{BOAST_Workflow.pdf}
 \caption{Structure and Workflow of the BOAST framework.}
@@
 
-\subparagraph{Porting kernels to BOAST}
+\paragraph{Porting kernels to BOAST}
 
@@
 
-\subparagraph{Porting run-time to \OCL} The second part of the port
+\paragraph{Porting run-time to \OCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
@@
 
-\subparagraph{Debugging \OCL Execution: \code{GPUTrace}}
+\paragraph{Debugging \OCL Execution: \code{GPUTrace}}
 
@@
 
-\subparagraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
+\paragraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
 merged in \Specfem's development tree and under test and extension by

### 2015-12-22 Paper2/BOAST.tex
Figures compressées...

@@
 \begin{lstlisting}
-void laplace(const int width,
-             const int height,
+void laplace(const int width, const int height,
              const unsigned char src[height][width][3],
@@
       for (int c = 0; c < 3; c++) {
-        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
+        int tmp =
+   -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
   - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
   - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
-        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
+        dst[j][i][c] = (tmp<0 ? 0 : (tmp>255 ? 255 : tmp));
       }
@@
 \begin{lstlisting}
-kernel laplace(const int width,
-               const int height,
+kernel laplace(const int width, const int height,
                global const uchar *src,
@@
 
-  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
-  uchar16 v12_ = vload16( 0, src + 3*width*(j-1) + 3*5*i     );
-  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
-  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
-  uchar16 v22_ = vload16( 0, src + 3*width*(j  ) + 3*5*i     );
-  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
-  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
-  uchar16 v32_ = vload16( 0, src + 3*width*(j+1) + 3*5*i     );
-  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
+  uchar16 v11_ =vload16(0,src + 3*width*(j-1) + 3*5*i -3);
+  uchar16 v12_ =vload16(0,src + 3*width*(j-1) + 3*5*i   );
+  uchar16 v13_ =vload16(0,src + 3*width*(j-1) + 3*5*i +3);
+  uchar16 v21_ =vload16(0,src + 3*width*(j  ) + 3*5*i -3);
+  uchar16 v22_ =vload16(0,src + 3*width*(j  ) + 3*5*i   );
+  uchar16 v23_ =vload16(0,src + 3*width*(j  ) + 3*5*i +3);
+  uchar16 v31_ =vload16(0,src + 3*width*(j+1) + 3*5*i -3);
+  uchar16 v32_ =vload16(0,src + 3*width*(j+1) + 3*5*i   );
+  uchar16 v33_ =vload16(0,src + 3*width*(j+1) + 3*5*i +3);
 
@@
 \begin{lstlisting}
-  uchar16 v11_ = vload16( 0, src + 3*width*(j-1) + 3*5*i - 3 );
-  uchar16 v13_ = vload16( 0, src + 3*width*(j-1) + 3*5*i + 3 );
-  uchar16 v12_ = uchar16( v11_.s3456789a, v13_.s56789abc );
-  uchar16 v21_ = vload16( 0, src + 3*width*(j  ) + 3*5*i - 3 );
-  uchar16 v23_ = vload16( 0, src + 3*width*(j  ) + 3*5*i + 3 );
-  uchar16 v22_ = uchar16( v21_.s3456789a, v23_.s56789abc );
-  uchar16 v31_ = vload16( 0, src + 3*width*(j+1) + 3*5*i - 3 );
-  uchar16 v33_ = vload16( 0, src + 3*width*(j+1) + 3*5*i + 3 );
-  uchar16 v32_ = uchar16( v31_.s3456789a, v33_.s56789abc );
+  uchar16 v11_ =vload16(0, src + 3*width*(j-1) + 3*5*i -3);
+  uchar16 v13_ =vload16(0, src + 3*width*(j-1) + 3*5*i +3);
+  uchar16 v12_ =uchar16(v11_.s3456789a, v13_.s56789abc);
+  uchar16 v21_ =vload16(0, src + 3*width*(j  ) + 3*5*i -3);
+  uchar16 v23_ =vload16(0, src + 3*width*(j  ) + 3*5*i +3);
+  uchar16 v22_ =uchar16(v21_.s3456789a, v23_.s56789abc);
+  uchar16 v31_ =vload16(0, src + 3*width*(j+1) + 3*5*i -3);
+  uchar16 v33_ =vload16(0, src + 3*width*(j+1) + 3*5*i +3);
+  uchar16 v32_ =uchar16(v31_.s3456789a, v33_.s56789abc);
 \end{lstlisting}
@@
         label={lst:BOAST_algebra},
-        numbers=left }
+        numbers=right }
 \begin{lstlisting}
@@
 
-\begin{minipage}[b]{0.50\linewidth}
+\begin{minipage}[b]{0.5\linewidth}
 \centering
@@
 i = 5
-k = (i + 5) * (2)
+k = (i+5)*(2)
 l(k) = 1.0_wp
-l(k + 1) = sin(i + 5)
+l(k+1) = sin(i+5)
 \end{lstlisting}
 \end{minipage}
-\hspace{0.04\linewidth}
-\begin{minipage}[b]{0.44\linewidth}
+\hspace{0.08\linewidth}
+\begin{minipage}[b]{0.40\linewidth}
 \centering
@@
 i = 5;
-k = (i + 5) * (2);
-l[k - (-5)] = 1.0;
-l[k + 1 - (-5)] = sin(i + 5);
+k = (i+5)*(2);
+l[k-(-5)] = 1.0;
+l[k+1-(-5)] = sin(i+5);
 \end{lstlisting}
@@
 \begin{lstlisting}
-j = 0
-do i = 0, 100, 1
-  if (modulo(i, 7) == 0) then
-    j = j + 1
+j=0
+do i=0, 100, 1
+  if(modulo(i, 7)==0) then
+    j=j+1
   end if
@@
 \end{minipage}
-\hspace{0.04\linewidth}
-\begin{minipage}[b]{0.47\linewidth}
+\hspace{0.08\linewidth}
+\begin{minipage}[b]{0.40\linewidth}
 \centering
@@
 \begin{lstlisting}
-j = 0;
-for (i = 0; i <= 100; i += 1) {
-  if (modulo(i, 7) == 0) {
-    j = j + 1;
+j=0;
+for(i=0; i<=100; i+=1){
+  if(modulo(i, 7)==0){
+    j=j+1;
   }
@@
         label={lst:BOAST_procedure},
-        numbers=left }
+        numbers=right }
 \begin{lstlisting}
-n = Int("n", :dir => :in)
-ndat = Int("ndat", :dir => :in)
-x = Real("x", :dir => :in, :dim => [ Dim(0, n-1), Dim(ndat) ])
-y = Real("y", :dir => :out, :dim => [ Dim(ndat), Dim(0, n-1) ])
+n = Int("n", :dir=>:in)
+ndat = Int("ndat", :dir=>:in)
+x = Real("x", :dir=>:in, :dim=>[Dim(0, n-1), Dim(ndat)])
+y = Real("y", :dir=>:out, :dim=>[Dim(ndat), Dim(0, n-1)])
 p = Procedure("magicfilter", [n, ndat, x, y])
@@
   integer(kind=4), intent(in) :: ndat
-  real(kind=8), intent(in), dimension(0:n - (1), ndat) :: x
-  real(kind=8), intent(out), dimension(ndat, 0:n - (1)) :: y
+  real(kind=8), intent(in), dimension(0:n-(1), ndat) :: x
+  real(kind=8), intent(out), dimension(ndat, 0:n-(1)) :: y
 END SUBROUTINE magicfilter
@@
 \begin{table}
+\caption{Best performance of ARM Laplace kernel.}
+\label{tbl:ARM}
 \centering
@@
 \end{tabular}
-\caption{Best performance of ARM Laplace kernel.}
-\label{tbl:ARM}
 \end{table}
@@
 \begin{table}
+\caption{Best performance of Laplace Kernel on several architectures.}
+\label{tbl:other}
 \centering
@@
 % \hline
-  Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NV & Ratio \\[4pt]
+  Image Size & ARM & Intel & Ratio & NV & Ratio \\[4pt]
   \hline&&&&\\[-8pt]
@@
 \end{tabular}
-\caption{Best performance of Laplace Kernel on several architectures.}
-\label{tbl:other}
 \end{table}

### 2015-12-22 Paper2/BOAST.tex
Fixed listing numbering.

@@
 
-\lstset{	language=C, 
+\lstset{language=C, 
   basicstyle=\scriptsize, 
@@
   caption={MagicFilter.}, 
-		label={lst:conv_example}}
+  label={lst:conv_example},
+  numbers=right}
 

### 2016-01-12 Paper/BOAST.tex
WIP: integrating Fred remarks.

--- 2015-07-06 Paper/BOAST.tex
@@
 %% or include affiliations in footnotes:
-\author[mymainaddress]{Brice Videau}
+\author[lig,cnrs]{Brice Videau}
 \ead{brice.videau@imag.fr}
-\author{Kevin Pouget}
-\author{Luigi Genovese}
-\author{Thierry Deutsch}
-\author{Dimitri Komatitsch}
-\author{Jean-François Méhaut}
-
-\address[mymainaddress]{LIG/CNRS}
-
+\author[lig,ujf]{Kevin Pouget}
+\author[cea]{Luigi Genovese}
+\author[cea]{Thierry Deutsch}
+\author[cnrs]{Dimitri Komatitsch}
+\author[lig,inria]{Frédéric Desprez}
+\author[lig,ujf]{Jean-François Méhaut}
+
+\address[lig]{Laboratoire d'Informatique de Grenoble}
+\address[cnrs]{Centre National de la Recherche Scientifique}
+\address[cea]{CEA/INAC/SP2M/L\_Sim}
+\address[ujf]{Université Joseph Fourier}
+\address[inria]{Inria}
 \begin{abstract}
@@
 \begin{keyword}
-Code Generation \sep Portability \sep Genericity \sep Productivity and Software
-Design \sep High Performance Computing \sep Autotuning \sep Non Regression
-Testing
+Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non
+Regression Testing \sep Productivity and Software Design \sep High Performance
+Computing
 \end{keyword}
@@
 Porting and tuning HPC applications to new platforms is tedious and costly in
-terms of human resources. Nonetheless, it is a very important aspect of the
-Mont-Blanc project. Indeed, for the project, more than ten applications were
-selected to be ported and optimized for the prototype platform.
-
-Unfortunately, portability efforts are often lost when migrating to a new
-architecture. Worse, code may lose maintainability because several versions of
-some functionalities coexist, usually with a lot of duplication.
-
-Thus productivity of porting and tuning efforts is low as a huge fraction of
-those developments are never used after the platform they were intended for is
-decommissioned.
-
-Genericity of HPC codes is often limited. One of the reason is that producing
-generic code in Fortran 90/95 is difficult as the language does not really fit
-for it. Sometimes, adding genericity degrades performance as optimization
-opportunities that come from over-specification are lost.
-
-Functionality of HPC codes is tied to the previous point. Without genericity,
-adding new functionalities can be quite costly.
+term of human resources.  Many developer teams are more concentrated on solving
+the physical problems faster and better with new algorithms and simulating new
+physics. This is why optimization, while important, is a task which is
+sometimes neglected or hidden for many developers.  In the same way, Domain
+Specific Language (DSLs) have exactly this goal, the complete separation of the
+high-level code dedicated to simulate new physics and the low-level one which
+is more dependent on the HPC architecture. Nevertheless, decoupling the high
+and low levels don't avoid the need of optimization which becomes more and more
+unavoidable with new architectures based on many cores and accelerators.
+
+Unfortunately, portability efforts of the low level i.e. computing kernels are
+often lost when migrating to a new architecture. Worse, code may lose
+maintainability because several versions of some functionalities coexist,
+usually with a lot of duplication.  Thus productivity of porting and tuning
+efforts is low as a huge fraction of those developments are never used after
+the platform they were intended for is decommissioned.
+
+The primary goal of the BOAST metaprogramming framework is to simplify the
+optimization of computing kernels for HPC applications. The main idea is to use
+a high level language to define an API of computing kernels i.e.  subroutines
+and their related arguments, but also describe the way to optimize it
+(unrolling, vectorization, parametrization, templating, tiling).  BOAST will
+test automatically the many possibilities of optimizations and choose the best
+one after validation from a reference version.  Then BOAST can generate pieces
+of Fortran, C, OpenCL, or CUDA codes, test it for a specific target
+architecture.
+
+With the definition of an API, BOAST can hide the implementation details at
+high level but also is a way to factorize different kernels depending only on
+parameters such as data types, length of filters or boundary conditions.  The
+maintainability and genericity of HPC codes is often limited. One of the reason
+is that producing generic code in Fortran 90/95 is difficult as the language
+does not really fit for it. Sometimes, adding genericity or functionalities can
+be quite costly and degrades performance as optimization opportunities that
+come from over-specification are lost. BOAST can provide a solution to handle
+genericity and add new functionalities at low levels.
+
+%For instance, the optimization of short convolutions depends strongly on the
+%length of the filters. This is why HPC codes using finite differences handle
+%only one order. BOAST allows handling any order and even more generalizing to
+%another numerical scheme.
+
+The paper is organized as follows. First we further motivate the need for
+metaprogramming frameworks.  Secondly we illustrate the optimization problems
+related to scientific computing with the Laplace equation. Thirdly we present
+our solution, called BOAST, its architecture and its functionalities. Fourthly
+we apply BOAST to three use cases: the Laplace equation motivating example;
+short multi-dimensional convolutions from the BigDFT quantum chemistry
+application; and the seismic wave propagation simulator SPECFEM3D where 80 GPU
+computing kernels have been ported improving, considerably the maintainability
+and providing a single code base for OpenCL and CUDA kernel versions. Finally
+we compare BOAST with other related works.
 
@@
 
+The motivation of the BOAST project comes from the need of the HPC communities
+which have full and large applications.  These applications are difficult to
+port and maintain for emerging architectures. Especially since new
+architectures are based on accelerators, longer vector units, or low power
+processors.
+
   \subsection{Evolution of HPC Architectures}
 
-Evolution of HPC Architectures is rapid and also diverse: in the last 5
-years no less than 6 architectures have been number one in the Top500:
+Evolution of HPC Architectures is rapid and also diverse: in the last 7 years
+no less than 6 architectures have been number one in the
+Top500~\cite{webtop500}:
 \begin{itemize}
-\item Intel Processor + Xeon Phi (Tianhe-2)
-\item AMD Processor + NVIDIA GPU (Titan)
-\item IBM BlueGene/Q (Sequoia)
-\item Fujitsu SPARC64 (K computer)
-\item Intel Processor + NVIDIA GPU (Tianhe-1)
-\item AMD Processor (Jaguar)
+\item Intel Processor + Xeon Phi (Tianhe-2, 2013)
+\item AMD Processor + NVIDIA GPU (Titan, 2012)
+\item IBM BlueGene/Q (Sequoia, 2012)
+\item Fujitsu SPARC64 (K computer, 2011)
+\item Intel Processor + NVIDIA GPU (Tianhe-1, 2010)
+\item AMD Processor (Jaguar, 2009)
 \end{itemize}
-Being able to efficiently use those architectures on such a small
-time-frame is challenging.
+Being able to efficiently use those architectures on such a small period is
+challenging for scientists and application developers.
 
 The race to exascale is not going to simplify the environment. All of the above
-architectures can be considered. Network architectures also can be very
-diverse.  For instance European FP7 project DEEP considers using Accelerators
-(XEON Phi) while the European FP7 project Mont-Blanc considers using low-power
-embedded processor with integrated GPU.
-
-Running existing applications on those new architectures is an open research
+machine architectures can be considered to build these supercomputers and other
+will appear soon. For instance European FP7 project DEEP considers using
+Accelerators (XEON Phi) while the European FP7 project Mont-Blanc considers
+using low-power embedded processor with integrated GPU.
+
+Running existing applications on these new architectures is an open research
 subject as well as an ongoing porting effort.  Thus, those projects have work
-packages dedicated to applications. Those work packages are dedicated to porting
-and optimizing Scientific applications on those new architectures. In the DEEP
-project six applications were selected for porting and optimizing, eleven were
-selected in the Mont-Blanc project.
+packages dedicated to applications. Those work packages are dedicated to
+porting and optimizing Scientific applications on those new architectures. In
+the DEEP project six applications were selected for porting and optimizing,
+eleven were selected in the Mont-Blanc project.
 
@@
 
-Scientific Computing Applications are usually developed by physicist, chemist or
-meteorologist. Those codes are usually written in Fortran for historical and
-performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
-with lots of functionalities. Nonetheless, they are usually based on computing
-kernels.  Computing kernels are resource intensive and well-defined parts of a
-program that usually work on precisely defined data. Those kernels represent the most
-time-consuming part of an HPC application, and consequently, they are the prime target for
-optimization.
+Scientific Computing Applications are usually developed by other science
+researchers such as physicists, chemists or meteorologists. Those codes are
+usually written in Fortran for historical and performance reasons. Codes can be
+quite huge (several thousands lines of code [LOC]) with lots of
+functionalities. Nonetheless, they are usually based on computing kernels.
+Computing kernels are resource intensive and well-defined parts of a program
+that usually work on precisely defined data. Those kernels represent the most
+time-consuming part of an HPC application, and consequently, they are the prime
+target for optimization.
 
@@
 language and architecture. C compilers are usually the first available while
-Fortran may sometimes arrive later.
+Fortran may sometimes come later.
 
@@
 The problematic here is to obtain computing kernels that present good
-performances on the architectures encountered by the application while still
-being portable after the optimization process took place. Indeed the
-application might encounter one of the many architectures that can be found in
-HPC. Investing manpower to optimize the application for a new architecture is
+performance on the architectures encountered by the application while staying
+portable after the optimization process took place. Indeed the application
+might encounter one of the many architectures that can be found in HPC.
+Investing manpower to optimize the application for a new architecture is
 reasonable, suffering hindrance from previous optimization work is not. Thus
@@
 
-What we want is computing kernels that are written:
-\begin{itemize}
-\item in a \emph{portable} manner,
-\item in a way that raises developer \emph{productivity},
-\item and, present good \emph{performance}.
-\end{itemize}
-
-\section{Motivating Example: OpenCL Laplace}
+What we want is computing kernels that are written in a \emph{portable} manner,
+in a way that raises developer \emph{productivity}, and they must present good
+\emph{performance}.
+
+\section{Basic Example: OpenCL Laplace}
 \label{sec:laplace}
@@
 During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
-on the Mali GPU was given. One of the case study was a Laplace filter on the
-GPU~\cite{opencl_arm_training}.
+on the Mali GPU was given. One of the case study was a Laplace filter~\cite{opencl_arm_training}.
 
@@
 \lstset{style=customc,
-        caption={Laplace C99 Filter},
+        caption={Laplace C99 Filter.},
         breaklines=true,
@@
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter},
+        caption={Laplace OpenCL Filter.},
         breaklines=true,
@@
 \paragraph{Vectorization}
-The first proposed optimization involves computing five pixels instead of one using the
-vectors available to the Mali architecture. The vectors have to reside in memory to
-ensure an efficient execution. Using vectors of 16 elements, 15 useful components
-are simultaneously loaded and can be used in computation. This optimization
-yields speedups between 1.5 and 6 depending on the image size.
+The first proposed optimization involves computing five pixels instead of one
+using the vectors available on the Mali architecture. The vectors have to be
+kept in memory to ensure an efficient execution. Using vectors of 16 elements,
+15 useful components are simultaneously loaded and can be used in computation.
+This optimization yields speedups between 1.5 and 6 depending on the image
+size.
 
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized},
+        caption={Laplace OpenCL Filter Vectorized.},
         breaklines=true,
@@
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads},
+        caption={Laplace OpenCL Filter Vectorized with Synthesized Loads.},
         breaklines=true,
@@
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized Using \textit{short}},
+        caption={Laplace OpenCL Filter Vectorized Using \textit{short}.},
         breaklines=true,
@@
 Nonetheless, the process is frustrating. The optimizations are never evaluated
-independently from one another. Some were arbitrarily configured (the number of
-pixels chosen for instance). Testing the different combinations of those
-optimizations and the different parameters would be very costly in developer
-time and, moreover, like all repetitive and tedious tasks, error prone. Thus
-every created kernel would have to be thoroughly tested to ensure no error was
-done.
+independently from one another. Some were arbitrarily configured (like the
+number of pixels chosen for instance). Testing the different combinations of
+those optimizations and the different parameters would be very costly in
+developer time and, moreover, like all repetitive and tedious tasks, error
+prone. Thus every created kernel would have to be thoroughly tested to ensure
+no error was done.
 
@@
 are \emph{decl}, \emph{pr}, \emph{opn} and \emph{close}. As the language is an
-EDSL, these four keywords are methods in the BOAST namespace. Sample usage of
-these keywords will be found in the next Figures.
+Embedded Domain Specific Language (EDSL), these four keywords are methods in
+the BOAST namespace. Sample usage of these keywords will be found in the next
+Figures.
 %Usage of these keywords can be found on Figure~\ref{fig:keywords}.
@@
 \lstset{style=BOAST,
-        caption={BOAST code}, 
+        caption={BOAST code.}, 
         label={lst:BOAST_algebra},
@@
 \lstset{style=BOAST,
-        caption={BOAST code}, 
+        caption={BOAST code.}, 
         label={lst:BOAST_control},
@@
 \lstset{style=BOAST,
-        caption={BOAST code}, 
+        caption={BOAST code.}, 
         label={lst:BOAST_procedure},

### 2016-01-13 Paper/BOAST.tex
Finished implementing Fred remarks.

@@
 
-\newcommand{\Specfem}{\productname{Specfem 3D}\xspace}
-\newcommand{\Cuda}{\productname{Cuda}\xspace}
-\newcommand{\OCL}{\productname{OpenCL}\xspace}
 \newcommand{\code}[1]{\texttt{#1}}
@@
 
-\paragraph{Temporary Variables Size} Using \textit{int} to store intermediary
+\paragraph{Temporary Variables Size} Using \emph{int} to store intermediary
 results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
@@
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized Using \textit{short}.},
+        caption={Laplace OpenCL Filter Vectorized Using \emph{short}.},
         breaklines=true,
@@
 In this section we will present the different aspects of BOAST run-time that
-allow this automation. Those aspects include:
-  \begin{itemize}
-  \item multi target language generation ( performance, portability )
-  \item compilation ( productivity, performance )
-  \item execution ( productivity, performance )
-  \item tracer, dumper and replay for non regression tests ( productivity )
-  \end{itemize}
+allow this automation. Those aspects are multi-target language generation
+(\emph{performance}, \emph{portability}), kernel compilation
+(\emph{productivity}, \emph{performance}), kernel execution
+(\emph{productivity}, \emph{performance}) and last, kernel tracer, dumper and
+replay for non regression tests (\emph{productivity}).
 
@@
 trace format. Listing~\ref{lst:GPUTrace} presents a sample output
-generated during \Specfem tracing.
+generated during SPECFEM3D tracing.
 
@@
 In this section we will present the benefits of using BOAST on the Laplace
-motivating example as well as two scientific applications that use BOAST. The
-first one, BigDFT~\cite{genovese2008daubechies}, uses BOAST in order to develop new
+motivating example as well as on two scientific applications. The first one,
+BigDFT~\cite{genovese2008daubechies}, uses BOAST in order to develop new
 functionalities with performance portability in mind. The second one,
@@
 \begin{itemize}
-\item \textit{x\_component\_number}: a positive integer
-\item \textit{y\_component\_number}: a positive integer
-\item \textit{vector\_length}: 1, 2, 4, 8 or 16
-\item \textit{temporary\_size}: 2 or 4
-\item \textit{synthesize\_loads}: \textit{true} or \textit{false}
-\item \textit{vector\_recompute}: \textit{true} or \textit{false}
+\item \emph{x\_component\_number}: a positive integer
+\item \emph{y\_component\_number}: a positive integer
+\item \emph{vector\_length}: 1, 2, 4, 8 or 16
+\item \emph{temporary\_size}: 2 or 4
+\item \emph{synthesize\_loads}: \emph{true} or \emph{false}
+\item \emph{vector\_recompute}: \emph{true} or \emph{false}
 \end{itemize}
-The last parameter is used when \textit{x\_component\_number} is not divisible by
-\textit{vector\_length}. Two solutions are possible then, divide the remainder
-of the division in vectors of smaller sizes (\textit{vector\_recompute} =
-\textit{false}) or load more data and compute useless values in vectors of the
-specified size (\textit{vector\_recompute} = \textit{true}).
+The last parameter is used when \emph{x\_component\_number} is not divisible by
+\emph{vector\_length}. Two solutions are possible then, divide the remainder
+of the division in vectors of smaller sizes (\emph{vector\_recompute} =
+\emph{false}) or load more data and compute useless values in vectors of the
+specified size (\emph{vector\_recompute} = \emph{true}).
 This last option mimics the behavior of the ARM implementation, although when
@@
 are disappointing: the same kernel configuration is the best for all image
-sizes. This kernel uses \textit{x\_component\_number} = 16,
-\textit{y\_component\_number} = 1, \textit{vector\_length} = 16,
-\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false.
-\textit{vector\_recompute} because \textit{x\_component\_number}
-== \textit{vector\_length}. Those results show that, when working on full
+sizes. This kernel uses \emph{x\_component\_number} = 16,
+\emph{y\_component\_number} = 1, \emph{vector\_length} = 16,
+\emph{temporary\_size} = 2 and \emph{synthesize\_loads} = false.
+\emph{vector\_recompute} because \emph{x\_component\_number}
+== \emph{vector\_length}. Those results show that, when working on full
 vectors, synthesizing the loads is harmful to performance and the programmer is
@@
 Indeed, the Intel CPU favors kernels that have the parameters,
-\textit{x\_compo-nent\_number} = 16, \textit{vector\_length} = 8,
-\textit{temporary\_size} = 2 and \textit{synthesize\_loads} = false. Once again
-\textit{vector\_recompute} does not apply. \textit{y\_component\_number} varies
+\emph{x\_compo-nent\_number} = 16, \emph{vector\_length} = 8,
+\emph{temporary\_size} = 2 and \emph{synthesize\_loads} = false. Once again
+\emph{vector\_recompute} does not apply. \emph{y\_component\_number} varies
 from 4 to 2 when image size increase, thus decreasing task granularity as the
@@
 happen could be interesting. The NVIDIA GPU favors processing square tiles:
-\textit{x\_component\_number} = 4, \textit{y\_component\_number} = 4,
-\textit{vector\_length} = 4, \textit{temporary\_size} = 2 and
-\textit{synthesize\_loads} = false. Once more, \textit{vector\_recompute} has no
+\emph{x\_component\_number} = 4, \emph{y\_component\_number} = 4,
+\emph{vector\_length} = 4, \emph{temporary\_size} = 2 and
+\emph{synthesize\_loads} = false. Once more, \emph{vector\_recompute} has no
 meaning.
@@
 have been added, making the BigDFT more and more powerful in terms of
-scientific applications.  However, the cost of the maintenance and of
-development of the convolutions is always a delicate point to be considered
-while including a new functionality.  The convolution patterns are usually
-rather similar, leading to code duplication and difficulties in code
-maintainability.
-
-It appears therefore very interesting to benefit from an automatic tool to drive
+scientific applications.  However, the cost of maintenance and development of
+the convolutions is always a delicate point to be considered while including a
+new functionality.  The convolution patterns are usually rather similar,
+leading to code duplication and difficulties in code maintainability.
+
+Therefore it appears very interesting to benefit from an automatic tool to drive
 the implementation and the generation of new convolutions.  This would lead to
@@
 The interest in having a robust and optimally tuned library in this scientific
-field is therefore evident. Techniques are under investigation to provide also
-fine-tuned binaries to the end-user rather than the code sources, such that
-more aggressive inter procedural optimization can be performed. Indeed, BOAST
-finds the optimal source code for a given kernel and compiler configuration but
-one could imagine using binary optimizer to the compiled binaries. Those
-optimizers could be coupled to BOAST and the output of the process would be the
-final binary rather than the source code. This experience therefore would be a
-first step toward the release of a tunable optimised convolution library
-oriented to computational physics communities.
-
-    \subsubsection{Performance report}
+field is therefore evident. Techniques are under investigation to also provide
+end users with fine-tuned binaries rather than the source codes, such that more
+aggressive inter procedural optimization can be performed. Indeed, BOAST finds
+the optimal source code for a given kernel and compiler configuration but one
+could imagine using binary optimizer to the compiled binaries. Those optimizers
+could be coupled to BOAST and the output of the process would be the final
+binary rather than the source code. This experience therefore would be a first
+step toward the release of a tunable optimized convolution library oriented to
+computational physics communities.
+
+    \subsubsection{Performance Report}
 
@@
 \end{center}
-\caption{Impact of Unrolling, Language and OpenMP on a Wavelet Transform}
+\caption{Impact of Unrolling, Language Used and OpenMP on a Wavelet Transform Code.}
 \label{fig:synthesis}
@@
 
-In this last subsection, we study a free seismic wave propagation
-simulator, \Specfem, in its \productname{Globe}
-flavor\footnote{Specfem3d Globe --- CIG
-  \url{http://www.geodynamics.org/cig/software/specfem3d-globe}}. \Specfem
-simulates seismic wave propagation at the local or regional scale
-based upon spectral-element method (SEM), with very good accuracy and
-convergence properties. It is a reference application for super
-computer benchmarking, thanks to its good scaling capabilities.
+In this last subsection, we study a free seismic wave propagation simulator,
+SPECFEM3D, in its GLOBE flavor\footnote{Specfem3d Globe --- CIG
+\url{http://www.geodynamics.org/cig/software/specfem3d-globe}}. SPECFEM3D
+simulates seismic wave propagation at the local or regional scale based upon
+spectral-element method (SEM), with very good accuracy and convergence
+properties. It is a reference application for supercomputer benchmarking,
+thanks to its good scaling capabilities.
 
@@
 When we started to work on the project (version v2.1 of July 2013) it
-supported graphics card GPU acceleration through NVidia \Cuda. This
+supported graphics card GPU acceleration through NVidia CUDA. This
 GPU support came in addition to the MPI support implemented to enable
-multi-CPU parallel computing. Most of \Specfem code-base is written in
-\productname{Fortran2003}, only the GPU related parts are written in
+multi-CPU parallel computing. Most of SPECFEM3D code-base is written in
+Fortran 2003 and only the GPU related parts are written in
 C.  The split between CPU and GPU code was done at a rather fine
@@
 
-As part of the Mont-Blanc project, we had to port \Specfem to \OCL, so
+As part of the Mont-Blanc project, we had to port SPECFEM3D to OpenCL, so
 that it could be used to benchmark Mont-Blanc HPC platforms.
@@
 
-Nvidia \Cuda and \OCL are based on the same programming model: a
+Nvidia CUDA and OpenCL are based on the same programming model: a
 massively parallel accelerator running in disjoint, non addressable
@@
 out most of the porting task with only a limited knowledge about
-\Specfem internal physics.
-
-This lack of \Specfem internal knowledge led us to be particularly
+SPECFEM3D internal physics.
+
+This lack of SPECFEM3D internal knowledge led us to be particularly
 careful to the path we undertook for the port, as we would have
@@
 Hence, our first milestone in the port was the translation of
-\Specfem's \Cuda kernels into BOAST EDSL. This way, we could ask BOAST
+SPECFEM3D's CUDA kernels into BOAST EDSL. This way, we could ask BOAST
 framework to generate a CUDA version of the kernels, plug them back
-into \Specfem and get (after fixing compile-time errors---prototypes
-and naming mistakes mainly) a first set of \Specfem seismograms.
+into SPECFEM3D and get (after fixing compile-time errors---prototypes
+and naming mistakes mainly) a first set of SPECFEM3D seismograms.
 
@@
 of shell scripts and BOAST framework ability to store and provide the
-kernels' original source code, we built a set of \Specfem binaries
+kernels' original source code, we built a set of SPECFEM3D binaries
 including only \emph{one} BOAST-generated kernel, with all others
@@
 
-\subparagraph{Porting run-time to \OCL} The second part of the port
+\subparagraph{Porting run-time to OpenCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
-\Cuda API to \OCL API. Most of the functions of the interfaces are
+CUDA API to OpenCL API. Most of the functions of the interfaces are
 very similar, with only naming-convention and data-structure
@@
 useful. To give an idea of the cost of a \emph{manual} rewriting, we
-can count (in \# of \OCL API function calls): 70 kernel ``function
+can count (in \# of OpenCL API function calls): 70 kernel ``function
 calls'', 790 arguments to set, 230 memory transfers, 160 buffer
@@
 Once the transformations were applied,
-compilation errors fixed, and \OCL unsuccessful function calls solved,
+compilation errors fixed, and OpenCL unsuccessful function calls solved,
 the application managed to complete its execution and generate a first
@@
 
-As we had already validated BOAST-generated kernels (and trusted \Cuda and \OCL
+As we had already validated BOAST-generated kernels (and trusted CUDA and OpenCL
 versions to be semantically identical), we knew that the bugs were now in the
-usage of the run-time, and we had to find a way to understand where \Specfem's
-\Cuda version of the code diverged from its \OCL counterpart. To help us in
+usage of the run-time, and we had to find a way to understand where SPECFEM3D's
+CUDA version of the code diverged from its OpenCL counterpart. To help us in
 that purpose, we had a strong assumption: both versions of the code were
 supposed to perform exactly the same operations, with the same ``logical''
-parameters (the APIs have \emph{implementation} differences, for instance \OCL
+parameters (the APIs have \emph{implementation} differences, for instance OpenCL
 has two memory transfer functions, \code{clEnqueueReadBuffer},
-\code{clEnqueueWriteBuffer}, whereas \Cuda has only one, with a direction
+\code{clEnqueueWriteBuffer}, whereas CUDA has only one, with a direction
 parameter \code{cudaMemcpy(..., dir)}, but above that, it is the same
@@
 Hence, our idea for locating the execution problems was to make sure
-that both execution actually did the same thing. As the \OCL results
+that both execution actually did the same thing. As the OpenCL results
 were invalid, we knew the executions would diverge at one or several
@@
 
-\subparagraph{Debugging \OCL Execution: \code{GPUTrace}}
-
-With the help of \code{GPUTrace} (Section~\ref{sec:gputrace}), we
-could confront \Cuda and \OCL execution traces with a graphical
-\code{diff} tool, and spot the different porting mistakes: some
-parameters reversed, offsets incorrectly applied, \etc{}.
+\subparagraph{Debugging OpenCL Execution: \code{GPUTrace}}
+
+With the help of \code{GPUTrace} (described in Section~\ref{sec:gputrace}), we
+could confront CUDA and OpenCL execution traces with a graphical \code{diff}
+tool, and spot the different porting mistakes: some parameters reversed,
+offsets incorrectly applied, \etc{}.
 
@@
 unclear where it started. We finally got it after hours of code review
-of BOAST kernels and \OCL code. One kernel was
+of BOAST kernels and OpenCL code. One kernel was
 \emph{three}-dimensional, whereas the others were two-dimensional. But
@@
 missing.
-%As it was the same person who coded the \OCL port and the GPU
+%As it was the same person who coded the OpenCL port and the GPU
 %tracer, this third dimension was forgotten in both codes.
 
-\subparagraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
-merged in \Specfem's development tree and under test and extension by
+\subparagraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
+merged in SPECFEM3D's development tree and under test and extension by
 different research teams. On a platform with two K40x GPUs and N Intel
 Xeon processors, we measured similar results between the original
-\Cuda version and our BOAST-\Cuda version. With the same set of
-optimization flags, BOAST \Cuda and \OCL version reported similar
-execution time spans. The best execution speed was achieved with \Cuda
-version though (25\% higher than \OCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
+CUDA version and our BOAST-CUDA version. With the same set of
+optimization flags, BOAST CUDA and OpenCL version reported similar
+execution time spans. The best execution speed was achieved with CUDA
+version though (25\% higher than OpenCL), as one optimization parameter
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to OpenCL run-time, as of version
 1.1. This parameter, in addition to specifying the work group size (which can
-be done in \OCL), also constrains the number of work group that must run in
-parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
+be done in OpenCL), also constrains the number of work group that must run in
+parallel on a multiprocessor. This value is set to 7 in CUDA. This means that
 the compiler must be very conservative on register usage in order to allow this
 parallelism which allows better overlapping of communications end computations.
-This functionality is not supported in \OCL.
+This functionality is not supported in OpenCL.
 
@@
 because of code duplication, also removing manually unrolled
-loops). This is beneficial for \Specfem as it improves the readability
+loops). This is beneficial for SPECFEM3D as it improves the readability
 and maintainability of its source code.
 
-We have also been able to enhance \Specfem's non-regression test-suite
+We have also been able to enhance SPECFEM3D's non-regression test-suite
 by adding per-kernel non-regression tests. This was done with the help
@@
 loads these buffer files, allocates GPU memory and initializes it through
-\Cuda or \OCL run-time, and triggers the kernel execution. A comparison
+CUDA or OpenCL run-time, and triggers the kernel execution. A comparison
 of the output values (for instance against a maximal error level)
@@
 
-In the same mindset, we provided \Specfem test-suite with kernel
+In the same mindset, we provided SPECFEM3D test-suite with kernel
 performance evaluation mechanisms. These tests will help developers to
@@
 SPIRAL~\cite{puschel2004spiral} is a decade old generation
-framework for signal processing. It uses a proprietary DSL, SPL (Signal
+framework for signal processing. It uses a proprietary DSL called SPL (Signal
 Processing Language), to describe a DSP algorithm. This DSL is then transformed
@@
 POET~\cite{yi2007poet} also uses a DSL to describe custom code transformations,
-like loop unrolling, loop blocking and loop interchange. Those transformation
+like loop unrolling, loop blocking and loop interchange. Those transformations
 can be parametrized in order to tune the application. Orio~\cite{Hart2009:Orio}
 can be compared to POET as it aims at describing possible code transformations
-using a DSL. All those approaches are very different from our as they put the
+using a DSL. All those approaches are very different from ours as they put the
 emphasis on compilation techniques whereas BOAST relies on the user to express
@@
 
-%\section*{References}
+\section*{References}
 

### 2016-01-13 Paper/BOAST.tex
Added reference to Heterogeneous Programming Library (HPL).

@@
 
+  Heterogeneous Programming Library~\cite{vinas2013exploiting} (HPL) is maybe
+the most similar to BOAST. It is an autotuning kernel library in C++ that targets
+OpenCL. It uses an EDSL in C++ to describe generic computing kernels. Like when
+using BOAST, optimization are left to the users and written directly using the
+EDSL. The syntax is very close to real C++ with the classical control
+structures having a trailing underscore. But whereas BOAST re-interprets the
+same code while changing the environment, in HPL each expression is stored in a
+tree and the tree is reevaluated in a different environment.
+
   \subsection{Optimization Space Pruner}

### 2016-01-27 Paper/BOAST.tex
Added Acknowledgments.
Tentative abstract and improved conclusion.

@@
 \newcommand{\code}[1]{\texttt{#1}}
+\newcommand*{\Num}{N\textsuperscript{o}\xspace}
 
@@
 \begin{abstract}
-The Abstract.
+The portability of real HPC applications on new platforms is an open and very
+delicate problem.  Especially the performance portability of the underlying
+computing kernels, is problematic as they need to be tuned for each and every
+platform the application encounters.  This paper presents BOAST, a
+metaprogramming framework dedicated computing kernels.  BOAST allows the
+description of a kernel and its possible optimizations using a DSL.  BOAST
+runtime will then compare the different version's performance as well as verify
+their exactness. BOAST is applied to three use cases: a Laplace kernel in
+OpenCL, and two HPC applications BigDFT (electronic density computation) and
+SPECFEM3D (seismic and wave propagation).
 \end{abstract}
@@
 
-In this paper we presented the BOAST infrastructure as well as its
-application to three use cases from the Mont-Blanc project. Results are
-encouraging as BOAST proved to be a powerful and flexible tool that allowed
-gains in performance and performance portability.
+In this paper we presented the BOAST infrastructure as well as its application
+to three use cases from the Mont-Blanc project. Results are encouraging as
+BOAST proved to be a powerful and flexible tool that allowed gains in
+performance compared to hand tuned codes. Performance portability of those
+codes is also improved.
 
 Future development will focus on three goals. Interfacing with binary analysis
-tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop to
-guide optimization. Interfacing with search space modellers/pruners in order to
-optimize the search of the optimal version of a kernel. Continue working on
-vector code. For instance producing a collection of small to medium useful
-vector patterns (transposition, for instance) in BOAST could really help develop
-vectorized version of our algorithm.
+tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop
+to guide optimization. Interfacing with search space modellers/pruners in order
+to optimize the search of the optimal version of a kernel. Work should also
+continue on improving the support for vector codes. For instance producing a
+collection of small to medium useful vector patterns (transposition, for
+instance) in BOAST could really help users develop vectorized version of their
+algorithm.
+
+\section*{Acknowledgments}
+
+The research leading to these results has received funding from the European
+Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement
+\Num 288777 and 610402.
 

### 2016-07-21 BigDFT/GenericConvolution.rb
Update to reflect syntax improvements in BOAST 1.3.0.

@@ -826,7 +826,7 @@ class ConvolutionOperator1d
     mods = get_mods(mod_arr)
     constants = get_constants
 
-    return Procedure(function_name, vars, constants ){
+    return Procedure(function_name, vars, :constants => constants ){
       decl_filters
       decl *iters
       decl l

### 2016-07-21 BigDFT/MFVec.rb

@@ -32,7 +32,7 @@ def comp_tile_n1( n1 = 1, n2 = 1, vector_length = 1, real_size = 8 )
   input =  Real( "input", :dir => :in, :align => vector_length * real_size, :dim => Dim() )
   output = Real( "output", :dir => :out, :align => vector_length * real_size, :dim => Dim() )
   return (1..n1).collect { |n|
-    Procedure( "comp_n1_#{n}_#{n2}_#{vector_length}_#{real_size}", [n1_stride_input, n1_stride_output, input, output], [filter], :inline => true, :local => true ) {
+    Procedure( "comp_n1_#{n}_#{n2}_#{vector_length}_#{real_size}", [n1_stride_input, n1_stride_output, input, output], :constants => [filter], :inline => true, :local => true ) {
       f = Real( "f", :vector_length => vector_length )
       decl f
       tt = ((n+vector_length-1)/vector_length).times.collect { |i|

### 2016-07-21 CacheTest/KernelRead.rb

@@ -39,7 +39,7 @@ module BOAST
     j = Variable::new("j",Int)
     @@output.print "#include <#{header}>\n"
     @@output.print "inline #{Int::new.decl} modulo( #{Int::new.decl} a, #{Int::new.decl} b) { return (a+b)%b;}\n"
-    p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], [], {:return => sum , :headers => [header]}) {
+    p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], :return => sum , :headers => [header]) {
       i.decl
       j.decl
       resultV.decl
@@ -106,7 +106,7 @@ module BOAST
     i = Variable::new("i",Int)
     j = Variable::new("j",Int)
     @@output.print "inline #{Int::new.decl} modulo( #{Int::new.decl} a, #{Int::new.decl} b) { return (a+b)%b;}\n"
-    p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], [], {:return => sum } ) {
+    p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], :return => sum ) {
       i.decl
       j.decl
       sum.decl

### 2016-07-21 CacheTest/KernelRead2.rb

@@ -20,7 +20,7 @@ module BOAST
     @@output.print "inline #{Int::new.decl} modulo( #{Int::new.decl} a, #{Int::new.decl} b) { return (a+b)%b;}\n"
     register_funccall("modulo")
     sum = Int("sum")
-    print p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], [], {:return => sum , :headers => [header]}) {
+    print p = Procedure::new(function_name, [m_start, m_cycles, m_stride, buffer_size, buffer], :return => sum, :headers => [header]) {
 
       resultV = Int("resultV", {:size => elem_size, :dimension => [Dim(length)], :local => true})
       sumV = Int("sumV", {:size => elem_size, :vector_length => length})

### 2016-07-21 RuntimeTest/Swann.rb

@@ -17,7 +17,7 @@ module BOAST
     j = Variable::new("j",Int)
     sum = Variable::new("sum",Real)
     vars = [colidx,a,p]
-    sw = Procedure::new(function_name, [colidx,a,p],[ncolidx,np]) {
+    sw = Procedure::new(function_name, [colidx,a,p], :constants => [ncolidx,np]) {
       i.decl
       j.decl
       sum.decl

### 2016-07-21 SPECFEM3D/inner_core_impl_kernel_forward.rb

@@ -59,7 +59,7 @@ module BOAST
     v.push sigma_xz = Real("sigma_xz", :dir => :out, :dim => [Dim()], :private => true )
     v.push sigma_yz = Real("sigma_yz", :dir => :out, :dim => [Dim()], :private => true )
 
-    p = Procedure( function_name, v, [], :local => true ) {
+    p = Procedure( function_name, v, :local => true ) {
       c = (0..5).collect { |indx1|
             (0..5).collect { |indx2|
               if(indx2 < indx1) then
@@ -132,7 +132,7 @@ module BOAST
     v.push sigma_xz               = Real("sigma_xz",           :dir => :out, :dim => [Dim()], :private => true )
     v.push sigma_yz               = Real("sigma_yz",           :dir => :out, :dim => [Dim()], :private => true )
 
-    p = Procedure( function_name, v, [], :local => true ) {
+    p = Procedure( function_name, v, :local => true ) {
       decl lambdal = Real("lambdal"), mul = Real("mul"), lambdalplus2mul = Real("lambdalplus2mul"), kappal = Real("kappal")
 
       print kappal === d_kappavstore[offset]
@@ -186,7 +186,7 @@ module BOAST
     v.push sigma_xz               = Real("sigma_xz",           :dir => :out, :dim => [Dim()], :private => true )
     v.push sigma_yz               = Real("sigma_yz",           :dir => :out, :dim => [Dim()], :private => true )
 
-    p = Procedure( function_name, v, [], :local => true ) {
+    p = Procedure( function_name, v, :local => true ) {
       decl kappavl = Real("kappavl"), muvl = Real("muvl"), kappahl = Real("kappahl"), muhl = Real("muhl")
       decl rhovpvsq = Real("rhovpvsq"), rhovphsq = Real("rhovphsq"), rhovsvsq = Real("rhovsvsq"), rhovshsq = Real("rhovshsq"), eta_aniso = Real("eta_aniso")
       decl costheta = Real("costheta"), sintheta = Real("sintheta"), cosphi = Real("cosphi"), sinphi = Real("sinphi")
@@ -556,7 +556,7 @@ module BOAST
       qualifiers = "__attribute__((reqd_work_group_size(#{ngll3_padded},1,1))) "
     end
 
-    p = Procedure(function_name, v, constants, :qualifiers => qualifiers)
+    p = Procedure(function_name, v, :constants => constants, :qualifiers => qualifiers)
[…]

### 2016-07-21 SPECFEM3D/outer_core_impl_kernel_forward.rb

@@ -16,7 +16,7 @@ module BOAST
 
     ngll3 = Int("NGLL3", :const => n_gll3)
 
-    p = Procedure(function_name, v, [], :local => true) {
+    p = Procedure(function_name, v, :local => true) {
       decl two_omega_deltat = Real("two_omega_deltat")
       decl cos_two_omega_t  = Real("cos_two_omega_t")
       decl sin_two_omega_t  = Real("sin_two_omega_t")
@@ -121,7 +121,7 @@ module BOAST
       #v.push(d_hprime_xx_oc_tex)
     end
 
-    p = Procedure(function_name, v, constants)
+    p = Procedure(function_name, v, :constants => constants)
     if(get_lang == CUDA and ref) then
       output.print File::read("references/#{function_name}.cu".gsub("_forward","").gsub("_adjoint",""))
     elsif(get_lang == CL or get_lang == CUDA) then

### 2016-12-26 Paper/BOAST.tex
abstract

@@
 delicate problem.  Especially the performance portability of the underlying
-computing kernels, is problematic as they need to be tuned for each and every
+computing kernels is problematic as they need to be tuned for each and every
 platform the application encounters. This paper presents BOAST, a
-metaprogramming framework dedicated computing kernels.  BOAST allows the
+metaprogramming framework dedicated to computing kernels.  BOAST allows the
 description of a kernel and its possible optimizations using a DSL. BOAST
-runtime will then compare the different version's performance as well as verify
+runtime will then compare the different versions' performance as well as verify
 their exactness. BOAST is applied to three use cases: a Laplace kernel in
-OpenCL, and two HPC applications BigDFT (electronic density computation) and
+OpenCL and two HPC applications BigDFT (electronic density computation) and
 SPECFEM3D (seismic and wave propagation).

### 2016-12-26 Paper/BOAST.tex
introduction

@@
 \begin{abstract}
-The portability of real HPC applications on new platforms is an open and very
-delicate problem.  Especially the performance portability of the underlying
-computing kernels is problematic as they need to be tuned for each and every
-platform the application encounters. This paper presents BOAST, a
-metaprogramming framework dedicated to computing kernels.  BOAST allows the
-description of a kernel and its possible optimizations using a DSL. BOAST
-runtime will then compare the different versions' performance as well as verify
-their exactness. BOAST is applied to three use cases: a Laplace kernel in
-OpenCL and two HPC applications BigDFT (electronic density computation) and
-SPECFEM3D (seismic and wave propagation).
+  The portability of real HPC applications on new platforms is an open and very delicate problem.  Especially the performance portability of the
+  underlying computing kernels is problematic as they need to be tuned for each and every platform the application encounters. This paper presents
+  BOAST, a metaprogramming framework dedicated to computing kernels.  BOAST allows the description of a kernel and its possible optimizations using a
+  DSL. BOAST runtime will then compare the different versions' performance as well as verify their exactness. BOAST is applied to three use cases: a
+  Laplace kernel in OpenCL and two HPC applications BigDFT (electronic density computation) and SPECFEM3D (seismic and wave propagation).
 \end{abstract}
@@
 \begin{keyword}
-Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non
-Regression Testing \sep Productivity and Software Design \sep High Performance
-Computing
+  Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non Regression Testing \sep Productivity and Software Design \sep High
+  Performance Computing
 \end{keyword}
@@
 
+%%% INTRODUCTION %%%
 \section{Introduction}
 
-Porting and tuning HPC applications to new platforms is tedious and costly in
-term of human resources.  Many developer teams are more concentrated on solving
-the physical problems faster and better with new algorithms and simulating new
-physics. This is why optimization, while important, is a task which is
-sometimes neglected or hidden for many developers.  In the same way, Domain
-Specific Language (DSLs) have exactly this goal, the complete separation of the
-high-level code dedicated to simulate new physics and the low-level one which
-is more dependent on the HPC architecture. Nevertheless, decoupling the high
-and low levels don't avoid the need of optimization which becomes more and more
-unavoidable with new architectures based on many cores and accelerators.
-
-Unfortunately, portability efforts of the low level i.e. computing kernels are
-often lost when migrating to a new architecture. Worse, code may lose
-maintainability because several versions of some functionalities coexist,
-usually with a lot of duplication.  Thus productivity of porting and tuning
-efforts is low as a huge fraction of those developments are never used after
-the platform they were intended for is decommissioned.
-
-The primary goal of the BOAST metaprogramming framework is to simplify the
-optimization of computing kernels for HPC applications. The main idea is to use
-a high level language to define an API of computing kernels i.e.  subroutines
-and their related arguments, but also describe the way to optimize it
-(unrolling, vectorization, parametrization, templating, tiling).  BOAST will
-test automatically the many possibilities of optimizations and choose the best
-one after validation from a reference version.  Then BOAST can generate pieces
-of Fortran, C, OpenCL, or CUDA codes, test it for a specific target
-architecture.
-
-With the definition of an API, BOAST can hide the implementation details at
-high level but also is a way to factorize different kernels depending only on
-parameters such as data types, length of filters or boundary conditions.  The
-maintainability and genericity of HPC codes is often limited. One of the reason
-is that producing generic code in Fortran 90/95 is difficult as the language
-does not really fit for it. Sometimes, adding genericity or functionalities can
-be quite costly and degrades performance as optimization opportunities that
-come from over-specification are lost. BOAST can provide a solution to handle
-genericity and add new functionalities at low levels.
+Porting and tuning HPC applications to new platforms is tedious and costly in term of human resources.  Many developer teams are more concentrated on
+solving the physical problems faster and better with new algorithms and simulating new physics. This is why optimization, while important, is a task
+which is sometimes neglected or hidden for many developers.  In the same way, Domain Specific Language (DSLs) have exactly this goal, the complete
+separation of the high-level code dedicated to simulate new physics and the low-level one which is more dependent on the HPC
+architecture. Nevertheless, decoupling the high and low levels does not avoid the need of optimizations which become more and more unavoidable with
+new architectures based on many cores and accelerators.
+
+Unfortunately, portability efforts of the low level codes (i.e. computing kernels) are often lost when migrating to a new architecture. Worse, code
+may lose maintainability because several versions of some functionalities coexist, usually with a lot of duplication.  Thus productivity of porting
+and tuning efforts is low as a huge fraction of those developments are never used after the platform they were intended for is decommissioned.
+
+The primary goal of the BOAST metaprogramming framework is to simplify the optimization of computing kernels for HPC applications. The main idea is to
+use a high level language to define an API of computing kernels i.e. subroutines and their related arguments, but also describe the way to optimize
+it (unrolling, vectorization, parametrization, templating, tiling). BOAST will then test automatically the many possibilities of optimizations and
+choose the best one after validation from a reference version. Then BOAST can generate pieces of Fortran, C, OpenCL, or CUDA codes, test them for a
+specific target architecture.
+
+With the definition of an API, BOAST can hide the implementation details at high level but also is a way to factorize different kernels depending only
+on parameters such as data types, length of filters, or boundary conditions. The maintainability and genericity of HPC codes is often limited. One of
+the reason is that producing generic code in Fortran 90/95 is difficult as the language does not really fit for it. Sometimes, adding genericity or
+functionalities can be quite costly and may degrade performance as optimization opportunities that come from over-specification are lost. BOAST can
+provide a solution to handle genericity and add new functionalities at low levels.
 
@@
 
-The paper is organized as follows. First we further motivate the need for
-metaprogramming frameworks.  Secondly we illustrate the optimization problems
-related to scientific computing with the Laplace equation. Thirdly we present
-our solution, called BOAST, its architecture and its functionalities. Fourthly
-we apply BOAST to three use cases: the Laplace equation motivating example;
-short multi-dimensional convolutions from the BigDFT quantum chemistry
-application; and the seismic wave propagation simulator SPECFEM3D where 80 GPU
-computing kernels have been ported improving, considerably the maintainability
-and providing a single code base for OpenCL and CUDA kernel versions. Finally
-we compare BOAST with other related works.
-
-\section{Background and Motivation}
+The paper is organized as follows. First in Section~\ref{sec:backg} we further motivate the need for metaprogramming frameworks.  Secondly in
+Section~\ref{sec:laplace} we illustrate the optimization problems related to scientific computing with the Laplace equation. Thirdly we present our
+solution, called BOAST, its architecture and its functionalities in Section~\ref{sec:boast}. Fourthly we apply BOAST to three use cases: the Laplace
+equation motivating example, short multi-dimensional convolutions from the BigDFT quantum chemistry application, and the seismic wave propagation
+simulator SPECFEM3D where 80 GPU computing kernels have been ported improving considerably the maintainability and providing a single code base for
+OpenCL and CUDA kernel versions. Finally, before a conclusion and future works, we compare BOAST with other related works.
+
+%%% BACKGROUND AND MOTIVATION %%%
+\section{Background and Motivation\label{sec:backg}}
 
@@
 
+%%%
 \subsection{Evolution of HPC Architectures}
@@
 
-\section{Basic Example: OpenCL Laplace}
-\label{sec:laplace}
+%%%
+\section{Basic Example: OpenCL Laplace\label{sec:laplace}}
 
@@
 
-\section{BOAST: Using Code Generation in Application Autotuning}
+\section{BOAST: Using Code Generation in Application Autotuning\label{sec:boast}}
 

### 2016-12-26 Paper/BOAST.tex
jusqu a section 4

@@
 simulator SPECFEM3D where 80 GPU computing kernels have been ported improving considerably the maintainability and providing a single code base for
-OpenCL and CUDA kernel versions. Finally, before a conclusion and future works, we compare BOAST with other related works.
+OpenCL and CUDA kernel versions. Finally, before a conclusion and future works, we compare BOAST with other related works in
+Section~\ref{sec:related}.
 
@@
 
-The motivation of the BOAST project comes from the need of the HPC communities
-which have full and large applications.  These applications are difficult to
-port and maintain for emerging architectures. Especially since new
-architectures are based on accelerators, longer vector units, or low power
-processors.
-
-%%%
+The motivation of the BOAST project comes from the need of the HPC communities which have full and large applications.  These applications are
+difficult to port and maintain for emerging architectures especially since new architectures are based on accelerators, longer vector units, or low
+power processors.
+
+%%% EVOLUTION OF HPC ARCHITECTURES %%%
 \subsection{Evolution of HPC Architectures}
 
-Evolution of HPC Architectures is rapid and also diverse: in the last 7 years
-no less than 6 architectures have been number one in the
+Evolution of HPC Architectures is rapid and also diverse. In the last seven years no less than 6 architectures have been number one in the
 Top500~\cite{webtop500}:
@@
 \end{itemize}
-Being able to efficiently use those architectures on such a small period is
-challenging for scientists and application developers.
-
-The race to exascale is not going to simplify the environment. All of the above
-machine architectures can be considered to build these supercomputers and other
-will appear soon. For instance European FP7 project DEEP considers using
-Accelerators (XEON Phi) while the European FP7 project Mont-Blanc considers
-using low-power embedded processor with integrated GPU.
-
-Running existing applications on these new architectures is an open research
-subject as well as an ongoing porting effort.  Thus, those projects have work
-packages dedicated to applications. Those work packages are dedicated to
-porting and optimizing Scientific applications on those new architectures. In
-the DEEP project six applications were selected for porting and optimizing,
-eleven were selected in the Mont-Blanc project.
-
+
+Being able to efficiently use those architectures on such a small period is challenging for scientists and application developers.
+
+The race to exascale is not going to simplify the environment. All of the above machine architectures can be considered to build these supercomputers
+and other will appear soon. For instance European FP7 project DEEP considers using Accelerators (XEON Phi) while the European FP7 project Mont-Blanc
+considers using low-power embedded processor with integrated GPU.
+
+Running existing applications on these new architectures is an open research subject as well as an ongoing porting effort.  Thus, those projects have
+work packages dedicated to applications. Those work packages are dedicated to porting and optimizing Scientific applications on those new
+architectures. In the DEEP project six applications were selected for porting and optimizing, eleven were selected in the Mont-Blanc project.
+
+%%% SCIENTIFIC COMPUTING APPLICATIONS %%%
 \subsection{Scientific Computing Applications}
 
-Scientific Computing Applications are usually developed by other science
-researchers such as physicists, chemists or meteorologists. Those codes are
-usually written in Fortran for historical and performance reasons. Codes can be
-quite huge (several thousands lines of code [LOC]) with lots of
-functionalities. Nonetheless, they are usually based on computing kernels.
-Computing kernels are resource intensive and well-defined parts of a program
-that usually work on precisely defined data. Those kernels represent the most
-time-consuming part of an HPC application, and consequently, they are the prime
-target for optimization.
-
-Those applications are often developed by several individuals. Sometimes, some
-of those developers only work a few months on the application. Maintaining
-optimized code written by someone else is quite a challenge. Several languages
-and programming paradigms can also be used in a project and thus the maintainer
-must be knowledgeable in several areas of expertise. Portability problems can
-also be caused by the availability of an optimizing compiler for a specific
-language and architecture. C compilers are usually the first available while
-Fortran may sometimes come later.
-
-In Section~\ref{use_cases} we will present two HPC applications that we used
-as use cases: SPECFEM3D and BigDFT. They are both based on computing
+Scientific Computing Applications are usually developed by other science researchers such as physicists, chemists or meteorologists. Those codes are
+usually written in Fortran for historical and performance reasons. Codes can be quite huge (several thousands Lines of Code, LoC) with lots of
+functionalities. Nonetheless, they are usually based on computing kernels.  Computing kernels are resource intensive and well-defined parts of a
+program that usually work on precisely defined data. Those kernels represent the most time-consuming part of an HPC application, and consequently,
+they are the prime target for optimizations.
+
+Those applications are often developed by several individuals. Sometimes, some of those developers only work a few months on the
+application. Maintaining optimized code written by someone else is quite a challenge. Several languages and programming paradigms can also be used in
+a project and thus the maintainer must be knowledgeable in several areas of expertise. Portability problems can also be caused by the availability of
+an optimizing compiler for a specific language and architecture. C compilers are usually first available while Fortran may sometimes come later.
+
+In Section~\ref{use_cases} we will present two HPC applications that we used as use cases: SPECFEM3D and BigDFT. They are both based on computing
 kernels and were selected as candidate applications in the Mont-Blanc project.
 
+%%% HOW SHOULD COMPUTING KERNELS BE WRITTEN? %%%
 \subsection{How Should Computing Kernels be Written?}
 
-The problematic here is to obtain computing kernels that present good
-performance on the architectures encountered by the application while staying
-portable after the optimization process took place. Indeed the application
-might encounter one of the many architectures that can be found in HPC.
-Investing manpower to optimize the application for a new architecture is
-reasonable, suffering hindrance from previous optimization work is not. Thus
-optimizations have to be as orthogonal as possible from one another so as to be
-easily activated and deactivated.
-
-If this paradigm is followed by developers then they will rapidly be confronted
-with a huge optimization space to search. They will need to be able to test
-easily the performance impact of the chosen optimizations without running the
-full application. The same reasoning implies that kernels should be tested for
-non regression without running the full application.
-
-What we want is computing kernels that are written in a \emph{portable} manner,
-in a way that raises developer \emph{productivity}, and they must present good
-\emph{performance}.
-
-%%%
+The problematic here is to obtain computing kernels that present good performance on the architectures encountered by the application while staying
+portable after the optimization process took place. Indeed the application might encounter one of the many architectures that can be found in HPC.
+Investing manpower to optimize the application for a new architecture is reasonable, suffering hindrance from previous optimization work is not. Thus
+optimizations have to be as orthogonal as possible from one another so as to be easily activated and deactivated.
+
+If this paradigm is followed by developers then they will rapidly be confronted with a huge optimization space to search. They will need to be able to
+test easily the performance impact of the chosen optimizations without running the full application. The same reasoning implies that kernels should be
+tested for non regression without running the full application.
+
+What we want is computing kernels that are written in a \emph{portable} manner, in a way that raises developer \emph{productivity}, and they must
+present good \emph{performance}.
+
+%%% BASIC EXAMPLE: OPENCL LAPLACE %%%
 \section{Basic Example: OpenCL Laplace\label{sec:laplace}}
 
-During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
-on the Mali GPU was given. One of the case study was a Laplace filter~\cite{opencl_arm_training}.
-
-This is a good example of an algorithm that is simple in its formulation but
-can be complex to optimize because many different optimizations are available
-and can be combined together. The correct combination of optimizations will
-depend on the target architecture.
-
+During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization on the Mali GPU was given. One of the case study was a Laplace
+filter~\cite{opencl_arm_training}. This is a good example of an algorithm that is simple in its formulation but can be complex to optimize because
+many different optimizations are available and can be combined together. The correct combination of optimizations will depend on the target
+architecture.
+
+%%% THE LAPLACE FILTER %%%
 \subsection{The Laplace Filter}
 
-Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace
-filter in C99. For clarity, boundary conditions have been omitted.
+Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace filter in C99. For sake of clarity, boundary conditions have been
+omitted.
 
@@
 
-A naive implementation in OpenCL is proposed in
-Listing~\ref{lst:laplace_opencl_filter}. Each work item processes one pixel of
-the resulting image.
+      A naive implementation in OpenCL is proposed in Listing~\ref{lst:laplace_opencl_filter}. Each work item processes one pixel of the resulting
+      image.
 
@@
 
+%%% POSSIBLE OPTIMIZATIONS OF THE OPENCL VERSION %%%
 \subsection{Possible Optimizations of the OpenCL Version}
 
-Several optimizations were proposed and successively applied to the previous
-implementation.
-
-\paragraph{Vectorization}
-The first proposed optimization involves computing five pixels instead of one
-using the vectors available on the Mali architecture. The vectors have to be
-kept in memory to ensure an efficient execution. Using vectors of 16 elements,
-15 useful components are simultaneously loaded and can be used in computation.
-This optimization yields speedups between 1.5 and 6 depending on the image
-size.
+Several optimizations were proposed and successively applied to the previous implementation.
+
+%%%
+\paragraph{Vectorization} The first proposed optimization involves computing five pixels instead of one using the vectors available on the Mali
+architecture. The vectors have to be kept in memory to ensure an efficient execution. Using vectors of 16 elements, 15 useful components are
+simultaneously loaded and can be used in computation.  This optimization yields speedups between 1.5 and 6 depending on the image size.
 
@@
 
-\paragraph{Synthesizing Loads} It is possible to reduce the number of loads
-since the vectors are overlapping.
-Listing~\ref{lst:laplace_opencl_filter_vector_synth} shows how the vectors are
-loaded in this case. This optimization yields marginal improvements.
+      \paragraph{Synthesizing Loads} It is possible to reduce the number of loads since the vectors are overlapping.
+      Listing~\ref{lst:laplace_opencl_filter_vector_synth} shows how the vectors are loaded in this case. This optimization yields marginal
+      improvements.
 
@@
 
-\paragraph{Temporary Variables Size} Using \emph{int} to store intermediary
-results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
-the code is modified to use smaller types. This yields a speedup of 1.3 for most
-image sizes.
+      \paragraph{Temporary Variables Size} Using \emph{int} to store intermediary results is
+      unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how the code is modified to use smaller types. This yields a speedup of 1.3 for
+      most image sizes.
 
@@
 
-\paragraph{More Optimizations} Several other optimizations can be attempted at
-this point. For instance reducing or increasing the number of pixels each work
-item is processing. They can yield improved performance in some cases, especially
-when avoiding memory alignment problems. These optimizations will not be shown
-here but can be found in~\cite{opencl_arm_training}. Others consist in tiling
-the problem space or improving the loads synthesis.
-
+      \paragraph{More Optimizations} Several other optimizations can be attempted at this point. For instance reducing or increasing the number of
+      pixels each work item is processing. They can yield improved performance in some cases, especially when avoiding memory alignment
+      problems. These optimizations will not be shown here but can be found in~\cite{opencl_arm_training}. Others consist in tiling the problem space
+      or improving the loads synthesis.
+
+%%% Improving the Methodology %%%
 \subsection{Improving the Methodology}
 
-The process described in the previous Subsection is quite tedious and requires
-some intimate knowledge of the target architecture. The results are impressive,
-as speedups of almost 10 are observed compared to the naive OpenCL
-implementation.
-
-Nonetheless, the process is frustrating. The optimizations are never evaluated
-independently from one another. Some were arbitrarily configured (like the
-number of pixels chosen for instance). Testing the different combinations of
-those optimizations and the different parameters would be very costly in
-developer time and, moreover, like all repetitive and tedious tasks, error
-prone. Thus every created kernel would have to be thoroughly tested to ensure
-no error was done.
-
-In the next Section we will present our answer to these issues: BOAST.  BOAST is
-a framework dedicated to kernel description, optimization, regression testing
-and autotuning.
-
+The process described in the previous Subsection is quite tedious and requires some intimate knowledge of the target architecture. The results are
+impressive, as speedups of almost 10 are observed compared to the naive OpenCL implementation.
+
+Nonetheless, the process is frustrating. The optimizations are never evaluated independently from one another. Some were arbitrarily configured (like
+the number of pixels chosen for instance). Testing the different combinations of those optimizations and the different parameters would be very costly
+in developer time and, moreover, like all repetitive and tedious tasks, error prone. Thus every created kernel would have to be thoroughly tested to
+ensure no error was done.
+
+In the next Section we will present our answer to these issues: BOAST.  BOAST is a framework dedicated to kernel description, optimization, regression
+testing, and autotuning.
+
+%%% BOAST: USING CODE GENERATION IN APPLICATION AUTOTUNING %%%
 \section{BOAST: Using Code Generation in Application Autotuning\label{sec:boast}}
@@
 
-\section{Related Work}
+%%%
+\section{Related Work\label{sec:related}}
 

### 2016-12-27 Paper2/BOAST.tex
Ported Fred modifications to IEEE style.

@@
 
-\newcommand{\Specfem}{\productname{Specfem 3D}\xspace}
-\newcommand{\Cuda}{\productname{Cuda}\xspace}
-\newcommand{\OCL}{\productname{OpenCL}\xspace}
 \newcommand{\code}[1]{\texttt{#1}}
+\newcommand*{\Num}{N\textsuperscript{o}\xspace}
 
@@
 
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for IEEE Journals}
-%\journal{Journal of Parallel and Distributed Computing}
-
-%%%%%%%%%%%%%%%%%%%%%%%
-%% Elsevier bibliography styles
-%%%%%%%%%%%%%%%%%%%%%%%
-%% To change the style, put a % in front of the second line of the current style and
-%% remove the % from the second line of the style you would like to use.
-%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Numbered
-%\bibliographystyle{model1-num-names}
-
-%% Numbered without titles
-%\bibliographystyle{model1a-num-names}
-
-%% Harvard
-%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
-
-%% Vancouver numbered
-%\usepackage{numcompress}\bibliographystyle{model3-num-names}
-
-%% Vancouver name/year
-%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
-
-%% APA style
-%\bibliographystyle{model5-names}\biboptions{authoryear}
-
-%% AMA style
-%\usepackage{numcompress}\bibliographystyle{model6-num-names}
-
-%% `Elsevier LaTeX' style
 \bibliographystyle{IEEEtran}
-%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%\begin{frontmatter}
-
 \title{BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
@@
 %% or include affiliations in footnotes:
-\author{Brice Videau, Kevin Pouget, Luigi Genovese, Thierry Deutsch, Dimitri Komatitsch, Frédéric Desprez, Jean-François Méhaut}
-
+\author{
+   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{2}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{2}Centre National de la Recherche Scientifique}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Joseph Fourier}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/SP2M/L\_Sim}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}
+}
+
+\markboth{International Journal of Scientific Computing}
+{Videau \MakeLowercase{\textit{et al.}}: BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
 %\address[mymainaddress]{LIG/CNRS}
@@
 %\begin{keyword}
-%Code Generation \sep Portability \sep Genericity \sep Productivity and Software
-%Design \sep High Performance Computing \sep Autotuning \sep Non Regression
-%Testing
+%  Code Generation \sep Portability \sep Genericity \sep Autotuning \sep Non Regression Testing \sep Productivity and Software Design \sep High Performance Computing
 %\end{keyword}
@@
 \begin{abstract}
-The Abstract.
+The portability of real HPC applications on new platforms is an open and very
+delicate problem.  Especially the performance portability of the underlying
+computing kernels is problematic as they need to be tuned for each and every
+platform the application encounters. This paper presents BOAST, a
+metaprogramming framework dedicated to computing kernels.  BOAST allows the
+description of a kernel and its possible optimizations using a DSL. BOAST
+runtime will then compare the different versions' performance as well as verify
+their exactness. BOAST is applied to three use cases: a Laplace kernel in
+OpenCL and two HPC applications BigDFT (electronic density computation) and
+SPECFEM3D (seismic and wave propagation).
 \end{abstract}
 
+%%% INTRODUCTION %%%
 \section{Introduction}
 
-Porting and tuning HPC applications to new platforms is tedious and costly in term of human resources.
-Many developer teams are more concentrated on solving the physical problems faster and better with new algorithms
-and simulating new physics. 
-This is why optimization even if it is important is a task which is sometimes neglected or hidden for many developers.
-In the same way, the domain specific language (DSL) has exactly this goal,  the complete separation of the high-level code dedicated to simulate new physics
-and the low-level one which is more dependent to the HPC architecture. 
-Nevertheless, decoupling the high and low levels don't avoid the need of optimization which becomes more and more unavoidable 
-with the new architectures based on many cores and accelerators.
-
-Unfortunately, portability efforts of the low level i.e. computing kernels are often lost when migrating to a new
-architecture. Worse, code may lose maintainability because several versions of
-some functionalities coexist, usually with a lot of duplication.
-Thus productivity of porting and tuning efforts is low as a huge fraction of
-those developments are never used after the platform they were intended for is
-decommissioned.
-
-The primary goal of the metaprogramming framework BOAST is to simplify the optimization of the computing kernels for HPC
-applications. The main idea is to use a high level language to define an API of computing kernels i.e. subroutines and
-their related arguments, but also describe the way to optimize it (unrolling, vectorization, parametrization, templating, tiling).
-BOAST will test automatically the many possibilities of optimizations and choose the best one after validation from a
-reference version.
-Then BOAST can generate pieces of Fortran, C, OpenCL or CUDA codes, test if for a specific target architecture.
-
-With the definition of an API, BOAST can hide the details of the implementation at high level but also is the way to
-factorize different kernels depending only on parameters as datatypes, the length of filters or the boundary conditions.
-The maintainability and genericity of HPC codes is often limited. One of the reason is that producing
-generic code in Fortran 90/95 is difficult as the language does not really fit
-for it. Sometimes, adding genericity or functionalities can be quite costly and degrades performance as optimization
-opportunities that come from over-specification are lost.
-BOAST can provide a solution to handle genericity and add new functionalities at low levels.
-For instance, the optimization of short convolutions depends strongly
-on the length of the filters. This is why HPC codes using finite differences handle only one order. BOAST allows
-handling any order and even more generalizing to another numerical scheme.
-
-In this paper, we illustrate the optimization problems related to scientific computing with the Laplace equation.
-Then we present our solution, BOAST, its architecture and its functionalities. We apply BOAST to three use cases, the
-Laplace equation, short multi-dimensional convolutions to perform differential operators and finally
-the wavelet transform which are used in a quantum chemistry code, BigDFT.
-Then BOAST is applied to a full application, SPECFEM3D dedicated to the seismic wave propagation simulator.
-80 GPU computing kernels have been ported improving considerably the maintanability and providing a single code base for
-OpenCL and CUDA kernel versions.
-Finally we compare BOAST with other related works.
-
-% Porting and tuning HPC applications to new platforms is tedious and costly in
-% terms of human resources. Nonetheless, it is a very important aspect of the
-% Mont-Blanc project. Indeed, for the project, more than ten applications were
-% selected to be ported and optimized for the prototype platform.
-% 
-% Unfortunately, portability efforts are often lost when migrating to a new
-% architecture. Worse, code may lose maintainability because several versions of
-% some functionalities coexist, usually with a lot of duplication.
-% 
-% Thus productivity of porting and tuning efforts is low as a huge fraction of
-% those developments are never used after the platform they were intended for is
-% decommissioned.
-% 
-% Genericity of HPC codes is often limited. One of the reason is that producing
-% generic code in Fortran 90/95 is difficult as the language does not really fit
-% for it. Sometimes, adding genericity degrades performance as optimization
-% opportunities that come from over-specification are lost.
-% 
-% Functionality of HPC codes is tied to the previous point. Without genericity,
-% adding new functionalities can be quite costly.
-
-\section{Background and Motivation}
-
-The motivation of the BOAST project comes from the need of the HPC communities which have full and large applications
-difficult to port and maintain for emerging architectures based on accelerators, longer vector lengths, or low power
-processors.
-
+Porting and tuning HPC applications to new platforms is tedious and costly in
+term of human resources.  Many developer teams are more concentrated on solving
+the physical problems faster and better with new algorithms and simulating new
+physics. This is why optimization, while important, is a task which is
+sometimes neglected or hidden for many developers.  In the same way, Domain
+Specific Language (DSLs) have exactly this goal, the complete separation of the
+high-level code dedicated to simulate new physics and the low-level one which
+is more dependent on the HPC architecture. Nevertheless, decoupling the high
+and low levels does not avoid the need of optimizations which become more and
+more unavoidable with new architectures based on many cores and accelerators.
+
+Unfortunately, portability efforts of the low level codes (i.e. computing
+kernels) are often lost when migrating to a new architecture. Worse, code may
+lose maintainability because several versions of some functionalities coexist,
+usually with a lot of duplication.  Thus productivity of porting and tuning
+efforts is low as a huge fraction of those developments are never used after
+the platform they were intended for is decommissioned.
+
+The primary goal of the BOAST metaprogramming framework is to simplify the
+optimization of computing kernels for HPC applications. The main idea is to use
+a high level language to define an API of computing kernels i.e. subroutines
+and their related arguments, but also describe the way to optimize it
+(unrolling, vectorization, parametrization, templating, tiling). BOAST will
+then test automatically the many possibilities of optimizations and choose the
+best one after validation from a reference version. Then BOAST can generate
+pieces of Fortran, C, OpenCL, or CUDA codes, test them for a specific target
+architecture.
+
+With the definition of an API, BOAST can hide the implementation details at
+high level but also is a way to factorize different kernels depending only on
+parameters such as data types, length of filters, or boundary conditions. The
+maintainability and genericity of HPC codes is often limited. One of the reason
+is that producing generic code in Fortran 90/95 is difficult as the language
+does not really fit for it. Sometimes, adding genericity or functionalities can
+be quite costly and may degrade performance as optimization opportunities that
+come from over-specification are lost. BOAST can provide a solution to handle
+genericity and add new functionalities at low levels.
+
+The paper is organized as follows. First in Section~\ref{sec:backg} we further
+motivate the need for metaprogramming frameworks.  Secondly in
+Section~\ref{sec:laplace} we illustrate the optimization problems related to
+scientific computing with the Laplace equation. Thirdly we present our
+solution, called BOAST, its architecture and its functionalities in
+Section~\ref{sec:boast}. Fourthly we apply BOAST to three use cases: the
+Laplace equation motivating example, short multi-dimensional convolutions from
+the BigDFT quantum chemistry application, and the seismic wave propagation
+simulator SPECFEM3D where 80 GPU computing kernels have been ported improving
+considerably the maintainability and providing a single code base for OpenCL
+and CUDA kernel versions. Finally, before a conclusion and future works, we
+compare BOAST with other related works in Section~\ref{sec:related}.
+
+%%% BACKGROUND AND MOTIVATION %%%
+\section{Background and Motivation\label{sec:backg}}
+
+The motivation of the BOAST project comes from the need of the HPC communities
+which have full and large applications.  These applications are difficult to
+port and maintain for emerging architectures especially since new architectures
+are based on accelerators, longer vector units, or low power processors.
+
+%%% EVOLUTION OF HPC ARCHITECTURES %%%
 \subsection{Evolution of HPC Architectures}
 
-Evolution of HPC Architectures is rapid and also diverse: in the last 5
-years no less than 6 architectures have been number one in the Top500:
+Evolution of HPC Architectures is rapid and also diverse. In the last seven
+years no less than 7 architectures have been number one in the
+Top500~\cite{webtop500}:
 \begin{itemize}
-\item Intel Processor + Xeon Phi (Tianhe-2)
-\item AMD Processor + NVIDIA GPU (Titan)
-\item IBM BlueGene/Q (Sequoia)
-\item Fujitsu SPARC64 (K computer)
-\item Intel Processor + NVIDIA GPU (Tianhe-1)
-\item AMD Processor (Jaguar)
+\item Sunway processor (TaihuLight, 2016)
+\item Intel Processor + Xeon Phi (Tianhe-2, 2013)
+\item AMD Processor + NVIDIA GPU (Titan, 2012)
+\item IBM BlueGene/Q (Sequoia, 2012)
+\item Fujitsu SPARC64 (K computer, 2011)
+\item Intel Processor + NVIDIA GPU (Tianhe-1, 2010)
+\item AMD Processor (Jaguar, 2009)
 \end{itemize}
-Being able to efficiently use those architectures on such a small
-time-frame is challenging.
+
+Being able to efficiently use those architectures on such a small period is
+challenging for scientists and application developers.
 
 The race to exascale is not going to simplify the environment. All of the above
-architectures can be considered. Network architectures also can be very
-diverse.  For instance European FP7 project DEEP considers using Accelerators
-(XEON Phi) while the European FP7 project Mont-Blanc considers using low-power
-embedded processor with integrated GPU.
-
-Running existing applications on those new architectures is an open research
+machine architectures can be considered to build these supercomputers and other
+will appear soon. For instance European FP7 project DEEP considers using
+Accelerators (XEON Phi) while the European FP7 project Mont-Blanc considers
+using low-power embedded processor with integrated GPU.
+
+Running existing applications on these new architectures is an open research
 subject as well as an ongoing porting effort.  Thus, those projects have work
-packages dedicated to applications. Those work packages are dedicated to porting
-and optimizing Scientific applications on those new architectures. In the DEEP
-project six applications were selected for porting and optimizing, eleven were
-selected in the Mont-Blanc project.
-
+packages dedicated to applications. Those work packages are dedicated to
+porting and optimizing Scientific applications on those new architectures. In
+the DEEP project six applications were selected for porting and optimizing,
+eleven were selected in the Mont-Blanc project.
+
+%%% SCIENTIFIC COMPUTING APPLICATIONS %%%
 \subsection{Scientific Computing Applications}
 
-Scientific Computing Applications are usually developed by physicist, chemist or
-meteorologist. Those codes are usually written in Fortran for historical and
-performance reasons. Codes can be quite huge (several thousands lines of code [LOC])
-with lots of functionalities. Nonetheless, they are usually based on computing
-kernels.  Computing kernels are resource intensive and well-defined parts of a
-program that usually work on precisely defined data. Those kernels represent the most
-time-consuming part of an HPC application, and consequently, they are the prime target for
-optimization.
+Scientific Computing Applications are usually developed by other science
+researchers such as physicists, chemists or meteorologists. Those codes are
+usually written in Fortran for historical and performance reasons. Codes can be
+quite huge (several thousands Lines of Code, LoC) with lots of functionalities.
+Nonetheless, they are usually based on computing kernels.  Computing kernels
+are resource intensive and well-defined parts of a program that usually work on
+precisely defined data. Those kernels represent the most time-consuming part of
+an HPC application, and consequently, they are the prime target for
+optimizations.
 
@@
 also be caused by the availability of an optimizing compiler for a specific
-language and architecture. C compilers are usually the first available while
-Fortran may sometimes arrive later.
-
-In Section~\ref{use_cases} we will present two HPC applications that we used
-as use cases: SPECFEM3D and BigDFT. They are both based on computing
-kernels and were selected as candidate applications in the Mont-Blanc project.
-
+language and architecture. C compilers are usually first available while
+Fortran may sometimes come later.
+
+In Section~\ref{use_cases} we will present two HPC applications that we used as
+use cases: SPECFEM3D and BigDFT. They are both based on computing kernels and
+were selected as candidate applications in the Mont-Blanc project.
+
+%%% HOW SHOULD COMPUTING KERNELS BE WRITTEN? %%%
 \subsection{How Should Computing Kernels be Written?}
@@
 
-What we want is computing kernels that are written:
-\begin{itemize}
-\item in a \emph{portable} manner,
-\item in a way that raises developer \emph{productivity},
-\item and, present good \emph{performance}.
-\end{itemize}
-
-\section{Basic Example: OpenCL Laplace}
-\label{sec:laplace}
+What we want is computing kernels that are written in a \emph{portable} manner,
+in a way that raises developer \emph{productivity}, and they must present good
+\emph{performance}.
+
+%%% BASIC EXAMPLE: OPENCL LAPLACE %%%
+\section{Basic Example: OpenCL Laplace\label{sec:laplace}}
 
 During one of the Mont-Blanc face to face meeting a talk on OpenCL optimization
-on the Mali GPU was given. One of the case study was a Laplace filter on the
-GPU~\cite{opencl_arm_training}.
-
-This is a good example of an algorithm that is simple in its formulation but
-can be complex to optimize because many different optimizations are available
-and can be combined together. The correct combination of optimizations will
-depend on the target architecture.
-
+on the Mali GPU was given. One of the case study was a Laplace
+filter~\cite{opencl_arm_training}. This is a good example of an algorithm that
+is simple in its formulation but can be complex to optimize because many
+different optimizations are available and can be combined together. The correct
+combination of optimizations will depend on the target architecture.
+
+%%% THE LAPLACE FILTER %%%
 \subsection{The Laplace Filter}
@@
 Listing~\ref{lst:laplace_c_filter} shows a simple implementation of the Laplace
-filter in C99. For clarity, boundary conditions have been omitted.
+filter in C99. For sake of clarity, boundary conditions have been omitted.
 
 \lstset{style=customc,
-        caption={Laplace C99 Filter},
+        caption={Laplace C99 Filter.},
         breaklines=true,
@@
 
+%%% POSSIBLE OPTIMIZATIONS OF THE OPENCL VERSION %%%
 \subsection{Possible Optimizations of the OpenCL Version}
@@
 
-\paragraph{Vectorization}
-The first proposed optimization involves computing five pixels instead of one using the
-vectors available to the Mali architecture. The vectors have to reside in memory to
-ensure an efficient execution. Using vectors of 16 elements, 15 useful components
-are simultaneously loaded and can be used in computation. This optimization
-yields speedups between 1.5 and 6 depending on the image size.
+\paragraph{Vectorization} The first proposed optimization involves computing
+five pixels instead of one using the vectors available on the Mali
+architecture. The vectors have to be kept in memory to ensure an efficient
+execution. Using vectors of 16 elements, 15 useful components are
+simultaneously loaded and can be used in computation.  This optimization yields
+speedups between 1.5 and 6 depending on the image size.
 
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized},
+        caption={Laplace OpenCL Filter Vectorized.},
         breaklines=true,
@@
 
-\paragraph{Temporary Variables Size} Using \textit{int} to store intermediary
+\paragraph{Temporary Variables Size} Using \emph{int} to store intermediary
 results is unnecessary. Listing~\ref{lst:laplace_opencl_filter_short} show how
-the code is modified to use smaller types. This yields a speedup of 1.3 for most
-image sizes.
+the code is modified to use smaller types. This yields a speedup of 1.3 for
+most image sizes.
 
 \lstset{style=customcl,
-        caption={Laplace OpenCL Filter Vectorized Using \textit{short}},
+        caption={Laplace OpenCL Filter Vectorized Using \emph{short}.},
         breaklines=true,
@@
 this point. For instance reducing or increasing the number of pixels each work
-item is processing. They can yield improved performance in some cases, especially
-when avoiding memory alignment problems. These optimizations will not be shown
-here but can be found in~\cite{opencl_arm_training}. Others consist in tiling
-the problem space or improving the loads synthesis.
-
+item is processing. They can yield improved performance in some cases,
+especially when avoiding memory alignment problems. These optimizations will
+not be shown here but can be found in~\cite{opencl_arm_training}. Others
+consist in tiling the problem space or improving the loads synthesis.
+
+%%% Improving the Methodology %%%
 \subsection{Improving the Methodology}
@@
 Nonetheless, the process is frustrating. The optimizations are never evaluated
-independently from one another. Some were arbitrarily configured (the number of
-pixels chosen for instance). Testing the different combinations of those
-optimizations and the different parameters would be very costly in developer
-time and, moreover, like all repetitive and tedious tasks, error prone. Thus
-every created kernel would have to be thoroughly tested to ensure no error was
-done.
-
-In the next Section we will present our answer to these issues: BOAST.  BOAST is
-a framework dedicated to kernel description, optimization, regression testing
-and autotuning.
-
-\section{BOAST: Using Code Generation in Application Autotuning}
+independently from one another. Some were arbitrarily configured (like the
+number of pixels chosen for instance). Testing the different combinations of
+those optimizations and the different parameters would be very costly in
+developer time and, moreover, like all repetitive and tedious tasks, error
+prone. Thus every created kernel would have to be thoroughly tested to ensure
+no error was done.
+
+In the next Section we will present our answer to these issues: BOAST.  BOAST
+is a framework dedicated to kernel description, optimization, regression
+testing, and autotuning.
+
+%%% BOAST: USING CODE GENERATION IN APPLICATION AUTOTUNING %%%
+\section{BOAST: Using Code Generation in Application Autotuning\label{sec:boast}}
 
@@
 
-%  \begin{itemize}
-%  \item Abstraction of programming concepts: variables, procedures (
-%portability, productivity )
-%  \item Operator overloading: simpler syntax ( productivity, portability )
-%  \item Two languages in one ( productivity, performance ) 
-%  \end{itemize}
-
 Usually computing kernels are hotspots of an HPC application, and they are most
@@
 operations is studied in~\cite{soliman2009performance}, and shows performance
-increase of a factor up to 5 when using correct loop imbrication. The importance
-of code transformation is stressed in~\cite{ye2011porting}, where a selection of GPU
-kernels are ported to CPU and optimized.
+increase of a factor up to 5 when using correct loop imbrication. The
+importance of code transformation is stressed in~\cite{ye2011porting}, where a
+selection of GPU kernels are ported to CPU and optimized.
 
@@
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[scale=0.8]{BOAST_DSL.pdf}
-%\caption{BOAST DSL intent.}
-%\label{fig:dsl}
-%\end{center}
-%\end{figure}
-
 In order to manipulate those abstractions we want to have a syntax resembling
-what programmers use. For instance, commonly used operators have to be available
-and behave as expected. It must also be possible to differentiate an action on
-an abstraction in the context of an expression and in the context of the
-management of this expression. $c = a + b$, an expression that affects the
-results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves the
-expression $a + b$ to a variable $c$. This is why an embedded domain specific
-language approach was selected~\cite{hudak1996building}. This feature allows for
-the coexistence of two languages: the host language and the Domain Specific
-Language (DSL).  In our case, DSL allows the description of the kernel ($c = a +
-b$) while the host language provides the meta programming of the kernel ($c
-\leftarrow a + b$). Operator overloading of the host language will provide the
-familiar syntax programmers are accustomed to.
+what programmers use. For instance, commonly used operators have to be
+available and behave as expected. It must also be possible to differentiate an
+action on an abstraction in the context of an expression and in the context of
+the management of this expression. $c = a + b$, an expression that affects the
+results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves
+the expression $a + b$ to a variable $c$. This is why an embedded domain
+specific language approach was selected~\cite{hudak1996building}. This feature
+allows for the coexistence of two languages: the host language and the Domain
+Specific Language (DSL).  In our case, DSL allows the description of the kernel
+($c = a + b$) while the host language provides the meta programming of the
+kernel ($c \leftarrow a + b$). Operator overloading of the host language will
+provide the familiar syntax programmers are accustomed to.
 
 It was also important to have our constructs like \emph{for} loops to have a
-syntax approaching those commonly found in programming languages. To this end, we
-needed a language which could seamlessly pass a block of code to a function.
+syntax approaching those commonly found in programming languages. To this end,
+we needed a language which could seamlessly pass a block of code to a function.
 Ruby~\cite{matsumoto2002ruby} is one such language. It has deep introspection
@@
 One of the added advantage of using a high-level scripting language like Ruby
-as the host language if its interfacing capabilities, providing an easy way
-to use the libraries needed during the development of the framework.
+as the host language if its interfacing capabilities, providing an easy way to
+use the libraries needed during the development of the framework.
 
@@
 
-Figure~\ref{fig:BOAST_control} shows some basic usages of the control structures.
-The example shows a C macro or function that behaves similarly to the Fortran modulo
-intrinsic. The sample script (inefficiently) computes and stores in \emph{j} the
-number of multiples of $7$ in the 0 to 100 range. It uses the For and If control
-structures. A Ruby block is passed to each of those constructs. This block is
-evaluated at the time the construct is printed. If several such constructs are
-needed (in an \emph{if elsif else} case, for instance) they can be explicitly
-passed as parameters using the \emph{lambda} Ruby keyword.
+Figure~\ref{fig:BOAST_control} shows some basic usages of the control
+structures.  The example shows a C macro or function that behaves similarly to
+the Fortran modulo intrinsic. The sample script (inefficiently) computes and
+stores in \emph{j} the number of multiples of $7$ in the 0 to 100 range. It
+uses the For and If control structures. A Ruby block is passed to each of those
+constructs. This block is evaluated at the time the construct is printed. If
+several such constructs are needed (in an \emph{if elsif else} case, for
+instance) they can be explicitly passed as parameters using the \emph{lambda}
+Ruby keyword.
 
@@
 \emph{opn} and \emph{close} keywords that are used to print a control structure
-without an associated Ruby block. This time we placed ourselves inside the BOAST
-namespace.
-
-The generated outputs in Fortran (Listing~\ref{lst:BOAST_procedure_FORTRAN}) and
-C (Listing~\ref{lst:BOAST_procedure_C}) show the difference in meta-information that is
-kept between both versions.
-
-%Considering source-to-source transformations that we investigate with the BOAST
-%tool in this paper, we can compare ourselves to POET~\cite{POET2012}  and
-%ABCLibScript~\cite{ABCLib2006}. However, both tools apply transformations
-%to the syntaxic tree of a program while BOAST intervenes well before in the
-%transformation process. Indeed, one may first use BOAST on order to generate a
-%program (a specific syntactic tree) and then benefit from the transformation
-%facilities offered by one of the above tools.
+without an associated Ruby block. This time we placed ourselves inside the
+BOAST namespace.
+
+The generated outputs in Fortran (Listing~\ref{lst:BOAST_procedure_FORTRAN})
+and C (Listing~\ref{lst:BOAST_procedure_C}) show the difference in
+meta-information that is kept between both versions.
 
@@
 In the previous Section we presented BOAST's language. This allows us to
-describe procedures and functions and to meta-program them using Ruby.
-Each version has to be compiled, linked, and executed to assess its performance
-in order to find the best version of a computing kernel.  This can be very
-time consuming if this process cannot be automated. By enabling more versions
-for evaluation, the automation will bring improved portability, better performance,
+describe procedures and functions and to meta-program them using Ruby.  Each
+version has to be compiled, linked, and executed to assess its performance in
+order to find the best version of a computing kernel.  This can be very time
+consuming if this process cannot be automated. By enabling more versions for
+evaluation, the automation will bring improved portability, better performance,
 and in the end will improve the productivity of the developer.
@@
 In this section we will present the different aspects of BOAST run-time that
-allow this automation. Those aspects include:
-  \begin{itemize}
-  \item multi target language generation ( performance, portability )
-  \item compilation ( productivity, performance )
-  \item execution ( productivity, performance )
-  \item tracer, dumper and replay for non regression tests ( productivity )
-  \end{itemize}
-
-  \subsubsection{Multi-target Language Generation}
-\label{sec:multitarget}
+allow this automation. Those aspects are multi-target language generation
+(\emph{performance}, \emph{portability}), kernel compilation
+(\emph{productivity}, \emph{performance}), kernel execution
+(\emph{productivity}, \emph{performance}) and last, kernel tracer, dumper and
+replay for non regression tests (\emph{productivity}).
+
+  \subsubsection{Multi-target Language Generation\label{sec:multitarget}}
 Language availability and performance varies between platforms. It is thus
@@
 Compilation of the generated kernels must also be very flexible because HPC
-application developers may encounter platforms with very diverse compilation environments.
-Proprietary and dedicated compilers are common on HPC infrastructures. Thus
-BOAST build system exhibits similar behavior to common build systems. Compilers
-and their compile/build options can be specified at several
-places. The list by increasing order of precedence includes: BOAST configuration
-file, environment variables, and at kernel build time.
+application developers may encounter platforms with very diverse compilation
+environments.  Proprietary and dedicated compilers are common on HPC
+infrastructures. Thus BOAST build system exhibits similar behavior to common
+build systems. Compilers and their compile/build options can be specified at
+several places. The list by increasing order of precedence includes: BOAST
+configuration file, environment variables, and at kernel build time.
 
@@
 available and performance study can include both kernel related parameters and
-compilation related parameters. This behavior is contained in the \emph{CKernel}
-class of BOAST. When instantiating this class, a BOAST \emph{Procedure}
-representing the entry point of the kernel is specified.
+compilation related parameters. This behavior is contained in the
+\emph{CKernel} class of BOAST. When instantiating this class, a BOAST
+\emph{Procedure} representing the entry point of the kernel is specified.
 
@@
 The next logical step is to benchmark the built kernel. BOAST offers a simple
-way to run a kernel that was successfully built without the need to fork another
-process.  A built BOAST \emph{CKernel} exposes a run method that accepts
-arguments corresponding to the BOAST \emph{Procedure} used to instantiate the
-kernel. Arrays must be instances of \emph{NArray} which are numerical arrays
-that use C arrays underneath.
+way to run a kernel that was successfully built without the need to fork
+another process.  A built BOAST \emph{CKernel} exposes a run method that
+accepts arguments corresponding to the BOAST \emph{Procedure} used to
+instantiate the kernel. Arrays must be instances of \emph{NArray} which are
+numerical arrays that use C arrays underneath.
 
@@
 instance, one such information includes the run time of the kernel and it is
-obtained using the system-wide real-time clock. Other probes can be inserted
-at compile-time, if needed.
-BOAST also supports PAPI~\cite{mucci1999papi} to capture hardware performance 
-counters during each kernel execution.
+obtained using the system-wide real-time clock. Other probes can be inserted at
+compile-time, if needed.  BOAST also supports PAPI~\cite{mucci1999papi} to
+capture hardware performance counters during each kernel execution.
 
@@
 
-Hence, the verification and validation of the new port can be
-narrowed down to asserting that both codes apply the same operations
-on the GPU. And the debugging part will consist in understanding what
-diverges. To that purpose, we developed \code{GPUTrace}, inspired by
-\code{strace} and \code{ltrace} tools: \code{GPUTrace} dynamically
-preloads a library between the application and the GPU library, and
-collects the function name, execution range and argument values (input
-\emph{and} output) of the relevant function calls. These information
-are traced in a unified format for all the APIs. This is a custom
-trace format. Listing~\ref{lst:GPUTrace} presents a sample output
-generated during \Specfem tracing.
-
-However, in contrast with \code{strace} and \code{ltrace},
-\code{GPUTrace} has to be \emph{state-full}. Indeed, most of API
-parameters are handles to opaque types. So, in order to generate
-meaningful traces, \code{GPUTrace} gathers information about these
-objects at creation time (handle value, buffer creation size and
-attributes, kernel name and prototype, \etc{}) and re-injects this
+Hence, the verification and validation of the new port can be narrowed down to
+asserting that both codes apply the same operations on the GPU. And the
+debugging part will consist in understanding what diverges. To that purpose, we
+developed \code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
+\code{GPUTrace} dynamically preloads a library between the application and the
+GPU library, and collects the function name, execution range and argument
+values (input \emph{and} output) of the relevant function calls. These
+information are traced in a unified format for all the APIs. This is a custom
+trace format. Listing~\ref{lst:GPUTrace} presents a sample output generated
+during SPECFEM3D tracing.
+
+However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
+to be \emph{state-full}. Indeed, most of API parameters are handles to opaque
+types. So, in order to generate meaningful traces, \code{GPUTrace} gathers
+information about these objects at creation time (handle value, buffer creation
+size and attributes, kernel name and prototype, \etc{}) and re-injects this
 information when the objects used. A state-less implementation of
@@
 section we know that a number of optimizations can have an impact on the
-performance of the kernel on the Mali architecture. But, what is the impact in other
-architectures? And, are there additional optimizations that can impact the
-performance?
+performance of the kernel on the Mali architecture. But, what is the impact in
+other architectures? And, are there additional optimizations that can impact
+the performance?
 
@@
 architectures. The chosen architectures include an Intel i7-2760QM CPU (Sandy
-Bridge architecture) that supports OpenCL 1.2 and a system with an NVIDIA gtx680 GPU that
-supports OpenCL 1.1. We see that the performance ratio between the different
-architectures is stable across image sizes.
+Bridge architecture) that supports OpenCL 1.2 and a system with an NVIDIA
+gtx680 GPU that supports OpenCL 1.1. We see that the performance ratio between
+the different architectures is stable across image sizes.
 
@@
 
-\paragraph{Porting run-time to \OCL} The second part of the port
+\paragraph{Porting run-time to OpenCL} The second part of the port
 consisted in the translation of the CPU-side of the application, from
-\Cuda API to \OCL API. Most of the functions of the interfaces are
+CUDA API to OpenCL API. Most of the functions of the interfaces are
 very similar, with only naming-convention and data-structure
@@
 
-\paragraph{Debugging \OCL Execution: \code{GPUTrace}}
-
-With the help of \code{GPUTrace} (Section~\ref{sec:gputrace}), we
-could confront \Cuda and \OCL execution traces with a graphical
-\code{diff} tool, and spot the different porting mistakes: some
-parameters reversed, offsets incorrectly applied, \etc{}.
+\paragraph{Debugging OpenCL Execution: \code{GPUTrace}}
+
+With the help of \code{GPUTrace} (described in Section~\ref{sec:gputrace}), we
+could confront CUDA and OpenCL execution traces with a graphical \code{diff}
+tool, and spot the different porting mistakes: some parameters reversed,
+offsets incorrectly applied, \etc{}.
 
@@
 missing.
-%As it was the same person who coded the \OCL port and the GPU
+%As it was the same person who coded the OpenCL port and the GPU
 %tracer, this third dimension was forgotten in both codes.
 
-\paragraph{Evaluation} Our \OCL/BOAST port of \Specfem is now
-merged in \Specfem's development tree and under test and extension by
+\paragraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
+merged in SPECFEM3D's development tree and under test and extension by
 different research teams. On a platform with two K40x GPUs and N Intel
 Xeon processors, we measured similar results between the original
-\Cuda version and our BOAST-\Cuda version. With the same set of
-optimization flags, BOAST \Cuda and \OCL version reported similar
-execution time spans. The best execution speed was achieved with \Cuda
-version though (25\% higher than \OCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed, to \OCL run-time, as of version
+CUDA version and our BOAST-CUDA version. With the same set of
+optimization flags, BOAST CUDA and OpenCL version reported similar
+execution time spans. The best execution speed was achieved with CUDA
+version though (25\% higher than OpenCL), as one optimization parameter
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to OpenCL run-time, as of version
 1.1. This parameter, in addition to specifying the work group size (which can
-be done in \OCL), also constrains the number of work group that must run in
-parallel on a multiprocessor. This value is set to 7 in \Cuda. This means that
+be done in OpenCL), also constrains the number of work group that must run in
+parallel on a multiprocessor. This value is set to 7 in CUDA. This means that
 the compiler must be very conservative on register usage in order to allow this
 parallelism which allows better overlapping of communications end computations.
-This functionality is not supported in \OCL.
+This functionality is not supported in OpenCL.
 
@@
 
-\section{Related Work}
+\section{Related Work\label{sec:related}}
 
@@
 
-In this paper we presented the BOAST infrastructure as well as its
-application to three use cases from the Mont-Blanc project. Results are
-encouraging as BOAST proved to be a powerful and flexible tool that allowed
-gains in performance and performance portability.
+In this paper we presented the BOAST infrastructure as well as its application
+to three use cases from the Mont-Blanc project. Results are encouraging as
+BOAST proved to be a powerful and flexible tool that allowed gains in
+performance compared to hand tuned codes. Performance portability of those
+codes is also improved.
 
 Future development will focus on three goals. Interfacing with binary analysis
-tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop to
-guide optimization. Interfacing with search space modellers/pruners in order to
-optimize the search of the optimal version of a kernel. Continue working on
-vector code. For instance producing a collection of small to medium useful
-vector patterns (transposition, for instance) in BOAST could really help develop
-vectorized version of our algorithm.
-
-%\section*{References}
+tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop
+to guide optimization. Interfacing with search space modellers/pruners in order
+to optimize the search of the optimal version of a kernel. Work should also
+continue on improving the support for vector codes. For instance producing a
+collection of small to medium useful vector patterns (transposition, for
+instance) in BOAST could really help users develop vectorized version of their
+algorithm.
+
+\section*{Acknowledgments}
+
+The research leading to these results has received funding from the European
+Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement
+\Num 288777 and 610402.
+
+\section*{References}
 

### 2016-12-27 Paper2/BOAST.tex
Fixed listing numbering.

@@
         label={lst:BOAST_algebra},
-        numbers=right }
+        numbers=left }
 \begin{lstlisting}
@@
         label={lst:BOAST_control},
-        numbers=left }
+        numbers=right }
 \begin{lstlisting}
@@
 
-\section*{References}
+%\section*{References}
 

### 2016-12-28 Paper2/BOAST.tex
section boast

@@
 \newcommand*{\Num}{N\textsuperscript{o}\xspace}
+
+%\usepackage[textwidth=17mm]{todonotes}
+\usepackage{todonotes}
+
+\newcommand{\FD}[2]{\todo[inline, size=\small]{\textbf{FD$\rightarrow$#1}: #2.}}
 
@@
 test application computing kernels~\cite{videau2013boast}.
+
 Figure~\ref{fig:boast_workflow} illustrates the envisioned workflow and program
 structure. The user starts from an application kernel (either designed or
-implemented), and writes it in a dedicated language (step 1).  The language
+implemented), and writes it in a dedicated language (Step 1). The language
 provides enough flexibility for the kernel to be meta-programmed with several
@@
 a target language. Those parameters define the output source code that will be
-generated by BOAST (step 2). The resulting code source is then built according
-to the user specified compiler and options (step 3). If input data are available
+generated by BOAST (Step 2). The resulting code source is then built according
+to the user specified compiler and options (Step 3). If input data are available
 then the kernel can be benchmarked and tested for non regression. Based on the
-results, other optimizations can be selected (step 4) or new optimizations can
+results, other optimizations can be selected (Step 4) or new optimizations can
 be added to the BOAST sources. The process can be repeated until a good
 candidate is found on the target platform. The resulting kernel is then added to
-the program (step 5).
+the program (Step 5).
 
 Several steps in the workflow can be automated: optimization and compiler flags
-exploration, non regression testing, benchmarking... This automation can be
+exploration, non regression testing, benchmarking, \ldots This automation can be
 scripted by the user or by interfacing other dedicated tools. In order
@@
 
+%%% KERNEL DESCRIPTION LANGUAGE %%%
 \subsection{Kernel Description Language}
 
-Usually computing kernels are hotspots of an HPC application, and they are most
+Usually computing kernels are hotspots of an HPC application, and they take most
 of the time based on loop nests. A lot of efforts are dedicated to their tuning
@@
 
+\FD{BV}{Bizarre la suite ici. Ca fait plus ``previous work''}
 The importance of  correct loop imbrication on BLAS~\cite{lawson1979basic}
@@
 \item Arbitrary number of variables have to be created and manipulated (types,
-attributes...).
+attributes, \ldots).
 \item Procedures have to be abstracted (reunite Fortran and C like languages,
-attributes...).
+attributes, \ldots).
 \item Functions must be available.
 \item Variables, constants and functions should be able to be composed in
-complex Expressions.
-\item Basic control structures (for, while, if/else...) have to be abstracted.
-\item Powerful array management (several dimensions, transformations,
-indexing...).
+complex expressions.
+\item Basic control structures (for, while, if/else, , \ldots) have to be abstracted.
+\item Powerful array management features (allowing several dimensions, transformations,
+indexing, \ldots).
 \end{itemize}
 
-In order to manipulate those abstractions we want to have a syntax resembling
+In order to manipulate those abstractions we want to have a syntax similar to 
 what programmers use. For instance, commonly used operators have to be
@@
 results of $a + b$ to $c$  is not equal to $c \leftarrow a + b$, which saves
-the expression $a + b$ to a variable $c$. This is why an embedded domain
-specific language approach was selected~\cite{hudak1996building}. This feature
+the expression $a + b$ to a variable $c$. This is why an embedded Domain
+Specific Language approach was selected~\cite{hudak1996building}. This feature
 allows for the coexistence of two languages: the host language and the Domain
@@
 
+%%% BOAST KEYWORDS %%%
 \subsubsection{BOAST Keywords}
@@
 The \emph{decl} method is used to declare variables or procedures and functions. 
-
 The \emph{pr} method calls the public \emph{pr} method of objects it is
@@
 called. Calling directly the \emph{pr} method of a BOAST object yields the same
-result.
-
-The \emph{opn} method can be used to print the beginning of a control
+result. The \emph{opn} method can be used to print the beginning of a control
 structure without an associated code block.
-
 The \emph{close} method is the counterpart to the \emph{opn} method. It is used
@@
 
+%%% BOAST ABSTRACTIONS %%%
 \subsubsection{BOAST Abstractions}
@@
 
+%%% ALGEBRA %%%
 \paragraph{Algebra}
@@
 produces the Fortran (Listing~\ref{lst:BOAST_algebra_FORTRAN}) and C
-(Listing~\ref{lst:BOAST_algebra_C}) output. At first we define 2 Variables
+(Listing~\ref{lst:BOAST_algebra_C}) output. First we define two variables
 \emph{i} and \emph{k} (note that \emph{k} is $64$ bit integer). The third
@@
 The last control structure is Procedure. It describes either procedures or
-functions. Figure~\ref{fig:BOAST_procedure} presents the use of this
+functions. Code in Figure~\ref{fig:BOAST_procedure} presents the use of this
 abstraction. It illustrates the signature of a real kernel from BigDFT. This
@@
 
+%%% BOAST RUN-TIME %%%
 \subsection{BOAST Run-time}
 
-In the previous Section we presented BOAST's language. This allows us to
+In the previous Section we presented the BOAST's language. This allows us to
 describe procedures and functions and to meta-program them using Ruby.  Each
@@
 
+%%% MULTI-TARGET LANGUAGE GENERATION %%%
 \subsubsection{Multi-target Language Generation\label{sec:multitarget}}
@@
 have additional features, such as languages that target GPUs (OpenCL, CUDA), or
-languages that support threading paradigms (OpenMP). The developer should be
+languages that support \FD{BV}{multi-threading?} threading paradigms (OpenMP). The developer should be
 given tools to determine what kind of language is currently used in order to be
@@
 
+%%% COMPILATION %%%
 \subsubsection{Compilation}
-
 Compilation of the generated kernels must also be very flexible because HPC
@@
 
+%%% EXECUTION %%%
 \subsubsection{Execution}
@@
 
+%%% KERNEL REPLAY %%%
 \subsubsection{Kernel Replay}
@@
 
-\subsection{Non Regression Testing Using Trace Debugging}
-\label{sec:gputrace}
+%%% NON REGRESSION TESTING USING TRACE DEBUGGING %%%
+\subsection{Non Regression Testing Using Trace Debugging\abel{sec:gputrace}}
 
@@
 
-\section{BOAST Use Cases}
-\label{use_cases}
+%%% BOAST USE CASES %%%
+\section{BOAST Use Cases\label{use_cases}}
 
@@
 
+%%% LAPLACE FILTER KERNEL %%%
 \subsection{Laplace Filter Kernel}
@@
 
+%%% OPTIMIZATION SPACE %%%
 \subsubsection{Optimization Space}

### 2016-12-28 Paper2/BOAST.tex
section boast use cases

@@
 performance of the kernel on the Mali architecture. But, what is the impact in
-other architectures? And, are there additional optimizations that can impact
+other architectures? And, are there any additional optimizations that can impact
 the performance?
@@
 
+%%% PERFORMANCE RESULTS %%%
 \subsubsection{Performance Results}
@@
 
+%%% LAPLACE CONCLUSION %%%
 \subsubsection{Laplace Conclusion}
@@
 
-  \subsection{Creating an Auto-Tuned Convolution Library for BigDFT using
-BOAST}
+%%% CREATING AN AUTO-TUNED CONVOLUTION LIBRARY FOR BIGDFT USING BOAST %%%
+\subsection{Creating an Auto-Tuned Convolution Library for BigDFT using BOAST}
 
@@
 
-Despite that the application is mainly written in Fortran, it currently includes
-360 kLOC and 70 kLOC of Fortran and C languages, respectively, accounting for
+Despite that the application is mainly written in Fortran (360 kLOC of Fortran), it 
+currently also includes 70 kLOC of C languages, accounting for
 more than 50\% of the code base.  It is a parallel application based on the
@@
 
+%%% BIGDFT %%%
 \subsubsection{BigDFT}
@@
 
+%%% A GENERIC CONVOLUTION LIBRARY %%%
 \subsubsection{A Generic Convolution Library}
@@
 
+%%% PERFORMANCE REPORT %%%
 \subsubsection{Performance Report}
@@
 
-  \subsection{Porting SPECFEM3D to OpenCL using BOAST}
-  \label{subsec:specfem}
+%%% Porting SPECFEM3D to OpenCL using BOAST %%%
+\subsection{Porting SPECFEM3D to OpenCL using BOAST\label{subsec:specfem}}
 
@@
 
+%%% SPECFEM3D %%%
 \subsubsection{SPECFEM3D}
@@
 
+%%% PORTING TO OPENCL %%%
 \subsubsection{Porting to OpenCL}
 
-\paragraph{Porting kernels to BOAST}
-
-Nvidia CUDA and OpenCL are based on the same programming model: a
+%%%
+\paragraph{Porting kernels to BOAST}  Nvidia CUDA and OpenCL are based on the same programming model: a
 massively parallel accelerator running in disjoint, non addressable
@@
 
+%%%
 \paragraph{Porting run-time to OpenCL} The second part of the port
@@
 
+%%%
 \paragraph{Debugging OpenCL Execution: \code{GPUTrace}} 
-
 With the help of \code{GPUTrace} (described in Section~\ref{sec:gputrace}), we
@@
 
+%%%
 \paragraph{Evaluation} Our OpenCL/BOAST port of SPECFEM3D is now
@@
 
+%%% RELATED WORK %%%
 \section{Related Work\label{sec:related}}
@@
 developments in hardware, and the HPC landscape being as diverse as it is now,
-there is a renewed interest in the field. This related work Section is split in
-three parts, focusing first on Auto-tuning frameworks. Tools that provide a DSL
+there is a renewed interest in the field. This related work section is split in
+three parts, focusing first on auto-tuning frameworks. Tools that provide a DSL
 to describe computing kernels will then be presented. Last, optimization space

### 2016-12-28 Paper2/BOAST.tex
section related work

@@
 
+%%% APPLICATION AUTO-TUNING %%%
 \subsection{Application Auto-Tuning}
@@
 Software} methodology that we implemented with BOAST. Their kernel generation is
-done using macro-functions in C.
+performed using macro-functions in C.
 
@@
 auto-tuning framework that has an approach close of that of BOAST but they are
-based on an annotated DSL describing loop transformation rather than a more
+based on an annotated DSL describing loop transformations rather than a more
 generic EDSL. Halide~\cite{ragan2013halide} is an auto-tuning framework
@@
 LGen~\cite{spampinato2014basic} is a compiler that generates linear algebra
-programs for small fixed size problems. Knowing the problem size it fully
+programs for small fixed size problems. Knowing the problem size, it fully
 unrolls and vectorizes loops, yielding better performance than state of the
@@
 
+%%% KERNEL DESCRIPTION DSL %%%
 \subsection{Kernel Description DSL}
@@
 
-  \subsection{Optimization Space Pruner}
+%%% OPTIMIZATION SPACE PRUNER %%%
+\subsection{Optimization Space Pruners}
 
@@
 
+%%% Conclusion and Future Works %%%
 \section{Conclusion and Future Works}

### 2016-12-28 Paper2/BOAST.tex
section related work
section ack

@@
 %%% NON REGRESSION TESTING USING TRACE DEBUGGING %%%
-\subsection{Non Regression Testing Using Trace Debugging\abel{sec:gputrace}}
+\subsection{Non Regression Testing Using Trace Debugging\label{sec:gputrace}}
 
@@
 The research leading to these results has received funding from the European
-Community's Seventh Framework Programme [FP7/2007-2013] under grant agreement
+Community's Seventh Framework Programme [FP7/2007-2013] under grant agreements
 \Num 288777 and 610402.

### 2016-12-31 Paper2/BOAST.tex
conclusion

@@
 
-%%% Conclusion and Future Works %%%
+%%% CONCLUSION AND FUTURE WORKS %%%
 \section{Conclusion and Future Works}
 
-In this paper we presented the BOAST infrastructure as well as its application
-to three use cases from the Mont-Blanc project. Results are encouraging as
-BOAST proved to be a powerful and flexible tool that allowed gains in
-performance compared to hand tuned codes. Performance portability of those
-codes is also improved.
-
-Future development will focus on three goals. Interfacing with binary analysis
-tools like MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop
-to guide optimization. Interfacing with search space modellers/pruners in order
-to optimize the search of the optimal version of a kernel. Work should also
-continue on improving the support for vector codes. For instance producing a
-collection of small to medium useful vector patterns (transposition, for
-instance) in BOAST could really help users develop vectorized version of their
-algorithm.
-
+Application portability is an important issue that should be solved efficiently, especially given the large number of different processors now
+available for today's supercomputers. The work needed to get performance portability is a tedious task, even for experienced programmers. The
+availability of semi-automatic tools is therefore mandatory for the development of large simulation applications. Computing kernels' identification
+and optimization has to be carefully performed as they usually consume most of the computing resources.
+
+In this paper we presented the BOAST infrastructure (DLS and runtime) that aims at describing kernels in a high level language and allows the
+comparison of the performance of different versions of the code in a simple and seamless way. We described its application to three use cases from the
+Mont-Blanc project. Results are encouraging as BOAST proved to be a powerful and flexible tool that allowed gains in performance compared to hand
+tuned codes. Performance portability of those codes is also improved.
+
+Future development will focus on three main goals. First we find interesting to interface it with binary analysis tools like
+MAQAO~\cite{djoudi2005exploring} in order to build a feedback loop to guide optimization. On the autotuning side, interfacing with search space
+modellers/pruners in order to optimize the search of the optimal version of a kernel will allow us to gain some time in the optimization
+process. Finally, work should also continue on improving the support for vector codes. For instance producing a collection of small to medium useful
+vector patterns (transposition, for instance) in BOAST could really help users develop vectorized version of their algorithm. Other computing kernels
+and applications will also be ported on various architectures.
+
+%%% ACKNOWLEDGMENTS %%%
 \section*{Acknowledgments}

### 2016-12-31 Paper2/BOAST.tex
Added BEAST reference.

@@
 
+BEAST~\cite{CPE:CPE3516} is an autotuning framework that uses macro-processing
+in source files. Inside the macros, BEAST offers a DSL that can be used to
+specify values derived from a set of iterators. Those iterators explore the
+search space and can be constrained. Great care has been used to ensure that
+the iterators enumeration as well as the pruning are as efficient as possible.
+
 Heterogeneous Programming Library~\cite{vinas2013exploiting} (HPL) is maybe

### 2016-12-31 Paper2/BOAST.tex
Minor corrections.

@@
 or loop interchange~\cite{allen1984automatic} (loop variables change places).
-
-\FD{BV}{Bizarre la suite ici. Ca fait plus ``previous work''}
-The importance of  correct loop imbrication on BLAS~\cite{lawson1979basic}
+The importance of  correct loop imbrication on BLAS~\cite{lawson1979basic} %\FD{BV}{Bizarre la suite ici. Ca fait plus ``previous work''}
 operations is studied in~\cite{soliman2009performance}, and shows performance
@@
 complex expressions.
-\item Basic control structures (for, while, if/else, , \ldots) have to be abstracted.
+\item Basic control structures (for, while, if/else, \ldots) have to be abstracted.
 \item Powerful array management features (allowing several dimensions, transformations,
@@
 have additional features, such as languages that target GPUs (OpenCL, CUDA), or
-languages that support \FD{BV}{multi-threading?} threading paradigms (OpenMP). The developer should be
+languages that support multi-threading paradigms (OpenMP). The developer should be
 given tools to determine what kind of language is currently used in order to be

### 2017-03-09 Paper2/BOAST.tex
Modified affiliations.

@@
 \author{
-   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{2}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
+   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{6}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{2}Centre National de la Recherche Scientifique}\\
-   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Joseph Fourier}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Grenoble Alpes}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/SP2M/L\_Sim}\\
-   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}
+   \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{6}Aix Marseille Univ., CNRS, Centrale Marseille, LMA, Marseille, France.}
 }

### 2017-03-28 hou10ni/Tutorial/hou10ni_profile.tex
Update hou10ni's kernel

\documentclass[10pt,twoside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{a4paper, bottom=2.5cm, top=2cm, left=2cm, right=2cm}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx,subfigure, array}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex	
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes}
\usepackage{listings}
\usepackage{color}
\usepackage{hyperref,url}
\usepackage[hang,  bf, margin=19pt, tableposition=top,textfont=it]{caption}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{multicol,multirow}
\usepackage{subfigure}
\usepackage{fancyvrb,newverbs,xcolor}

% COMMANDS
\newcommand{\N}{\mathbb{N}} 
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{property}{Property}
\newtheorem{proof}{Proof}

\newlength{\textlarg}
\newcommand{\barre}[1]{%
   \settowidth{\textlarg}{#1}
   #1\hspace{-\textlarg}\rule[0.5ex]{\textlarg}{0.5pt}}

\algnewcommand\algorithmicswitch{\textbf{switch}}
\algnewcommand\algorithmiccase{\textbf{case}}
\algnewcommand\algorithmicassert{\texttt{break}}
\algnewcommand\Break[1]{\State \algorithmicbreak(#1)}%
% New "environments"
\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
\algtext*{EndSwitch}%
\algtext*{EndCase}%

% COLORS
\definecolor{bleugris}{RGB}{43,78,103}
\definecolor{codegreen}{rgb}{0.56,0.74,0.56}
\definecolor{backcolor}{rgb}{0.95,0.95,0.96}
\definecolor{lightgray}{gray}{0.35}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codered}{rgb}{0.87,0.19,0.39}
\definecolor{cverbbg}{gray}{0.93}

\newenvironment{cverbatim}
 {\SaveVerbatim{cverb}}
 {\endSaveVerbatim
  \flushleft\fboxrule=0pt\fboxsep=.5em
  \colorbox{cverbbg}{\BUseVerbatim{cverb}}%
  \endflushleft
}
\newenvironment{lcverbatim}
 {\SaveVerbatim{cverb}}
 {\endSaveVerbatim
  \flushleft\fboxrule=0pt\fboxsep=.5em
  \colorbox{cverbbg}{%
    \makebox[\dimexpr\linewidth-2\fboxsep][l]{\BUseVerbatim{cverb}}%
  }
  \endflushleft
}

\newcommand{\ctexttt}[1]{\colorbox{cverbbg}{\texttt{#1}}}
\newverbcommand{\cverb}
  {\setbox\verbbox\hbox\bgroup}
  {\egroup\colorbox{cverbbg}{\box\verbbox}}

\lstdefinestyle{mystyle}{
	%backgroundcolor=\color{backcolor},
	commentstyle=\color{codegreen},
	keywordstyle=\color{codered},
	numberstyle=\tiny\color{codegray},
	basicstyle=\scriptsize,
	breaklines=true,
	breakatwhitespace=false,
	numbers=left,
	tabsize=2
}

\lstset{style=mystyle}

\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox[cmyk]{0.43, 0.35, 0.35,0.01}{\parbox{\textwidth}{\hspace{15pt}#3}}}
\DeclareCaptionFormat{listing2}{\colorbox[cmyk]{0.9, 0.9, 0.9,0.01}{\parbox{\textwidth}{\hspace{15pt}#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white,singlelinecheck=false,margin=0pt,font={bf}}

\title{Hou10ni Profiling with TAU}
\date{}

\begin{document}
\maketitle

This document presents some results obtained with the performance evaluation tool TAU on the application hou10ni.

\section{How to use TAU on hou10ni}

\begin{cverbatim}
export TAU_MAKEFILE=$TAU_DIR/Makefile.
export TAU_OPTIONS='-optPdtF95Opts="-R free" -optVerbose'

cd build
cmake ..
ccmake ..
\end{cverbatim}

\noindent $\rightarrow$ Press "t" to change the compiler and chose \texttt{tau\_f90.sh}. Then you can tape \texttt{make} and launch the application.

\begin{cverbatim}
make
cd .. 
mpirun -np 4 tau_exec -T mpi  build/hou10ni.out < HPC4E.dat
\end{cverbatim}

\noindent To visualize the results with paraprof, 
\begin{cverbatim}
paraprof &
\end{cverbatim}
or
\begin{cverbatim}
paraprof --pack app.ppk
\end{cverbatim}
This last command will create the file \texttt{app.ppk}.

\section{Trace Visualisation with Jumpshot tool}

Jumpshot user guide: \url{ftp://ftp.mcs.anl.gov/pub/mpi/slog2/js4-usersguide.pdf}\\

\noindent To generate a trace of an application, follow the steps presented below.

\begin{cverbatim}
export TAU_TRACE=1
mpirun -np 4 tau_exec -T mpi  build/hou10ni.out < HPC4E.dat
tau_treemerge.pl
\end{cverbatim}

\noindent $\rightarrow$ Two files are created:  tau.trc and tau.edf

\begin{cverbatim}
tau2slog2 tau.trc tau.edf  -o app.slog2
jumpshot app.slog2
\end{cverbatim}

Figure \ref{fig:trace} shows the trace obtained with 4 MPI processes, TAU instrumentation and displayed with the Jumpshot tool. Figure \ref{fig:trace2} presents a zoom on the big portion and Figure \ref{fig:trace3} depicts the end of the trace. The purple/pink color represents the \texttt{MPI\_Wait} function. In Figure \ref{fig:trace3} we can see the different time spent in \texttt{MPI\_Wait} for all processes.

\begin{figure}[ht!]
\centering
\includegraphics[width=13cm]{IMAGES/jumpshot1}
\caption{Trace Visualisation with Jumpshot}
\label{fig:trace}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=13cm]{IMAGES/jumpshot2}
\caption{Zoom on the big portion in Figure \ref{fig:trace}}
\label{fig:trace2}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[width=15cm]{IMAGES/jumpshot3}
\caption{Zoom on the end of the trace}
\label{fig:trace3}
\end{figure}

%\begin{figure}[ht!]
%\centering
%\includegraphics[width=10cm]{IMAGES/jumpshot4}
%\caption{Zoom on the beginning of the trace}
%\label{fig:trace3}
%\end{figure}

\section{Profile Visualisation with paraprof}

It is possible to get a detailed summary in a text format. To record all information in a file \textit{profile\_summary.txt}, the command is
\begin{cverbatim}
pprof > profile_summary.txt
\end{cverbatim}

\noindent The file created looks like the following

{\scriptsize
\begin{verbatim}
Reading Profile files in profile.*

NODE 0;CONTEXT 0;THREAD 0:
---------------------------------------------------------------------------------------
%Time    Exclusive    Inclusive       #Call      #Subrs  Inclusive Name
              msec   total msec                          usec/call
---------------------------------------------------------------------------------------
 72.9            0     2:02.995       12300           0      10000 .TAU application => [CONTEXT] .TAU application
 72.9            0     2:02.995       12300           0      10000 [CONTEXT] .TAU application
 54.0     1:31.155     1:31.155        9116           0      10000 [SAMPLE] UNRESOLVED /lib64/libpthread-2.11.3.so
 54.0     1:31.122     1:31.122        9113           0       9999 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 23.3       39,314       39,314        1644           0      23914 MPI_Wait()
 23.3            0       39,300        3929           0      10003 MPI_Wait() => [CONTEXT] MPI_Wait()
 23.3            0       39,300        3929           0      10003 [CONTEXT] MPI_Wait()
 14.5       24,542       24,542        2454           0      10001 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 14.5       24,542       24,542        2454           0      10001 [SAMPLE] UNRESOLVED /home/esaillard/hou10ni2d/build/hou10ni.out
 10.7       18,132       18,132        1813           0      10002 [SAMPLE] MPID_nem_mpich_blocking_recv
 10.3       17,368       17,368        1737           0       9999 MPI_Wait() => [CONTEXT] MPI_Wait() => [SAMPLE] MPID_nem_mpich_blocking_recv
....

USER EVENTS Profile :NODE 0, CONTEXT 0, THREAD 0
---------------------------------------------------------------------------------------
NumSamples   MaxValue   MinValue  MeanValue  Std. Dev.  Event Name
---------------------------------------------------------------------------------------
         4          8          4          6          2  Message size for all-reduce
        76  1.375E+07          4  6.716E+05  2.489E+06  Message size for broadcast
      2022  5.426E+04          4  1.111E+04  1.929E+04  Message size received from all nodes
       822  5.426E+04       1024  2.732E+04  2.174E+04  Message size received in wait

....

NODE 1;CONTEXT 0;THREAD 0:
---------------------------------------------------------------------------------------
%Time    Exclusive    Inclusive       #Call      #Subrs  Inclusive Name
              msec   total msec                          usec/call
---------------------------------------------------------------------------------------
 93.7            0     2:38.072       15805           0      10001 .TAU application => [CONTEXT] .TAU application
 93.7            0     2:38.072       15805           0      10001 [CONTEXT] .TAU application
 61.9     1:44.423     1:44.423       10444           0       9998 [SAMPLE] UNRESOLVED /lib64/libpthread-2.11.3.so
 61.9     1:44.405     1:44.405       10443           0       9998 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 31.7       53,555       53,555        5351           0      10008 .TAU application => [CONTEXT] .TAU application => [SAMPLE] UNRESOLVED ..
 31.7       53,555       53,555        5351           0      10008 [SAMPLE] UNRESOLVED /home/esaillard/hou10ni2d/build/hou10ni.out
  4.2        7,074        7,074         424           0      16685 MPI_Recv()
  4.2            0        7,071         707           0      10002 MPI_Recv() => [CONTEXT] MPI_Recv()
  4.2            0        7,071         707           0      10002 [CONTEXT] MPI_Recv()
  3.0        5,135        5,135         516           0       9953 [SAMPLE] MPID_nem_mpich_blocking_recv
  2.3        3,807        3,807         380           0      10021 MPI_Recv() => [CONTEXT] MPI_Recv() => [SAMPLE] MPID_nem_mpich_blocking_recv
  1.5            0        2,550         255           0      10004 MPI_Bcast() => [CONTEXT] MPI_Bcast()
  1.5            0        2,550         255           0      10004 [CONTEXT] MPI_Bcast()
  1.5        2,543        2,543          76           0      33468 MPI_Bcast()

....
and so on
\end{verbatim}
}

\noindent Figures \ref{fig:1} and \ref{fig:2} show the entire application time with paraprof.

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot2}
\caption{Paraprof main window with stack bars together}
\label{fig:1}
\end{figure}

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot2bis}
\caption{Paraprof main window }
\label{fig:2}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{IMAGES/ScreenShotWait}
\caption{Time in one of MPI\_Wait function}
\label{fig:wait}
\end{figure}

\noindent {\bf{Statistics per process}} \\

\begin{figure}[H]
\includegraphics[width=18cm]{IMAGES/ScreenShot5}
\caption{Statistics for processes 0 and 1}
\label{fig:5}
\end{figure}

\noindent {\bf{Comparison Window}} \\

To get a comparison window, click right on the text \textit{node 0} for example and chose \textit{"Add Thread to Comparison Window"}.

\begin{figure}[H]
\includegraphics[width=14cm]{IMAGES/ScreenShot3}
\caption{Comparison window}
\label{fig:3}
\end{figure}

\noindent {\bf{3D Visualization}} \\

\begin{figure}[H]
\centering
\includegraphics[width=10cm]{IMAGES/ScreenShot4}
\caption{3D Visualization}
\label{fig:4}
\end{figure}

A complete manual to use TAU and paraprof is available at \url{https://www.cs.uoregon.edu/research/tau/tau-usersguide.pdf}.

\section{Trace Analysis with Lucas and Arnaud's tool}

To use the tool, we first need to convert TAU traces to PAJE traces\footnote{For a tutorial on how to install tau2paje: https://github.com/schnorr/akypuera/wiki/TAUWithAkypuera}. After TAU generates tau.trc and tau.edf files, use tau2paje to generate tau.paje with the following command

\begin{cverbatim}
tau2paje --no-links --only-mpi --normalize-mpi tau.trc tau.edf > tau.paje
\end{cverbatim}

Then we use pj\_dump\footnote{For a tutorial on how to install pj dump: https://github.com/schnorr/pajeng/wiki/Install} to transform PAJE traces to CSV:

\begin{cverbatim}
pj_dump tau.paje | grep -e ^State -e ^Event > tau.csv
\end{cverbatim}

To check the head of the file:

\begin{cverbatim}
head tau.csv
\end{cverbatim}

\subsection*{Analysis of the CSV using R+dplyr+ggplot2}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :session *R* :exports both  
library(dplyr);
df <- read.csv("tau.csv", header=FALSE, strip.white=TRUE);
# get only lines starting with State
df <- df %>% filter(V1 == "State");
names(df) <- c("Type", "Thread", "Type2", "Start", "End", "Duration", "Imbric", "Region");
df <- df %>% select(-Type, -Type2, -Duration, -Imbric);
df$Thread <- as.integer(gsub("rank", "", as.character(df$Thread)));
head(df);
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
#+RESULTS:
:   Thread    Start      End        Region
: 1      3 0.013019 0.037004      MPI_Init
: 2      3 0.038628 0.038630 MPI_Comm_size
: 3      3 0.038637 0.038637 MPI_Comm_rank
: 4      3 0.041826 0.046533     MPI_Bcast
: 5      3 0.046534 0.046543     MPI_Bcast
: 6      3 0.046546 0.046550     MPI_Bcast
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :session *R* :exports both
library(ggplot2)
tstart = min(df$Start);
tend = max(df$End);

df %>%
  ggplot() + theme_bw(base_size=16) + xlab("Time [s]") +
  ylab("Thread") +
  theme(
     plot.margin = unit(c(0,0,0,0), "cm"),
     legend.margin = unit(.1, "line"),
     panel.grid = element_blank(),
     legend.position = "bottom",
     legend.title = element_blank()
   ) +
    coord_cartesian(xlim=c(tstart,tend)) +
    guides(fill = guide_legend(nrow = 2)) +
    geom_rect(alpha=1, aes(fill=Region,
                           xmin=Start,
                           xmax=End,
                           ymin=Thread,
                           ymax=Thread + 0.9))

tspan = tend - tstart;
tspan;
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
\end{lstlisting}
\vspace{-.5cm}

\begin{figure}[H]
\centering
\includegraphics[width=13cm]{IMAGES/tool1}
\caption{Trace}
\label{fig:5}
\end{figure}

\lstset{language=R}
\begin{lstlisting}[caption=Code in R that analyses the CVS file (written in emacs using org mode)]
#+begin_src R :results output :exports both :session *R* 
df %>%
  filter(Start >= tstart, End <= tend) %>%
  group_by(Thread, Region) %>%
  summarize(count = length(End), TDur = sum(End-Start)/tspan*100) %>%
  group_by(Thread) %>%
  summarize(t=sum(TDur)) %>%
  as.data.frame();
#+end_src
\end{lstlisting}

\lstset{language=R}
\begin{lstlisting}[caption=Results]
#+RESULTS:
:   Thread         t
: 1      0  3.221339
: 2      1 11.986336
: 3      2  6.962705
: 4      3 10.838378
\end{lstlisting}

\section{Impact of input parameters}

The results presented in the previous section have been obtained with the following input parameters (file HPC4E.txt):

\begin{cverbatim}
3 Dimension
"MAILLAGES/florian3D" Nom du maillage
.H (.dat or .model)
0 p-adaptivity (no=0)
1 Ordre des elements
0 transient
0.01 Simulation Time
0.01 Time Step for the snapshots
0 HDG formulation (1=yes, 0=no)
10 Plane wave ...
1
1
40 30 50
430 30 50
680 1380 50
1
2
0 Do you want curved elements
1 Do you want to use mumps with symmetry
0 seismograms 
1 Type de CLA (1 pour du/dn=iomega u)
.false.
0 type RTM
0 Analytic Solution (0=no)
\end{cverbatim}

\subsection{Scalability}

To show the impact of input parameters, we can change the elements order in HPC4E.txt (\textit{Ordre des elements}) or refine the mesh (\texttt{tail MAILLAGES/florian3D.1.ele} will show the command). Beware of the total time, it takes into account the time spent to write in a file (time in function \texttt{sub\_time\_loop}).

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements, {\bf{Simulation time=0.001}} (Time Step=0.001) - {\bf{Sans l'ecriture des instantannees}} - Beware!! I wasn't alone on idchire}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&  min 	& max	 & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				&  .7407	& .760 	& .746									&  	143.738 & 143.786	& 143.758 \\ \hline
 		8		                 				& .464	& .575  	& .512									&  	182.481 & 182.598 	& 182.554	\\ \hline		
 		16		                 				&  .142	& .174	 & .152									&  	215.348 & 215.398 	&  215.365\\ \hline
		 32		                 				& 1.288	&  5.884  & 4.011		  						        &  	293.542 &  293.639	& 293.619	\\ \hline
		64		                 				& 1.466	& 2.401 & 1.972									&  	360.580 & 360.936	& 360.886\\ \hline
\end{tabular}
}
\label{tab:1}
\end{table}

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements,  {\bf{Simulation time=0.01}} (Time Step=0.01)  - {\bf{Sans l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				& 5.504 & 5.543 & 5.525										& 156.409 & 156.481 & 156.438 	\\ \hline
 		8		                 				& 2.652  &  2.686 &  2.665									&   175.770  &  175.813 &  175.786	\\ \hline
 		16		                 				&  1.612 & 1.660  & 1.631 									& 217.008 &  217.055 &  217.026	\\ \hline
		 32		                 				& .924 & .948   & .940										& 292.669  &  292.749 & 292.730	\\ \hline
		64		                 				& .430 & .444  & .438										&  356.742  & 357.0359 &  357.001	\\ \hline
\end{tabular}
}
\label{tab:2}
\end{table}

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order =1, 859247 elements,    {\bf{Simulation time=0.1}} (Time Step=0.1) - {\bf{Sans l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Number of MPI processes}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg									& 	min & max & avg			\\ \hline \hline
 		4		                 				& 55.325	& 55.344 & 55.333									&  194.500 & 194.522 & 194.512	\\ \hline
 		8		                 				& 26.804	& 26.843  & 26.820									&  200.059 & 200.105 & 200.083	\\ \hline
 		16		                 				&  17.671	&  17.710 & 17.688									&  233.131 & 233.162 &  233.145	\\ \hline
		 32		                 				& 9.779 & 9.803  & 9.794										&  301.450 & 301.533 & 301.514	\\ \hline
		64		                 				& 4.561 & 4.576  & 4.570										&  363.928 & 364.221 & 364.177	\\ \hline
\end{tabular}
}
\label{tab:3}
\end{table}

\subsection{Elements Order Variation}

Table \ref{tab:5} shows the total time and the time spent in function \texttt{sub\_loop\_time} for different elements orders, with 4 MPI processes. The times are obtained with the fortran function \texttt{CPU\_TIME}\footnote{Returns the CPU time, in seconds. If CPU\_TIME is used without any setting, the function returns the total user and system time for the current process.}. 
%The total execution time is obtained with \texttt{usr/bin/time}\footnote{Returns the elapsed real time between invocation and termination. The command is used with the -p option and the output format \%e (=elapsed real time in seconds).}.

\begin{table}[ht!]
\centering
\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: 859247 elements,  Simulation time=0.001 (Time Step=0.001) - {\bf{Sans l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Elements order}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg							& 	min & max & avg			\\ \hline \hline
 		1		                 				& 1.273 & 1.374 & 1.298								& 156.247  & 156.435 &  156.309	\\ \hline
 		2		                 				& 3.384	&  3.390  & 3.388							&  171.825 & 171.855 & 171.843		\\ \hline
 		3		                 				& 59.873 	 &  61.902	& 61.060							&  578.569 & 579.954 & 579.564 	\\ \hline
		 4		                 				& 39.003	& 39.288  & 39.156							& 527.505  & 527.555 & 527.528		\\ \hline
		5		                 				& 	&  &											&  	 & 	& 		\\ \hline
\end{tabular}
}
\label{tab:5}
\end{table}

\subsection{Mesh refinement}

To refine the mesh, we launch the following command

\begin{cverbatim}
./tetgen -kna10000A florian3D.poly
\end{cverbatim}

\noindent and change the number in the option "-kna10000A".

Table \ref{tab:6} shows the total time and the time spent in function \texttt{sub\_loop\_time} for different mesh refinement, with 8 MPI processes.

\begin{table}[ht!]
\centering
\caption{TODO: Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order=1, Simulation time=..}
{\small
\begin{tabular}{|c|c|c|c|c|c|c|} \hline 
\multirow{2}{*}{ {{\bf{Mesh refinement}}} }	& \multicolumn{3}{c|}{ {{\bf{Time spent in \texttt{sub\_time\_loop}  }}} }	& \multicolumn{3}{c|}{  {{\bf{Total execution time}}} }\\ 
										&   		min & max & avg							& 	min & max & avg			\\ \hline \hline
 		10\ 000		                 				& 29.832	& 31.821 & 								&  	 & 	&  173.77	\\ \hline
 		5\ 000		                 				& 62.568	&  66.309  &	 							&  	 & 	 & 351.35	\\ \hline
 		2\ 500		                 				&  141.110 &  147.306	&							&  	& &  735.548	\\ \hline
\end{tabular}
}
\label{tab:6}
\end{table}

\section{Port hou10ni's kernel with BOAST}

This section presents how to install BOAST and use it to port Houdini's kernel.

\subsection{How to install BOAST}

\begin{cverbatim}
git clone https://github.com/Nanosim-LIG/boast.git
cd boast
gem build BOAST.gemspec
gem install --user-install BOAST-1.3.5.gem 
\end{cverbatim}

\subsection{Original version}

The kernel is the loop over the fluid inner elements in \texttt{lib/libtransient/sub\_time\_loop.f90}. The loop is written outside of the code, in a new subroutine as presented below.

\lstset{language=FORTRAN}
\begin{lstlisting}[caption=kernel.f90]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

  integer, intent(in) :: Nflu_inner
  integer, intent(in) :: Nflusol_inner
  integer, intent(in) :: nb_rhs
  integer, intent(in) :: afluSize
  integer, intent(in) :: pSize
  integer, intent(in) :: vecSize_1
  integer, intent(in) :: vecSize_2
  integer, intent(in) :: matSize
  INTEGER, INTENT(in), DIMENSION(vecSize_1,vecSize_2) :: idx_vec_flu
  INTEGER, INTENT(in), DIMENSION(matSize) :: idx_mat_flu
  REAL(kind=dp), INTENT(inout), DIMENSION(pSize,nb_rhs) :: P_new
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_inter
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_old
  REAL(kind=dp), INTENT(in), DIMENSION(afluSize) :: A_flu

  INTEGER :: I,J,I_tmp1,I_tmp2, I1, I2, I3, I4
  REAL(kind=dp), DIMENSION(20,nb_rhs) :: P_aux

  DO I=1,Nflu_inner+Nflusol_inner
    I_tmp1=1
    DO J=1,idx_vec_flu(1,I)
      I1=idx_vec_flu(2*J,I)
      I2=idx_vec_flu(2*J+1,I)
      I_tmp2=I_tmp1+I2-I1
      P_aux(I_tmp1:I_tmp2,1:nb_rhs)=P_inter(I1:I2,1:nb_rhs)
      I_tmp1=I_tmp2+1
    END DO
    I1=idx_vec_flu(2,I)
    I2=idx_vec_flu(3,I)
    I3=idx_mat_flu(I)
    I4=idx_mat_flu(I+1)-1

   P_new(I1:I2,1:nb_rhs)=MATMUL(RESHAPE(A_flu(I3:I4), (/I2-I1+1, I_tmp2/)),P_aux(1:I_tmp2,1:nb_rhs))-P_old(I1:I2,1:nb_rhs)
  END DO

END SUBROUTINE fluid_inner_elt_ref
\end{lstlisting}

\begin{lstlisting}[caption=sub\_time\_loop.f90]
SUBROUTINE sub_time_loop
...
CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
...
END SUBROUTINE 
\end{lstlisting}

To have an idea of the time spent in the kernel, table \ref{tab:7} shows the time of the entire function \texttt{sub\_time\_loop} (without the writing at the end) compared to the time spent in the kernel. Simulation time=0.01.

\begin{table}[H]
\centering
\caption{Time spent in the kernel (loop over fluid elements in {\tt sub\_time\_loop}), only one run - with an allocation at each step - {\bf{Avec l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|} \hline 
 {{\bf{Process ID}}} 	& {{\bf{Time spent in \texttt{sub\_time\_loop} }}}	& {{\bf{Time spent in \texttt{fluid\_inner\_elt()} }}}\\ \hline \hline
		0		&   5.388336							&  5.1683220	(95,9\%)		\\ \hline 
 		1		&  5.352341							&  5.108320	(95,4\%)		\\ \hline
 		2		 &  5.360336							&  5.1203189	(95,5\%)		\\ \hline
		3		 &  5.412338 							&  5.1843260	(95,8\%)		\\ \hline	
\end{tabular}
}
\label{tab:7}
\end{table}

\begin{table}[H]
\centering
\caption{Time spent in the kernel (loop over fluid elements in {\tt sub\_time\_loop}), only one run - without allocation at each step - {\bf{Avec l'ecriture des instantannees}}}
{\small
\begin{tabular}{|c|c|c|} \hline 
 {{\bf{Process ID}}} 	& {{\bf{Time spent in \texttt{sub\_time\_loop} }}}	& {{\bf{Time spent in \texttt{fluid\_inner\_elt()} }}}\\ \hline \hline
		0		&   4.776291							&  4.54828700	(95,2\%)		\\ \hline 
 		1		&  4.74029							&  4.5162959	(95,3\%)		\\ \hline
 		2		 &  4.744293							&  4.5242530	(95,4\%)		\\ \hline
		3		 & 4.768295 							&  4.53228600	(95\%)		\\ \hline	
\end{tabular}
}
\label{tab:8}
\end{table}

\subsection{Reference version}

We generate the exact same kernel as a reference version and run it with BOAST in order to compare the results in term of correctness and performances. KRef defines the reference version. This version simply need to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST.

\lstset{language=RUBY}
\begin{lstlisting}[caption=KRef.rb]
class KRef
 attr_reader :kernel

 def initialize(options)
  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
  @opts.update(options)
  # Parameters 
  @Nflu_inner    = Int("Nflu_inner", :dir => :in )
  @Nflusol_inner = Int("Nflusol_inner", :dir => :in )
  @nb_rhs        = Int("nb_rhs", :dir => :in )
  @afluSize      = Int("afluSize", :dir => :in )
  @pSize         = Int("pSize", :dir => :in )
  @vecSize_1     = Int("vecSize_1", :dir => :in )
  @vecSize_2     = Int("vecSize_2", :dir => :in )
  @matSize       = Int("matSize", :dir => :in )
  @idx_vec_flu   = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
  @idx_mat_flu   = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
  @P_new         = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_inter       = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @P_old         = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
  @A_flu         = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
 end

 def generate
  codeF = File::read("./kernel.f90")

  push_env(:lang => FORTRAN) {
     @kernel = CKernel:: new()
     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
     get_output.print codeF
  }
   return @kernel
 end

end
\end{lstlisting}

The file run.rb is used to launch the kernel.

\begin{lstlisting}[caption=run.rb]
require 'BOAST'
require './KRef.rb'
require 'narray_ffi'

include BOAST

class Experiment

 def self.run()
  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}

  set_lang(FORTRAN)
  kernels={}

  # Creating ref kernel
  kernels[k_ref_params] = KRef::new(k_ref_params)
  kernels[k_ref_params].generate
 # Building ref kernel
  puts kernels[k_ref_params].kernel
  kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )

  # Parameters for loop iteration 1 and MPI Process 0 (in the directory fluid_inner_elt_ref/sub_time_loop_1/)
  inputs = kernels[k_ref_params].kernel.load_ref_inputs()
  puts "** inputs =", inputs

  inputs.each_key { |key|
    puts key
    puts kernels[k_ref_params].kernel.run(*(inputs[key])).inspect
    puts kernels[k_ref_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
  }

  return kernels
 end
end

Experiment.run
\end{lstlisting}

\begin{lstlisting}[caption=Results (obtained with the command ruby run.rb)]
SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
&ecSize_1, vecSize_2, matSize)
  implicit none
  INTEGER, PARAMETER :: dq=8 !< quadruple precision
  INTEGER, PARAMETER :: dp=4  !< double precision

  integer, intent(in) :: Nflu_inner
  integer, intent(in) :: Nflusol_inner
  integer, intent(in) :: nb_rhs
  integer, intent(in) :: afluSize
  integer, intent(in) :: pSize
  integer, intent(in) :: vecSize_1
  integer, intent(in) :: vecSize_2
  integer, intent(in) :: matSize
  INTEGER, INTENT(in), DIMENSION(vecSize_1,vecSize_2) :: idx_vec_flu
  INTEGER, INTENT(in), DIMENSION(matSize) :: idx_mat_flu
  REAL(kind=dp), INTENT(inout), DIMENSION(pSize,nb_rhs) :: P_new
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_inter
  REAL(kind=dp), INTENT(in), DIMENSION(pSize,nb_rhs) :: P_old
  REAL(kind=dp), INTENT(in), DIMENSION(afluSize) :: A_flu

  !! Variables declaration
  INTEGER :: I,J,I_tmp1,I_tmp2, I1, I2, I3, I4
  REAL(kind=dp), DIMENSION(20,nb_rhs) :: P_aux

  DO I=1,Nflu_inner+Nflusol_inner
    I_tmp1=1
    DO J=1,idx_vec_flu(1,I)
      I1=idx_vec_flu(2*J,I)
      I2=idx_vec_flu(2*J+1,I)
      I_tmp2=I_tmp1+I2-I1
      P_aux(I_tmp1:I_tmp2,1:nb_rhs)=P_inter(I1:I2,1:nb_rhs)
      I_tmp1=I_tmp2+1
    END DO
    I1=idx_vec_flu(2,I)
    I2=idx_vec_flu(3,I)
    I3=idx_mat_flu(I)
    I4=idx_mat_flu(I+1)-1

   P_new(I1:I2,1:nb_rhs)=MATMUL(RESHAPE(A_flu(I3:I4), (/I2-I1+1, I_tmp&
&2/)),P_aux(1:I_tmp2,1:nb_rhs))-P_old(I1:I2,1:nb_rhs)
  END DO

END SUBROUTINE fluid_inner_elt_ref

** inputs =
{"./fluid_inner_elt_ref/sub_time_loop_1"=>[210025, 0, 1, NArray.int(2577732): 
[ 5, 1, 4, 538597, 538600, 319233, 319236, 363785, 363788, 665493, ... ], NArray.int(214812): 
[ 1, 81, 161, 209, 289, 369, 449, 529, 609, 689, 769, 849, 913, 993, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(439866): 
[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(8476280): 
[ 1.1026e-18, 9.61982e-17, 1.0165, 9.2055e-17, 9.13418e-19, 9.01442e-17, ... ], 16952560, 879732, 12, 214811, 214812]}

{:duration=>0.153594073}
\end{lstlisting}

\subsection{BOAST version}

TODO

\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
TODO

\end{document}

### 2017-03-28 hou10ni/KRef.rb

@@ -1,4 +1,4 @@
-class KOssRef
+class KRef
  attr_reader :kernel
 
  def initialize(options)
@@ -16,12 +16,12 @@ class KOssRef
   @vecSize_2          = Int("vecSize_2", :dir => :in )
   @matSize            = Int("matSize", :dir => :in )
 
- @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [@vecSize_1,@vecSize_2] )
- @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [@matSize] )
- @P_new              = Real("P_new",:dir => :inout, :dim => [@pSize,@nb_rhs])
- @P_inter            = Real("P_inter",:dir => :in, :dim => [@pSize,@nb_rhs])
- @P_old              = Real("P_old",:dir => :in, :dim => [@pSize,@nb_rhs])
- @A_flu              = Real("A_flu",:dir => :in, :dim => [@afluSize])
+ @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+ @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+ @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
 
  end
  
@@ -31,6 +31,7 @@ class KOssRef
    push_env(:lang => FORTRAN) {
      @kernel = CKernel:: new()
      @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+		 #puts "Parameters = ", @kernel.procedure.parameters
      get_output.print codeF
    }
    return @kernel

### 2017-03-28 hou10ni/kernel.f90

@@ -21,11 +21,7 @@ SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, i
 
   !! Variables declaration
   INTEGER :: I,J,I_tmp1,I_tmp2, I1, I2, I3, I4
-  !! P_aux allocation
-  !REAL(kind=dp),ALLOCATABLE :: P_aux(:,:)
-  !ALLOCATE(P_aux((Nfacesperelem+1)*Nphi(MAXVAL(degree_tri_loc)),nb_rhs))
-  !! To evaluate P_aux allocation
-  REAL(kind=dp), DIMENSION(20,1) :: P_aux
+  REAL(kind=dp), DIMENSION(20,nb_rhs) :: P_aux
 
   DO I=1,Nflu_inner+Nflusol_inner
     I_tmp1=1

### 2017-03-28 hou10ni/run.rb

@@ -1,48 +1,50 @@
 require 'BOAST'
-include BOAST
-require './KOssRef.rb'
+require './KRef.rb'
 require 'narray_ffi'
+
+include BOAST
 
 class Experiment
 
  def self.run()
-  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran"}  
+  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}  
 
   set_lang(FORTRAN)
   kernels={}
 
   # Creating ref kernel
-  kernels[k_ref_params] = KOssRef::new(k_ref_params)
+  kernels[k_ref_params] = KRef::new(k_ref_params)
   kernels[k_ref_params].generate
+	puts "*********** Building kernel ***********\n"
   puts kernels[k_ref_params].kernel
-  kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS])
+  kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )
  
-  # Setting the values of arguments
-  seed = 10
-  ANArray.srand(seed) unless seed.nil?
-  ## Ces valeurs different selon le processus MPI
-  ## P0 AfluSize=16952560  Size=879732  Nflu_inner=210025  idx_vec_flu(12,214811) idx_mat_flu(214812)
-  ## P1 AfluSize=16955792  Size=881612  Nflu_inner=209629  idx_vec_flu(12,214805) idx_mat_flu(214806)
-  ## P2 AfluSize=16950864  Size=880460  Nflu_inner=209830  idx_vec_flu(12,214820) idx_mat_flu(214821)
-  ## P3 AfluSize=16953312  Size=880240  Nflu_inner=209986  idx_vec_flu(12,214811) idx_mat_flu(214812)
-  #@@Nflu_inner=210025
-  @@Nflu_inner=2
-  @@Nflusol_inner=0
-  @@nb_rhs=1
-  @@afluSize=16952560
-  @@pSize=879732
-  @@vecSize_2=214811
-  @@vecSize_1=12
-  @@matSize=214812
[…]

### 2017-03-29 Paper2/BOAST.tex
Modified affiliations.
Rewrote workflow description. Answers point 1) from reviewer.

@@
    \IEEEauthorblockA{\IEEEauthorrefmark{3}Université Grenoble Alpes}\\
-   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/SP2M/L\_Sim}\\
+   \IEEEauthorblockA{\IEEEauthorrefmark{4}CEA/INAC/MEM/L\_Sim}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{5}Inria}\\
@@
 
-Figure~\ref{fig:boast_workflow} illustrates the envisioned workflow and program
-structure. The user starts from an application kernel (either designed or
-implemented), and writes it in a dedicated language (Step 1). The language
-provides enough flexibility for the kernel to be meta-programmed with several
-orthogonal optimizations. Then the developer selects a kernel configuration and
-a target language. Those parameters define the output source code that will be
-generated by BOAST (Step 2). The resulting code source is then built according
-to the user specified compiler and options (Step 3). If input data are available
-then the kernel can be benchmarked and tested for non regression. Based on the
-results, other optimizations can be selected (Step 4) or new optimizations can
-be added to the BOAST sources. The process can be repeated until a good
-candidate is found on the target platform. The resulting kernel is then added to
-the program (Step 5).
-
-Several steps in the workflow can be automated: optimization and compiler flags
-exploration, non regression testing, benchmarking, \ldots This automation can be
-scripted by the user or by interfacing other dedicated tools. In order
-to achieve those goals three aspects should be considered: code description,
-code generation, and kernel execution run-time.
+Figure~\ref{fig:boast_workflow} illustrates the workflow and program structure.
+The user starts from an application kernel (either designed or implemented),
+and writes it in a dedicated language (Step 1). The language provides enough
+flexibility for the kernel to be meta-programmed with several orthogonal
+optimizations. From this set of optimizations, possible languages targets, and
+compilation options the user can design an optimization space to explore. This
+optimization space can contains rules to remove infeasible candidates. BOAST
+provides the mechanisms to specify those optimization spaces and enforce the
+users rules.
+
+Once this optimization space is designed, the user selects an optimization
+strategy, brute force and genetic algorithm are provided by BOAST, or can
+design his own. BOAST will then evaluate the different candidates generated by
+the optimization program. The candidate's parameters define the output source
+code that will be generated by BOAST (Step 2). The resulting code source is
+then built according to the specified compiler and options (Step 3). The kernel
+can be benchmarked and tested for non regression. Based on the results, other
+optimizations can be selected (Step 4).The process can be repeated until a good
+candidate is found on the target platform. The resulting kernel is then added
+to the program (Step 5).
+
+Of course, this workflow if not the only possible use of BOAST.  Especially,
+during the kernel description phase the user can use the BOAST framework to
+test his existing optimizations and analyse the results in order to design new
+optimizations. He can possibly do it using external tools.
+
+In order too achieve those results three aspects should be considered: code
+description, code generation, and kernel execution run-time.
 

### 2017-03-29 Paper2/BOAST.tex
Better explain vectorization. Answers to point 2) of reviewer.

@@
 five pixels instead of one using the vectors available on the Mali
-architecture. The vectors have to be kept in memory to ensure an efficient
-execution. Using vectors of 16 elements, 15 useful components are
-simultaneously loaded and can be used in computation.  This optimization yields
-speedups between 1.5 and 6 depending on the image size.
+architecture. This is done using the native OpenCL support for vectorization.
+For instance \emph{vload16} load 16 consecutive elements from memory and stores
+it a vector.  Here the loaded vector is of type \emph{uchar16} meaning that it
+is a vector of 16 consecutive bytes. In order to do computations without
+overflowing, those vectors have to be converted (using \emph{convert\_int16}) to
+a signed and bigger integer type.  Here they are converted to signed 4 bytes
+integers, leading to a vector size of 64 bytes.
+
+Once those vector are loaded they can be used in Single Instruction Multiple
+Data (SIMD) paradigm. Operations are done component-wise on the vector. For
+instance, \emph{v22 * (int16)9} multiplies each component of the vector by 9.
+The same thing goes for the difference operator. Once computed, the result is
+clamped and converted back to bytes using \emph{convert\_uchar16}.
+
+Using vectors of 16 elements, 5 pixels (15 components) can be simultaneously
+loaded and can be used in computation. Here the developer chose a pixel based
+approach and thus decided to only save the 15 relevant components. These
+components are saved in a vector manner using the power of two decomposition
+of 15. The first 8 components (0-7) are saved using \emph{vstore8} then the
+next 4 (8,9,a,b) using \emph{vstore4} and so on.
+
+This optimization yields speedups between 1.5 and 6 depending on the image size.
 
@@
 
-  int16 res = v22 * (int)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
+  int16 res = v22 * (int16)9 - v11 - v12 - v13 - v21 - v23 - v31 - v32 - v33;
         res = clamp(res, (int16)0, (int16)255);
@@
   vstore8(res_.s01234567, 0, dst + 3*width*j + 3*5*i);
-  vstore8(res_.s89ab,     0, dst + 3*width*j + 3*5*i + 8);
-  vstore8(res_.scd,       0, dst + 3*width*j + 3*5*i + 12);
+  vstore4(res_.s89ab,     0, dst + 3*width*j + 3*5*i + 8);
+  vstore2(res_.scd,       0, dst + 3*width*j + 3*5*i + 12);
   dst[3*width*j + 3*5*i + 14] = res_.se;

### 2017-03-29 Paper2/BOAST.tex
Answers point 3) of the review.

@@
 from 4 to 2 when image size increase, thus decreasing task granularity as the
-global work size increase. This result is unexpected and understanding why this
-happen could be interesting. The NVIDIA GPU favors processing square tiles:
-\emph{x\_component\_number} = 4, \emph{y\_component\_number} = 4,
+global work size increase. This result is interesting because its unexpected
+nature strongly backs the use of autotuning. The NVIDIA GPU favors processing
+square tiles: \emph{x\_component\_number} = 4, \emph{y\_component\_number} = 4,
 \emph{vector\_length} = 4, \emph{temporary\_size} = 2 and

### 2017-03-30 Paper2/BOAST.tex
Modification of SPECFEM3D section. Answers point 4) and 5) of reviewer.

@@
 %%% Porting SPECFEM3D to OpenCL using BOAST %%%
-\subsection{Porting SPECFEM3D to OpenCL using BOAST\label{subsec:specfem}}
-
-In this last subsection, we study a free seismic wave propagation simulator,
-SPECFEM3D, in its GLOBE flavor\footnote{Specfem3d Globe --- CIG
-\url{http://www.geodynamics.org/cig/software/specfem3d-globe}}. SPECFEM3D
-simulates seismic wave propagation at the local or regional scale based upon
-spectral-element method (SEM), with very good accuracy and convergence
-properties. It is a reference application for supercomputer benchmarking,
-thanks to its good scaling capabilities.
-
+\subsection{Porting Application Kernels: Porting SPECFEM3D from CUDA to
+OpenCL using BOAST\label{subsec:specfem}}
+
+In this last subsection, we present an alternative use-case of the BOAST
+framework, focused on portability. As part of the Mont-Blanc project,
+we had to port a scientific application so that it could be used to
+benchmark the Mont-Blanc HPC cluster. It was composed of 40 complex
+and hand-tuned CUDA kernels, and we had to port them to OpenCL. This
+scenario also highlights the factorization and maintainability
+improvements of the BOAST framework, as we obtained a single BOAST
+implementation of the kernels (instead of CUDA plus OpenCL) and this
+version of the code is still used and updated in the upstream project.
+
+The challenges we faced in this case-study were more related to the
+complexity of the scientific application and its kernels than to the
+BOAST framework.
 %%% SPECFEM3D %%%
 \subsubsection{SPECFEM3D}
+
+SPECFEM3D GLOBE\footnote{SPECFEM3D GLOBE --- CIG
+  \url{http://www.geodynamics.org/cig/software/specfem3d-globe}} is a
+free seismic wave propagation simulator. It simulates seismic wave
+propagation at the local or regional scale based upon spectral-element
+method (SEM), with very good accuracy and convergence properties. It
+is a reference application for supercomputer benchmarking, thanks to
+its good scaling capabilities.
 
@@
 massively parallel scale), while at the other side of the spectrum,
-some complex kernels took more than 80 parameters and perform very
+some complex kernels took more than 80 parameters and performed very
 specific physical transformations.
@@
 %%%
-\paragraph{Porting kernels to BOAST}  Nvidia CUDA and OpenCL are based on the same programming model: a
+\paragraph{Porting kernels to BOAST}  NVidia CUDA and OpenCL are based on the same programming model: a
 massively parallel accelerator running in disjoint, non addressable
@@
 merged in SPECFEM3D's development tree and under test and extension by
-different research teams. On a platform with two K40x GPUs and N Intel
-Xeon processors, we measured similar results between the original
-CUDA version and our BOAST-CUDA version. With the same set of
-optimization flags, BOAST CUDA and OpenCL version reported similar
+different research teams. On a platform with two K40x GPUs and 24 Intel
+Xeon processors, we measured identical performance between the original
+CUDA version and our BOAST-CUDA version. This result was expected, as
+the BOAST-generated version is identical to the original, apart from
+naming differences.
+
+With the same set of
+optimization flags, BOAST CUDA and OpenCL versions reported similar
 execution time spans. The best execution speed was achieved with CUDA
 version though (25\% higher than OpenCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to OpenCL run-time, as of version
+(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to the OpenCL run-time, as of version
 1.1. This parameter, in addition to specifying the work group size (which can
@@
 
-By refactoring GPU kernel code in BOAST EDSL, the size of kernel code
-shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, partly
-because of code duplication, also removing manually unrolled
+By refactoring GPU kernels' code in BOAST EDSL, the size of the kernels' source code
+shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, mainly
+because of the removal of code duplication and manually unrolled
 loops). This is beneficial for SPECFEM3D as it improves the readability
-and maintainability of its source code.
+and maintainability of its source code. This gain was confirmed by the
+SPECFEM3D community as the subsequent developments have been carried out
+on the BOAST version of the kernels, and not on the CUDA or OpenCL generated code.
 
@@
 performance evaluation mechanisms. These tests will help developers to
-try new optimizations in kernel code and measure their impact, without
-executing the whole application.
+try new optimizations in kernels' code and measure their impact,
+without executing the whole application.
 

### 2017-03-30 Paper2/BOAST.tex
Modified title.
fix and improve SPECFEM3D modifications

@@
 framework, focused on portability. As part of the Mont-Blanc project,
-we had to port a scientific application so that it could be used to
+we ported a scientific application so that it could be use to
 benchmark the Mont-Blanc HPC cluster. It was composed of 40 complex
-and hand-tuned CUDA kernels, and we had to port them to OpenCL. This
+and hand-tuned CUDA kernels, that we ported to OpenCL framework. This
 scenario also highlights the factorization and maintainability
 improvements of the BOAST framework, as we obtained a single BOAST
-implementation of the kernels (instead of CUDA plus OpenCL) and this
+implementation of the kernels (instead of two, CUDA and OpenCL). This
 version of the code is still used and updated in the upstream project.
@@
 
-Because of the complexity of wave propagation (and of the application
-architecture), it is hard to impossible to define unit tests. Hence,
-the application is validated by the accuracy of the results it
-produces, that is, the accuracy of seismograms of the simulated
-earthquakes, in comparison with the actual ones. Non-regression
-testing (after new developments) is also based on these seismograms,
-with measurements of the relative error between two identical
-simulations.
-
-As part of the Mont-Blanc project, we had to port SPECFEM3D to OpenCL, so
-that it could be used to benchmark Mont-Blanc HPC platforms.
+Because of the complexity of the wave propagation kernels, it was
+impossible for us to understand the kernel source kernel. Hence, we
+treated them as black-boxes, and aimed at obtaining identical
+performances. Furthermore, the kernels' complexity prevents the
+specification of unit tests, thus the application is validated by the
+accuracy of the results it produces (the seismograms of the simulated
+earthquakes), against the actual ones. Non-regression testing (after
+new developments) is also based on these seismograms, with
+measurements of the relative error between two identical simulations.
 
@@
 %%%
-\paragraph{Porting kernels to BOAST}  NVidia CUDA and OpenCL are based on the same programming model: a
+\paragraph{Porting kernels to BOAST} 
+NVidia CUDA and OpenCL are based on the same programming model: a
 massively parallel accelerator running in disjoint, non addressable
@@
 consisted in the translation of the CPU-side of the application, from
-CUDA API to OpenCL API. Most of the functions of the interfaces are
-very similar, with only naming-convention and data-structure
+CUDA API to OpenCL'. Most of the functions of the interfaces are very
+similar, with only naming-convention and data-structure
 distinctions. Hence, it was clear that automatic rewriting tools
@@
 CUDA version and our BOAST-CUDA version. This result was expected, as
-the BOAST-generated version is identical to the original, apart from
+the BOAST-generated version is identical to the original, except from
 naming differences.
 
-With the same set of
-optimization flags, BOAST CUDA and OpenCL versions reported similar
-execution time spans. The best execution speed was achieved with CUDA
-version though (25\% higher than OpenCL), as one optimization parameter
-(\code{CUDA\_LAUNCH\_BOUND}) cannot be passed to the OpenCL run-time, as of version
-1.1. This parameter, in addition to specifying the work group size (which can
-be done in OpenCL), also constrains the number of work group that must run in
-parallel on a multiprocessor. This value is set to 7 in CUDA. This means that
-the compiler must be very conservative on register usage in order to allow this
-parallelism which allows better overlapping of communications end computations.
-This functionality is not supported in OpenCL.
-
-By refactoring GPU kernels' code in BOAST EDSL, the size of the kernels' source code
-shrank by a factor of 1.8 (from 7500 to less than 4000 LOC, mainly
-because of the removal of code duplication and manually unrolled
-loops). This is beneficial for SPECFEM3D as it improves the readability
-and maintainability of its source code. This gain was confirmed by the
-SPECFEM3D community as the subsequent developments have been carried out
-on the BOAST version of the kernels, and not on the CUDA or OpenCL generated code.
+With the same set of optimization flags, BOAST CUDA and OpenCL
+versions reported similar execution time spans. The best execution
+speed was achieved with CUDA version though (25\% higher than OpenCL),
+as one optimization parameter (\code{CUDA\_LAUNCH\_BOUND}) cannot be
+passed to the OpenCL run-time, as of version 1.1. This parameter, in
+addition to specifying the work group size (which can be done in
+OpenCL), also constrains the number of work group that must run in
+parallel on a multiprocessor. This value is set to 7 in CUDA. This
+means that the compiler must be very conservative on register usage in
+order to allow this parallelism which allows better overlapping of
+communications end computations.  This functionality is not supported
+in OpenCL.
+
+Thanks to the porting of the SPECFEM3D GPU kernels to BOAST EDSL, the size of the
+kernels' source code shrank by a factor of 1.8 (from 7500 to less than
+4000 LOC, mainly because of the removal of code duplication and
+manually unrolled loops). This is beneficial for SPECFEM3D as it
+improves the readability and maintainability of its source code. This
+gain was confirmed by the SPECFEM3D community as the subsequent
+developments have been carried out on the BOAST version of the
+kernels, and not on the CUDA or OpenCL generated code.
 

### 2017-04-07 Paper2/BOAST.tex
passe FD-BV

@@
 architecture. This is done using the native OpenCL support for vectorization.
-For instance \emph{vload16} load 16 consecutive elements from memory and stores
-it a vector.  Here the loaded vector is of type \emph{uchar16} meaning that it
+For instance \emph{vload16} loads 16 consecutive elements from memory and stores
+it in a vector.  Here the loaded vector is of type \emph{uchar16} meaning that it
 is a vector of 16 consecutive bytes. In order to do computations without
@@
 Once this optimization space is designed, the user selects an optimization
-strategy, brute force and genetic algorithm are provided by BOAST, or can
+strategy, brute force and genetic algorithms are provided by BOAST, or can
 design his own. BOAST will then evaluate the different candidates generated by
@@
 %%% Porting SPECFEM3D to OpenCL using BOAST %%%
-\subsection{Porting Application Kernels: Porting SPECFEM3D from CUDA to
-OpenCL using BOAST\label{subsec:specfem}}
+\subsection{Porting SPECFEM3D Application Kernels: from CUDA to OpenCL
+using BOAST\label{subsec:specfem}}
 

### 2017-04-07 Paper2/BOAST.tex
passe FD-BV affiliations

@@
 \author{
-   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}, Thierry Deutsch\IEEEauthorrefmark{4}, Dimitri Komatitsch\IEEEauthorrefmark{6}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
+   \IEEEauthorblockN{Brice Videau\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, Kevin Pouget\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Luigi Genovese\IEEEauthorrefmark{4}\IEEEauthorrefmark{3}, Thierry Deutsch\IEEEauthorrefmark{4}\IEEEauthorrefmark{3}, Dimitri Komatitsch\IEEEauthorrefmark{6}, Frédéric Desprez\IEEEauthorrefmark{1}\IEEEauthorrefmark{5}, Jean-François Méhaut\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}}\\
    \IEEEauthorblockA{\IEEEauthorrefmark{1}Laboratoire d'Informatique de Grenoble}\\

### 2017-04-25 Paper2/answer-reviewers.tex
ajout

\documentclass[10pt]{article}

\usepackage{fullpage}
\usepackage{url}

\begin{document}

\noindent
{\Large\bf ANSWERS TO THE REVIEWERS}\\~\\

We thank the  reviewers for their interest in our work and for their helpful comments that will greatly improve  the manuscript. We have  tried  to do  our  best to  respond to  the  points  raised. The reviewers have brought  up  some  good  points  and  we  appreciate  the  opportunity  to  clarify  our research objectives and results. As  indicated  below,  we  have  checked  all  the  general  and  specific  comments  provided  by  the reviewers and have made necessary changes accordingly to their indications.\\

\noindent{\large\bf Anonymous reviewers \#1}\\~\\

\end{document}
%%%

### 2017-04-28 Paper2/answer-reviewers.tex
Improved answer.

@@
 
-\noindent{\large\bf Anonymous reviewers \#1}\\~\\
+\noindent{\large\bf Anonymous reviewers \#1}\\
+
+\noindent
+{\em 1) The usage workflow is not that clear. For instance how optimisations can be selected ? Is there a list of such optimisations or is it the user
+  to design the optimisation sets. In Section V.A how different values are selected? Is there any way for the framework to guide the user towards
+  optimisations or to give feedback on what is working, what is not working, etc. The authors mention MAQAO in the conclusion... Overall, this part is
+  really fuzzy.}\\
+
+\noindent
+{\em 2) Listing 3 should be explained in details or removed. Not every reader is a specialist of vectorisation}\\
+
+\noindent
+{\em 3) section V.A page 8.
+You state that you have an unexpected results that should be investigate.
+A journal paper is not a conference paper.
+I think you should investigate and publish what you have found in ***this*** paper.} \\
+
+\noindent
+{\em 4) Section V.C.
+is not easy to follow.
+It is hard to understand where the problems you faced came from : from the kernels, from the design of BOAST, from a bug in the implementation, etc.
+What kind of conclusion should be taken from your experience in the usage of BOAST?
+Such discussion should help the reader in doing better science.
+The writing of this section does not go in that direction.
+What can be generalised from your experience?
+what are your advices?}\\
+
+\noindent
+{\em 5) in Section V.C why did you not generate the CUDA version as well and compare it with the hand-tuned existing version? For the revised version, I need to see that experiment added to paper!}\\
 
@@
 %%%
+
+Reviewer: 1
+
+Comments to the Author
+
+This paper present a framework, called BOAST,  for developing computational kernels in a portable efficient and productive way.
+It provides a DSL and a set of tools to generate different source code with different optimisations and to autotune the code in order to produce  an optimised version for the target architecture.
+
+I think this is a good work with several examples showing how the framework works and how it yields to interesting results.
+
+I have several remarks in order to make this paper better.

### 2017-04-28 Paper2/BOAST.tex
changed numbering and figure's position

@@
         label={lst:BOAST_procedure},
-        numbers=right }
+        numbers=left }
 \begin{lstlisting}
@@
 Section~\ref{subsec:specfem}.
-
-The idea behind this tool is based on the assumption that the different GPU
-ports of the code should perform the very same operations, at least at the
-logical level (the APIs will have \emph{implementation} differences, but they
-should offer nonetheless same functionalities). The usage of BOAST framework
-sustains this assumption, as both sets of kernels should be generated from the
-same source code.
-
-Hence, the verification and validation of the new port can be narrowed down to
-asserting that both codes apply the same operations on the GPU. And the
-debugging part will consist in understanding what diverges. To that purpose, we
-developed \code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
-\code{GPUTrace} dynamically preloads a library between the application and the
-GPU library, and collects the function name, execution range and argument
-values (input \emph{and} output) of the relevant function calls. These
-information are traced in a unified format for all the APIs. This is a custom
-trace format. Listing~\ref{lst:GPUTrace} presents a sample output generated
-during SPECFEM3D tracing.
-
-However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
-to be \emph{state-full}. Indeed, most of API parameters are handles to opaque
-types. So, in order to generate meaningful traces, \code{GPUTrace} gathers
-information about these objects at creation time (handle value, buffer creation
-size and attributes, kernel name and prototype, \etc{}) and re-injects this
-information when the objects used. A state-less implementation of
-\code{GPUTrace} would highly rely on the GPU libraries introspection
-capabilities, which seem not possible at the moment.
 
@@
 
+The idea behind this tool is based on the assumption that the different GPU
+ports of the code should perform the very same operations, at least at the
+logical level (the APIs will have \emph{implementation} differences, but they
+should offer nonetheless same functionalities). The usage of BOAST framework
+sustains this assumption, as both sets of kernels should be generated from the
+same source code.
+
+Hence, the verification and validation of the new port can be narrowed down to
+asserting that both codes apply the same operations on the GPU. And the
+debugging part will consist in understanding what diverges. To that purpose, we
+developed \code{GPUTrace}, inspired by \code{strace} and \code{ltrace} tools:
+\code{GPUTrace} dynamically preloads a library between the application and the
+GPU library, and collects the function name, execution range and argument
+values (input \emph{and} output) of the relevant function calls. These
+information are traced in a unified format for all the APIs. This is a custom
+trace format. Listing~\ref{lst:GPUTrace} presents a sample output generated
+during SPECFEM3D tracing.
+
+However, in contrast with \code{strace} and \code{ltrace}, \code{GPUTrace} has
+to be \emph{state-full}. Indeed, most of API parameters are handles to opaque
+types. So, in order to generate meaningful traces, \code{GPUTrace} gathers
+information about these objects at creation time (handle value, buffer creation
+size and attributes, kernel name and prototype, \etc{}) and re-injects this
+information when the objects used. A state-less implementation of
+\code{GPUTrace} would highly rely on the GPU libraries introspection
+capabilities, which seem not possible at the moment.
+
 Once two call traces have been generated by \code{GPUTrace}, the user can
@@
   label={lst:conv_example},
-  numbers=right}
+  numbers=left}
 

### 2017-05-04 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni's kernel - Ref version Ok

@@
 To show the impact of input parameters, we can change the elements order in HPC4E.txt (\textit{Ordre des elements}) or refine the mesh (\texttt{tail MAILLAGES/florian3D.1.ele} will show the command).
-Beware of the total time, it takes into account the time spent to write in a file (time in function \texttt{sub\_time\_loop}).
+Beware of the total execution time, it takes into account the time spent to write in a file (time in function \texttt{sub\_time\_loop}).
 
@@
 \caption{Averaged of execution-Time (5 to 10 runs).
-{\underline{Input parameters}}: elements order =1, 859247 elements, {\bf{Simulation time=0.001}} (Time Step=0.001) - {\bf{Sans l'ecriture des instantannees}} - Beware!!
-I wasn't alone on idchire}
+{\underline{Input parameters}}: elements order =1, 859247 elements, {\bf{Simulation time=0.001}} (Time Step=0.001) - {\bf{without writting in sub\_time\_loop}}}
 {\small
@@
 \caption{Averaged of execution-Time (5 to 10 runs).
-{\underline{Input parameters}}: elements order =1, 859247 elements,  {\bf{Simulation time=0.01}} (Time Step=0.01)  - {\bf{Sans l'ecriture des instantannees}}}
+{\underline{Input parameters}}: elements order =1, 859247 elements,  {\bf{Simulation time=0.01}} (Time Step=0.01)  - {\bf{without writting in sub\_time\_loop}}}
 {\small
@@
 \caption{Averaged of execution-Time (5 to 10 runs).
-{\underline{Input parameters}}: elements order =1, 859247 elements,    {\bf{Simulation time=0.1}} (Time Step=0.1) - {\bf{Sans l'ecriture des instantannees}}}
+{\underline{Input parameters}}: elements order =1, 859247 elements,    {\bf{Simulation time=0.1}} (Time Step=0.1) - {\bf{without writting in sub\_time\_loop}}}
 {\small
@@
 \centering
-\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: 859247 elements,  Simulation time=0.001 (Time Step=0.001) - {\bf{Sans l'ecriture des instantannees}}}
+\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: 859247 elements,  Simulation time=0.001 (Time Step=0.001) - {\bf{without writting in sub\_time\_loop}}}
 {\small
@@
 
-\noindent and change the number in the option "-kna10000A".
+\noindent and change the number in the option "-kna10000A".\\
 
@@
 \centering
-\caption{TODO: Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order=1, Simulation time=..}
+\caption{Averaged of execution-Time (5 to 10 runs). {\underline{Input parameters}}: elements order=1, Simulation time=0.001 (Time Step=0.001)}
 {\small
@@
 										&   		min & max & avg							& 	min & max & avg			\\ \hline \hline
- 		10\ 000		                 				& 29.832	& 31.821 & 								&  	 & 	&  173.77	\\ \hline
- 		5\ 000		                 				& 62.568	&  66.309  &	 							&  	 & 	 & 351.35	\\ \hline
- 		2\ 500		                 				&  141.110 &  147.306	&							&  	& &  735.548	\\ \hline
+ 		10\ 000 (859\ 247 elt)		           	& .285 & .2973 &  .2917						&  175.0226 & 175.0522 & 175.041	\\ \hline
+ 		5\ 000 (1\ 704\ 052 elt)		                	& .8287 &  .8860 & .8590						&  348.888 & 379.0077 & 366.995	\\ \hline
+ 		2\ 500 (3\ 385\ 940 elt)	                		&  1.9168 &  1.9248	& 1.9214					&  708.008 & 707.994 &  707.993 \\ \hline
 \end{tabular}
@@
 \end{table}
+
+%% Lancer avec 4 processus?? pour mieux comparer avec le reste
 
@@
 \begin{tabular}{|c|c|c|} \hline 
- {{\bf{Process ID}}} 	& {{\bf{Time spent in \texttt{sub\_time\_loop} }}}	& {{\bf{Time spent in \texttt{fluid\_inner\_elt()} }}}\\ \hline \hline
+ {{\bf{Process ID}}} 	& {{\bf{Time in \texttt{sub\_time\_loop} }}}	& {{\bf{Time in \texttt{fluid\_inner\_elt} }}}\\ \hline \hline
 		0		&   5.388336							&  5.1683220	(95,9\%)		\\ \hline 
@@
 \begin{tabular}{|c|c|c|} \hline 
- {{\bf{Process ID}}} 	& {{\bf{Time spent in \texttt{sub\_time\_loop} }}}	& {{\bf{Time spent in \texttt{fluid\_inner\_elt()} }}}\\ \hline \hline
+ {{\bf{Process ID}}} 	& {{\bf{Time in \texttt{sub\_time\_loop} }}}	& {{\bf{Time in \texttt{fluid\_inner\_elt} }}}\\ \hline \hline
 		0		&   4.776291							&  4.54828700	(95,2\%)		\\ \hline 

### 2017-05-04 hou10ni/run_boast.rb

@@ -1,0 +1,41 @@
+require 'BOAST'
+require './KBoast.rb'
+require 'narray_ffi'
+
+include BOAST
+
+class Experiment
+
+ def self.run()
+
+ # set_default_int_size(8)
+  set_default_real_size(4)
+  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}  
+
+  set_lang(FORTRAN)
+  kernels={}
+
+  # Creating ref kernel
+  kernels[k_boast_params] = KBoast::new(k_boast_params)
+  kernels[k_boast_params].generate
+	puts "*********** Building kernel ***********\n"
+  puts kernels[k_boast_params].kernel
+  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+ 
+	puts "\n*********** Testing boast kernel with fluid_inner_elt_boast/sub_time_loop_1/ inputs parameters ***********\n"
+  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
+#  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
+
+
+  inputs.each_key { |key|
+		puts key
+# 		puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect 
+#		puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
+	}
+
+	puts "Testing done\n"
+  return kernels
+ end
+end
+
+Experiment.run

### 2017-05-04 Paper2/answer-reviewers.tex

@@
 
+The workflow has been better described (see diff file attached).\\
+
 \noindent
@@
 
+Listing 3 and vectorization have been described carefully in Section B.a.\\
+
 \noindent
-{\em 3) section V.A page 8.
-You state that you have an unexpected results that should be investigate.
-A journal paper is not a conference paper.
-I think you should investigate and publish what you have found in ***this*** paper.} \\
+{\em 3) section V.A page 8. You state that you have an unexpected results that should be investigate. A journal paper is not a conference paper. I
+  think you should investigate and publish what you have found in ***this*** paper.} \\
+
+Autotuning is useful because it can challenge our assumptions. We modified a sentence accordingly. \\
 
@@
 
+Our approach's objectives have been better described at the beginning of the section. Then several paragraphs of this section have been modified to
+better emphasize the results and expectations of this part of our work.\\
+
 \noindent
@@
 
+Indeed this experience was already performed. We rewrote this section to better mention our results around CUDA.
+
 \end{document}
 %%%
-
-Reviewer: 1
-
-Comments to the Author
-
-This paper present a framework, called BOAST,  for developing computational kernels in a portable efficient and productive way.
-It provides a DSL and a set of tools to generate different source code with different optimisations and to autotune the code in order to produce  an optimised version for the target architecture.
-
-I think this is a good work with several examples showing how the framework works and how it yields to interesting results.
-
-I have several remarks in order to make this paper better.

### 2017-05-04 hou10ni/KBoast.rb
Hou10ni's kernel : Starting BOAST version

@@ -1,0 +1,72 @@
+class KBoast
+ attr_reader :kernel
+
+
+ def initialize(options)
+
+  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
+  @opts.update(options)
+
+  # Parameters 
+  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
+  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
+  @nb_rhs             = Int("nb_rhs", :dir => :in )
+  @afluSize           = Int("afluSize", :dir => :in )
+  @pSize              = Int("pSize", :dir => :in )
+  @vecSize_1          = Int("vecSize_1", :dir => :in )
+  @vecSize_2          = Int("vecSize_2", :dir => :in )
+  @matSize            = Int("matSize", :dir => :in )
+
+ @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+ @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+ @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+ @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
+
+ end
+ 
+ def generate
+
+   push_env(:lang => FORTRAN) {
+
+     @kernel = CKernel:: new()
+     @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+		 #puts "Parameters = ", @kernel.procedure.parameters
+		 opn @kernel.procedure
+
+		# locals
+ 		i = Variable::new('I', Int)
[…]

### 2017-06-14 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni's kernel : Optimisation

@@
 gem build BOAST.gemspec
-gem install --user-install BOAST-1.3.5.gem 
+gem install --user-install BOAST-2.0.3.gem 
 \end{cverbatim}
@@
 KRef defines the reference version.
-This version simply need to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST.
+This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST.
+To put it in a nutshell, a reference version should follow the template bellow.
+
+\lstset{language=RUBY}
+\begin{lstlisting}[caption=KRef\_template.rb]
+def KRef
+  lang = BOAST::get_lang
+  BOAST::set_lang(BOAST::FORTRAN)
+  kernel = BOAST::CKernel::new
+  function_name = "function_ref"
+
+  VARIABLES HERE
+  FOR EXAMPLE : 
+  geocode = BOAST::Int("geocode", :dir => :in )
+  n01 = BOAST::Int("n01", :dir => :in)
+  u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )
+
+   p = BOAST::Procedure::new(function_name, [geocode,n01,u])
+
+   kernel.code.print <<EOF
+  ORIGINAL C or FORTRAN CODE HERE
+EOF
+
+  kernel.procedure = p
+  BOAST::set_lang(lang)
+  return kernel
+end
+
+geocode = ...
+n01 = ...
+u = ...
+k = KRef
+stats = k.run(geocode, n01, u)
+puts "#{k.procedure.name}: #{stats[:duration]*1.0e3}"
+\end{lstlisting}
+
+We wrote the reference version a bite differently as shown bellow.
 
@@
 
-The file run.rb is used to launch the kernel.
+The file run\_ref.rb is used to launch the kernel.
 
-\begin{lstlisting}[caption=run.rb]
+\begin{lstlisting}[caption=run\_ref.rb]
 require 'BOAST'
@@
 
-\begin{lstlisting}[caption=Results (obtained with the command ruby run.rb)]
+\begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
 SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
@@
 
-TODO
+\lstset{language=RUBY}
+\begin{lstlisting}[caption=KBoast.rb]
+class KBoast
+ attr_reader :kernel
+
+ def initialize(options)
+
+  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
+  @opts.update(options)
+
+  # Parameters 
+  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
+  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
+  @nb_rhs             = Int("nb_rhs", :dir => :in )
+  @afluSize           = Int("afluSize", :dir => :in )
+  @pSize              = Int("pSize", :dir => :in )
+  @vecSize_1          = Int("vecSize_1", :dir => :in )
+  @vecSize_2          = Int("vecSize_2", :dir => :in )
+  @matSize            = Int("matSize", :dir => :in )
+
+  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
+
+ end
+
+ def generate
+
+   push_env(:lang => FORTRAN) {
+
+    @kernel = CKernel:: new()
+    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+
+    #..  FUNCTIONS ..#
+    register_funccall("matmul") if get_lang == FORTRAN
+    call_matmul = lambda{|x,y|
+      if get_lang == FORTRAN
+        return matmul(x,y)
+      end
+    }
+
+    #.. KERNEL ..#
+
+    opn @kernel.procedure
+
+    # local variables
+    i = Variable::new('I', Int)
+    j = Variable::new('J', Int)
+    i_tmp1 = Variable::new('I_tmp1', Int)
+    i_tmp2 = Variable::new('I_tmp2', Int)
+    i1 = Variable::new('I1', Int)
+    i2 = Variable::new('I2', Int)
+    i3 = Variable::new('I3', Int)
+    i4 = Variable::new('I4', Int)
+    p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])
+
+    decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
+
+    pr For(i,1,@Nflu_inner+@Nflusol_inner){
+      pr i_tmp1 === 1
+      pr For(j,1,@idx_vec_flu[1,i]){
+        pr i1 === @idx_vec_flu[2*j,i]
+        pr i2 === @idx_vec_flu[2*j+1,i]
+        pr i_tmp2 === i_tmp1 + i2 - i1
+        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs) 
+        pr i_tmp1 === i_tmp2 +1
+      }
+      pr i1 === @idx_vec_flu[2,i]
+      pr i2 === @idx_vec_flu[3,i]
+      pr i3 === @idx_mat_flu[i]
+      pr i4 === @idx_mat_flu[i+1] - 1
+      reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
+      pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
+    }
+    close @kernel.procedure
+   }
+
+   return @kernel
+ end
+end
+\end{lstlisting}
+
+\noindent The file run\_boast.rb is used to launch the BOAST version of the kernel. We run 3 times the kernel (repeat). \\
+
+\begin{lstlisting}[caption=run\_boast.rb]
+require 'BOAST'
+require './KBoast.rb'
+require 'narray_ffi'
+
+include BOAST
+
+class Experiment
+
+ def self.run()
+
+ # set_default_int_size(8)
+  set_default_real_size(4)
+  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}
+
+  set_lang(FORTRAN)
+  kernels = {}
+  stats_svg = {}
+  stats = []
+  repeat = 3
+
+  # Creating ref kernel
+  kernels[k_boast_params] = KBoast::new(k_boast_params)
+  kernels[k_boast_params].generate
+  puts "*********** Building kernel ***********\n"
+  puts kernels[k_boast_params].kernel
+  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+
+  puts "\n*********** Testing boast kernel with inputs parameters ***********\n"
+  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
+  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
+
+  repeat.times {
+    inputs.each_key { |key|
+      puts key
+      stats.push kernels[k_boast_params].kernel.run(*(inputs[key]))
+      puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect
+      puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
+    }
+  }
+  stats.sort_by! { |a| a[:duration] }
+  stats = stats.first
+
+  puts "#{kernels[k_boast_params].kernel.procedure.name}: #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
+  return kernels
+ end
+end
+
+Experiment.run
+\end{lstlisting}
+
+\begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
+SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
+&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
+&ecSize_1, vecSize_2, matSize)
+
+  integer, parameter :: wp=kind(1.0d0)
+  integer(kind=4), intent(in) :: Nflu_inner
+  integer(kind=4), intent(in) :: Nflusol_inner
+  integer(kind=4), intent(in) :: nb_rhs
+  ...
+
+END SUBROUTINE fluid_inner_elt_boast
+
+*********** Testing boast kernel with inputs parameters ***********
+./fluid_inner_elt_boast/sub_time_loop_1
+{:duration=>0.13263683, :start=>-4004705691753970176, :end=>-3872068861753970176}
+{"P_new"=>0.0}
+./fluid_inner_elt_boast/sub_time_loop_1
+{:duration=>0.140625219, :start=>-3642235174753970176, :end=>-3501609955753970176}
+{"P_new"=>0.0}
+./fluid_inner_elt_boast/sub_time_loop_1
+{:duration=>0.147113606, :start=>-3274643277753970176, :end=>-3127529671753970176}
+{"P_new"=>0.0}
+fluid_inner_elt_boast: 0.15179012 s ->  151.79012 ms  
+\end{lstlisting}
+
+\subsubsection*{Optimisations}
 

### 2017-06-14 hou10ni/KBoast.rb

@@ -33,9 +33,28 @@ class KBoast
      @kernel = CKernel:: new()
      @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
 		 #puts "Parameters = ", @kernel.procedure.parameters
-		 opn @kernel.procedure
 
-		# locals
+
+		#..  FUNCTIONS ..#
+
+		register_funccall("matmul") if get_lang == FORTRAN
+		#reecrire matmul pour le C @matmul = Procedure("matmul",[..,..], :return => ..){ ... }
+		call_matmul = lambda{|x,y|
+			if get_lang == FORTRAN
+				return matmul(x,y)
+	    #else
+			#	return @functions[:matmul].call(x,y)
+ 			end
+		}
+
+		register_funccall(:modulo) if get_lang == FORTRAN
+
+
+		#.. KERNEL ..#
+
+		opn @kernel.procedure
+
+		# local variables
  		i = Variable::new('I', Int)
  		j = Variable::new('J', Int)
  		i_tmp1 = Variable::new('I_tmp1', Int)
@@ -48,37 +67,63 @@ class KBoast
 
 		decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
 
+
+
+   	l = lambda { |unroll|
+
+			loop_1 = lambda { |indice|
+				pr i1 === @idx_vec_flu[2*indice,i]
+				pr i2 === @idx_vec_flu[2*indice+1,i]
+				pr i_tmp2 === i_tmp1 + i2 - i1
+				pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs)
+				pr i_tmp1 === i_tmp2 + 1.to_var
+			}
+		
+			f = For(j,1,@idx_vec_flu[1,i]-modulo(@idx_vec_flu[1,i],unroll), step: unroll){
+				unroll.times { |k|
+					loop_1[j+k]
+      	}
+			}
+			f_arr = [f]
+			if unroll > 1 then
[…]

### 2017-06-14 hou10ni/run.rb

@@ -17,7 +17,7 @@ class Experiment
   set_default_real_size(4)
 
   k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}
-  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}  
+  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O3"}  
 
   set_lang(FORTRAN)
   kernels={}
@@ -41,8 +41,8 @@ class Experiment
 
   # Run (inputs are the same for ref and boast kernels) 
 	puts "\n* Testing boast and ref versions of the kernel with fluid_inner_elt_boast/sub_time_loop_1/ inputs parameters ...\n"
-  inputs = kernels[k_ref_params].kernel.load_ref_inputs()
-
+  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
+  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
 
 10.times{|i|
   inputs.each_key { |key|
@@ -50,7 +50,7 @@ class Experiment
  		#puts kernels[k_ref_params].kernel.run(*(inputs[key])).inspect 
  		stats[k_ref_params][:time][i] = kernels[k_ref_params].kernel.run(*(inputs[key]))[:duration]
  		stats[k_boast_params][:time][i] = kernels[k_boast_params].kernel.run(*(inputs[key]))[:duration]
-#		puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
+		puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
 	}
 }
 

### 2017-06-14 hou10ni/run_boast.rb

@@ -13,7 +13,10 @@ class Experiment
   k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}  
 
   set_lang(FORTRAN)
-  kernels={}
+  kernels = {}
+  stats_svg = {}
+	stats = []
+  repeat = 3
 
   # Creating ref kernel
   kernels[k_boast_params] = KBoast::new(k_boast_params)
@@ -26,14 +29,20 @@ class Experiment
   inputs = kernels[k_boast_params].kernel.load_ref_inputs()
   outputs = kernels[k_boast_params].kernel.load_ref_outputs()
 
-
-  inputs.each_key { |key|
-		puts key
- 		puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect 
-		puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
+	repeat.times {
+  	inputs.each_key { |key|
+			puts key
+ 			stats.push kernels[k_boast_params].kernel.run(*(inputs[key])) 
+ 			puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect 
+			puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
+		}
 	}
+  stats.sort_by! { |a| a[:duration] }
+	stats = stats.first
 
-	puts "Testing done\n"
+
+
+  puts "#{kernels[k_boast_params].kernel.procedure.name}: #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
   return kernels
  end
 end

### 2017-06-19 hou10ni/KBoast.rb
Hou10ni's kernel : optimizations update

@@ -128,8 +128,11 @@ class KBoast
  		i4 = Variable::new('I4', Int)
  		p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])
 
-		decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
+ 		jmax = Variable::new('jmax', Int)
+ 		imax = Variable::new('imax', Int)
 
+		decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
+		decl jmax,imax
 
 
 		##.. Nested loop optimization
@@ -144,6 +147,8 @@ class KBoast
 		}
 
 		#  Option 1 : Unroll 'unroll' iterations
+
+
    	optim_1 = lambda { |unroll,i|
 			loop_1 = lambda { |indice|
 				pr i1 === @idx_vec_flu[2*indice,i]
@@ -153,7 +158,8 @@ class KBoast
 				pr i_tmp1 === i_tmp2 + 1.to_var
 			}
 		
-			f_nested = For(j,1,@idx_vec_flu[1,i]-modulo(@idx_vec_flu[1,i],unroll), step: unroll){
+			pr jmax === @idx_vec_flu[1,i]-(unroll-1)
+			f_nested = For(j,1,jmax, step: unroll){
 				unroll.times { |k|
 					loop_1[j+k]
       	}
@@ -196,10 +202,11 @@ class KBoast
       	pr i3 === @idx_mat_flu[indice]
       	pr i4 === @idx_mat_flu[indice+1] - 1
 				reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
-      	pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
+      	#pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
 			}
 			
-			f_main = For(i,1,@Nflu_inner+@Nflusol_inner-modulo(@Nflu_inner+@Nflusol_inner,unroll), step: unroll){
[…]

### 2017-06-19 hou10ni/run.rb

@@ -11,49 +11,67 @@ include BOAST
 
 class Experiment
 
- def self.run()
+	def self.run()
 
- # set_default_int_size(8)
+  # set_default_int_size(8)
   set_default_real_size(4)
 
-  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}
-  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O3"}  
-
+	opt_space = OptimizationSpace::new( optim_nested: 1..10,
+                                    optim_main: 1..10,
+                                    OFLAGS: ["-O2", "-O3"]
+                                    )
   set_lang(FORTRAN)
   kernels={}
   stats={}
+	repeat = 10
 
-  # Creating ref kernel
+  #.. REF KERNEL ..#
+
+  k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
   kernels[k_ref_params] = KRef::new(k_ref_params)
   kernels[k_ref_params].generate
-	puts "* Building Ref kernel ...\n"
-  #puts kernels[k_ref_params].kernel
   kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] ) 
   stats[k_ref_params]={:time => []}
 
-  # Creating boast kernel
-  kernels[k_boast_params] = KBoast::new(k_boast_params)
-  kernels[k_boast_params].generate
-	puts "* Building BOAST kernel ...\n"
-  #puts kernels[k_boast_params].kernel
-  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
-  stats[k_boast_params]={:time => []}
 
-  # Run (inputs are the same for ref and boast kernels) 
-	puts "\n* Testing boast and ref versions of the kernel with fluid_inner_elt_boast/sub_time_loop_1/ inputs parameters ...\n"
-  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
-  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
[…]

### 2017-06-19 hou10ni/run_boast.rb

@@ -9,13 +9,13 @@ class Experiment
  def self.run()
 
 	optim_nested = 3
-	optim_main = 7
+	optim_main = 1
  
   #set_default_int_size(8)
   set_default_real_size(4)
 
 	# Parameters - optim = number of unroll of the nested loop per iteration
-  k_boast_params = {:kernel => :boast, :optim_nested => optim_nested, :optim_main => optim_main, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}  
+  k_boast_params = {:kernel => :boast, optim_nested: optim_nested, optim_main: optim_main, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}  
 
   set_lang(FORTRAN)
   kernels = {}
@@ -33,17 +33,16 @@ class Experiment
   inputs = kernels[k_boast_params].kernel.load_ref_inputs()
   outputs = kernels[k_boast_params].kernel.load_ref_outputs()
 
-	repeat.times {
-  	inputs.each_key { |key|
-			puts key
+  inputs.each_key { |key|
+		repeat.times {
  			stats.push kernels[k_boast_params].kernel.run(*(inputs[key])) 
- 			puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect 
-			puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
 		}
+		puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
 	}
-  stats.sort_by! { |a| a[:duration] }
-	stats = stats.first
 
+  stats.sort_by! { |a| a[:duration] }
+	p stats
+	stats = stats.first
 
   puts "#{kernels[k_boast_params].kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
   return kernels

### 2017-06-19 hou10ni/run_ref.rb

@@ -7,7 +7,7 @@ include BOAST
 class Experiment
 
  def self.run()
-  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}  
+  k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -O2"}  
 
   set_lang(FORTRAN)
   kernels={}

### 2017-06-19 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni's kernel : Clean

@@
 
-class Experiment
-
- def self.run()
-  k_ref_params = {:kernel => :ref, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check"}
-
  set_lang(FORTRAN)
-  kernels={}
 
 # Creating ref kernel
-  kernels[k_ref_params] = KRef::new(k_ref_params)
-  kernels[k_ref_params].generate
- # Building ref kernel
-  puts kernels[k_ref_params].kernel
-  kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )
-
-  # Parameters for loop iteration 1 and MPI Process 0 (in the directory fluid_inner_elt_ref/sub_time_loop_1/)
-  inputs = kernels[k_ref_params].kernel.load_ref_inputs()
+k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -O2"}
+k = KRef::new(k_ref_params)
+k.generate
+puts k.kernel
+k.kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )
+
+# Parameters for loop iteration 1 and MPI Process 0 
+# in the directory fluid_inner_elt_ref/sub_time_loop_1/
+inputs = k.kernel.load_ref_inputs()
 puts "** inputs =", inputs
@@
 inputs.each_key { |key|
-    puts key
-    puts kernels[k_ref_params].kernel.run(*(inputs[key])).inspect
-    puts kernels[k_ref_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
-  }
-
-  return kernels
- end
-end
-
-Experiment.run
+  puts k.kernel.run(*(inputs[key]))
+  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
+}
+
+puts "Testing done\n"
 \end{lstlisting}
@@
 
-  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
-  @opts.update(options)
-
   # Parameters 
@@
   @matSize            = Int("matSize", :dir => :in )
-
   @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
@@
     call_matmul = lambda{|x,y|
-      if get_lang == FORTRAN
         return matmul(x,y)
-      end
     }
@@
 
-class Experiment
-
- def self.run()
-
- # set_default_int_size(8)
 set_default_real_size(4)
-  k_boast_params = {:kernel => :boast, :preprocessor => false, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none"}
+
+k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}
 
 set_lang(FORTRAN)
-  kernels = {}
-  stats_svg = {}
 stats = []
-  repeat = 3
-
-  # Creating ref kernel
-  kernels[k_boast_params] = KBoast::new(k_boast_params)
-  kernels[k_boast_params].generate
-  puts "*********** Building kernel ***********\n"
-  puts kernels[k_boast_params].kernel
-  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
-
-  puts "\n*********** Testing boast kernel with inputs parameters ***********\n"
-  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
-  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
-
+repeat = 5
+
+# Creating boast kernel
+k = KBoast::new(k_boast_params)
+k.generate
+puts k.kernel
+k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+
+inputs = k.kernel.load_ref_inputs()
+outputs = k.kernel.load_ref_outputs()
+
+inputs.each_key { |key|
   repeat.times {
-    inputs.each_key { |key|
-      puts key
-      stats.push kernels[k_boast_params].kernel.run(*(inputs[key]))
-      puts kernels[k_boast_params].kernel.run(*(inputs[key])).inspect
-      puts kernels[k_boast_params].kernel.compare_ref(outputs[key], inputs[key]).inspect
-    }
-  }
+    stats.push k.kernel.run(*(inputs[key]))
+  }
+  puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
+}
+
 stats.sort_by! { |a| a[:duration] }
+p stats
 stats = stats.first
 
-  puts "#{kernels[k_boast_params].kernel.procedure.name}: #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
-  return kernels
- end
-end
-
-Experiment.run
+puts "#{k.kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
 \end{lstlisting}
@@
 
-*********** Testing boast kernel with inputs parameters ***********
-./fluid_inner_elt_boast/sub_time_loop_1
-{:duration=>0.13263683, :start=>-4004705691753970176, :end=>-3872068861753970176}
 {"P_new"=>0.0}
-./fluid_inner_elt_boast/sub_time_loop_1
-{:duration=>0.140625219, :start=>-3642235174753970176, :end=>-3501609955753970176}
-{"P_new"=>0.0}
-./fluid_inner_elt_boast/sub_time_loop_1
-{:duration=>0.147113606, :start=>-3274643277753970176, :end=>-3127529671753970176}
-{"P_new"=>0.0}
-fluid_inner_elt_boast: 0.15179012 s ->  151.79012 ms  
-\end{lstlisting}
-
-\subsubsection*{Optimisations}
+[{:duration=>0.011553553000000001, :start=>-8096482666880853504, :end=>-8084929113880853504}, {:duration=>0.011923121, :start=>-8122491033880853504, :end=>-8110567912880853504}, {:duration=>0.013262399000000001, :start=>-8135791916880853504, :end=>-8122529517880853504}, {:duration=>0.013991319, :start=>-8110525936880853504, :end=>-8096534617880853504}, {:duration=>0.014984313, :start=>-8150824959880853504, :end=>-8135840646880853504}]
+fluid_inner_elt_boast: optim_nested = 3  0.011553553000000001 s ->  11.553553 ms
+\end{lstlisting}
+
+\subsection{Optimisations}
+
+Unroll both loops.
+
+\lstset{language=RUBY}
+\begin{lstlisting}[caption=KBoast.rb]
+class KBoast
+ attr_reader :kernel
+
+ def initialize(options)
+
+  @opts = {:optim_nested => 1, :optim_main => 1}
+  @opts.update(options)
+
+  # Parameters 
+  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
+  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
+  @nb_rhs             = Int("nb_rhs", :dir => :in )
+  @afluSize           = Int("afluSize", :dir => :in )
+  @pSize              = Int("pSize", :dir => :in )
+  @vecSize_1          = Int("vecSize_1", :dir => :in )
+  @vecSize_2          = Int("vecSize_2", :dir => :in )
+  @matSize            = Int("matSize", :dir => :in )
+  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
+
+ end
+
+  def generate
+
+   push_env(:lang => FORTRAN) {
+
+   @kernel = CKernel:: new()
+   @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+
+    #..  FUNCTIONS ..#
+
+    register_funccall("matmul") if get_lang == FORTRAN
+    call_matmul = lambda{|x,y|
+        return matmul(x,y)
+    }
+
+    register_funccall(:modulo) if get_lang == FORTRAN
+
+    #.. KERNEL ..#
+
+    opn @kernel.procedure
+
+    # local variables
+    i = Variable::new('I', Int)
+    j = Variable::new('J', Int)
+    i_tmp1 = Variable::new('I_tmp1', Int)
+    i_tmp2 = Variable::new('I_tmp2', Int)
+    i1 = Variable::new('I1', Int)
+    i2 = Variable::new('I2', Int)
+    i3 = Variable::new('I3', Int)
+    i4 = Variable::new('I4', Int)
+    p_aux = Variable::new("P_aux",Real, :dimension => [Dim(20),Dim(@nb_rhs)])
+
+    jmax = Variable::new('jmax', Int)
+    imax = Variable::new('imax', Int)
+
+    decl i,j,i_tmp1,i_tmp2, i1,i2,i3,i4,p_aux
+    decl jmax,imax
+
+    ##.. Nested loop optimization
+    optim_1 = lambda { |unroll,i|
+      loop_1 = lambda { |indice|
+        pr i1 === @idx_vec_flu[2*indice,i]
+        pr i2 === @idx_vec_flu[2*indice+1,i]
+        pr i_tmp2 === i_tmp1 + i2 - i1
+        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs)
+        pr i_tmp1 === i_tmp2 + 1.to_var
+      }
+
+      pr jmax === @idx_vec_flu[1,i]-(unroll-1)
+      f_nested = For(j,1,jmax, step: unroll){
+        unroll.times { |k|
+          loop_1[j+k]
+        }
+      }
+      f_arr_nested = [f_nested]
+      if unroll > 1 then
+          f2_nested = For(j,(@idx_vec_flu[1,i]-modulo(@idx_vec_flu[1,i],unroll))+1, @idx_vec_flu[1,i]){
+            loop_1[j]
+          }
+          f_arr_nested.push f2_nested
+      end
+      f_arr_nested
+   }
+
+    ##.. Main loop optimization
+    optim_2 = lambda { |unroll|
+      loop_2 = lambda { |indice|
+        pr i_tmp1 === 1
+        # Nested loop optimized - No optimization -> = optim_1[1,indice].each{ |f|  pr f }
+        optim_1[@opts[:optim_nested],indice].each { |f_nested|
+          pr f_nested
+        }
+        pr i1 === @idx_vec_flu[2,indice]
+        pr i2 === @idx_vec_flu[3,indice]
+        pr i3 === @idx_mat_flu[indice]
+        pr i4 === @idx_mat_flu[indice+1] - 1
+        reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
+        pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)
+      }
+
+      pr imax === @Nflu_inner+@Nflusol_inner-(unroll-1)
+      f_main = For(i,1,imax, step: unroll){
+        unroll.times { |k|
+          loop_2[i+k]
+        }
+      }
+      f_arr_main = [f_main]
+      if unroll > 1 then
+        f2_main = For(i,(@Nflu_inner+@Nflusol_inner-modulo(@Nflu_inner+@Nflusol_inner,unroll))+1, @Nflu_inner+@Nflusol_inner){
+          loop_2[i]
+        }
+        f_arr_main.push f2_main
+      end
+      f_arr_main
+    }
+
+    ## Procedure
+    optim_2[@opts[:optim_main]].each { |f_main|
+        pr f_main
+    }
+
+    close @kernel.procedure
+   }
+
+   return @kernel
+ end    
+end 
+\end{lstlisting}
 

### 2017-06-19 hou10ni/KBoast.rb

@@ -4,7 +4,7 @@ class KBoast
 
  def initialize(options)
 
-  @opts = {:optim_nested => 1, :optim_main => 1, :preprocessor => false, :unroll => false, :inline => :included}
+  @opts = {:optim_nested => 1, :optim_main => 1}
   @opts.update(options)
 
   # Parameters 
@@ -95,7 +95,6 @@ class KBoast
 
      @kernel = CKernel:: new()
      @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
-		 #puts "Parameters = ", @kernel.procedure.parameters
 
 
 		#..  FUNCTIONS ..#
@@ -137,18 +136,6 @@ class KBoast
 
 		##.. Nested loop optimization
 
-		#  Option 0 :
-		no_optim_nested = For(j,1,@idx_vec_flu[1,i]){
-        pr i1 === @idx_vec_flu[2*j,i]
-        pr i2 === @idx_vec_flu[2*j+1,i]
-        pr i_tmp2 === i_tmp1 + i2 - i1
-        pr p_aux.slice(i_tmp1..i_tmp2,1..@nb_rhs) === @P_inter.slice(i1..i2,1..@nb_rhs) 
-        pr i_tmp1 === i_tmp2 +1
-		}
-
-		#  Option 1 : Unroll 'unroll' iterations
-
-
    	optim_1 = lambda { |unroll,i|
 			loop_1 = lambda { |indice|
 				pr i1 === @idx_vec_flu[2*indice,i]
@@ -177,19 +164,6 @@ class KBoast
 
 		##.. Main loop optimization
 
-		#  Option 0 : No optimization
-		no_optim_main = For(i,1,@Nflu_inner+@Nflusol_inner){
-			pr i_tmp1 === 1
-			pr no_optim_nested
-			pr i1 === @idx_vec_flu[2,i]
-      pr i2 === @idx_vec_flu[3,i]
-      pr i3 === @idx_mat_flu[i]
-      pr i4 === @idx_mat_flu[i+1] - 1
[…]

### 2017-06-19 hou10ni/run_boast.rb

@@ -4,49 +4,38 @@ require 'narray_ffi'
 
 include BOAST
 
-class Experiment
-
- def self.run()
-
-	optim_nested = 3
-	optim_main = 1
+optim_nested = 3
+optim_main = 1
  
-  #set_default_int_size(8)
-  set_default_real_size(4)
+#set_default_int_size(8)
+set_default_real_size(4)
 
-	# Parameters - optim = number of unroll of the nested loop per iteration
-  k_boast_params = {:kernel => :boast, optim_nested: optim_nested, optim_main: optim_main, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}  
+# Parameters - optim = number of unroll of the nested loop per iteration
+k_boast_params = {:kernel => :boast, optim_nested: optim_nested, optim_main: optim_main, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}  
 
-  set_lang(FORTRAN)
-  kernels = {}
-	stats = []
-  repeat = 5
+set_lang(FORTRAN)
+stats = []
+repeat = 5
 
-  # Creating boast kernel
-  kernels[k_boast_params] = KBoast::new(k_boast_params)
-  kernels[k_boast_params].generate
-	puts "*********** Building kernel ***********\n"
-  puts kernels[k_boast_params].kernel
-  kernels[k_boast_params].kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+# Creating boast kernel
+k = KBoast::new(k_boast_params)
+k.generate
+puts k.kernel
+k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
  
-	puts "\n*********** Testing boast kernel with fluid_inner_elt_boast/sub_time_loop_1/ inputs parameters ***********\n"
-  inputs = kernels[k_boast_params].kernel.load_ref_inputs()
-  outputs = kernels[k_boast_params].kernel.load_ref_outputs()
+inputs = k.kernel.load_ref_inputs()
+outputs = k.kernel.load_ref_outputs()
 
-  inputs.each_key { |key|
-		repeat.times {
- 			stats.push kernels[k_boast_params].kernel.run(*(inputs[key])) 
-		}
[…]

### 2017-06-19 hou10ni/run_ref.rb

@@ -4,49 +4,37 @@ require 'narray_ffi'
 
 include BOAST
 
-class Experiment
-
- def self.run()
-  k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -O2"}  
-
-  set_lang(FORTRAN)
-  kernels={}
-
-  # Creating ref kernel
-  kernels[k_ref_params] = KRef::new(k_ref_params)
-  kernels[k_ref_params].generate
-	puts "*********** Building kernel ***********\n"
-  puts kernels[k_ref_params].kernel
-  kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] )
- 
-  #puts "Parameters = ", kernels[k_ref_params].kernel.procedure.parameters
-
-	# Parameters for loop iteration 1 and MPI Process 0
-  puts "** inputs from the application ="
-  puts "{ ./fluid_inner_elt_ref/sub_time_loop_1 =>[210025, 0, 1, NArray.int(2577732): "
-  puts "[ 5, 1, 4, 538597, 538600, 319233, ...], NArray.int(214812):"
-  puts "[ 1, 81, 161, 209, 289, 369, ...],  NArray.float(879732):"
-  puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
-  puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
-  puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(16952560):"
-  puts "[1.877937, 1.1007047E-02, 1.0670856E-02, 1.9009672E-02, 1.1590085E-02, ...], 16952560, 879732, 12, 214811, 214812]} " 
-
-	puts "\n*********** Testing ref kernel with fluid_inner_elt_ref/sub_time_loop_1/ inputs parameters ***********\n"
-  inputs = kernels[k_ref_params].kernel.load_ref_inputs()
[…]

### 2017-06-28 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni's kernel : Add OpenMP

@@
 
-\title{Hou10ni Profiling with TAU}
+\title{Hou10ni}
 \date{}
@@
 
-This document presents some results obtained with the performance evaluation tool TAU on the application hou10ni.
+This document presents a study of hou10ni's kernel. We present results obtained with the performance evaluation tools TAU and Paje on the application.
+
+\tableofcontents
+
+\newpage
+
+\section{Build and launch hou10ni}
+
+Compilation:
+\begin{cverbatim}
+mkdir build && cd build
+cmake ..
+ccmake ..
+make
+cd .. 
+\end{cverbatim}
+
+\noindent Execution with 4 MPI processes:
+\begin{cverbatim}
+mpirun -np 4 build/hou10ni.out < HPC4E.dat
+\end{cverbatim}
 
@@
 
-\section{Trace Analysis with Lucas and Arnaud's tool}
+\section{Trace Analysis with PAJE}
 

### 2017-06-28 hou10ni/KBoast.rb

@@ -169,7 +169,10 @@ class KBoast
 				pr i_tmp1 === 1
 				# Nested loop optimized - No optimization -> = optim_1[1,indice].each{ |f|  pr f }
 				optim_1[@opts[:optim_nested],indice].each { |f_nested|
+					pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads]){
+					pr OpenMP::For()
           pr f_nested
+					}
         }
 				pr i1 === @idx_vec_flu[2,indice]
       	pr i2 === @idx_vec_flu[3,indice]
@@ -199,7 +202,10 @@ class KBoast
 
 		## Procedure
 		optim_2[@opts[:optim_main]].each { |f_main|
+				#pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads]){
+				#pr OpenMP::For()
         pr f_main
+				#}
     }
 
 		close @kernel.procedure

### 2017-06-28 hou10ni/run.rb

@@ -16,18 +16,20 @@ class Experiment
   # set_default_int_size(8)
   set_default_real_size(4)
 
-	opt_space = OptimizationSpace::new( optim_nested: 1..10,
-                                    optim_main: 1..10,
-                                    OFLAGS: ["-O2", "-O3"]
+	opt_space = OptimizationSpace::new( optim_nested: 5..10,
+                                    optim_main: 5..10,
+                                    #OFLAGS: ["-O2", "-O3"], 
+                                    OFLAGS: ["-O2"], 
+																		:omp_num_threads => 1..4
                                     )
   set_lang(FORTRAN)
   kernels={}
   stats={}
-	repeat = 10
+	repeat = 1
 
   #.. REF KERNEL ..#
 
-  k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
+  k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fimplicit-none #{opt[:OFLAGS]}"}
   kernels[k_ref_params] = KRef::new(k_ref_params)
   kernels[k_ref_params].generate
   kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] ) 
@@ -35,10 +37,12 @@ class Experiment
 
 
 	# input and output parameters to check the kernel
+	# found in fluid_inner_elt_ref
   inputs = kernels[k_ref_params].kernel.load_ref_inputs()
-  outputs = kernels[k_ref_params].kernel.load_ref_outputs()
+  output = kernels[k_ref_params].kernel.load_ref_outputs()
 
 
+	# As many keys as directories in fluid_inner_elt_ref/
   inputs.each_key { |key|
 		repeat.times{|i|
  			stats[k_ref_params][:time][i] = kernels[k_ref_params].kernel.run(*(inputs[key]))[:duration]
@@ -54,7 +58,7 @@ class Experiment
 	optimizer = BruteForceOptimizer::new(opt_space, :randomize => true)
 	puts optimizer.optimize { |opt|
 		p opt
[…]

### 2017-06-28 hou10ni/run_boast.rb

@@ -6,12 +6,13 @@ include BOAST
 
 optim_nested = 3
 optim_main = 1
- 
+num_threads = 2
+
 #set_default_int_size(8)
 set_default_real_size(4)
 
 # Parameters - optim = number of unroll of the nested loop per iteration
-k_boast_params = {:kernel => :boast, optim_nested: optim_nested, optim_main: optim_main, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}  
+k_boast_params = {:kernel => :boast, omp_num_threads: num_threads, optim_nested: optim_nested, optim_main: optim_main, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}  
 
 set_lang(FORTRAN)
 stats = []
@@ -21,10 +22,13 @@ repeat = 5
 k = KBoast::new(k_boast_params)
 k.generate
 puts k.kernel
-k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+k.kernel.build(:OPENMP => true, :LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
  
 inputs = k.kernel.load_ref_inputs()
 outputs = k.kernel.load_ref_outputs()
+puts "** inputs =", inputs
+puts "** outputs=",outputs
+
 
 inputs.each_key { |key|
 	repeat.times {

### 2017-06-28 hou10ni/run_ref.rb

@@ -16,16 +16,15 @@ k.kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCF
 
 
 # Parameters for loop iteration 1 and MPI Process 0
-puts "** inputs from the application ="
-puts "{ ./fluid_inner_elt_ref/sub_time_loop_1 =>[210025, 0, 1, NArray.int(2577732): "
-puts "[ 5, 1, 4, 538597, 538600, 319233, ...], NArray.int(214812):"
-puts "[ 1, 81, 161, 209, 289, 369, ...],  NArray.float(879732):"
-puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
-puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
-puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(16952560):"
-puts "[1.877937, 1.1007047E-02, 1.0670856E-02, 1.9009672E-02, 1.1590085E-02, ...], 16952560, 879732, 12, 214811, 214812]} " 
+#puts "** inputs from the application ="
+#puts "{ ./fluid_inner_elt_ref/sub_time_loop_1 =>[210025, 0, 1, NArray.int(2577732): "
+#puts "[ 5, 1, 4, 538597, 538600, 319233, ...], NArray.int(214812):"
+#puts "[ 1, 81, 161, 209, 289, 369, ...],  NArray.float(879732):"
+#puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
+#puts "[ 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.0, ... ], NArray.float(879732):" 
[…]

### 2017-06-28 Paper2/BOAST.tex
Updated Bibliography to use Harvard style.

@@
 \usepackage{xspace}
+\usepackage{natbib}
 
@@
 
-\bibliographystyle{IEEEtran}
+\bibliographystyle{SageH}
 
@@
 
-\markboth{International Journal of Scientific Computing}
+\markboth{The International Journal of High Performance Computing Applications}
 {Videau \MakeLowercase{\textit{et al.}}: BOAST: a Metaprogramming Framework to Produce Portable and Efficient Computing Kernels for HPC Applications}
@@
 short, separable filters.  The detailed description of how these operations are
-defined is beyond the scope of this paper and can be found in the BigDFT
+defined can be found in the BigDFT
 reference paper~\cite{genovese2008daubechies}. Convolutions are basic operations
@@
 binary rather than the source code. This experience therefore would be a first
-step toward the release of a tunable optimized convolution library oriented to
-computational physics communities.
+step toward the release of a tunable optimized convolution library oriented toward
+computational physics.
 
@@
 Software} methodology that we implemented with BOAST. Their kernel generation is
-performed using macro-functions in C.
+performed using C macro-functions.
 

### 2017-08-30 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni: kernel update

@@
 \begin{lstlisting}[caption=kernel.f90]
-SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
+SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
   implicit none
@@
 ...
-CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
+CALL fluid_inner_elt(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
 ...
@@
 
-We generate the exact same kernel as a reference version and run it with BOAST in order to compare the results in term of correctness and performances.
-KRef defines the reference version.
+We generate the same kernel as a reference version (KRef ) and run it with BOAST in order to compare the results in term of correctness and performances.
 This version simply needs to keep the Fortran code (referred as \texttt{codeF}) and wrap it with BOAST.
-To put it in a nutshell, a reference version should follow the template bellow.
+A reference version should follow the template bellow.
 
@@
   kernel = BOAST::CKernel::new
-  function_name = "function_ref"
-
-  VARIABLES HERE
-  FOR EXAMPLE : 
+  kernel_name = "function_ref"
+
+  # ARGUMENTS OF KERNEL HERE
+  # FOR EXAMPLE : 
   geocode = BOAST::Int("geocode", :dir => :in )
-  n01 = BOAST::Int("n01", :dir => :in)
   u = BOAST::Real("u", :dir => :in, :dim => [ BOAST::Dim(0, n01),  BOAST::Dim(0, n02))] )
 
-   p = BOAST::Procedure::new(function_name, [geocode,n01,u])
+   p = BOAST::Procedure::new(kernel_name, [geocode,u])
 
    kernel.code.print <<EOF
-  ORIGINAL C or FORTRAN CODE HERE
+  # ORIGINAL C or FORTRAN CODE HERE
 EOF
@@
  def initialize(options)
-  @opts = {:preprocessor => false, :unroll => false, :inline => :included}
-  @opts.update(options)
   # Parameters 
@@
      @kernel = CKernel:: new()
-     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
      get_output.print codeF
@@
 # Creating ref kernel
-k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -O2"}
+k_ref_params = {:kernel => :ref, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
 k = KRef::new(k_ref_params)
@@
 
-# Parameters for loop iteration 1 and MPI Process 0 
-# in the directory fluid_inner_elt_ref/sub_time_loop_1/
+# Input parameters 
+# found in the directory fluid_inner_elt_ref/sub_time_loop_1/ for example (created by the user)
 inputs = k.kernel.load_ref_inputs()
@@
 
+# Launch the kernel fluid_inner_elt_ref
+# As many key as directories in 
 inputs.each_key { |key|
@@
 \begin{lstlisting}[caption=Results (obtained with the command ruby run\_ref.rb)]
-SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
-&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
-&ecSize_1, vecSize_2, matSize)
+SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, afluSize, pSize, vecSize_1, &
+&vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
   implicit none
@@
     @kernel = CKernel:: new()
-    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+    @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
 
@@
 require 'narray_ffi'
-
 include BOAST
 
+k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-O2"}
+
 set_default_real_size(4)
-
-k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fbounds-check -fimplicit-none -O2"}
-
 set_lang(FORTRAN)
@@
 
+# Get input and output parameters # found in the directory fluid_inner_elt_boast/
 inputs = k.kernel.load_ref_inputs()
@@
 
+# Launch the kernel
 inputs.each_key { |key|
@@
 
+# Output
 puts "#{k.kernel.procedure.name}: optim_nested = #{k_boast_params[:optim_nested]}  #{stats[:duration]} s ->  #{stats[:duration]*1.0e3} ms"
@@
 \begin{lstlisting}[caption=Results (obtained with the command ruby run\_boast.rb)]
-SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, nb_rhs, idx_&
-&vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, v&
-&ecSize_1, vecSize_2, matSize)
-
+SUBROUTINE fluid_inner_elt_boast(Nflu_inner, Nflusol_inner, afluSize, pSize, vecSize_1, &
+&vecSize_2, matSize, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
   integer, parameter :: wp=kind(1.0d0)
@@
   ... 
-
 END SUBROUTINE fluid_inner_elt_boast
 
-{"P_new"=>0.0}
+{"P_new"=>0.0}  # No difference with the output
+
 [{:duration=>0.011553553000000001, :start=>-8096482666880853504, :end=>-8084929113880853504}, {:duration=>0.011923121, :start=>-8122491033880853504, :end=>-8110567912880853504}, {:duration=>0.013262399000000001, :start=>-8135791916880853504, :end=>-8122529517880853504}, {:duration=>0.013991319, :start=>-8110525936880853504, :end=>-8096534617880853504}, {:duration=>0.014984313, :start=>-8150824959880853504, :end=>-8135840646880853504}]
@@
 
-Unroll both loops.
+The following code shows how we optimized the kernel with loop unrolling and using OpenMP.
 
@@
    @kernel = CKernel:: new()
-   @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+   @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
 
@@
     optim_2[@opts[:optim_main]].each { |f_main|
+        pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads]){
+        pr OpenMP::For(:schedule => "static")
         pr f_main
+        }
     }
@@
 
-\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
-TODO
+\subsubsection{Results}
+
+The results were obtained using a Macbook Pro Intel Core i7 and the GNU fortran compiler. Values correspond to the minimum value of 20. Optimization flag: -O2
+When \texttt{unroll main} and \texttt{unroll nested} are equal to 1, this means no unrolling is done.
+
+\begin{table}[ht!]
+\centering
+\caption{Time depending on the number of OpenMP threads used for the main loop. No loop unrollment.}
+{\scriptsize
+\begin{tabular}{|c|c|c|} \hline 
+\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	  &  {{\bf{\# Threads}}}    &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
+					 		  & 	 	[1..6]   &  				\\ \hline \hline
+ref  &         -  &  {\bf{83,769018}} \\ \hline
+boast &     1 & 90,608141 \\ \hline
+boast &      2 & 87,766917 \\  \hline
+boast &    3 & 90,69217800 \\ \hline
+boast &     4 & 87,29976 \\  \hline
+boast &    5 & 84,527453 \\  \hline
+boast &   6 & {\bf{81,92376500}} \\ \hline
+\end{tabular}
+}
+\label{tab:1}
+\end{table}
+
+\begin{table}[ht!]
+\centering
+\caption{Statistics on time depending on the number of iterations unrolled for the main loop.}
+{\scriptsize
+\begin{tabular}{|c|c|c|c|c|} \hline 
+\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 &   {{\bf{unroll main}}}   & {{\bf{unroll nested}}}   &    \multirow{2}{*}{ {{\bf{\# Threads}}} }  &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\ 
+							 &    	[1..3]		 	    &   [1..10]	 		 & 							      &  				\\ \hline \hline
+boast &    1 &  &  & 90,608141 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 1 & 1 & 88,888357 \\  \cline{1-2}  \cline{5-5}
+boast &    3 &  & & 89,98507500 \\ \hline \hline
+
+boast &   1 &  &  & 88,516961 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 2 & 1 & 87,892416 \\ \cline{1-2}  \cline{5-5}
+boast &    3 &  &  & 86,117465 \\ \hline \hline
+
+boast &    1 &  &  & 86,232686 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 3 & 1 & 86,40083400 \\ \cline{1-2}  \cline{5-5}
+boast &   3 &  &  & 82,334101 \\ \hline \hline
+
+boast &    1 &  &  & 88,345071000 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 4 & 1 & 82,89916300 \\ \cline{1-2}  \cline{5-5}
+boast &    3 &  &  & 82,646548 \\ \hline \hline
+
+boast &    1 & &  & 91,76817400 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 5 & 1 & 89,60873700 \\ \cline{1-2}  \cline{5-5}
+boast &   3 &  &  & 83,658755 \\ \hline \hline
+
+boast &   1 & &  & 92,428135000 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 6 & 1 & 84,5453400 \\ \cline{1-2}  \cline{5-5}
+boast &  3 &  &  & 87,82848900 \\ \hline \hline
+
+boast &    1 &  &  & 85,81414200 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 7 & 1 & 89,83438 \\ \cline{1-2}  \cline{5-5}
+boast &   3 &  & & {\bf{81,77698900}}  \\ \hline \hline
+
+boast &    1 &  &  & 85,221406 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 8 & 1 & 85,728041 \\ \cline{1-2}  \cline{5-5}
+boast &   3 &  &  & 87,331661 \\ \hline \hline
+
+boast &   1 &  &  & 88,48286500 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 9 & 1 & 87,05563300 \\ \cline{1-2}  \cline{5-5}
+boast &    3 &  &  & 86,669093 \\ \hline \hline
+
+boast &    1 &  &  & 92,284591 \\ \cline{1-2}  \cline{5-5}
+boast &    2 & 10 & 1 & 85,477758 \\ \cline{1-2}  \cline{5-5}
+boast &   3 & &  & 85,488478 \\ \hline
+\end{tabular}
+}
+\label{tab:2}
+\end{table}
+
+\begin{table}[ht!]
+\centering
+\caption{Time depending on the number of iterations unrolled for the nested loop.}
+{\scriptsize
+\begin{tabular}{|c|c|c|c|c|} \hline 
+\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  {{\bf{unroll nested}}}   &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time (ms)}}} }  \\
+						 		 &   		 					    &  	 		[1..10]				           & 	 						      &  				\\ \hline \hline
+boast &     & 1 &  & 90,608141 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &     & 2 &  & 88,516961 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &    & 3 &  & 86,232686 \\ \cline{1-1}  \cline{3-3}  \cline{5-5} 
+boast &   & 4 &  & 88,345071000 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &   1 & 5 & 1 & 91,76817400 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &     & 6 &  & 92,42813500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &    & 7 &  & 85,81414200 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &     & 8 &  & {\bf{85,221406}} \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &    & 9 &  & 88,48286500 \\ \cline{1-1}  \cline{3-3}  \cline{5-5}
+boast &     & 10 &  & 92,284591 \\ \hline
+\end{tabular}
+}
+\label{tab:3}
+\end{table}
+
+\begin{table}[ht!]
+\centering
+\caption{All values under the reference version.}
+{\scriptsize
+\begin{tabular}{|c|c|c|c|c|} \hline 
+\multirow{2}{*}{ {{\bf{Kernel vers.}}} }	 	 &  \multirow{2}{*}{ {{\bf{unroll main}}} }  &  \multirow{2}{*}{ {{\bf{unroll nested}}} }  &  \multirow{2}{*}{ {{\bf{\# Threads}}} }   &  \multirow{2}{*}{ {{\bf{Time}}} }  \\
+						 		 &    							    &  	 							  & 	 						      &  				\\ \hline \hline
+ref    &   &   &   &  0.083769018 \\ \hline \hline
+
+boast &    3 & 7 & 1 & {\bf{81,77698900}} \\ \hline
+boast &   2 & 3 & 5 & 82,006891 \\ \hline
+boast &   3 & 4 & 3 & 82,17091900\\ \hline
+boast &   3 & 3 & 1 & 82,334101 \\ \hline
+boast &   3 & 4 & 1 & 82,646548 \\ \hline
+boast &    2 & 4 & 5 & 82,809038 \\ \hline
+boast &    2 & 4 & 4 & 82,81120700 \\ \hline
+boast &    3 & 3 & 2 & 82,818522 \\ \hline
+boast  &    1 & 5 & 4 & 82,866756 \\ \hline
+boast &    2 & 4 & 1 & 82,89916300 \\ \hline
+boast &    3 & 2 & 3 & 82,949804 \\ \hline
+boast  &    1 & 3 & 4 & 82,994519 \\ \hline
+boast &    1 & 2 & 3 & 83,047285 \\ \hline
+boast &    3 & 7 & 2 & 83,08922 \\ \hline
+boast  &    2 & 3 & 6 & 83,09479700 \\ \hline
+boast &    3 & 4 & 6 & 83,137816 \\ \hline
+boast &    3 & 1 & 3 & 83,15446500 \\ \hline
+boast &    3 & 5 & 4 & 83,33983 \\ \hline
+boast &    2 & 7 & 6 & 83,558517 \\ \hline
+boast &    2 & 7 & 2 & 83,57951500 \\ \hline
+\end{tabular}
+}
+\label{tab:4}
+\end{table}
+
+%\section{Comparison between MPI and MPI+OpenMP versions of hou10ni}
+%TODO
 

### 2017-08-30 Alya/src/boast/ASGS/ASGS_Arguments.rb

@@ -1,3 +1,0 @@
-module ASGS_Arguments
-  
-end

### 2017-08-30 Alya/src/boast/ASGS/Debug.rb

@@ -1,37 +1,0 @@
-require '../Common/CommonArgs.rb'
-require './KASGSRef.rb'
-require './KASGSBoast.rb'
-
-class Debug
-  include CommonArgs
-  def self.run
-    vector_length = 1
-    seed = 1
-    nests = (1..23).to_a
-    nb_instances = 1
-
-    pgaus = 6
-    pnode = 6
-    dimension = 2
-
-    kernels = []
-
-    CommonArgs.init(pgaus,pnode,dimension,vector_length,nb_instances,seed)
-    k_orig_params = {:vector_length => vector_length, :preprocessor => false, :nests => nests}
-    
-    # set_lang(FORTRAN)
-    # set_fortran_line_length(300)
-    # kernels[0] = KASGSRef::new(k_orig_params)
-    # kernels[0].generate
-    # kernels[0].kernel.build(:FCFLAGS => "-ffree-line-length-none -x f95-cpp-input -O3")
-
-    set_lang(FORTRAN)
-    kernels[1] = KASGSBoast::new(k_orig_params)
-    kernels[1].generate
-    puts kernels[1].kernel
-    kernels[1].kernel.build(:FCFLAGS => "-ffree-line-length-none -x f95-cpp-input -O3")
-    
-  end
-end
-
-Debug.run

### 2017-08-30 Alya/src/boast/ASGS/KASGS.rb

@@ -1,97 +1,0 @@
-require '../Common/Parameters.rb'
-
-class KASGS
-  include Parameters
-  attr_reader :kernel
-
-  def initialize(options)
-    @opts = {:vector_length => 1, :preprocessor => false, :nests => (1..10).to_a, :unroll => false, :inline => :included}
-    @opts.update(options)
-    
-    Parameters.initialize(@opts[:vector_length])
-
-    @args = []
-    @args.push @ndime              = $ndime
-    @args.push @kfl_lumped         = $kfl_lumped
-    @args.push @mnode              = $mnode
-    @args.push @ntens              = $ntens
-    @args.push @kfl_duatss         = $kfl_duatss
-    @args.push @fact_duatss        = $fact_duatss
-    @args.push @kfl_stabi_nsi      = $kfl_stabi_nsi
-    @args.push @fvins_nsi          = $fvins_nsi
-    @args.push @fcons_nsi          = $fcons_nsi
-    @args.push @bemol_nsi          = $bemol_nsi
-    @args.push @kfl_regim_nsi      = $kfl_regim_nsi
-    @args.push @fvela_nsi          = $fvela_nsi
-    @args.push @kfl_rmom2_nsi      = $kfl_rmom2_nsi
-    @args.push @kfl_press_nsi      = $kfl_press_nsi
-    @args.push @kfl_p1ve2_nsi      = $kfl_p1ve2_nsi
-    @args.push @kfl_linea_nsi      = $kfl_linea_nsi
-    @args.push @pabdf_nsi          = $pabdf_nsi
-    @args.push @kfl_confi_nsi      = $kfl_confi_nsi
-    @args.push @nbdfp_nsi          = $nbdfp_nsi
-    @args.push @kfl_sgsti_nsi      = $kfl_sgsti_nsi
-    @args.push @kfl_nota1_nsi      = $kfl_nota1_nsi
-    @args.push @kfl_limit_nsi      = $kfl_limit_nsi
-    @args.push @kfl_penal_nsi      = $kfl_penal_nsi
-    @args.push @penal_nsi          = $penal_nsi
-    @args.push @kfl_bubbl_nsi      = $kfl_bubbl_nsi
-
-    @args.push @pnode              = $pnode
-    @args.push @pgaus              = $pgaus
-    @args.push @gpden              = $gpden
-    @args.push @gpvis              = $gpvis
-    @args.push @gppor              = $gppor
-    @args.push @gpgvi              = $gpgvi
[…]

### 2017-08-30 Alya/src/boast/ASGS/KASGSBoast.rb

@@ -1,145 +1,0 @@
-require_relative "./KASGS.rb"
-
-class KASGSBoast < KASGS
-  def set_arr(arr,value)
-    if get_lang == FORTRAN then
-      pr arr === value
-    else
-      get_output.print <<EOF
-      memset(#{arr}, 0, sizeof(#{arr.type.decl}) * #{arr.dimension.join("*")});
-EOF
-    end
-  end
-
-  def copy_arr(dst,src)
-    if get_lang == FORTRAN then
-      pr dst === src
-    else
-      get_output.print <<EOF
-      memcpy(#{dst}, #{src}, sizeof(#{dst.type.decl}) * #{dst.dimension.join("*")});
-EOF
-    end
-  end
-
-  def decl_max
-
-  end
-
-  def decl_local
-    @inode = $inode
-    @jnode = $jnode
-    @idofv = $idofv
-    @idof = $idof
-    @jdime = $jdime
-    @ivect = $ivect
-    @igaus = $igaus
-    @idime = $idime
-    @jdofv = $jdofv
-    @kdime = $kdime
-    @xvis2 = $xvis2
-    @xvisc = $xvisc
-    @one_rp = $one_rp
-    @p1ve2 = $p1ve2
-    @p1vec = $p1vec
-    @p2vec = $p2vec
-    @p2sca = $p2sca
-    @wgrvi = $wgrvi
-    @fact = $fact
-    @factx = $factx
-    @facx1 = $facx1
-    @ugraN = $ugraN
-    @gramugraN = $gramugraN
-    @penal = $penal
-    @gprhh = $gprhh
-    @taupr = $taupr
-    @gpveo = $gpveo
-    @gpcar1ji = $gpcar1ji
-    @gpcar2ji = $gpcar2ji
-    @gpcar3ji = $gpcar3ji
-    @p2sca_bub = $p2sca_bub
-    @p2vec_bub = $p2vec_bub
-    @elauu_a = $elauu_a
-    @factvec = $factvec
-
-    decl @inode,@jnode,@idofv,@idof,@jdime,@ivect,@igaus,@idime,
-         @jdofv,@kdime,@xvis2,@xvisc,@one_rp,@p1ve2,@p1vec,@p2vec,
-         @p2sca,@wgrvi,@fact,@factx,@facx1,@ugraN,@gramugraN,
-         @penal,@gprhh,@taupr,@gpveo,@gpcar1ji,@gpcar2ji,@gpcar3ji,
[…]

### 2017-08-30 Alya/src/boast/ASGS/KASGSRef.rb

@@ -1,1295 +1,0 @@
-require './KASGS.rb'
-
-class KASGSRef < KASGS
-  def gen_def_parameters
-    parameters =<<EOF
-      integer,     parameter  :: ip    = 4               ! 4-byte integer
-      integer,     parameter  :: rp    = 8               ! Double precision 
-
-      real(rp),    parameter  :: zeror = epsilon(1.0_rp) ! Almost zero
-      integer(ip), parameter  :: TET04 = 30              ! 3D 
-      integer(ip), parameter  :: TET10 = 31              ! 3D 
-      integer(ip), parameter  :: PYR05 = 32              ! 3D 
-      integer(ip), parameter  :: PYR14 = 33              ! 3D 
-      integer(ip), parameter  :: PEN06 = 34              ! 3D  
-      integer(ip), parameter  :: PEN15 = 35              ! 3D 
-      integer(ip), parameter  :: PEN18 = 36              ! 3D 
-      integer(ip), parameter  :: HEX08 = 37              ! 3D 
-      integer(ip), parameter  :: HEX20 = 38              ! 3D 
-      integer(ip), parameter  :: HEX27 = 39              ! 3D 
-      integer(ip), parameter  :: HEX64 = 40              ! 3D 
-      integer(ip), parameter  :: SHELL = 51              ! 3D shell element
-      integer(ip), parameter  :: BAR3D = 52              ! 3D bar element
-EOF
-    return parameters
-  end
-
-  def generate
-    nests = []
-    
-    macro =<<EOF
-#define VECTOR_SIZE #{@opts[:vector_length]}
-#ifdef OPENACC
-#define DEF_VECT ivect
-#else
-#define DEF_VECT 1:VECTOR_SIZE
-#endif
-EOF
-
-    subroutine_prototype =<<EOF
-    subroutine nsi_element_assembly_asgs_oss(ndime,     &
-       pnode,pgaus,gpden,gpvis,gppor,gpgvi,gpsp1,       &
-       gptt1,gpsp2,gptt2,gpvol,gpsha,gpcar,gplap,gphes, &
[…]

### 2017-08-30 Alya/src/boast/Common/CommonArgs.rb

@@ -1,128 +1,0 @@
-module CommonArgs
-  def self.init(pgaus,pnode,ndime,vector_length,nb_instances,seed = nil)
-    ANArray.srand(seed) unless seed.nil?
-    @@pgaus = pgaus
-    @@pnode = pnode
-    @@ndime = ndime
- 
-    @@mnode     = 8
-    @@ntens     = 3*@@ndime-3
-    @@ncomp_nsi = 3
-
-    @@kfl_lumped    = 0
-    @@kfl_duatss    = 0
-    @@fact_duatss   = 0.0
-    @@kfl_stabi_nsi = 2
-    @@fvins_nsi     = 1.0
-
-    @@fcons_nsi     = 0.0
-    @@bemol_nsi     = 0.0
-    @@kfl_regim_nsi = 0
-    @@fvela_nsi     = ANArray.float(8*vector_length,3).random!
-    @@kfl_rmom2_nsi = 0
-    @@kfl_press_nsi = 1
-    @@kfl_p1ve2_nsi = 0
-    @@kfl_linea_nsi = 2
-    @@pabdf_nsi     = 1.0
-    @@kfl_confi_nsi = 0
-    @@nbdfp_nsi     = @@ncomp_nsi-1
-    @@kfl_sgsti_nsi = 1
-    @@kfl_nota1_nsi = 0
-    @@kfl_limit_nsi = 0
-    @@kfl_penal_nsi = 0
-    @@penal_nsi     = 0.0
-    @@kfl_bubbl_nsi = 0
-
-    @@gpden     = ANArray.float(8*vector_length,vector_length,@@pgaus).random!
-    @@gpvis     = ANArray.float(8*vector_length,vector_length,@@pgaus).random!
-    @@gppor     = ANArray.float(8*vector_length,vector_length,@@pgaus).random! 
-    @@gpgvi     = ANArray.float(8*vector_length,vector_length,@@ndime,@@pgaus).random!
-    @@gpsp1     = ANArray.float(8*vector_length,vector_length,@@pgaus).random! 
-    @@gptt1     = ANArray.float(8*vector_length,vector_length,@@pgaus).random! 
-    @@gpsp2     = ANArray.float(8*vector_length,vector_length,@@pgaus).random!
-    @@gptt2     = ANArray.float(8*vector_length,vector_length,@@pgaus).random! 
-    @@gpvol     = ANArray.float(8*vector_length,vector_length,@@pgaus).random! 
[…]

### 2017-08-30 Alya/src/boast/Common/Parameters.rb

@@ -1,173 +1,0 @@
-
-require 'BOAST'
-include BOAST
-
-module Parameters
-  $ndime         = Int("ndime", :dir => :in )
-  $kfl_lumped    = Int("kfl_lumped",    :dir => :in)
-  $mnode         = Int("mnode",         :dir => :in)
-  $ntens         = Int("ntens",         :dir => :in)
-  $kfl_duatss    = Int("kfl_duatss",    :dir => :in)
-  $fact_duatss   = Int("fact_duatss",   :dir => :in)
-  $kfl_stabi_nsi = Int("kfl_stabi_nsi", :dir => :in)
-  $fvins_nsi     = Real("fvins_nsi",    :dir => :in)
-  $fcons_nsi     = Real("fcons_nsi",    :dir => :in)
-  $bemol_nsi     = Real("bemol_nsi",    :dir => :in)
-  $kfl_regim_nsi = Int("kfl_regim_nsi", :dir => :in)
-  $fvela_nsi     = Real("fvela_nsi",    :dir => :in, :dim => [Dim(3)])
-  $kfl_rmom2_nsi = Int("kfl_rmom2_nsi", :dir => :in)
-  $kfl_press_nsi = Int("kfl_press_nsi", :dir => :in)
-  $kfl_p1ve2_nsi = Int("kfl_p1ve2_nsi", :dir => :in)
-  $kfl_linea_nsi = Int("kfl_linea_nsi", :dir => :in)
-  $pabdf_nsi     = Real("pabdf_nsi",    :dir => :in, :dim => [Dim()])
-  $kfl_confi_nsi = Int("kfl_confi_nsi", :dir => :in)
-  $nbdfp_nsi     = Int("nbdfp_nsi",     :dir => :in)
-  $kfl_sgsti_nsi = Int("kfl_sgsti_nsi", :dir => :in)
-  $kfl_nota1_nsi = Int("kfl_nota1_nsi", :dir => :in)
-  $kfl_limit_nsi = Int("kfl_limit_nsi", :dir => :in)
-  $kfl_penal_nsi = Int("kfl_penal_nsi", :dir => :in)
-  $penal_nsi     = Real("penal_nsi",    :dir => :in)
-  $kfl_bubbl_nsi = Int("kfl_bubbl_nsi", :dir => :in)
-
-  $pnode     = Int("pnode",             :dir => :in)
-  $pgaus     = Int("pgaus",             :dir => :in)
-
-  $inode     = Int("inode")
[…]

### 2017-08-30 Alya/src/boast/Common/subroutine.rb

@@ -1,35 +1,0 @@
-
-require_relative './Parameters.rb'
-class Subroutine
-  include Parameters
-  attr_accessor :code
-  def initialize(name,options,functions)
-    opts = {:vector_length => 1, :unroll => false, :inline => :included}
-    opts.update(options)
-    @name = name
-    @functions = functions
-    @args = []
-    @code = nil
-    @vector_length = options[:vector_length]
-    @usage = opts[:inline]
-    @unroll = opts[:unroll]
-    @ndime = opts[:ndime]
-    @nb_original_param = instance_variables.length + 1
-  end
-  def construct(block)
-    if @usage == :included then
-      @code = block
-    else
-      functions = @functions.nil? ? nil : @functions.values
-      inlined = @usage == :inlined ? true : false
-      @code = Procedure(@name, @args, :inline => inlined, :functions => functions, &block)
-    end
-  end
-  def call
-    if @usage == :included then
-      @code.call
-    else
-      eval "pr @code.call( #{instance_variables[@nb_original_param...instance_variables.length].join(',')} )"
-    end
-  end
-end

### 2017-08-30 Alya/src/boast/Split/KSplitOss.rb

@@ -1,95 +1,0 @@
-
-require_relative '../Common/Parameters.rb'
-class KSplitOss
-  include Parameters
-  attr_reader :kernel
-
-  def declare_parameters(ndime = nil)
-    Parameters.initialize(@opts[:vector_length],ndime)
-    @args = []
-    if ndime.nil? then
-      @args.push @ndime       = $ndime
-    else
-      @ndime                  = ndime
-    end
-    @args.push @kfl_lumped    = $kfl_lumped
-    @args.push @mnode         = $mnode
-    @args.push @ntens         = $ntens
-    @args.push @kfl_duatss    = $kfl_duatss
-    @args.push @fact_duatss   = $fact_duatss
-    @args.push @kfl_stabi_nsi = $kfl_stabi_nsi
-    @args.push @fvins_nsi     = $fvins_nsi
-    @args.push @fcons_nsi     = $fcons_nsi
-    @args.push @bemol_nsi     = $bemol_nsi
-    @args.push @kfl_regim_nsi = $kfl_regim_nsi
-    @args.push @fvela_nsi     = $fvela_nsi
-    @args.push @kfl_rmom2_nsi = $kfl_rmom2_nsi
-    @args.push @kfl_press_nsi = $kfl_press_nsi
-    @args.push @kfl_p1ve2_nsi = $kfl_p1ve2_nsi
-    @args.push @kfl_linea_nsi = $kfl_linea_nsi
-    @args.push @kfl_confi_nsi = $kfl_confi_nsi
-    @args.push @nbdfp_nsi     = $nbdfp_nsi
-    @args.push @kfl_sgsti_nsi = $kfl_sgsti_nsi
-    @args.push @kfl_nota1_nsi = $kfl_nota1_nsi
-    @args.push @kfl_limit_nsi = $kfl_limit_nsi
-    @args.push @kfl_penal_nsi = $kfl_penal_nsi
-    @args.push @penal_nsi     = $penal_nsi
-    @args.push @kfl_bubbl_nsi = $kfl_bubbl_nsi
-  
-    @args.push @pnode     = $pnode
-    @args.push @pgaus     = $pgaus
-    @args.push @gpden     = $gpden
-    @args.push @gpvis     = $gpvis
-    @args.push @gppor     = $gppor
-    @args.push @gpsp1     = $gpsp1
-    @args.push @gpsp2     = $gpsp2
-    @args.push @gpvol     = $gpvol
[…]

### 2017-08-30 Alya/src/boast/Split/KSplitOssBoast.rb

@@ -1,1033 +1,0 @@
-
-require_relative '../Common/subroutine.rb'
-require_relative './KSplitOss.rb'
-
-class Nest1 < Subroutine
-  def initialize(options,functions = nil)
-    super("nest1",options,functions)
-    # ins
-    @mnode = Parameters.copy($mnode,:in)
-    @pnode = Parameters.copy($pnode,:in)
-    @pgaus = Parameters.copy($pgaus,:in)
-    @gpden = Parameters.copy($gpden,:in)
-    @gpcar = Parameters.copy($gpcar,:in)
-    @gpadv = Parameters.copy($gpadv,:in)
-
-    # inouts
-    @wgrgr = Parameters.copy($wgrgr,:inout)
-    @agrau = Parameters.copy($agrau,:inout)
-
-    @args.push @mnode,@pnode,@pgaus,@gpden,@gpcar,@gpadv,@wgrgr,@agrau
-  end
- 
-  def generate
-    # locals
-    inode = $inode
-    jnode = $jnode
-    igaus = $igaus
-
-    block = lambda{
-      decl inode,jnode,igaus unless @usage == :included
-
-      exp1 = ""
-      exp2 = ""
-      @ndime.times{|i|
-        exp1 += "@gpadv[#{i+1},igaus]*@gpcar[#{i+1},inode,igaus]"
-        exp2 += "@gpcar[#{i+1},inode,igaus]*@gpcar[#{i+1},jnode,igaus]"
-        exp1 +=  " + " if i+1 < @ndime
-        exp2 +=  " + " if i+1 < @ndime
-      }
-      
-      form_agrau = "pr @agrau[inode,igaus] === @gpden[igaus] * (#{exp1})"
-      form_wgrgr = "pr @wgrgr[inode,jnode,igaus] === #{exp2}"
-      
-      pr For(igaus,1,@pgaus){
-        pr For(inode,1,@pnode){
-          eval form_agrau
-          pr For(jnode,1,@pnode){                    
-            eval form_wgrgr
-          }
-        }
-      }
-    }
-    
-    construct(block)
-  end
-end
-
-class Nest2 < Subroutine
-  def initialize(options, functions = nil)
-    super("nest2",options,functions)
-
-    # ins
-    @mnode = Parameters.copy($mnode,:in)
-    @pnode = Parameters.copy($pnode,:in)
-    @pgaus = Parameters.copy($pgaus,:in)
-    @gpden = Parameters.copy($gpden,:in)
[…]

### 2017-08-30 Alya/src/boast/Split/KSplitOssRef_v1.rb

@@ -1,592 +1,0 @@
-
-require_relative './KSplitOss.rb'
-
-def generate_mocks
-  mocks = <<EOF
-  function pabdf_nsi(x) result(y)
-    integer,intent(in) :: x 
-    real :: y
-    y = 1.0
-  end function pabdf_nsi
-EOF
-return mocks  
-end
-def generate_ref_declaration
-     decl_ref =<<EOF
-     subroutine nsi_element_assembly_split_oss_ref(ndime,&
-     kfl_lumped,mnode,ntens,kfl_duatss,fact_duatss,&
-     kfl_stabi_nsi,fvins_nsi,fcons_nsi,bemol_nsi,&
-     kfl_regim_nsi,fvela_nsi,kfl_rmom2_nsi,kfl_press_nsi,&
-     kfl_p1ve2_nsi,kfl_linea_nsi,kfl_confi_nsi,nbdfp_nsi,&
-     kfl_sgsti_nsi,kfl_nota1_nsi,kfl_limit_nsi,kfl_penal_nsi,&
-     penal_nsi,kfl_bubbl_nsi,pnode,pgaus,gpden,gpvis,gppor,&
-     gpsp1,gpsp2,gpvol,gpsha,gpcar,gpadv,gpvep,gpgrp,gprhs,&
-     gprhc,gpvel,gpsgs,elvel,elpre,elbub,wgrgr,agrau,elauu,&
-     elaup,elapp,elapu,elrbu,elrbp,dtinv_loc,dtsgs,pbubl,&
-     gpsha_bub,gpcar_bub,elauq,elapq,elaqu,elaqp,elaqq,elrbq)
-
-     #{gen_def_parameters}
-
-     integer(ip), intent(in)    :: ndime 
-     integer(ip), intent(in)    :: kfl_lumped
-     integer(ip), intent(in)    :: mnode
-     integer(ip), intent(in)    :: ntens
-     integer(ip), intent(in)    :: kfl_duatss
-     integer(ip), intent(in)    :: fact_duatss
-     integer(ip), intent(in)    :: kfl_stabi_nsi
-     real(rp),    intent(in)    :: fvins_nsi
-     real(rp),    intent(in)    :: fcons_nsi
-     real(rp),    intent(in)    :: bemol_nsi
-     integer(ip), intent(in)    :: kfl_regim_nsi
-     real(rp),    intent(in)    :: fvela_nsi(3)
[…]

### 2017-08-30 Alya/src/boast/Split/KSplitOssRef_v2.rb

@@ -1,523 +1,0 @@
-
-require_relative './KSplitOss.rb'
- 
-def generate_mocks
-  mocks = <<EOF
-  function pabdf_nsi(x) result(y)
-    integer,intent(in) :: x 
-    real :: y
-    y = 1.0
-  end function pabdf_nsi
-EOF
-return mocks  
-end
-def generate_ref_declaration
-     decl_ref =<<EOF
-     subroutine nsi_element_assembly_split_oss_ref(ndime,&
-     kfl_lumped,mnode,ntens,kfl_duatss,fact_duatss,&
-     kfl_stabi_nsi,fvins_nsi,fcons_nsi,bemol_nsi,&
-     kfl_regim_nsi,fvela_nsi,kfl_rmom2_nsi,kfl_press_nsi,&
-     kfl_p1ve2_nsi,kfl_linea_nsi,kfl_confi_nsi,nbdfp_nsi,&
-     kfl_sgsti_nsi,kfl_nota1_nsi,kfl_limit_nsi,kfl_penal_nsi,&
-     penal_nsi,kfl_bubbl_nsi,pnode,pgaus,gpden,gpvis,gppor,&
-     gpsp1,gpsp2,gpvol,gpsha,gpcar,gpadv,gpvep,gpgrp,gprhs,&
-     gprhc,gpvel,gpsgs,elvel,elpre,elbub,wgrgr,agrau,elauu,&
-     elaup,elapp,elapu,elrbu,elrbp,dtinv_loc,dtsgs,pbubl,&
-     gpsha_bub,gpcar_bub,elauq,elapq,elaqu,elaqp,elaqq,elrbq)
-
-     #{gen_def_parameters}
-
-     integer(ip), intent(in)    :: ndime 
-     integer(ip), intent(in)    :: kfl_lumped
-     integer(ip), intent(in)    :: mnode
-     integer(ip), intent(in)    :: ntens
-     integer(ip), intent(in)    :: kfl_duatss
-     integer(ip), intent(in)    :: fact_duatss
-     integer(ip), intent(in)    :: kfl_stabi_nsi
-     real(rp),    intent(in)    :: fvins_nsi
-     real(rp),    intent(in)    :: fcons_nsi
-     real(rp),    intent(in)    :: bemol_nsi
-     integer(ip), intent(in)    :: kfl_regim_nsi
-     real(rp),    intent(in)    :: fvela_nsi(3)
[…]

### 2017-08-30 Alya/src/boast/Split/debug.rb

@@ -1,124 +1,0 @@
-require './KSplitOssRef_v1.rb'
-require './KSplitOssRef_v2.rb'
-require './KSplitOssBoast.rb'
-require 'narray_ffi'
-require '../Common/CommonArgs.rb'
-
-class Debug 
-  include CommonArgs
-  def self.run
-    nests = [1,2,3,4,5,6,7,8,9,10,11]
-    vector_size=2
-    seed = 10
-
-    pgaus = 8
-    pnode = 8
-    ndime = 3
-    kernels = []
-
-    k_orig_params = {:vector_length => vector_size, :preprocessor => false, :nests => nests}
-    k_boast_params = {:vector_length => vector_size, :nests => nests, :unroll => false, :inline => :inlined}
-
-    set_fortran_line_length(100)
-    kernels[0] = KSplitOssRef_v2::new(k_orig_params)
-    kernels[0].generate
-    kernels[0].kernel.build(:FCFLAGS => "-cpp", :LDFLAGS => "-lgfortran")
-
-    set_lang(C)
-    kernels[1] = KSplitBoast::new(k_boast_params,ndime)
-    kernels[1].generate
-    puts kernels[1].kernel
-    
-    stats_boast = []
-    stats_ref = []
-
-    # # # 100.times{|i|
-    i = 1
-    CommonArgs.init(pgaus,pnode,ndime,vector_size,2,seed)
-    @@kfl_lumped = 1 # 2
-    @@kfl_limit_nsi = 1 # 2
-    @@kfl_stabi_nsi = 1 # -1
-
-    kernels[0].kernel.run(@@ndime,@@kfl_lumped,@@mnode,@@ntens,@@kfl_duatss,@@fact_duatss,@@kfl_stabi_nsi,
-                       @@fvins_nsi,@@fcons_nsi,@@bemol_nsi,@@kfl_regim_nsi,@@fvela_nsi,@@kfl_rmom2_nsi,
-                       @@kfl_press_nsi,@@kfl_p1ve2_nsi,@@kfl_linea_nsi,@@kfl_confi_nsi,@@nbdfp_nsi,
-                       @@kfl_sgsti_nsi,@@kfl_nota1_nsi,@@kfl_limit_nsi,@@kfl_penal_nsi,@@penal_nsi,
-                       @@kfl_bubbl_nsi,@@pnode,@@pgaus,@@gpden,@@gpvis,@@gppor,@@gpsp1,@@gpsp2,@@gpvol,
[…]

### 2017-08-30 Alya/src/boast/Split/run.rb

@@ -1,212 +1,0 @@
-require_relative './KSplitOssRef_v1.rb'
-require_relative './KSplitOssRef_v2.rb'
-require_relative './KSplitOssBoast.rb'
-require 'narray_ffi'
-require_relative '../Common/CommonArgs.rb'
-require 'yaml'
-require 'pp'
-require 'csv'
-require 'optparse'
-require_relative '../Tools/LogInfo.rb'
-
-class Experiment
-  include CommonArgs
-  def self.run(output_info,dimension)
-    nests = (1..11).to_a
-    LogInfo.init(output_info)
-    LogInfo.get_info
-    vector_size=2
-    seed = 10
-    charac = {:pgaus => 8, :pnode => 8}
-    epsilon = 10e-15
-
-    stats = {}
-    k = {}
-    nests.each{|n|
-      opts1 = {:vector_length => vector_size, :preprocessor => false, :nests => [n], :unroll => true, :inline => :included, :CFLAGS => "-O3"}
-      opts2 = {:vector_length => vector_size, :preprocessor => false, :nests => [n], :unroll => true, :inline => :inlined, :CFLAGS => "-O3"}
-      opts3 = {:vector_length => vector_size, :preprocessor => false, :nests => [n], :unroll => true, :inline => :call, :CFLAGS => "-O3"}
-
-      set_lang(C)
-      set_fortran_line_length(100)
-      # k[h_ref] = KSplitOssRef_v2::new(opts)
-      k[opts1] = KSplitBoast::new(opts1,dimension)
-      k[opts1].generate
-      LogInfo.register_kernel_info(opts1, k[opts1].kernel.to_s)
-      # k[opts1].kernel.build(:FCFLAGS => "-O3")
-      k[opts1].kernel.build(:CFLAGS => opts1[:CFLAGS])
-
-      set_lang(C)
-      k[opts2] = KSplitBoast::new(opts2,dimension)
-      k[opts2].generate
-      LogInfo.register_kernel_info(opts2, k[opts2].kernel.to_s)
-      k[opts2].kernel.build(:CFLAGS => opts2[:CFLAGS])
-
[…]

### 2017-08-30 Alya/src/boast/Split/tests/BOAST_kernel_test.rb

@@ -1,233 +1,0 @@
-gem "minitest"
-require 'minitest/autorun'
-require 'BOAST'
-include BOAST
-require '../KSplitOssRef_v1.rb'
-require '../KSplitOssRef_v2.rb'
-require '../KSplitOssBoast.rb'
-require '../../Common/CommonArgs.rb'
-
-class TestFor < Minitest::Test
-  include CommonArgs
-  def test_nests_orig_vs_boast_fortran
-    epsilon = 10e-15
-    set_fortran_line_length(100)
-    nests = [1,2,3,4,5,6,7,8,9,10,11]
-
-    pgaus = 4
-    pnode = 1
-    ndime = 3
-
-    kernels = {}
-    
-    (1..2).each{|vector_size|
-      k_ref_params = {:vector_length => vector_size, :preprocessor => false, :nests => nests}
-      k_boast_params = {:vector_length => vector_size, :nests => nests}
-
-      kernels[:ref] = KSplitOssRef_v2::new(k_ref_params)
-      kernels[:ref].generate
-      kernels[:ref].kernel.build(:FCFLAGS => "-cpp")
-      
-      [:included,:call,:inlined].each{|inlining|
-        k_boast_params[:inline] = inlining
-        [false,true].each{|unroll|
-          k_boast_params[:unroll] = unroll
-          (2..3).each{|dim|
-            CommonArgs.init(pgaus,pnode,dim,vector_size,2)
-
-            (1..2).each{|kfl_lumped|
-              next if kfl_lumped == 1 and dim == 2
-              @@kfl_lumped = kfl_lumped
-
-              [-1,1].each{|kfl_stabi_nsi|
-                @@kfl_stabi_nsi = kfl_stabi_nsi
-
-                (1..2).each{|kfl_limit_nsi|
-                  @@kfl_limit_nsi = kfl_limit_nsi
-                  [FORTRAN].each{|lang|
-                    set_lang(lang)
-                    kernels[:boast] = KSplitBoast::new(k_boast_params,dim)
-                    kernels[:boast].generate
-                    kernels[:boast].kernel.build
-                    kernels[:ref].kernel.run(dim,@@kfl_lumped,@@mnode,@@ntens,@@kfl_duatss,@@fact_duatss,@@kfl_stabi_nsi,
[…]

### 2017-08-30 Alya/src/boast/Tools/LogInfo.rb

@@ -1,41 +1,0 @@
-require 'yaml'
-
-module LogInfo
-  @doc = {}
-  @path = ""
-  def self.init(path)
-    @path = path 
-    @doc[:header] = {}
-    @doc[:body] = {}
-    @doc[:body][:environment] = {}
-    @doc[:body][:environment][:hardware] = {}
-    @doc[:body][:environment][:software] = {} 
-    @doc[:body][:kernel_info] = {}
-  end
-  def self.get_info
-    @doc[:header][:title] = "Experiment information"
-    @doc[:header][:time] = Time.now.utc
-    @doc[:header][:machine] = `hostname`
-
-    @doc[:body][:environment][:hardware][:cpu] = `less /proc/cpuinfo`
-    @doc[:body][:environment][:hardware][:cpu_frequency] = `cat /sys/devices/system/cpu/cpu0/cpufreq/scaling_cur_freq`
-    begin
-      @doc[:body][:environment][:hardware][:gpu] = `nvidia-smi -q`
-    rescue Errno::ENOENT
-      @doc[:body][:environment][:hardware][:gpu] = `lshw -numeric -C display`
-    end
-    @doc[:body][:environment][:software][:linux] = `cat /proc/version` if File.exists?("/proc/version")
-    @doc[:body][:environment][:software][:environment_variables] = `env`
-    @doc[:body][:environment][:software][:cpu_governor] = `cat /sys/devices/system/cpu/cpu0/cpufreq/scaling_governor`
-    @doc[:body][:environment][:software][:running_softwares] = `ps -le`
-    @doc[:body][:environment][:software][:users] = `who`
-  end
-  def self.register_kernel_info(key,infos)
-    @doc[:body][:kernel_info][key] = infos
-  end
-  def self.dump_info
-    File::open( @path, "w") { |f|
-      f.print YAML::dump(@doc)
-    }
-  end
-end

### 2017-08-30 Alya/src/kernel/grenoble1/def_parameters.f90

@@ -1,27 +1,0 @@
-module def_parameters
-  !
-  ! Parameters that do not change
-  !
-  integer,     parameter  :: ip    = 4               ! 4-byte integer
-  integer,     parameter  :: rp    = 8               ! Double precision 
-  integer(ip), parameter  :: ndime = 3               ! Spatial dimension  
-  real(rp),    parameter  :: zeror = epsilon(1.0_rp) ! Almost zero
-  integer(ip), parameter  :: TET04 = 30              ! 3D 
-  integer(ip), parameter  :: TET10 = 31              ! 3D 
-  integer(ip), parameter  :: PYR05 = 32              ! 3D 
-  integer(ip), parameter  :: PYR14 = 33              ! 3D 
-  integer(ip), parameter  :: PEN06 = 34              ! 3D  
-  integer(ip), parameter  :: PEN15 = 35              ! 3D 
-  integer(ip), parameter  :: PEN18 = 36              ! 3D 
-  integer(ip), parameter  :: HEX08 = 37              ! 3D 
-  integer(ip), parameter  :: HEX20 = 38              ! 3D 
-  integer(ip), parameter  :: HEX27 = 39              ! 3D 
-  integer(ip), parameter  :: HEX64 = 40              ! 3D 
-  integer(ip), parameter  :: SHELL = 51              ! 3D shell element
-  integer(ip), parameter  :: BAR3D = 52              ! 3D bar element
-  !
-  ! Size of vector: "VECTOR_SIZE" elements are assembled at the same time
-  !
-  integer(ip), parameter  :: VECTOR_SIZE = 4         ! Size for vectorization
-  
-end module def_parameters

### 2017-08-30 Alya/src/kernel/grenoble1/grenoble.f90

@@ -1,43 +1,0 @@
-program grenoble
-
-  use def_parameters
-  use mod_nsi_assembly, only : nsi_assembly
-  implicit none  
-  integer(ip) :: pnode,pgaus,itype,mnode,ntens,ncomp_nsi
-  integer(ip) :: ielem
-  real(rp)    :: amatr(1000),rhsid(1000)
-  !
-  ! Choose element type
-  !
-  itype     = HEX08
-
-  mnode     = 8
-  ntens     = 3*ndime-3
-  ncomp_nsi = 3
-
-  if(      itype == HEX08 ) then
-     ! HEX08
-     pnode = 8
-     pgaus = 8
-  else if( itype == TET04 ) then
-     ! TET04
-     pnode = 4
-     pgaus = 1
-  else if( itype == PYR05 ) then
-     ! PYR05
-     pnode = 5
-     pgaus = 5
-  else if( itype == PEN06 ) then
-     ! PEN06
-     pnode = 6
-     pgaus = 6
-  end if
-  !
-  ! Assembly
-  !
-  do ielem = 1,20000
-     call nsi_assembly(mnode,pnode,pgaus,ntens,ncomp_nsi,amatr,rhsid)
-  end do
-
-end program grenoble
-

### 2017-08-30 Alya/src/kernel/grenoble1/mod_nsi_assembly.f90

@@ -1,275 +1,0 @@
-module mod_nsi_assembly 
-
-  use def_parameters
-  use mod_nsi_element_assembly, only : nsi_element_assembly_asgs_oss
-  implicit none
-
-  private
-
-  public :: nsi_assembly
-
-contains
-
-  !-----------------------------------------------------------------------
-  !>
-  !> @author  Guillaume Houzeaux
-  !> @date    6/10/2016
-  !> @brief   Call to element assembly
-  !> @details Call to element assembly
-  !>
-  !-----------------------------------------------------------------------
-
-  subroutine nsi_assembly(mnode,pnode,pgaus,ntens,ncomp_nsi,amatr,rhsid)
-    !
-    ! Dimensions
-    !
-    integer(ip), intent(in)  :: mnode                                             !< Max number of nodes
-    integer(ip), intent(in)  :: pnode                                             !< Number of nodes
-    integer(ip), intent(in)  :: pgaus                                             !< Number of Gauss points
-    integer(ip), intent(in)  :: ntens                                             !< Hessian size
-    integer(ip), intent(in)  :: ncomp_nsi                                         !< Number of temporal components
-
-    real(rp),    intent(out) :: amatr(*)                                          !< Simulated output
-    real(rp),    intent(out) :: rhsid(*)                                          !< Simulated output
-    ! 
-    ! Element matrices and vectors 
-    !
-    real(rp)                 :: elauu(VECTOR_SIZE,pnode*ndime,pnode*ndime)        ! Auu
-    real(rp)                 :: elaup(VECTOR_SIZE,pnode*ndime,pnode)              ! Aup
-    real(rp)                 :: elapp(VECTOR_SIZE,pnode,pnode)                    ! App
-    real(rp)                 :: elapu(VECTOR_SIZE,pnode,pnode*ndime)              ! Apu
[…]

### 2017-08-30 Alya/src/kernel/grenoble1/mod_nsi_element_assembly.f90

@@ -1,1383 +1,0 @@
-module mod_nsi_element_assembly
-
-  !------------------------------------------------------------------------
-  !> @addtogroup NastinMatrixAssembly
-  !> @{
-  !> @file    nsi_elmmat.f90
-  !> @author  Guillaume Houzeaux
-  !> @brief   Navier-Stokes system element assembly and other element 
-  !>          calculations
-  !> @details Compute element matrices
-  !>
-  !>          \verbatim
-  !> 
-  !>          Without enrichement
-  !>          -------------------
-  !>            +-        +  +- -+     +-  -+
-  !>            | Auu Aup |  | u |     | bu |
-  !>            |         |  |   |  =  |    |
-  !>            | Apu App |  | p |     | bp |
-  !>            +-       -+  +- -+     +-  -+
-  !>
-  !>          With enrichement
-  !>          ----------------
-  !>            +-            +  +- -+     +-  -+
-  !>            | Auu Aup Auq |  | u |     | bu |
-  !>            |             |  |   |     |    |
-  !>            | Apu App Apq |  | p |  =  | bp |
-  !>            |             |  |   |     |    |
-  !>            | Aqu Aqp Aqq |  | q |     | bq | 
-  !>            +-           -+  +- -+     +-  -+
-  !>
-  !>            q = Aqq^-1 ( bq - Aqu u - Aqp p )
-  !>
-  !>            Auu <= Auu - Auq Aqq^-1 Aqu
-  !>            Aup <= Aup - Auq Aqq^-1 Aqp
-  !>            bu  <= bu  - Auq Aqq^-1 bq
-  !>
-  !>            Apu <= Apu - Apq Aqq^-1 Aqu 
-  !>            App <= App - Apq Aqq^-1 Aqp
-  !>            bp  <= bp  - Apq Aqq^-1 bq
-  !>
-  !>          \endverbatim
-  !>
-  !> @{
-  !------------------------------------------------------------------------
-   
-  use def_parameters
-  implicit none
-
-  private 
-
-  public :: nsi_element_assembly_asgs_oss     
-
-contains
-
[…]

### 2017-08-30 Alya/src/kernel/grenoble2/compile.sh

@@ -1,6 +1,0 @@
-rm -rf *.o *.mod
-gfortran -c -J./ -I./ -ffree-line-length-none -x f95-cpp-input -O3 ./def_parameters.f90
-gfortran -c -J./ -I./ -ffree-line-length-none -x f95-cpp-input -O3 ./mod_nsi_element_assembly.f90
-gfortran -c -J./ -I./ -ffree-line-length-none -x f95-cpp-input -O3 ./mod_nsi_assembly.f90
-gfortran -c -J./ -I./ -ffree-line-length-none -x f95-cpp-input -O3 ./grenoble.f90
-gfortran -o grenoble.x *.o 

### 2017-08-30 Alya/src/kernel/grenoble2/grenoble.f90

@@ -1,44 +1,0 @@
-program grenoble
-
-  use def_parameters
-  use mod_nsi_assembly, only : nsi_assembly
-  implicit none  
-  integer(ip) :: pnode,pgaus,itype,mnode,ntens,ncomp_nsi
-  integer(ip) :: ielem,iassembly
-  real(rp)    :: amatr(1000),rhsid(1000)
-  !
-  ! Choose element type
-  !
-  itype     = HEX08
-
-  mnode     = 8
-  ntens     = 3*ndime-3
-  ncomp_nsi = 3
-  iassembly = 1              ! 1=short, 2=long
-
-  if(      itype == HEX08 ) then
-     ! HEX08
-     pnode = 8
-     pgaus = 8
-  else if( itype == TET04 ) then
-     ! TET04
-     pnode = 4
-     pgaus = 1
-  else if( itype == PYR05 ) then
-     ! PYR05
-     pnode = 5
-     pgaus = 5
-  else if( itype == PEN06 ) then
-     ! PEN06
-     pnode = 6
-     pgaus = 6
-  end if
-  !
-  ! Assembly
-  !
-  do ielem = 1,20000
-     call nsi_assembly(iassembly,mnode,pnode,pgaus,ntens,ncomp_nsi,amatr,rhsid)
-  end do
-
-end program grenoble
-

### 2017-08-30 Alya/src/kernel/grenoble2/mod_nsi_assembly.f90

@@ -1,306 +1,0 @@
-module mod_nsi_assembly 
-
-  use def_parameters
-  use mod_nsi_element_assembly, only : nsi_element_assembly_asgs_oss
-  use mod_nsi_element_assembly, only : nsi_element_assembly_split_oss
-  implicit none
-
-  private
-
-  public :: nsi_assembly
-
-contains
-
-  !-----------------------------------------------------------------------
-  !>
-  !> @author  Guillaume Houzeaux
-  !> @date    6/10/2016
-  !> @brief   Call to element assembly
-  !> @details Call to element assembly
-  !>
-  !-----------------------------------------------------------------------
-
-  subroutine nsi_assembly(iassembly,mnode,pnode,pgaus,ntens,ncomp_nsi,amatr,rhsid)
-    !
-    ! Dimensions
-    !
-    integer(ip), intent(in)  :: iassembly                                         !< Assembly type
-    integer(ip), intent(in)  :: mnode                                             !< Max number of nodes
-    integer(ip), intent(in)  :: pnode                                             !< Number of nodes
-    integer(ip), intent(in)  :: pgaus                                             !< Number of Gauss points
-    integer(ip), intent(in)  :: ntens                                             !< Hessian size
-    integer(ip), intent(in)  :: ncomp_nsi                                         !< Number of temporal components
-
-    real(rp),    intent(out) :: amatr(*)                                          !< Simulated output
-    real(rp),    intent(out) :: rhsid(*)                                          !< Simulated output
-    ! 
-    ! Element matrices and vectors 
-    !
-    real(rp)                 :: elauu(VECTOR_SIZE,pnode*ndime,pnode*ndime)        ! Auu
-    real(rp)                 :: elaup(VECTOR_SIZE,pnode*ndime,pnode)              ! Aup
[…]

### 2017-08-30 Alya/src/kernel/grenoble2/mod_nsi_element_assembly.f90

@@ -1,1975 +1,0 @@
-module mod_nsi_element_assembly
-
-  !------------------------------------------------------------------------
-  !> @addtogroup NastinMatrixAssembly
-  !> @{
-  !> @file    nsi_elmmat.f90
-  !> @author  Guillaume Houzeaux
-  !> @brief   Navier-Stokes system element assembly and other element 
-  !>          calculations
-  !> @details Compute element matrices
-  !>
-  !>          \verbatim
-  !> 
-  !>          Without enrichement
-  !>          -------------------
-  !>            +-        +  +- -+     +-  -+
-  !>            | Auu Aup |  | u |     | bu |
-  !>            |         |  |   |  =  |    |
-  !>            | Apu App |  | p |     | bp |
-  !>            +-       -+  +- -+     +-  -+
-  !>
-  !>          With enrichement
-  !>          ----------------
-  !>            +-            +  +- -+     +-  -+
-  !>            | Auu Aup Auq |  | u |     | bu |
-  !>            |             |  |   |     |    |
-  !>            | Apu App Apq |  | p |  =  | bp |
-  !>            |             |  |   |     |    |
-  !>            | Aqu Aqp Aqq |  | q |     | bq | 
-  !>            +-           -+  +- -+     +-  -+
-  !>
-  !>            q = Aqq^-1 ( bq - Aqu u - Aqp p )
-  !>
-  !>            Auu <= Auu - Auq Aqq^-1 Aqu
-  !>            Aup <= Aup - Auq Aqq^-1 Aqp
-  !>            bu  <= bu  - Auq Aqq^-1 bq
-  !>
-  !>            Apu <= Apu - Apq Aqq^-1 Aqu 
-  !>            App <= App - Apq Aqq^-1 Aqp
-  !>            bp  <= bp  - Apq Aqq^-1 bq
-  !>
-  !>          \endverbatim
-  !>
-  !> @{
-  !------------------------------------------------------------------------
-   
-  use def_parameters
-  implicit none
-
-  private 
-
-  public :: nsi_element_assembly_asgs_oss     
[…]

### 2017-08-30 Alya/src/kernel/mod_nsi_element_assembly.f90

@@ -1,3167 +1,0 @@
-module mod_nsi_element_assembly
-
-  !------------------------------------------------------------------------
-  !> @addtogroup NastinMatrixAssembly
-  !> @{
-  !> @file    nsi_elmmat.f90
-  !> @author  Guillaume Houzeaux
-  !> @brief   Navier-Stokes system element assembly and other element 
-  !>          calculations
-  !> @details Compute element matrices
-  !>
-  !>          \verbatim
-  !> 
-  !>          Without enrichement
-  !>          -------------------
-  !>            +-        +  +- -+     +-  -+
-  !>            | Auu Aup |  | u |     | bu |
-  !>            |         |  |   |  =  |    |
-  !>            | Apu App |  | p |     | bp |
-  !>            +-       -+  +- -+     +-  -+
-  !>
-  !>          With enrichement
-  !>          ----------------
-  !>            +-            +  +- -+     +-  -+
-  !>            | Auu Aup Auq |  | u |     | bu |
-  !>            |             |  |   |     |    |
-  !>            | Apu App Apq |  | p |  =  | bp |
-  !>            |             |  |   |     |    |
-  !>            | Aqu Aqp Aqq |  | q |     | bq | 
-  !>            +-           -+  +- -+     +-  -+
-  !>
-  !>            q = Aqq^-1 ( bq - Aqu u - Aqp p )
-  !>
-  !>            Auu <= Auu - Auq Aqq^-1 Aqu
-  !>            Aup <= Aup - Auq Aqq^-1 Aqp
-  !>            bu  <= bu  - Auq Aqq^-1 bq
-  !>
-  !>            Apu <= Apu - Apq Aqq^-1 Aqu 
-  !>            App <= App - Apq Aqq^-1 Aqp
-  !>            bp  <= bp  - Apq Aqq^-1 bq
-  !>
-  !>          \endverbatim
-  !>
-  !> @{
-  !------------------------------------------------------------------------
-
-  use def_kintyp, only :  ip,rp
-#ifndef VECTOR_SIZE
-  use def_master, only :  VECTOR_SIZE
-#endif
[…]

### 2017-08-30 Alya/src/scripts/format_data_comparison.rb

@@ -1,29 +1,0 @@
-require 'yaml'
-require 'pp'
-require 'csv'
-
-input = ARGV[0]
-output = ARGV[1]
-
-header = [:kernel,:nest,:time]
-body = []
-
-h = YAML::load(File::open(input).read)
-h.each{|k1,v1|
-  v1.each{|e|
-    t = []
-    k1.each{|k2,v2|
-      t.push(v2)
-    }
-    t.push(e)
-    body.push(t)
-  }
-}
-
-CSV.open(output, "w"){ |f|
-  f << header
-  body.each{ |e|
-    f << e
-  }
-}
-

### 2017-08-30 Alya/src/scripts/format_data_comparison2.rb

@@ -1,31 +1,0 @@
-require 'yaml'
-require 'pp'
-require 'csv'
-
-input = ARGV[0]
-output = ARGV[1]
-
-header = [:vector_length,:preprocessor,:nest,:unroll,:usage,:CFLAGS,:pgaus,:pnode,:time]
-body = []
-
-h = YAML::load(File::open(input).read)
-h.each{|k1,v1|
-  v1[:time].each{|e1|
-    t = []
-    k1.each{|k2,v2|
-      t.push v2.class == Array ? v2[0] : v2
-    }
-    t.push v1[:characteristics][:pgaus]
-    t.push v1[:characteristics][:pnode]
-    t.push(e1)
-    body.push(t)
-  }
-}
-
-CSV.open(output, "w"){ |f|
-  f << header
-  body.each{ |e|
-    f << e
-  }
-}
-

### 2017-08-30 Alya/src/scripts/run_experiment.sh

@@ -1,49 +1,0 @@
-#!/bin/bash
-
-BASE="../.."
-EXP_DIR="$BASE/experiments"
-SRC_DIR="$BASE/src"
-
-help_script()
-{
-    cat << EOF
-Usage: $0 [options] /path_to_data_directory
-
-Script for to get machine information before doing the experiment
-
-OPTIONS:
-   -h      Show this message
-EOF
-}
-
-while getopts "d:" opt; do
-    case $opt in
-	d)
-	    DATADIR="$OPTARG"
-	    ;;
-	h)
-	    help_script
-	    exit 4
-	    ;;
-	\?)
-	    echo "Invalid option: -$OPTARG"
-	    help_script
-	    exit 3
-	    ;;
-    esac
-done
-
-DATA_FOLD_DAY=`date +%Y_%m_%d`
-DATA_FOLD_DAY="$EXP_DIR/$DATA_FOLD_DAY"
-BKUP=`date +%H_%M_%S`
-DATA_FOLD_HOST=`hostname`
-DATA_FOLD_HOST="$DATA_FOLD_DAY/$DATA_FOLD_HOST"
-DATA_FOLD_TIME="$DATA_FOLD_HOST/$BKUP"
-INFO_FILE="$DATA_FOLD_TIME/Info.yaml"
-DATA_FILE="$DATA_FOLD_TIME/Data.yaml"
-
-mkdir -p $DATA_FOLD_DAY
-mkdir -p $DATA_FOLD_HOST
-mkdir -p $DATA_FOLD_TIME
-
-ruby $SRC_DIR/boast/Split/run.rb --data=$DATA_FILE --info=$INFO_FILE --dimension=3

### 2017-08-30 hou10ni/KBoast.rb

@@ -4,8 +4,8 @@ class KBoast
 
  def initialize(options)
 
-  @opts = {:optim_nested => 1, :optim_main => 1}
-  @opts.update(options)
+  #@opts = {:optim_nested => 1, :optim_main => 1}
+  #@opts.update(options)
 
   # Parameters 
   @Nflu_inner         = Int("Nflu_inner", :dir => :in )
@@ -30,11 +30,11 @@ class KBoast
  ### Non-optimized version of the kernel ###
  ###########################################
 
- def generate_orig
+ def generate
 
 	push_env(:lang => FORTRAN) {
 		@kernel = CKernel:: new()
-     @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+     @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
 
     #..  FUNCTIONS ..#
     register_funccall("matmul") if get_lang == FORTRAN
@@ -62,7 +62,7 @@ class KBoast
 
 		pr For(i,1,@Nflu_inner+@Nflusol_inner){
 			pr i_tmp1 === 1
-			For(j,1,@idx_vec_flu[1,i]){
+			pr For(j,1,@idx_vec_flu[1,i]){
 				pr i1 === @idx_vec_flu[2*j,i]
         pr i2 === @idx_vec_flu[2*j+1,i]
         pr i_tmp2 === i_tmp1 + i2 - i1
@@ -75,148 +75,12 @@ class KBoast
       pr i4 === @idx_mat_flu[i+1] - 1
       reshape_code ="RESHAPE(#{@A_flu.slice(i3..i4)}, (/#{i2}-#{i1}+1, #{i_tmp2}/))"
       pr @P_new.slice(i1..i2,1..@nb_rhs) === call_matmul.call(reshape_code , p_aux.slice(1..i_tmp2,1..@nb_rhs) ) - @P_old.slice(i1..i2,1..@nb_rhs)		
-
 		}
[…]

### 2017-08-30 hou10ni/KBoastCOST.rb

@@ -1,0 +1,97 @@
+class KBoastCOST
+ attr_reader :kernel
+
+
+ def initialize(options)
+
+
+  # Parameters 
+  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
+  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
+  @nb_rhs             = Int("nb_rhs", :dir => :in )
+  @afluSize           = Int("afluSize", :dir => :in )
+  @pSize              = Int("pSize", :dir => :in )
+  @vecSize_1          = Int("vecSize_1", :dir => :in )
+  @vecSize_2          = Int("vecSize_2", :dir => :in )
+  @matSize            = Int("matSize", :dir => :in )
+  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+  @cost            		= Int("cost", :dir => :out )
+  #@cost_mem           = Int("cost_mem", :dir => :out )
+
+  #@P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  #@P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  #@P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  #@A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
+
+ end
+
+
+ ###########################################
+ ### Non-optimized version of the kernel ###
+ ###########################################
+
+ def generate
+
+	push_env(:lang => FORTRAN) {
+		@kernel = CKernel:: new()
+     @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @cost] ,:functions => nil)
[…]

### 2017-08-30 hou10ni/KBoastOPT.rb

@@ -1,0 +1,160 @@
+class KBoastOPT
+ attr_reader :kernel
+
+
+ def initialize(options)
+
+  @opts = {:optim_nested => 1, :optim_main => 1}
+  @opts.update(options)
+
+  # Parameters 
+  @Nflu_inner         = Int("Nflu_inner", :dir => :in )
+  @Nflusol_inner      = Int("Nflusol_inner", :dir => :in )
+  @nb_rhs             = Int("nb_rhs", :dir => :in )
+  @afluSize           = Int("afluSize", :dir => :in )
+  @pSize              = Int("pSize", :dir => :in )
+  @vecSize_1          = Int("vecSize_1", :dir => :in )
+  @vecSize_2          = Int("vecSize_2", :dir => :in )
+  @matSize            = Int("matSize", :dir => :in )
+  @idx_vec_flu        = Int("idx_vec_flu", :dir => :in, :dim => [Dim(@vecSize_1),Dim(@vecSize_2)] )
+  @idx_mat_flu        = Int("idx_mat_flu", :dir => :in, :dim => [Dim(@matSize)] )
+  @P_new              = Real("P_new",:dir => :inout, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_inter            = Real("P_inter",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @P_old              = Real("P_old",:dir => :in, :dim => [Dim(@pSize),Dim(@nb_rhs)])
+  @A_flu              = Real("A_flu",:dir => :in, :dim => [Dim(@afluSize)])
+
+ end
+
+
+ ######################################n
+ ### Optimized version of the kernel ###
+ #######################################
+
+ def generate
+
+   push_env(:lang => FORTRAN) {
+
+     @kernel = CKernel:: new()
+     @kernel.procedure = Procedure("fluid_inner_elt_boast",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
+
+
+		#..  FUNCTIONS ..#
+
[…]

### 2017-08-30 hou10ni/KRef.rb

@@ -30,7 +30,7 @@ class KRef
 
    push_env(:lang => FORTRAN) {
      @kernel = CKernel:: new()
-     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize] ,:functions => nil)
+     @kernel.procedure = Procedure("fluid_inner_elt_ref",[@Nflu_inner, @Nflusol_inner, @nb_rhs, @afluSize, @pSize,@vecSize_1,@vecSize_2,@matSize, @idx_vec_flu, @idx_mat_flu, @P_new, @P_inter, @P_old, @A_flu] ,:functions => nil)
 		 #puts "Parameters = ", @kernel.procedure.parameters
      get_output.print codeF
    }

### 2017-08-30 hou10ni/dump_file.c

@@ -1,0 +1,20 @@
+/*
+  File used to record all input parameters in binary format for BOAST 
+*/
+
+#include <stdlib.h>
+#include <stdio.h>
+
+void dump_file_(char* filename, void* mem, size_t sz){
+        FILE *f = fopen(filename, "wb");
+        fwrite(mem,1,sz,f);
+        fclose(f);
+}
+
+int main(void){
+        int cost;
+        size_t sz=sizeof(cost);
+        cost=0;
+        dump_file_("cost.in",&cost,sz);
+        return 0;
+}

### 2017-08-30 hou10ni/kernel.f90

@@ -1,4 +1,4 @@
-SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu, afluSize, pSize, vecSize_1, vecSize_2, matSize)
+SUBROUTINE fluid_inner_elt_ref(Nflu_inner, Nflusol_inner, nb_rhs,afluSize, pSize, vecSize_1, vecSize_2, matSize, idx_vec_flu, idx_mat_flu, P_new, P_inter, P_old, A_flu)
   implicit none
   INTEGER, PARAMETER :: dq=8 !< quadruple precision
   INTEGER, PARAMETER :: dp=4  !< double precision

### 2017-08-30 hou10ni/run_COMP.rb

@@ -1,6 +1,6 @@
 require 'BOAST'
 require './KRef.rb'
-require './KBoastOPT.rb'
+require './KBoast.rb'
 require 'narray_ffi'
 require 'yaml'
 require 'pp'
@@ -16,20 +16,14 @@ class Experiment
   # set_default_int_size(8)
   set_default_real_size(4)
 
-	svg_opt_space = OptimizationSpace::new( optim_nested: 1..10, 
-                                    optim_main: 1..10,
-                                    OFLAGS: ["-O2", "-O3"], 
-																		:omp_num_threads => 1..6
-                                    )
-	opt_space = OptimizationSpace::new( optim_nested: 1..10, 
-                                    optim_main: 1..3,
-                                    OFLAGS: ["-O2"], 
-																		:omp_num_threads => 1..6
-                                    )
+
+  #generer la sortie avec la ref et comparer avec ce qu'a boast
+  # utiliser des numeros pour les inputs
+
   set_lang(FORTRAN)
   kernels={}
   stats={}
-	repeat = 20
+	repeat = 1
 
   #.. REF KERNEL ..#
 
@@ -39,49 +33,34 @@ class Experiment
   kernels[k_ref_params].kernel.build(:LDFLAGS => k_ref_params[:LDFLAGS], :FCFLAGS => k_ref_params[:FCFLAGS] ) 
   stats[k_ref_params]={:time => []}
 
-
-	# input and output parameters to check the kernel
-	# found in fluid_inner_elt_ref
   inputs = kernels[k_ref_params].kernel.load_ref_inputs()
   outputs = kernels[k_ref_params].kernel.load_ref_outputs()
 
-
-	# As many keys as directories in fluid_inner_elt_ref/
   inputs.each_key { |key|
-		repeat.times{|i|
- 			stats[k_ref_params][:time][i] = kernels[k_ref_params].kernel.run(*(inputs[key]))[:duration]
-		}
+ 		stats[k_ref_params].push kernels[k_ref_params].kernel.run(*(inputs[key]))
   }
-	p k_ref_params
-	p stats[k_ref_params][:time].min
-
-
 
   #.. BOAST KERNEL ..#
   
-	optimizer = BruteForceOptimizer::new(opt_space, :randomize => true)
-	puts optimizer.optimize { |opt|
-		p opt
[…]

### 2017-08-30 hou10ni/run_boastCOST.rb

@@ -1,5 +1,5 @@
 require 'BOAST'
-require './KBoast.rb'
+require './KBoastCOST.rb'
 require 'narray_ffi'
 
 include BOAST
@@ -11,36 +11,36 @@ num_threads = 1
 #set_default_int_size(8)
 set_default_real_size(4)
 
-#k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran -L/usr/lib/ -lblas", :FCFLAGS => "-fimplicit-none -fexternal-blas -O3"}  
-k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fimplicit-none -O3"}  
+k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran -L/usr/lib/ -lblas", :FCFLAGS => "-fimplicit-none -fexternal-blas -O3"}  
 
 set_lang(FORTRAN)
 stats = []
 repeat = 5
 
 # Creating boast kernel
-k = KBoast::new(k_boast_params)
+k = KBoastCOST::new(k_boast_params)
 k.generate
 puts k.kernel
 k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
  
 inputs = k.kernel.load_ref_inputs()
-outputs = k.kernel.load_ref_outputs()
-puts "** inputs =", inputs
-puts "** outputs=",outputs
-
-
+#outputs = k.kernel.load_ref_outputs()
+#puts "** inputs =", inputs
+#puts "** outputs=",outputs
 
 inputs.each_key { |key|
-	repeat.times {
+	 k.kernel.run(*(inputs[key]))
  		stats.push k.kernel.run(*(inputs[key])) 
-	}
-	puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
+	#puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
 }
 
 stats.sort_by! { |a| a[:duration] }
 p stats
 stats = stats.first
 
+
 puts "#{k.kernel.procedure.name}: #{stats[:duration]*1.0e3} ms    #{ 68914896 / (stats[:duration]*1.0e9)} GFlops"
 
+#  5GFlops * 3,1GHz / 3 = 5,166
+
+#puts "#{k.kernel.procedure.name}: #{stats[:duration]*1.0e3} #{ k.kernel.cost() / (stats[:duration]*1.0e9)} GFlops"

### 2017-10-06 hou10ni/Tutorial/hou10ni_profile.tex
Hou10ni: new scripts + Papi

@@
 \begin{cverbatim}
-export TAU_MAKEFILE=$TAU_DIR/Makefile.
+export TAU_MAKEFILE=$TAU_DIR/Makefile.tau...  (ex.: Makefile.tau-icpc-mpi-pdt)
 export TAU_OPTIONS='-optPdtF95Opts="-R free" -optVerbose'
@@
 
-\noindent $\rightarrow$ Press "t" to change the compiler and chose \texttt{tau\_f90.sh}. Then you can tape \texttt{make} and launch the application.
+\noindent $\rightarrow$ Press "t" to change the compiler and chose \texttt{tau\_f90.sh} and add TAU lib\\
+ (e.g., \texttt{/home/esaillard/TOOLS/TAU/tau-2.26/build/x86\_64/lib/libTAU\_traceinput-icpc-mpi-pdt.a}) in the linker flag. Then you can tape \texttt{make} and launch the application.
 
@@
 cd .. 
-mpirun -np 4 tau_exec -T mpi  build/hou10ni.out < HPC4E.dat
+mpirun -np 4 tau_exec -T mpi  -io -memory build/hou10ni.out < HPC4E.dat
 \end{cverbatim}

### 2017-10-06 hou10ni/KBoastOPT.rb

@@ -140,8 +140,9 @@ class KBoastOPT
 
 		## Procedure
 		optim_2[@opts[:optim_main]].each { |f_main|
-				pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads], :shared => [@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu], :private => [i1, i2, i3, i4, i, j, i_tmp1, i_tmp2, @A_flu, @P_old, @P_new]){
-				pr OpenMP::For(:schedule => "static")
+				# :default => none
+				pr OpenMP::Parallel(:num_threads => @opts[:omp_num_threads], :shared => [@Nflu_inner, @Nflusol_inner, @nb_rhs, @idx_vec_flu, @idx_mat_flu, @A_flu, @P_old, @P_new], :private => [i1, i2, i3, i4, i, j, i_tmp1, i_tmp2, @P_aux]){
+				#pr OpenMP::For(:schedule => "static")
         pr f_main
 				}
     }

### 2017-10-06 hou10ni/parse_cvs.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+
+FILE=test.cvs
+RES_FILE=res.dat
+
+echo "=> Parsing file $FILE"
+LINE_NUMBER=0
+TIME_REF=0
+
+printf "#Options\t Time\n" >> $RES_FILE
+printf "\"\"\t 0.0000000\n" >> $RES_FILE
+
+while read Line; do
+
+	LINE_NUMBER="$(($LINE_NUMBER + 1))"
+  OPTIONS=`echo $Line | cut -d',' -f1,2,3,4`
+  TIME=$(echo $Line | cut -d',' -f5)
+  printf -v TIME "%.7f" "$TIME"
+
+	if [[ $LINE_NUMBER -eq 1 ]]; then
+  	printf -v TIME_REF "%.7f" "$TIME"
+  	printf "\"$OPTIONS\"\t $TIME\n " >> $RES_FILE
+	fi
+
+	if [[ $(echo "$TIME < $TIME_REF" | bc) -eq 1 ]]; then
+  	printf "\"$OPTIONS\"\t $TIME\n " >> $RES_FILE
+	fi
+
+
+done < $FILE
+
+echo "=> Done"

### 2017-10-06 hou10ni/parse_data_yaml.rb

@@ -1,0 +1,33 @@
+require 'yaml'
+require 'pp'
+require 'csv'
+
+
+# USAGE : ruby parse_data_yaml.rb ../experiments/../Data.yaml test.cvs
+input = ARGV[0]
+output = ARGV[1]
+
+#header = [:kernel,:optim_main, :optim_nested, :omp_num_threads,:time]
+#header = []
+body = []
+
+h = YAML::load(File::open(input).read)
+h.each{|k1,v1|
+  #v1[:time].each{|e|
+  	e = v1[:time].min
+    t = []
+    t.push k1[:kernel]
+    t.push k1[:optim_main]
+    t.push k1[:optim_nested]
+    t.push k1[:omp_num_threads]
+    t.push(e)
+    body.push(t)
+  #}
+}
+
+CSV.open(output, "w"){ |f|
+  f << header
+  body.each{ |e|
+    f << e
+  }
+}

### 2017-10-06 hou10ni/run_boast.rb

@@ -36,6 +36,7 @@ inputs.each_key { |key|
 	repeat.times {
  		stats.push k.kernel.run(*(inputs[key])) 
 	}
+
 	#puts k.kernel.compare_ref(outputs[key], inputs[key]).inspect
 }
 

### 2017-10-06 hou10ni/run_boastPAPI.rb

@@ -1,0 +1,73 @@
+require 'BOAST'
+require './KBoast.rb'
+require 'narray_ffi'
+
+include BOAST
+
+set_default_real_size(4)
+
+#k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran -L/usr/lib/ -lblas", :FCFLAGS => "-fimplicit-none -fexternal-blas -O3"}  
+k_boast_params = {:kernel => :boast, :LDFLAGS => "-lgfortran", :FCFLAGS => "-fimplicit-none -O3"}
+
+set_lang(FORTRAN)
+stats = []
+repeat = 5
+
+# Creating boast kernel
+k = KBoast::new(k_boast_params)
+k.generate
+puts k.kernel
+k.kernel.build(:LDFLAGS => k_boast_params[:LDFLAGS], :FCFLAGS => k_boast_params[:FCFLAGS] )
+
+inputs = k.kernel.load_ref_inputs()
+outputs = k.kernel.load_ref_outputs()
+
+
+l=[]
+
+inputs.each_key { |key|
+
+  k.kernel.run(*(inputs[key]))
+  repeat.times {
+    stats.push k.kernel.run(*(inputs[key]))
+  }
+
+  l[1]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_L1_DCM"])
+  l[2]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_L2_DCM"])
+  l[3]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_TOT_CYC"])
+  l[4]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_TOT_INS"])
+  l[5]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_BR_MSP"])
+  l[6]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_L1_LDM"])
+  l[7]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_L1_STM"])
+  l[8]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_VEC_SP"])
+  l[9]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_BR_INS"])
+  l[10]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_L2_TCA"])
+  l[11]=k.kernel.run(*(inputs[key]), :PAPI =>["PAPI_LD_INS"])
[…]

### 2017-10-06 hou10ni/run_compOPT.rb

@@ -22,9 +22,9 @@ class Experiment
 																		:omp_num_threads => 1..6
                                     )
 	opt_space = OptimizationSpace::new( optim_nested: 1..10, 
-                                    optim_main: 1..3,
-                                    OFLAGS: ["-O2"], 
-																		:omp_num_threads => 1..6
+                                    optim_main: 1..10,
+                                    OFLAGS: ["-O2", "-O3"], 
+																		:omp_num_threads => 1..2
                                     )
   set_lang(FORTRAN)
   kernels={}

### 2017-10-06 hou10ni/run_experiment.sh

@@ -51,4 +51,4 @@ mkdir -p $DATA_FOLD_TIME
 #ruby $SRC_DIR/run.rb --data=$DATA_FILE --kernel=$KERNEL_FILE
 
 echo "EXECUTING ruby $SRC_DIR/run.rb --data=$DATA_FILE --kernel=$KERNEL_FILE"
-ruby $SRC_DIR/run.rb --data=$DATA_FILE --kernel=$KERNEL_FILE
+ruby $SRC_DIR/run_compOPT.rb --data=$DATA_FILE --kernel=$KERNEL_FILE

### 2018-03-20 Tutorial/presentation/Tutorial-BOAST-2.tex
Moved talk to presentation directory.

--- 2015-07-22 Tutorial/Tutorial-BOAST-2.tex
@@
   basicstyle=\tiny\ttfamily,
+  keywordstyle=\color{javapurple}\bfseries,
+  stringstyle=\color{javared},
+  commentstyle=\color{javagreen},
   backgroundcolor=\color{white},
-  frame=single,
-  captionpos=b,
   morekeywords={BOAST, pr, decl, opn, close}

### 2022-02-23 Talks/2022-02-26-SIAMPP22/Talk-BOAST-SIAMPP22.tex
Added SIAMPP22 presentation draft.

\documentclass{beamer}
\usepackage{etex}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amssymb}
\usepackage{color}
\usepackage{verbatim}
\usepackage{beamerthemesplit}
\usepackage{appendixnumberbeamer}
\usepackage{graphicx}
\usepackage{xspace}
\usepackage{algorithm}
\usepackage{listings}
\usepackage{figlatex}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{alltt}

\beamertemplatetransparentcovereddynamic

%\setbeamertemplate{background canvas}[vertical shading][bottom=red!10,top=blue!10]
%\usetheme{Warsaw}

\usepackage{../style/beamercolorthemeprogressbar}
\usepackage{../style/beamerfontthemeprogressbar}
\usepackage{../style/beamerouterthemeprogressbar}
\usepackage{../style/beamerinnerthemeprogressbar}

\setbeamertemplate{navigation symbols}{}
\beamertemplatetransparentcovereddynamic

\definecolor{javared}{rgb}{0.6,0,0} % for strings
\definecolor{javagreen}{rgb}{0.30,0.55,0.40} % comments
\definecolor{javapurple}{rgb}{0.5,0,0.35} % keywords
\definecolor{javadocblue}{rgb}{0.25,0.35,0.75} % java

\lstdefinestyle{BOAST}{
  language=Ruby,
  basicstyle=\tiny\ttfamily,
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  backgroundcolor=\color{white},
  numbers=left,
  morekeywords={BOAST, pr, decl, opn, close, For, Procedure, Real, Int, If, out, inout, Dim}
}
\lstdefinestyle{BFortran}{
  language=Fortran,
  basicstyle=\tiny\ttfamily,
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  numbers=left,
  tabsize=4,
  showspaces=false,
  showstringspaces=false
}
\lstdefinestyle{BC}{
  language=C,
  basicstyle=\tiny\ttfamily,
  keywordstyle=\color{javapurple}\bfseries,
  stringstyle=\color{javared},
  commentstyle=\color{javagreen},
  numbers=left,
  tabsize=4,
  showspaces=false,
  showstringspaces=false,
  morekeywords={kernel, global, uchar, short16, uchar16, int16, int32\_t, int64\_t},
}

\graphicspath{{../figures/}}

\title{BOAST}
\subtitle{A Computing Kernel Metaprogramming and Autotuning Framework for Heterogeneous Architectures}
\author[B. V.]{\textbf{Brice~Videau}~\inst{1}, Kevin~Pouget~\inst{2}, Luigi~Genovese~\inst{3},
                    Thierry~Deutsch~\inst{3}, Julien Bigot~\inst{3}, Guillaume Latu~\inst{3}, Virginie Grandgirard~\inst{3}, Dimitri~Komatitsch~\inst{4}, Fr\'ed\'eric~Desprez~\inst{2}, Jean-François~Méhaut~\inst{2}}
\institute[ANL]{
\inst{1} ANL,
\inst{2} INRIA/LIG,
\inst{3} CEA,
\inst{4} CNRS
}

\date{\textbf{SIAMPP22} - Minisymposium on Code Generation and Transformation in HPC on Heterogeneous Platforms\\February 26, 2022}

%\logo{
%   \includegraphics[scale=.09]{logo_eocoe}
%   }

\bibliographystyle{plain}
\begin{document}

\frame{\titlepage}

\section{Introduction}

\subsection{Context}

\begin{frame}
  \frametitle{Scientific Application Portability}

  \begin{block}{\footnotesize Limited Portability}
    \begin{itemize}
      \item \scriptsize Huge codes (more than 100 000 lines), Written in FORTRAN or C++
      \item \scriptsize Collaborative efforts
      \item \scriptsize Use many different programming paradigms (OpenMP, OpenCL, CUDA, ...)
    \end{itemize}
  \end{block}

  \begin{columns}

  \column{0.45\linewidth}
  \begin{block}{\footnotesize But Based on \emph{Computing Kernels}}
    \begin{itemize}
      \item \scriptsize Well defined parts of a program
      \item \scriptsize Compute intensive
      \item \scriptsize Prime target for optimization
    \end{itemize}
  \end{block}

  \column{0.58\linewidth}
  \begin{block}{\footnotesize Kernels Should Be Written}
    \begin{itemize}
      \item \scriptsize In a \emph{portable} manner
      \item \scriptsize In a way that raises developer \emph{productivity}
      \item \scriptsize To present good \emph{performance}
    \end{itemize}
  \end{block}

  \end{columns}

\end{frame}

\subsection{Related Work}

\begin{frame}
  \frametitle{Related Work}

  \begin{itemize}
    \item \textbf{Ad hoc autotuners (usually for libraries):}
    \begin{itemize}
      \item \textbf{Atlas}~\cite{whaley04} (C macro processing)
      \item \textbf{SPIRAL}~\cite{puschel2004spiral} (DSL)
      \item ...
    \end{itemize}
    \item \textbf{Generic frameworks using annotation systems:}
    \begin{itemize}
      \item \textbf{POET}~\cite{yi2007poet} (external annotation file)
      \item \textbf{Orio}~\cite{Hart2009:Orio} (source annotation)
      \item \textbf{BEAST}~\cite{CPE:CPE3516} (Python preprocessor based, embedded DSL for optimization space definition/pruning)
    \end{itemize}
    \item \textbf{Generic frameworks using embedded DSL:}
    \begin{itemize}
      \item \textbf{Halide}~\cite{ragan2013halide} (C++, not very generic, 2D stencil targeted)
      \item \textbf{Heterogeneous Programming Library}~\cite{F.Fabeiro:2016:WPM:2894387.2894576} (C++)
    \end{itemize}
  \end{itemize}

\end{frame}

\section{A Parametrized Generator}
\begin{frame}
\frametitle{BOAST: a Parametrized Generator}
\end{frame}

\subsection{Motivation and Principles}

\begin{frame}

  \frametitle{Classical Tuning of Computing Kernels}

  \begin{center}%
    \includegraphics<1>[scale=1]{Workflow1-1}%
    \includegraphics<2>[scale=1]{Workflow1-2}%
    \includegraphics<3>[scale=1]{Workflow1-3}%
    \includegraphics<4>[scale=1]{Workflow1-4}%
  \end{center}%
  \begin{itemize}%
\only<1>{    \item Kernel optimization workflow}%
\only<1>{    \item Usually performed by a knowledgeable developer}%
\only<2>{    \item Compilers perform optimizations}%
\only<2>{    \item Architecture specific or generic optimizations}%
\only<3>{    \item Performance data hint at source transformations}%
\only<3>{    \item Architecture specific or generic hints}%
\only<4>{    \item Multiplication of kernel versions and/or loss of versions}%
\only<4>{    \item Difficulty to benchmark versions against each-other}%
  \end{itemize}%

\end{frame}

\begin{frame}
  \frametitle{BOAST Workflow \vphantom{Cp}}
  \begin{center}
    \includegraphics<1>[scale=1]{Workflow2-1}
    \includegraphics<2>[scale=1]{Workflow2-2}
    \includegraphics<3>[scale=1]{Workflow2-3}
  \end{center}
  \begin{itemize}%
\only<1>{    \item Meta-programming of optimizations in BOAST }
\only<1>{    \item High level object oriented language }
\only<2>{    \item Generate combination of optimizations }
\only<2>{    \item C, OpenCL, FORTRAN and CUDA are supported }
\only<3>{    \item Compilation and analysis are automated }
\only<3>{    \item Selection of best version can also be automated \vphantom{Cp}}
  \end{itemize}%
\end{frame}

\begin{frame}
\frametitle{BOAST Architecture}
 \begin{center}
   \includegraphics[scale=0.5]{BOAST_Workflow}
 \end{center}
\end{frame}

\subsection{Simple Illustration}

\begin{frame}[fragile]
\frametitle{Kernel Definition}
\lstset{style=BOAST}
\begin{lstlisting}
require 'narray'
require 'BOAST'
include BOAST

set_array_start(0)
set_default_real_size(4)

def vector_add
  n = Int("n",:dir => :in)
  a = Real("a",:dir => :in, :dim => [ Dim(n)] )
  b = Real("b",:dir => :in, :dim => [ Dim(n)] )
  c = Real("c",:dir => :out, :dim => [ Dim(n)] )
  p = Procedure("vector_add", [n,a,b,c]) {
    decl i = Int("i")
    expr = c[i] === a[i] + b[i]
    if (get_lang == CL or get_lang == CUDA) then
      pr i === get_global_id(0)
      pr expr
    else
      pr For(i,0,n-1) {
        pr expr
      }
    end
  }
  return p.ckernel
end
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Kernel Testing}
\lstset{style=BOAST}
\begin{lstlisting}
n = 1024*1024
a = NArray.sfloat(n).random!
b = NArray.sfloat(n).random!
c = NArray.sfloat(n)
c_ref = a + b

epsilon = 1e-8

[FORTRAN, C, CL, CUDA].each { |l|
  set_lang( l )
  puts "#{get_lang_name}:"
  k = vector_add
  puts k
  c.random!
  puts k.run(n, a, b, c, global_work_size: [n,1,1], local_work_size: [32,1,1])
  diff_max = (c_ref - c).abs.max
  raise "Error: max error too big: #{diff_max}!" if diff_max > epsilon
}
puts "Success!"
\end{lstlisting}
\end{frame}

\subsection{Evaluation}

\begin{frame}
\frametitle{Strength and Weakness of the Approach}
Major benefits:
\begin{itemize}
\item Powerful meta-programming capabilities, Ruby is your macro-processor;
\item Generated kernels are readable, can be debugged, and can be integrated in applications;
\item Kernel correctness and performance evaluation is efficient:
  \begin{itemize}
  \item Kernels are self contained,
  \item and ran inside BOAST memory space, avoiding repeated setup.
  \end{itemize}
\end{itemize}
Major hurdles:
\begin{itemize}
\item Upfront cost of kernel extraction and setting up validation can be important if data shape is complex;
\item Debugging syntax errors in lambdas can be difficult;
\item No semantic validation beside basic type checking;
\end{itemize}

\end{frame}

\section{Case Studies}

\begin{frame}
\frametitle{Case Studies}
\end{frame}

\subsection{BigDFT}

\begin{frame}
  \frametitle{BigDFT}
  \begin{center}
    \includegraphics[scale=0.35]{Res_synthesis}
  \end{center}

 \begin{itemize}
  \item Novel approach for DFT computation based on Daubechies wavelets
  \item Fortran and C code, MPI, OpenMP, supports CUDA and OpenCL
  \item Reference is hand tuned code on target architecture (Nehalem)
  \item BLAS-like library for wavelets: \url{https://github.com/luigigenovese/libconv}
  \end{itemize}
\end{frame}

\subsection{SPECFEM3D}

\begin{frame}
  \frametitle{SPECFEM3D}

  \begin{itemize}
   \item Seismic wave propagation simulator
    \item SPECFEM3D ported to OpenCL using BOAST, and now HIP
    \begin{itemize}
      \item Unified code base (CUDA/OpenCL/HIP)
      \item Refactoring: kernel code base reduced by 40\%
      \item Similar performance on NVIDIA Hardware
      \item Non regression test for GPU kernels
    \end{itemize}
    \item On the Mont-Blanc prototype:
    \begin{itemize}
      \item OpenCL+MPI runs
      \item Speedup of 3 for the GPU version
    \end{itemize}
  \end{itemize}
\end{frame}

\subsection{Laplacian Filter}

\begin{frame}[fragile]
  \frametitle{Laplace Kernel Port to ARM Mali GPU - 1}
\lstset{style=BC}
\begin{lstlisting}
void laplace(const int width,
             const int height,
             const unsigned char src[height][width][3],
                   unsigned char dst[height][width][3]){
  for (int j = 1; j < height-1; j++) {
    for (int i = 1; i < width-1; i++) {
      for (int c = 0; c < 3; c++) {
        int tmp = -src[j-1][i-1][c] -   src[j-1][i][c] - src[j-1][i+1][c]\
                 - src[j  ][i-1][c] + 9*src[j  ][i][c] - src[j  ][i+1][c]\
                 - src[j+1][i-1][c] -   src[j+1][i][c] - src[j+1][i+1][c];
        dst[j][i][c] = (tmp < 0 ? 0 : (tmp > 255 ? 255 : tmp));
      }
    }
  }
}
\end{lstlisting}
\begin{itemize}
\item C reference implementation
\item Many possible optimization strategies in OpenCL
\item ARM GPU Mali 604 within the Mont-Blanc project
\end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Laplace Kernel Port to ARM Mali GPU - 2}
  Hand tuned code by specialist uses vectorization, temporary data-type change, and register swizzling to avoid loads
  \begin{columns}
  \column{0.55\textwidth}
  \begin{itemize}
    \item Hyperparameters used in BOAST:
    \begin{itemize}
      \item \emph{tile\_x}: a positive integer
      \item \emph{tile\_y}: a positive integer
      \item \emph{vector\_length}: 1, 2, 4, 8 or 16
      \item \emph{temporary\_size}: 2 or 4 bytes
      \item \emph{register\_swizzle}: true or false
    \end{itemize}
  \end{itemize}
  \column{0.45\textwidth}
  \begin{itemize}
    \item Optimal parameter values:
    \begin{itemize}
      \item tile\_x: 16
      \item tile\_y: 1
      \item vector\_length: 16
      \item temporary\_size: 2
      \item register\_swizzle: \emph{false}
    \end{itemize}
  \end{itemize}
  \end{columns}

  \begin{center}
  \scriptsize
  \begin{tabular}{c|r|r|c|r|c}
    Image Size  & Naive (s) & Best (s) & Acceleration & BOAST (s) & Acceleration \\
    \hline&&&&\\
  768 x 432   & 0.0107    & 0.00669  & x1.6         & 0.000639  & x16.7 \\
  2560 x 1600 & 0.0850    & 0.0137   & x6.2         & 0.00687   & x12.4 \\
  2048 x 2048 & 0.0865    & 0.0149   & x5.8         & 0.00715   & x12.1 \\
  5760 x 3240 & 0.382     & 0.0449   & x8.5         & 0.0325    & x11.8 \\
  7680 x 4320 & 0.680     & 0.0747   & x9.1         & 0.0581    & x11.7
  \end{tabular}
  \end{center}

\end{frame}

\begin{frame}
  \frametitle{Laplace Kernel Port to ARM Mali GPU - 3}
  \begin{center}
  \scriptsize
  \begin{tabular}{c|r|r|c|r|c}
    Image Size & BOAST ARM (s) & BOAST Intel & Ratio & BOAST NVIDIA & Ratio \\
    \hline&&&&\\
  768 x 432   & 0.000639     & 0.000222    & x2.9  & 0.0000715    & x8.9  \\
  2560 x 1600 & 0.00687      & 0.00222     & x3.1  & 0.000782     & x8.8  \\
  2048 x 2048 & 0.00715      & 0.00226     & x3.2  & 0.000799     & x8.9  \\
  5760 x 3240 & 0.0325       & 0.0108      & x3.0  & 0.00351      & x9.3  \\
  7680 x 4320 & 0.0581       & 0.0192      & x3.0  & 0.00623      & x9.3
  \end{tabular}
  \end{center}

  \begin{columns}
  \column{0.5\textwidth}
  \begin{itemize}
    \item \footnotesize Optimal parameter values Intel\\(I7 2760):
    \begin{itemize}
      \item \scriptsize tile\_x: 16
      \item \scriptsize tile\_y: 4..2
      \item \scriptsize vector\_length: 8
      \item \scriptsize temporary\_size: 2
      \item \scriptsize register\_swizzle: false
    \end{itemize}
  \end{itemize}
  \column{0.5\textwidth}
  \begin{itemize}
    \item \footnotesize Optimal parameter values nVidia\\(GTX 680):
    \begin{itemize}
      \item \scriptsize tile\_x: 4
      \item \scriptsize tile\_y: 4
      \item \scriptsize vector\_length: 4
      \item \scriptsize temporary\_size: 2
      \item \scriptsize register\_swizzle: false
    \end{itemize}
  \end{itemize}
  \end{columns}
  \vspace{1cm}
  \centering Performance \emph{portability} among several different architectures.
\end{frame}

\subsection{Gysela}

\begin{frame}
\frametitle{Gysela 2d avection}

\scalebox{1.}{
\begin{minipage}{\textwidth}
Gysela: Gyrokinetic Semi-Lagrangian\\
Tokamak plasma simulation for fusion (ITER)
\begin{itemize}
\item Chosen Auto-tuning parameters:
\begin{itemize}
\item directive based inlining / BOAST driven inlining
\item BOAST driven loop unrolling
\item C or Fortran code generated
\item scan versions of gfortran/gcc/icc/ifort
\item loop blocking parameter (one of the most internal loop)
\item explicit vectorization: BOAST generates INTEL intrinsincs, e.g.
\scalebox{.6}{
\begin{minipage}{1.5\textwidth}
\begin{alltt}
ftmp1 = \_mm256\_setzero\_pd( );\\
ftmp2 = \_mm256\_setzero\_pd( );\\
ftmp1 = \_mm256\_fmadd\_pd(  base1[0], \_mm256\_load\_pd( \&ftransp[(0) * (4)]);\\
ftmp2 = \_mm256\_fmadd\_pd( base1[0 + 1], \_mm256\_load\_pd( \&ftransp[(0 + 1)  * (4)] ), ftmp2 );
\end{alltt}
\end{minipage}}
\end{itemize}
\item Final result: Ruby code of 200 lines for the 2d advection kernel (vs 300 original FORTRAN)
\item between 1 min and 20 min for the {parameter scan} on 1 machine
\end{itemize}
\end{minipage}}

\end{frame}

\begin{frame}
\frametitle{Auto-tuning on INTEL Westmere (2011)}
\begin{columns}
\column{.3\textwidth}
\scalebox{0.76}{
\begin{minipage}{6cm}
Auto-tuning for 2D advection\\
Computing center at Marseille\\
12-cores node -\\
\hspace*{.4cm} Intel X5675, 3.07GHz\\[1cm]
Nb of runs in this scan: 609\\
Runs sorted from quickest to slowest\\
Result of the scan \\
\hspace*{.4cm}\ (best parameters):\\[.2cm]
\end{minipage}
}
\scalebox{0.7}{
\begin{minipage}{6cm}
\begin{alltt}
:lang: FORTRAN\\
:unroll: true\\
:force.inline: true\\
:intrinsic: false\\
:blocking.size: 4\\
:module: intel/16.0.2\\
\end{alltt}

Speedup: \textcolor{red}{1.9}
\end{minipage}
}
\column{.70\textwidth}
\includegraphics[width=1.0\textwidth]{rheticus.pdf}\\
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Auto-tuning on INTEL Sandy-Bridge (2012)}
\begin{columns}
\column{.3\textwidth}
\scalebox{0.8}{
\begin{minipage}{6cm}
Auto-tuning for 2D advection\\
Computing center at Orsay\\
16-cores node - \\
\hspace*{0.4cm}Intel E5-2670 v1, 2.60GHz\\[1cm]
Result of the scan \\
\hspace*{.4cm}\ (best parameters):\\[.2cm]
\end{minipage}
}
\scalebox{0.7}{
\begin{minipage}{6cm}
\begin{alltt}
:lang: FORTRAN\\
:unroll: \textcolor{red}{false}\\
:force.inline: \textcolor{red}{false}\\
:intrinsic: false\\
:blocking.size: 2\\
:module: intel/15.0.0
\end{alltt}

Speedup: \textcolor{red}{1.7}
\end{minipage}
}
\column{.70\textwidth}
\includegraphics[width=1.0\textwidth]{poincare.pdf}\\
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Auto-tuning on INTEL Haswell (2015)}
\begin{columns}
\column{.3\textwidth}
\scalebox{0.8}{
\begin{minipage}{6cm}
Auto-tuning for 2D advection\\
Computing center at Montpellier\\
24-cores node - \\
\hspace*{0.4cm}Intel E5-2690 v3, 2.60GHz\\[1cm]
Result of the scan \\
\hspace*{.4cm}\ (best parameters):\\[.2cm]
\end{minipage}
}
\scalebox{0.7}{
\begin{minipage}{6cm}
\begin{alltt}
:lang: FORTRAN\\
:unroll: true\\
:force.inline: true\\
:intrinsic: false\\
:blocking.size: 4\\
:module: \textcolor{red}{intel/14.0.4.211}\\
\end{alltt}

Speedup: \textcolor{red}{2.0}
\end{minipage}
}
\column{.70\textwidth}
\includegraphics[width=1.0\textwidth]{occigen.pdf}\\
\end{columns}
\end{frame}

\begin{frame}
\frametitle{Auto-tuning on INTEL KNL (Phi 2016)}
\begin{columns}
\column{.3\textwidth}
\scalebox{0.8}{
\begin{minipage}{6cm}
Auto-tuning for 2D advection\\
Computing center at Montpellier\\
64-cores node - \\
\hspace*{0.4cm}Intel 7210  1.30GHz\\[1cm]
Result of the scan \\
\hspace*{.4cm}\ (best parameters):\\[.2cm]
\end{minipage}
}
\scalebox{0.7}{
\begin{minipage}{6cm}
\begin{alltt}
:lang: FORTRAN\\
:unroll: true\\
:force.inline: true\\
:intrinsic: false\\
:blocking.size: \textcolor{red}{32}\\
:module: \textcolor{red}{intel/17.0}\\
\end{alltt}

Speedup: \textcolor{red}{3.6}
\end{minipage}
}
\column{.70\textwidth}
\includegraphics[width=1.0\textwidth]{knl.pdf}\\
\end{columns}
\end{frame}

\begin{comment}
\section{Real Applications}
\end{comment}

\section{Conclusion}

\begin{frame}
\frametitle{Conclusion}
\end{frame}

\begin{frame}
  \frametitle{Conclusion}
  \begin{itemize}
    \item A kernel tuning framework based on code transformation and generation proved:
    \begin{itemize}
      \item Flexible enough to cover a broad range of applications and targets,
      \item While still outperforming hand tuned or other compiler based optimization strategies,
      \item Maintainable as BOAST is still used in production after a decade.
    \end{itemize}
    \item But also useful for:
    \begin{itemize}
      \item Auto-tuning research,
      \item Architecture and compiler performance evaluation,
      \item Non-regression testing through kernel replay and validation.
    \end{itemize}
  \end{itemize}
  Hopefully, this approach (if not BOAST itself) will be picked up in the future and further generalized and improved.
\end{frame}

\appendix
\subsection{Bibliography}

\begin{frame}[shrink=50]
  \frametitle{Bibliography}
  \bibliography{../biblio/BOAST}
\end{frame}

\subsection{Acknowledgements}
\begin{frame}
  \frametitle{Acknowledgements}
  \begin{itemize}
    \item This project has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement Nos 288777, 610402, and EINFRA-676629
    \item The research leading to these results has received funding from the European Union’s Horizon 2020 Programme (2014-2020) and from Brazilian Ministry of Science, Technology and Innovation through Rede Nacional de Pesquisa (RNP) under the HPC4E project, grant agreement No 689772
    \item This research used resources of the Argonne Leadership Computing Facility, which is a DOE Office of Science User Facility supported under Contract DE-AC02-06CH11357.
  \end{itemize}
\end{frame}

\end{document}
